```{r setup_plates, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(parallel)
library(tidyverse)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Eschweilera_Paracou/"
```

```{r data}
nanodrop <- lapply(list.files(file.path(path, "NanoDrop", "Extraction"), 
                  full.names = T), read.delim2) %>% 
  bind_rows(.id = "Plate_extraction") %>% 
  mutate(Conc. = ifelse(Conc. < 0, 0, Conc.)) %>% 
  dplyr::rename(Position_extraction = Well,
                concentration = Conc.) %>% 
  mutate(Plate_extraction = as.numeric(Plate_extraction)) %>% 
  select(Plate_extraction, Position_extraction, concentration)
extraction <- googlesheets::gs_title("Parvicapture") %>% 
  googlesheets::gs_read("Extraction")
purification <- lapply(list.files(file.path(path, "NanoDrop", "Purification"), 
                  full.names = T), read.delim2) %>% 
  bind_rows(.id = "Plate_extraction") %>% 
  mutate(Conc. = ifelse(Conc. < 0, 0, Conc.)) %>% 
  dplyr::rename(Position_extraction = Well,
                Concentration_Purified = Conc.) %>% 
  mutate(Plate_extraction = as.numeric(Plate_extraction)) %>% 
  select(Plate_extraction, Position_extraction, Concentration_Purified)
qubit <- googlesheets::gs_title("Parvicapture") %>% 
  googlesheets::gs_read("Qubit")
```

# Extraction

After extraction, we prepared plates for library preparation. We proceed as follow :

1. __Dosage__: plate has been dosed with NanoDrop including sample repeats
1. __Reorganization__: sample repeats have been used to replace original sample if they included more DNA
1. __Fragments size assesment__: plates 1 & 5 have been used to test fragment sizes with electrophoresis
1. __Purification__: due too degraded DNA and undesired product as phenol, we puridied all samples
1. __Dosage__: purification resulting in variable concentration not linear with input, we dosed again all samples by NanoDrop and adjusted the estimate by Qubit as concentration were low
1. __Concentration__: concentration being low, we concentrated samples by SpeedVac to a minimum of 20 $ng. \mu L ^{-1}$
1. __Library test__: we tested library preparation on 2 samples
1. __Normalization__: we normalized library plates to an input of ca  100 $ng$ in 6 $\mu L$

## Dosage

NanoDrop was used to assess DNA concentration with $2 \mu L$ of samples (figure \@ref(fig:nanodrop)).

```{r nanodrop, fig.cap="Samples concentration."}
nanodrop %>% 
  mutate(row = substr(Position_extraction, 1, 1)) %>% 
  mutate(col = substr(Position_extraction, 2, 3)) %>% 
  group_by(Plate_extraction) %>% 
  ggplot(aes(y = factor(row, levels = rev(LETTERS[1:8])),
             x = factor(col, levels = 1:12))) + 
  geom_point(aes(col = log(concentration)), size = 4) +
  geom_text(aes(label = round(concentration)), 
            col = "white", size = 2.5) +
  facet_wrap(~ Plate_extraction, ncol = 2) +
  theme_bw() +
  labs(x=NULL, y = NULL)
```

## Reorganization

Over 42 repeated extractions, 36 had a better result (around 10 folds better). And samples that were not better extracted had a mean of 10 $ng. \mu L^{-1}$. We replaced succesfull repeat original by their repeated extractions and get rid of the others.

```{r repeats, eval=F}
success <- extraction %>% 
  left_join(nanodrop) %>% 
  group_by(IdGenetic) %>% 
  filter(n() > 1) %>% 
  ungroup() %>% 
  reshape2::dcast(IdGenetic ~ Repeat, value.var = "concentration") %>% 
  filter(`0` < `1`) %>% 
  select(IdGenetic) %>% 
  unlist()
extraction %>% 
  filter(IdGenetic %in% success) %>% 
  mutate(Position = paste0("P", Plate_extraction, "_", Position_extraction)) %>% 
  mutate(Repeat = ifelse(Repeat == 1, "Repeat", "Origin")) %>% 
  reshape2::dcast(IdGenetic ~ Repeat, value.var = "Position") %>% 
  select(IdGenetic, Repeat, Origin) %>% 
  kable(col.names = c("Id", "Repeat position", "Origin position (destination)"), 
        caption = "Succesfull repeats.")
```


## Fragments size assesment

Plates 1 & 5 have been used to test fragment sizes with electrophoresis. They globally had both high size DNA (band with high width) and degraded DNA (smir), see [electrophoresis presentation](https://drive.google.com/open?id=12MUcfVZ6PvAA3HxVi7g7Tkr-qUIv2E1jk5Edf45Ij7k).

## Purification

Due too degraded DNA and undesired product as phenol, we puridied all samples. We purified the 4.5 plates with PGTB Beads batch L at 1.8X.

```{r purification, fig.cap="Relation between samples concentration dosed by NanoDrop before and after purification for plate P1."}
model <- lm(Concentration_Purified ~ concentration, data = extraction %>% 
     left_join(nanodrop) %>% 
     left_join(purification) %>% 
       filter(Concentration_Purified < 1000) %>% 
       filter(Concentration_Purified > 1))
extraction %>% 
  left_join(nanodrop) %>% 
  left_join(purification) %>% 
  filter(Concentration_Purified < 1000) %>% 
  filter(Concentration_Purified > 1) %>% 
  ggplot(aes(concentration, Concentration_Purified, 
             col = Species, group = NA)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm") +
  ggtitle(paste("Concentration purified =", round(model$coefficients[1]), "+", 
                round(model$coefficients[2], 2), "* Concentration +/-", round(mean(abs(model$residuals))))) 
```

## Dosage

Purification resulting in variable concentration not linear with input (Figure \@ref(fig:purification)), we dosed again all samples by NanoDrop and adjusted the estimate by Qubit as concentration were low (Figure \@ref(fig:extractionQubit)).

```{r extractionQubit, fig.cap="Samples concentration after purification. Concentration was assessed by NanoDrop and ajusted by Qubit with the relation in the subplot."}
model <- lm(Concentration_Qubit ~ Concentration_Purified, data = extraction %>% 
              left_join(nanodrop) %>% 
              left_join(purification) %>% 
              left_join(qubit) %>% 
              filter(!is.na(Concentration_Qubit)))
p1 <- extraction %>% 
  left_join(nanodrop) %>% 
  left_join(purification) %>% 
  left_join(qubit) %>% 
  filter(!is.na(Concentration_Qubit)) %>% 
  ggplot(aes(Concentration_Purified, Concentration_Qubit, col = Species, group = NA)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_color_discrete(guide = "none") +
  scale_x_log10() +
  xlab("Concentration NanoDrop") +
  ylab("Concentration Qubit")
data <- extraction %>% 
  left_join(nanodrop) %>% 
  left_join(purification) %>% 
  left_join(qubit) %>% 
  mutate(Concentration_est = model$coefficients[1] + model$coefficients[2]*Concentration_Purified)
ggplot(data, aes(Concentration_est, fill = Species)) +
  geom_histogram(position = "dodge") +
  scale_x_log10() +
  xlab("Concentration estimated") +
  ggtitle(paste(nrow(filter(data, Concentration_est > 25)), "samples > 25 ng/uL"),
          paste("Concentration estimated =", round(model$coefficients[1]), "+", 
                round(model$coefficients[2], 2), "* Concentration NanoDrop +/-", round(mean(abs(model$residuals))))) +
  geom_vline(xintercept = 48, col = "red", linetype = "dashed") +
  annotation_custom(grob = ggplotGrob(p1), xmin = 1.8, xmax = 2.47, ymin = 10, ymax = 90)
```

## Concentration

Concentration being low, we concentrated samples by SpeedVac to a minimum of 25 $ng. \mu L ^{-1}$. It means 4 time reduction (from 39 $\mu L$ to 10 $\mu L$).

## Library test

We tested library preparation on 2 samples with 1 $\mu L$ input:

```{r libTestIV}
data %>% 
  mutate(Concentration_est = Concentration_est*4) %>% 
  top_n(2, Concentration_est) %>% 
  select(IdGenetic, Plate_extraction, Position_extraction, Genus, Species, Concentration_est) %>% 
   kable(col.names = c("Id", "Plate", "Position", "Genus", "Species", "Estimated concentration"), 
        caption = "Samples used for tests.")
```

## Normalization

We normalized library plates to an input of ca  100 $ng$ in 6 $\mu L$...
