```{r setup_capture, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(parallel)
library(tidyverse)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
path <- "../../../data/Eschweilera_Paracou"
```

# Capture

After library preparation, we prepared plates for capture. We proceed as follow :

1. __Dosage__: Obtained libraries were dosed by Qubit after range assessment by NanoDrop (16 libraries from P5)
1. __Repeats__: X libraries were repeated from extraction to replace least concentrated libraries
1. __Equimolar pools preparation__: Equimolar pools were built for capture
1. __Pools fragment size evaluation__: Pool fragents sizes were evalutaed by Bioanalyzer
1. __Pools fragment size selection__: Pool fragments size were filtered by Pippin between 400 and 600 bp
1. __Capture__: ...

## Dosage

Obtained libraries were dosed by Qubit after range assessment by NanoDrop (16 libraries from P5). Most of libraries had a concentration above 10 $ng.\mu L^{-1}$ and libraries had consistent degradation pattern with electrophoresis. Thus we decided to repeat the X libraries below X $ng.\mu L^{-1}$ and pooled every libraries in 4 equimolar pools to be size selected by Pippin.

```{r quantit, fig.cap="Library concentration measured by QuantIt."}
quantit1 <- read_csv(file.path(path, "QuantIt", 
                   "PARVICAPTURE_library_P1P2P5_02072019_fluo.csv"), skip = 53)[2:4] %>% 
  dplyr::rename(position = Well, fluo = `dosage_fluo_384_CB:501,527`) %>% 
  dplyr::select(position, fluo) %>% 
  filter(!is.na(fluo)) %>% 
  left_join(read_csv(file.path(path, "QuantIt", 
                               "PARVICAPTURE_library_P1P2P5_02072019_plates.csv"),  skip = 36) %>% 
              dplyr::rename(row = X1) %>% 
              reshape2::melt(id.vars = "row", variable.name = "col", value.name = "library") %>% 
              mutate(position = paste0(row, col)) %>% 
              select(position, library)) %>% 
  separate(library, c("Plate_library", "Position_library"), "_") %>% 
  mutate(Plate_library = as.numeric(gsub("P", "", Plate_library))) %>% 
  select(Plate_library, Position_library, fluo) %>% 
  arrange(Plate_library, Position_library) %>% 
  left_join(data.frame(Plate_library = 0,
                       Position_library = as.vector(sapply(1:6, function(x) paste0(LETTERS[1:4], x))),
                       concentration = rep(c(0, 10, 20, 40, 80, 120), each = 4)))
data1 <- filter(quantit1, Plate_library == 0)
quantit1 <- quantit1 %>% 
  mutate(concentration_est = lm(concentration ~ fluo, data = data1)$coefficients[2]*fluo)

ggplot(quantit1, aes(concentration_est, fill = concentration_est > 10)) +
  geom_histogram() +
  scale_x_log10() +
  xlab("Concentration") +
  ylim(0, 45) +
  annotation_custom(grob = ggplotGrob(ggplot(data1, aes(fluo, concentration)) +
                                        geom_point() +
                                        geom_smooth(method = "lm") +
                                        xlab("Fluorescence") +
                                        ylab("Concentration")), 
                    xmin = log10(10), xmax = log10(100), ymin = 25, ymax = 45)
```

## Repeats

X libraries were repeated from extraction to replace least concentrated libraries

## Equimolar pools preparation

Equimolar pools were built for capture

## Pools fragment size evaluation

Pool fragents sizes were evalutaed by Bioanalyzer

## Pools fragment size selection

Pool fragments size were filtered by Pippin between 400 and 600 bp

## Capture

...
