```{r setup_filtering, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(parallel)
library(tidyverse)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = F, cache.lazy = F)
pathCluster <- "~/Remotes/genotoul/work/PhD/data/Eschweilera_Paracou"
path <- "../../../data/Eschweilera_Paracou/"
```

```{r filtData}
individuals <- googlesheets::gs_title("Parvicapture") %>% 
  googlesheets::gs_read("Extraction") %>% 
  filter(!is.na(Seq)) %>% 
  dplyr::select(IdGenetic, Genus, Species, Plot, SubPlot, TreeFieldNum) %>% 
  unique() %>% 
  mutate(cat = recode(Species, "wachenheimii" = "other", "grandiflora_form2" = "other",
                      "simiorum" = "outgroup", "collina"  = "other", "persistens" = "outgroup",
                      "squamata" = "other", "pedicellata" = "other", "grandiflora" = "other",
                      "micrantha" = "other", "parviflora" = "other", 
                      "congestiflora" = "outgroup"))
```

# Filtering

In this chapter, we filtered raw snps with different automatic filters on libraries missingness, snps missingness, and outgroups.

## Library filtering

After investigating missing data per libraries count and associated number of SNPs (Fig. \@ref(fig:libmiss)), we decided to try 10 values of libraries filtering from 50 to 286  (Fig. \@ref(fig:libFilterPCA)).

```{r libmiss, fig.cap="Missing data per libraries count."}
libmiss <- read_delim(file.path(path, "Sequences", "denovo", "raw", "raw.imiss"), delim = " ") %>% 
  dplyr::rename_all(funs(gsub(" ", "", .))) %>% 
  mutate(missingness = as.numeric(F_MISS)) %>% 
  mutate(snps = (1-missingness)*704485) %>% 
  arrange(missingness) %>% 
  mutate(libraries = 1:nrow(.))
Nlibs <- round(seq(50, max(libmiss$libraries), length.out = 10))
cowplot::plot_grid(
  ggplot(libmiss, aes(missingness)) + 
    geom_histogram() +
    geom_vline(xintercept = libmiss$missingness[Nlibs], col = "red") +
    ggtitle("704 485 SNPs"),
  ggplot(libmiss, aes(libraries, missingness)) + geom_point() + geom_line()
  + geom_vline(xintercept = Nlibs, col = "red") 
  + geom_hline(yintercept = libmiss$missingness[Nlibs], col = "red"),
  ggplot(libmiss, aes(libraries, snps)) + geom_point() + geom_line()
  + geom_vline(xintercept = Nlibs, col = "red")  
  + geom_hline(yintercept = libmiss$snps[Nlibs], col = "red"),
  nrow = 1)
cat("N ibraries", Nlibs)
cat("--mind", libmiss$missingness[Nlibs])
```

```{bash libFilter, eval=F, echo=T}
minds=(0.9686 0.974 0.9782 0.9811 0.9831 0.9855 0.9873 0.9891 0.9905 0.9928)
plink=~/Tools/plink_linux_x86_64_20190617/plink 
for mind in "${minds[@]}" ; 
  do 
  mkdir filtered.$mind
  $plink \
    --bfile ../../denovo/raw/raw \
    --allow-extra-chr \
    --mind $mind \
    --missing --het --freqx --pca var-wts \
    --make-bed --out filtered.$mind/filtered
done
```

```{r libFilterPCA, fig.cap="PCA for SNPs obtained after filtering libraries based on missingness."}
lapply(list.files(file.path(path, "Sequences", "variants", "libraries_filtered")), function(f)
  read_delim(file.path(path, "Sequences", "variants", "libraries_filtered", f, "filtered.eigenvec"),
             delim = " ", col_names = c("X1", "Sample", paste0("PCA", 1:20))) %>% 
    mutate(filter = f)) %>% 
  bind_rows() %>% 
  separate(Sample, paste0("X", 1:7), "[:punct:]", remove = F) %>% 
  dplyr::rename(IdGenetic = X4, Lane = X5) %>% 
  left_join(individuals) %>% 
  ggplot(aes(x = PCA1, y = PCA2, col = cat, 
             label = Sample)) +
  geom_line(aes(group = IdGenetic)) +
  geom_point(size=2, alpha = 0.5) +
  scale_color_discrete("") +
  theme(legend.position = "bottom") +
  facet_wrap(~ filter, scales = "free")
```

## SNPs filtering

After investigating missing data per snps count and associated number of libraries (Fig. \@ref(fig:snpmiss)), we decided to try 2 values of snps filtering corresponding to at least 5 and 10 libraries represented (varying `--geno`), resulting in 20 combination of filtering (Fig. \@ref(fig:snpFilterPCA)).

```{r snpmiss, fig.cap="Missing data per snp count."}
names(Nlibs) <- list.files(file.path(path, "Sequences", "variants", "libraries_filtered"))
snpmiss <- lapply(list.files(file.path(path, "Sequences", "variants", "libraries_filtered")), function(f)
  read_delim(file.path(path, "Sequences", "variants", "libraries_filtered", f, "filtered.lmiss"), delim = " ") %>% 
    dplyr::rename_all(funs(gsub(" ", "", .))) %>% 
    mutate(missingness = as.numeric(F_MISS)) %>% 
    arrange(missingness) %>%
    mutate(Nlibs = Nlibs[f]) %>% 
    mutate(libraries = (1-missingness)*Nlibs) %>% 
    mutate(snps = 1:nrow(.)) %>% 
    mutate(filter = f) %>% 
    filter(snps %in% seq(1, 704485, length.out = 100))) %>% 
  bind_rows()
snpmissN <- lapply(list.files(file.path(path, "Sequences", "variants", "libraries_filtered")), function(f)
  read_delim(file.path(path, "Sequences", "variants", "libraries_filtered", f, "filtered.lmiss"), delim = " ") %>% 
    dplyr::rename_all(funs(gsub(" ", "", .))) %>% 
    mutate(missingness = as.numeric(F_MISS)) %>% 
    arrange(missingness) %>%
    mutate(Nlibs = Nlibs[f]) %>% 
    mutate(libraries = round((1-missingness)*Nlibs)) %>% 
    mutate(snps = 1:nrow(.)) %>% 
    mutate(filter = f) %>% 
    filter(libraries %in% c(5,10))) %>% 
  bind_rows()
snpmissN <- group_by(snpmissN, filter, libraries) %>% 
  summarise(snps = max(snps), missingness = max(missingness))
ggplot(snpmiss, aes(snps, libraries, col = filter)) + 
  geom_point() + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  geom_hline(yintercept = c(5, 10)) +
  geom_vline(aes(xintercept = snps, col = filter), linetype = "dotted", data = snpmissN)
cat("Nmin=5: --geno", filter(snpmissN, libraries == 5)$missingness)
cat("Nmin=10: --geno", filter(snpmissN, libraries == 10)$missingness)
```

```{bash snpFilter, eval=F, echo=T}
minds=(0.9686 0.974 0.9782 0.9811 0.9831 0.9855 0.9873 0.9891 0.9905 0.9928)
genos5=(0.898 0.9342 0.9505 0.9615 0.9677 0.9722 0.9756 0.9785 0.9807 0.9825)
genos10=(0.7959 0.8684 0.901 0.9231 0.9355 0.9444 0.9512 0.9571 0.9614 0.96498)
plink=~/Tools/plink_linux_x86_64_20190617/plink 
for i in $(seq 0 9) ; 
  do 
  mkdir filtered.${minds[$i]}.n5
  $plink \
    --bfile ../libraries_filtered/filtered.${minds[$i]}/filtered \
    --allow-extra-chr \
    --geno ${genos5[$i]} \
    --missing --het --freqx --pca var-wts \
    --make-bed --out filtered.${minds[$i]}.n5/filtered
done
for i in $(seq 0 9) ; 
  do 
  mkdir filtered.${minds[$i]}.n10
  $plink \
    --bfile ../libraries_filtered/filtered.${minds[$i]}/filtered \
    --allow-extra-chr \
    --geno ${genos10[$i]} \
    --missing --het --freqx --pca var-wts \
    --make-bed --out filtered.${minds[$i]}.n10/filtered
done
```

```{r snpFilterPCA, fig.cap="PCA for SNPs obtained after filtering libraries based on missingness."}
pcas <- lapply(list.files(file.path(path, "Sequences", "variants", "snps_filtered")), function(f)
  read_delim(file.path(path, "Sequences", "variants", "snps_filtered", f, "filtered.eigenvec"),
             delim = " ", col_names = c("X1", "Sample", paste0("PCA", 1:20))) %>% 
    mutate(filter = f)) %>% 
  bind_rows() %>% 
  separate(Sample, paste0("X", 1:7), "[:punct:]", remove = F) %>% 
  dplyr::rename(IdGenetic = X4, Lane = X5) %>% 
  left_join(individuals)
ggplot(pcas, aes(x = PCA1, y = PCA2, col = cat, 
                 label = Sample)) +
  geom_line(aes(group = IdGenetic)) +
  geom_point(size=2, alpha = 0.5) +
  scale_color_discrete("") +
  theme(legend.position = "bottom") +
  facet_wrap(~ filter, scales = "free") +
  theme(strip.text.x = element_text(size = 6))
```

## Outgroup filtering

We removed spurious PCA with no spatial clustering. Then, we decided to automatically removed outgroups with clustering (Fig. \@ref(fig:outgroupsDetection)), to avoid biasing the decision. And we ran new PCAs to determine population structure (Fig. \@ref(fig:outgroupFilterPCA)).

```{r outgroupsDetection, fig.cap="Outgroup detection in not spurious PCAs."}
pcasoutgroup <- group_by(pcas, filter) %>% 
  filter(sum(PCA1) != 0) %>% 
  cbind(group_by(pcas, filter) %>% 
          filter(sum(PCA1) != 0) %>% 
          dplyr::select(., PCA1, PCA2) %>% 
          do(hc = cutree(hclust(dist(as.matrix(.))), k = 2)) %>% 
          broom::tidy(hc)) %>% 
  dplyr::rename(cluster = x) %>% 
  group_by(filter, cluster) %>% 
  mutate(N = n()) %>%
  group_by(filter) %>% 
  mutate(Nmin = min(N)) %>% 
  mutate(cluster = ifelse(N == Nmin, "outgroup", "parvifolia")) %>% 
  dplyr::select(-N, -Nmin) %>% 
  ungroup()
ggplot(pcasoutgroup, aes(x = PCA1, y = PCA2, col = cluster, 
                 label = Sample)) +
      geom_line(aes(group = IdGenetic)) +
      geom_point(size=2, alpha = 0.5) +
      scale_color_discrete("") +
      theme(legend.position = "bottom") +
      facet_wrap(~ filter, scales = "free") +
      theme(strip.text.x = element_text(size = 6))
```

```{r outgroupFilterR, eval=F, echo=T}
lapply(unique(pcasoutgroup$filter), function(f){
  dir.create(file.path(path, "Sequences", "variants", "outgroup_filtered", f))
  filter(pcasoutgroup, filter == f) %>% 
    filter(cluster == "parvifolia") %>% 
    mutate(FID = 0) %>% 
    dplyr::select(FID, Sample) %>% 
    write_tsv(file.path(path, "Sequences", "variants", "outgroup_filtered", f, "keep.list"), col_names = F)
})
```

```{bash outgroupFilterSh, eval=F, echo=T}
plink=~/Tools/plink_linux_x86_64_20190617/plink 
for filter in $(ls ./) ; 
  do
  $plink \
    --bfile ../snps_filtered/$filter/filtered \
    --allow-extra-chr \
    --keep $filter/keep.list \
    --missing --het --freqx --pca var-wts \
    --make-bed --out $filter/filtered
done
```

```{r outgroupFilterPCA, fig.cap="PCA for SNPs obtained after filtering outgroups."}
lapply(list.files(file.path(path, "Sequences", "variants", "outgroup_filtered")), function(f)
  read_delim(file.path(path, "Sequences", "variants", "outgroup_filtered", f, "filtered.eigenvec"),
             delim = " ", col_names = c("X1", "Sample", paste0("PCA", 1:20))) %>% 
    mutate(filter = f)) %>% 
  bind_rows() %>% 
  separate(Sample, paste0("X", 1:7), "[:punct:]", remove = F) %>% 
  dplyr::rename(IdGenetic = X4, Lane = X5) %>% 
  left_join(individuals) %>% 
  ggplot(aes(x = PCA1, y = PCA2, col = cat, 
                 label = Sample)) +
  geom_line(aes(group = IdGenetic)) +
  geom_point(size=2, alpha = 0.5) +
  scale_color_discrete("") +
  theme(legend.position = "bottom") +
  facet_wrap(~ filter, scales = "free") +
  theme(strip.text.x = element_text(size = 6))
```
