```{r setupCompdist, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
library(parallel)
library(tidyverse)
library(ggfortify)
library(ggrepel)
library(raster)
library(leaflet)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F, eval=T)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
```

```{r treesCompdist}
d <- 20
trees <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(Plot %in% c(1, 6, 11, 13:16)) %>% # control and biodiv only
  filter(CensusYear == 2015) %>%
  filter(Xfield > d, Xfield < 250-d, Yfield > d, Yfield < 250-d) %>% 
  filter(Species != "Indet.") %>% 
  mutate(DBH = CircCorr/pi) %>% 
  collect()
```

```{r envCompdist}
# env <- trees %>%
#   dplyr::select(idTree, Xutm, Yutm) %>%
#   unique()
# coordinates(env) <- ~Xutm + Yutm
# proj4string(env) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
# env <- spTransform(env, CRSobj = crs)
# topo <- stack(
#   raster(file.path(path, "topography", "DEM_1m_2015.tif")),
#   raster(file.path(path, "topography", "RelativeElevation_1m.tif")),
#   raster(file.path(path, "topography", "slope_1m.tif")),
#   raster(file.path(path, "topography", "curvature_1m.tif")),
#   raster(file.path(path, "topography", "TRI_1m.tif")),
#   raster(file.path(path, "topography", "TWI_1m.tif"))
# )
# names(topo) <- c("DEM", "RelativeElevation", "Slope", "Curvature", "TRI", "TWI")
# topo <- projectRaster(topo, crs = crs)
# env <- data.frame(cbind(env@data, raster::extract(topo, env)))
# rm(topo)
# save(env, file = "./distribution_save/env.Rdata")
load("./distribution_save/env.Rdata")
```


```{r CompetitionCompdist}
# cl <- makeCluster(4)
# clusterExport(cl, list("data", "path"))
# Competition <- clusterMap(cl,
#   function(id, plot, census, x, y, genus){
#   library(dplyr)
#   src_sqlite(file.path(path, "trees/Paracou.sqlite")) %>%
#     tbl("Paracou") %>%
#     filter(CensusYear == census) %>%
#     filter(Plot == plot) %>%
#     filter(idTree != id) %>%
#     mutate(DBH = CircCorr/pi) %>%
#     filter(DBH >= 10) %>% # != INRA plots
#     mutate(dij = sqrt((x - Xutm)^2+(y - Yutm)^2)) %>%
#     filter(dij < 10) %>%
#     summarise(BA = sum(0.25*pi*DBH^2),
#               BAgenus =  sum(0.25*pi*DBH^2*as.numeric(Genus == genus))) %>%
#     mutate(idTree = id) %>%
#     dplyr::select(idTree, BA, BAgenus) %>%
#     collect()},
#   id = trees$idTree,
#   plot = trees$Plot,
#   x = trees$Xutm,
#   y = trees$Yutm,
#   census = trees$CensusYear,
#   genus = trees$Genus,
#   SIMPLIFY = F)
# stopCluster(cl)
# rm(cl)
# Competition <- bind_rows(Competition)
# save(Competition, file = "./distribution_save/Competition.Rdata")
load("./distribution_save/Competition.Rdata")
Competition <- Competition %>% 
  mutate(dij = ifelse(dij <1, 1, dij)) %>% 
  group_by(idTree) %>% 
  summarise(NCI = log(sum(DBHj^2*exp(-0.25*dij))+1))
```

```{r complexesCompdist}
Complexes <- bind_rows(
  data.frame(Complex = "Parvifolia", Genus = "Eschweilera",
             Species = c("pedicellata", "coriacea", "decolorans", "sagotiana",
                         "wachenheimii", "grandiflora_form2")),
  data.frame(Complex = "Licania", Genus = "Licania",
             Species = c("alba", "membranacea", "canescens", "micrantha",
                         "ovalifolia", "sprucei", "densiflora",
                         "laxiflora", "parvifructa")),
  data.frame(Complex = "Iryanthera", Genus = "Iryanthera",
             Species = c("hostmannii", "sagotiana")),
  data.frame(Complex = "Talisia", Genus = "Talisia",
             Species = c("hexaphylla", "praealta", "simaboides")),
  data.frame(Complex = "Symphonia", Genus = "Symphonia",
             Species = c("globulifera", "sp.1")))
```

```{r dataCompdist}
data <- trees %>% 
  left_join(env) %>% 
  left_join(Competition) %>% 
  left_join(Complexes) %>% 
  filter(!is.na(Plot)) %>% 
  filter(TWI >= 0) %>% 
  mutate(Complex = ifelse(is.na(Complex), "no", Complex))
```

```{r mdataCompdist}
complexes <- unique(Complexes$Complex)
mdata <- lapply(complexes, function(complex) {
  list(N = nrow(data),
       K = 2,
       Y = as.numeric(data$Complex == complex),
       X = dplyr::select(data, TWI, NCI) %>%
         as.matrix(),
       N_pred = 2*100,
       X_pred = data.frame(
         TWI = c(seq(min(data$TWI), 
                   max(data$TWI), length.out = 100), 
                 rep(mean(data$TWI), 100)),
         NCI = c(rep(mean(data$NCI), 100),
                 seq(min(data$NCI), 
                     max(data$NCI), length.out = 100))
       ) %>% as.matrix()
       )
})
names(mdata) <- complexes
save(mdata, file = "./distribution_save/complexesData.Rdata")
```

```{bash fitCompCluster, eval=F}
complexes=(Parvifolia Licania Iryanthera Talisia Symphonia)
for complex in "${complexes[@]}" ; do echo "module purge ; module load compiler/gcc-7.2.0 ; module load system/R-3.5.3 ; R_LIBS_USER=\" \" Rscript complex.R $complex" ; done > complex.sh
sarray -J complex -o out/%j.complex.out -e out/%j.complex.err -t 36:00:00 --constraint=broadwell --cpus-per-task=2 --mail-type=BEGIN,END,FAIL complex.sh
```

```{r fitComp}
fits <- list()
for(complex in complexes){
  load(paste0("./distribution_save/complexes", complex, ".Rdata"))
  fits[[complex]] <- fit
}
preds <- lapply(names(fits), function(complex)
  as.data.frame(fits[[complex]], pars = "Y_pred") %>% 
    reshape2::melt(NULL) %>% 
    group_by(variable) %>% 
    summarise(mean = mean(value), 
              q5 = quantile(value, 0.05), q95 = quantile(value, 0.95)) %>% 
    dplyr::select(-variable) %>% 
    cbind(data.frame(mdata[[complex]]$X_pred) %>% 
                mutate(variable = rep(c("TWI", "NCI"), each = 100))) %>% 
    mutate(complex = complex) %>% 
    filter(TWI <= max(filter(data, Complex == complex)$TWI),
           TWI >= min(filter(data, Complex == complex)$TWI),
           NCI <= max(filter(data, Complex == complex)$NCI),
           NCI >= min(filter(data, Complex == complex)$NCI))
  ) %>% bind_rows() %>% 
  mutate(value = ifelse(variable == "TWI", TWI, NCI))
```

# Complexes distributions

The aim of this document is to study the distribution at micro-environmental scale with biotic and abiotic environment of whole species complexes in Paracou (not species within complexes).

## Material and Methods

### Study site

Paracou

### Individuals

Ddefined in document [A04-Complexes](A04-Complexes.html):

* *E. charatcaea*
* *E. Parvifolia*
* *Licania*
* *Iryanthera*
* *Talisia*
* *Symphonia*

### Environmental variables

Defined in document [A03-Descriptors](A03-Descriptors.html):

* TWI
* $NCI$

### Analysis

We analysed species complex distribution by comparing species complex occurrence, considered as presence, against all occurrences of tree belonging to other species, considered as pseudo-absence. Species complex distribution has been investigating through logistic regression following a Bernoulli distribution (model with the best likelihood and results see document [A01-SingleModel](A01-SingleModel.html)):

$$Presence \sim \mathcal{Bernoulli}[logit^{-1}(\alpha + \beta*X+\gamma*X^2)]$$
with $X$ the matrix of environmental descriptors of size $[N,K]$, $\alpha$ the intercept, $\beta$ a vector of size $K$ representing the slope and $\gamma$ a vector of size $K$ representing the quadratic form.

Environmental descriptors were all scaled for the model inference, in order to ease model inference and later compare strength of effects between traits and between effects. A Bayesian method was used to infer parameters of the model regarding each leaf traits using `stan` language and `rstan` package in the R environment (see supplementary material [S1: Model] for model equations and `stan` code).

## Results

The model inference correctly converged without divergent transition and resulted in a correct predictions density distribution (see supplementary material [S2: Diagnostic]).

```{r summaryCompdist}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = c("alpha", "beta", "gamma"), 
                  droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "complex") %>% 
  mutate(descriptor = gsub("([[:alpha:]])", "", term)) %>% 
  dplyr::select(complex, term, descriptor, estimate) %>% 
  mutate(descriptor = gsub("([[:punct:]])", "", descriptor)) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "_{TWI}$", `2` = "_{NCI}$")) %>%
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(term = recode_factor(term, alpha = "$\\alpha$", 
                              beta = "$\\beta", 
                              gamma = "$\\gamma")) %>% 
  mutate(term = paste0(term, descriptor)) %>% 
  reshape2::dcast(complex ~ term, value.var = "estimate") %>% 
  kable(caption = "Model parameters for each complex with scaled descriptors.",
        escape = F, digits = 3, format = "pandoc") %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1)  
```


```{r posteriorComplex, fig.cap="Model parameters posterior and predictions.", fig.height=8, fig.width=8}
col <- RColorBrewer::brewer.pal(4, name = "Set1")
names(col) <- c("pp", "pn", "nn", "np")
annotation_custom2 <- function (grob, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, data) 
  layer(data = data, stat = StatIdentity, position = PositionIdentity, 
        geom = ggplot2:::GeomCustomAnn,
        inherit.aes = TRUE, params = list(grob = grob, 
                                          xmin = xmin, xmax = xmax, 
                                          ymin = ymin, ymax = ymax))
subg_pp <- ggplot(data.frame(x= c(0, 1), y = c(0,1)), aes(x, y)) +
  stat_function(fun = function(x) arm::invlogit(x + x^2), col = col[1]) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())
subg_pn <- ggplot(data.frame(x= c(0, 1), y = c(0,1)), aes(x, y)) +
  stat_function(fun = function(x) arm::invlogit(x - x^2), col = col[2]) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())
subg_nn <- ggplot(data.frame(x= c(0, 1), y = c(0,1)), aes(x, y)) +
  stat_function(fun = function(x) arm::invlogit(- x - x^2), col = col[3]) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())
subg_np <- ggplot(data.frame(x= c(0, 1), y = c(0,1)), aes(x, y)) +
  stat_function(fun = function(x) arm::invlogit(- x + x^2), col = col[4]) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())
g1 <- lapply(lapply(fits, as.array, pars = c("beta", "gamma")), 
       bayesplot::mcmc_intervals_data, prob_outer = 0.8) %>% 
  bind_rows(.id = "complex") %>% 
  mutate(descriptor = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(descriptor = gsub("([[:punct:]])", "", descriptor)) %>% 
  mutate(descriptor = recode_factor(descriptor, `2` = "NCI", `1` = "TWI")) %>%
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>%
  dplyr::select(complex, parameter, descriptor, ll, l, m, h, hh) %>% 
  reshape2::melt(id.vars = c("complex", "parameter", "descriptor")) %>% 
  mutate(parameter = paste0(parameter, "_", variable)) %>%
  reshape2::dcast(complex + descriptor ~ parameter) %>% 
  mutate(form = 10*as.numeric(beta_m > 0) + as.numeric(gamma_m > 0)) %>% 
  mutate(form = recode(form, "11" = "pp", "1" = "np", 
                       "10" = "pn", "0" = "nn")) %>% 
  ggplot(aes(label = complex, col = form)) +
  geom_hline(yintercept = 0, linetype = "dashed", col = "lightgrey") +
  geom_vline(xintercept = 0, linetype = "dashed", col = "lightgrey") +
  geom_point(aes(x = beta_m, y = gamma_m), alpha = 1) +
  geom_segment(aes(x = beta_ll, xend = beta_hh, 
                   y = gamma_m, yend = gamma_m), alpha = 0.3) +
  geom_segment(aes(x = beta_m, xend = beta_m, 
                   y = gamma_ll, yend = gamma_hh), alpha = 0.3) +
  geom_text_repel(aes(x = beta_m, y = gamma_m), 
                  size = 2.5, point.padding = 1) +
  facet_wrap(~ descriptor) +
  xlab(expression(beta)) + ylab(expression(gamma)) +
  scale_color_manual(guide = "none", values = col) +
  coord_equal() +
  annotation_custom2(ggplotGrob(subg_pp),
                     data = data.frame(complex = "Iryanthera",
                                       form = "pp",
                                       descriptor = "TWI"),
                     xmin = 0.5, xmax = 1.5, ymin = 0.5, ymax = 1.5) +
  annotation_custom2(ggplotGrob(subg_np),
                     data = data.frame(complex = "Iryanthera",
                                       form = "pp",
                                       descriptor = "NCI"),
                     xmin = -3, xmax = -2, ymin = 0.5, ymax = 1.5) +
  annotation_custom2(ggplotGrob(subg_pn),
                     data = data.frame(complex = "Iryanthera",
                                       form = "pp",
                                       descriptor = "TWI"),
                     xmin = 0.5, xmax = 1.5, ymin = -2.4, ymax = -1.4) +
  annotation_custom2(ggplotGrob(subg_nn),
                     data = data.frame(complex = "Iryanthera",
                                       form = "pp",
                                       descriptor = "NCI"),
                     xmin = -3, xmax = -2, ymin = -2.4, ymax = -1.4)
g2 <- preds %>% 
  left_join(group_by(data, Complex) %>% 
              summarise(Yo = n()/nrow(.)),
            by = c("complex" = "Complex")) %>% 
  ggplot(aes(x = value, col = complex)) +
  geom_ribbon(aes(ymin = q5, ymax = q95), alpha = 0.2) +
  geom_hline(aes(yintercept = Yo, col = complex), lty = "dashed") +
  geom_line(aes(y = mean)) +
  facet_wrap(~ variable, nrow = 1, scales = "free_x") + 
  xlab("") + ylab("Species complex relative abundance") +
  scale_y_sqrt() +
  scale_color_discrete("") +
  theme(legend.position = "bottom",
        strip.background = element_blank(),
        strip.text.x = element_blank())
cowplot::plot_grid(g1, g2, nrow = 2,
                   labels = LETTERS[1:2], label_x = 0.1)
```

## Supplementray materials

### S1: Diagnostic

Model inference converged ($\hat R = 1$, figure \@ref(fig:mcmc)) for every complex without divergent transition, and relatively uncorrelated parameters.

```{r mcmc_compdist, fig.cap="Markov chains trace plot after warmup for leaf thickness model."}
# lapply(fits, check_divergences)
mcmc_trace(as.array(fits$Parvifolia, pars = c("beta", "gamma", "lp__")),
             facet_args = list(labeller = label_parsed))
# mcmc_pairs(as.array(fits$invSLA), regex_pars = pars)
```
