```{r setup_species2, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
library(parallel)
library(tidyverse)
library(ggfortify)
library(raster)
library(leaflet)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 10, fig.width = 10,
  cache = T, cache.lazy = F, eval=F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
```

```{r trees_species2}
d <- 20
trees <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(Plot %in% c(1, 6, 11, 13:16)) %>% # control and biodiv only
  filter(CensusYear == 2015) %>%
  filter(Xfield > d, Xfield < 250-d, Yfield > d, Yfield < 250-d) %>% 
  filter(Species != "Indet.") %>% 
  mutate(DBH = CircCorr/pi) %>% 
  collect()
```

```{r env_species2}
# env <- trees %>%
#   dplyr::select(idTree, Xutm, Yutm) %>%
#   unique()
# coordinates(env) <- ~Xutm + Yutm
# proj4string(env) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
# env <- spTransform(env, CRSobj = crs)
# topo <- stack(
#   raster(file.path(path, "topography", "DEM_1m_2015.tif")),
#   raster(file.path(path, "topography", "RelativeElevation_1m.tif")),
#   raster(file.path(path, "topography", "slope_1m.tif")),
#   raster(file.path(path, "topography", "curvature_1m.tif")),
#   raster(file.path(path, "topography", "TRI_1m.tif")),
#   raster(file.path(path, "topography", "TWI_1m.tif"))
# )
# names(topo) <- c("DEM", "RelativeElevation", "Slope", "Curvature", "TRI", "TWI")
# topo <- projectRaster(topo, crs = crs)
# env <- data.frame(cbind(env@data, raster::extract(topo, env)))
# rm(topo)
# save(env, file = "./distribution_save/env.Rdata")
load("./distribution_save/env.Rdata")
```

```{r Competition_species2}
# cl <- makeCluster(4)
# clusterExport(cl, list("data", "path"))
# Competition <- clusterMap(cl,
#   function(id, plot, census, x, y, genus){
#   library(dplyr)
#   src_sqlite(file.path(path, "trees/Paracou.sqlite")) %>%
#     tbl("Paracou") %>%
#     filter(CensusYear == census) %>%
#     filter(Plot == plot) %>%
#     filter(idTree != id) %>%
#     mutate(DBH = CircCorr/pi) %>%
#     filter(DBH >= 10) %>% # != INRA plots
#     mutate(dij = sqrt((x - Xutm)^2+(y - Yutm)^2)) %>%
#     filter(dij < 10) %>%
#     summarise(BA = sum(0.25*pi*DBH^2),
#               BAgenus =  sum(0.25*pi*DBH^2*as.numeric(Genus == genus))) %>%
#     mutate(idTree = id) %>%
#     dplyr::select(idTree, BA, BAgenus) %>%
#     collect()},
#   id = trees$idTree,
#   plot = trees$Plot,
#   x = trees$Xutm,
#   y = trees$Yutm,
#   census = trees$CensusYear,
#   genus = trees$Genus,
#   SIMPLIFY = F)
# stopCluster(cl)
# rm(cl)
# Competition <- bind_rows(Competition)
# save(Competition, file = "./distribution_save/Competition.Rdata")
load("./distribution_save/Competition.Rdata")
Competition <- Competition %>% 
  mutate(dij = ifelse(dij <1, 1, dij)) %>% 
  group_by(idTree) %>% 
  summarise(NCI = log(sum(DBHj^2*exp(-0.25*dij))+1))
```

```{r complexes_species2}
Complexes <- bind_rows(
  data.frame(Complex = "Parvifolia", Genus = "Eschweilera",
             Species = c("pedicellata", "coriacea", "decolorans", "sagotiana",
                         "wachenheimii", "grandiflora_form2")),
  data.frame(Complex = "Licania", Genus = "Licania",
             Species = c("alba", "membranacea", "canescens", "micrantha",
                         "ovalifolia", "sprucei", "densiflora",
                         "laxiflora", "parvifructa")),
  data.frame(Complex = "Iryanthera", Genus = "Iryanthera",
             Species = c("hostmannii", "sagotiana")),
  data.frame(Complex = "Talisia", Genus = "Talisia",
             Species = c("hexaphylla", "praealta", "simaboides")),
  data.frame(Complex = "Symphonia", Genus = "Symphonia",
             Species = c("globulifera", "sp.1")))
```

```{r data_species2}
data <- trees %>% 
  left_join(env) %>% 
  left_join(Complexes) %>% 
  left_join(Competition) %>% 
  filter(!is.na(Plot)) %>% 
  filter(!is.na(Complex)) %>% 
  group_by(Complex, Species) %>% 
  filter(n() > 10) %>%
  ungroup()
```

```{r mdata_species2}
complexes <- unique(Complexes$Complex)
mdata <- lapply(complexes, function(complex) {
  data<- filter(data, Complex == complex)
  list(N = nrow(data),
       S = length(unique(data$Species)),
       K = 2,
       Y = sapply(levels(as.factor(data$Species)),
                  function(sp) as.numeric(data$Species == sp)),
       X = dplyr::select(data, TWI, NCI) %>%
         mutate_all(funs(scale)) %>%
         as.matrix())
})
names(mdata) <- complexes
```

```{r fit_species2}
Model <- stan_model("./distribution_models/JointModel.stan")
fits <- lapply(mdata, function(data) sampling(Model, chains = 2, data = data))
names(fits) <- complexes
save(fits, file = "./distribution_save/speciesUnlogged2.Rdata")
load("./distribution_save/speciesUnlogged2.Rdata")
pars <- c("alpha", "beta", "gamma")
species <- lapply(mdata, function(data) 
  data.frame(species = 1:length(colnames(data$Y)),
             Species = colnames(data$Y))
  ) %>% bind_rows(.id = "complex") %>% 
  mutate(SpeciesLong = paste(substr(complex, 1, 1), Species))
```

```{r predSpecies2}
predSpecies <- lapply(complexes, function(complex){

  N <- 25
  TWI <- seq(from = min(mdata[[complex]]$X[,1]), 
             to = max(mdata[[complex]]$X[,1]), length.out = N)
  NCI <- seq(from = min(mdata[[complex]]$X[,2]), 
             to = max(mdata[[complex]]$X[,2]), length.out = N)
  
  S <- mdata[[complex]]$S
  K <- mdata[[complex]]$K
  alpha <- as.matrix(fits[[complex]], pars = "alpha")
  beta0 <- as.matrix(fits[[complex]], pars = "beta")
  gamma0 <- as.matrix(fits[[complex]], pars = "gamma")
  beta <- array(dim = c(nrow(alpha), S, K))
  gamma <- array(dim = c(nrow(alpha), S, K))
  for(k in 1:K){
    beta[,,k] <- beta0[,seq.int(1, 3*S, S)[k]:(seq.int(1, 3*S, S)+(S-1))[k]]
    gamma[,,k] <- gamma0[,seq.int(1, 3*S, S)[k]:(seq.int(1, 3*S, S)+(S-1))[k]]
  }
    
  Yp <- abind::abind(sapply(1:length(TWI), function(i) 
    sapply(1: nrow(alpha), function(n)
      tsensembler:::softmax(alpha[n,] + beta[n,,1] * TWI[i] + beta[n,,2]*NCI[i] + gamma[n,,1]*TWI[i]^2 + gamma[n,,2]*NCI[i]^2)
    ), simplify = F), along = 3)
  Ytwi <- abind::abind(sapply(1:length(TWI), function(i) 
    sapply(1: nrow(alpha), function(n)
      tsensembler:::softmax(alpha[n,] + beta[n,,1]*TWI[i] + beta[n,,2]*mean(NCI) + gamma[n,,1]*TWI[i]^2 + gamma[n,,2]*mean(NCI)^2)
    ), simplify = F), along = 3)
  Ynci <- abind::abind(sapply(1:length(TWI), function(i) 
    sapply(1: nrow(alpha), function(n)
      tsensembler:::softmax(alpha[n,] + beta[n,,1]*mean(TWI) + beta[n,,2]*NCI[i] + gamma[n,,1]*mean(TWI)^2 + gamma[n,,2]*NCI[i]^2)
    ), simplify = F), along = 3)

  bind_rows(
    lapply(1:S, function(sp)
      data.frame(
        complex = complex,
        species = colnames(mdata[[complex]]$Y)[sp],
        variable = "TWI",
        value = TWI,
        Yp = apply(Yp[sp,,], 2, mean),
        Y = apply(Ytwi[sp,,], 2, mean),
        Y5 = apply(Ytwi[sp,,], 2, quantile, probs = 0.05),
        Y95 = apply(Ytwi[sp,,], 2, quantile, probs = 0.95))) %>% 
      bind_rows(),
    lapply(1:S, function(sp)
      data.frame(
        complex = complex,
        species = colnames(mdata[[complex]]$Y)[sp],
        variable = "NCI",
        value = NCI,
        Yp = apply(Ynci[sp,,], 2, mean),
        Y = apply(Ynci[sp,,], 2, mean),
        Y5 = apply(Ynci[sp,,], 2, quantile, probs = 0.05),
        Y95 = apply(Ynci[sp,,], 2, quantile, probs = 0.95))) %>% 
      bind_rows())
}) %>% bind_rows()
```
# Species distributions

The aim of this document is to study the distribution at micro-environmental scale with biotic and abiotic environment of species within species complexes in Paracou (not the whole species complex).

## Material and Methods

### Study site

Paracou

### Individuals

Defined in document [A04-Complexes](A04-Complexes.html):

* *E. charatcaea*
* *E. Parvifolia*
* *Licania*
* *Iryanthera*
* *Talisia*
* *Symphonia*

### Environmental variables

Defined in document [A03-Descriptors](A03-Descriptors.html):

* TWI
* $NCI_{hetero}$
* $NCI_{con}$

### Analysis

We analysed species distribution within species complex by comparing to which species belonged occurrences of the species complex. Species distribution within species complex has been investigating through softmax regression following a conjugated Dirchlet and Multinomial distribution (see document [A02-JointModel](A01-JointModel.html)):

$$Presence_{species} \sim \mathcal{Dirichlet~Multinomial}(softmax(\alpha + \beta*X + \gamma*X^2))$$
$${Log~Probability~Mass~Function}_{Dirichlet~Multinomial}(y|\alpha) = \Gamma(\sum \alpha)  + \sum(\Gamma(\alpha + y)) - \Gamma(\sum \alpha+\sum y) - \sum\Gamma(\alpha)$$
with $X$ the matrix of environmental descriptors of size $[N,K]$, $\alpha$ the intercept, $\beta$ a vector of size $K$ representing the slope and $\gamma$ a vector of size $K$ representing the quadratic form.

Environmental descriptors were all scaled for the model inference, in order to ease model inference and later compare strength of effects between traits and between effects. A Bayesian method was used to infer parameters of the model regarding each leaf traits using `stan` language [ref] and `rstan` package [ref] in the R environment [ref] (see supplementary material [S1: Model] for model equations and `stan` code).

## Results

```{r summary_species2}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = c(pars), droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "complex") %>% 
  mutate(index = gsub("([[:alpha:]])", "", term)) %>% 
  mutate(index = gsub("([[:punct:]])", "", index)) %>% 
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(species = as.numeric(substr(index, 1, 1))) %>% 
  left_join(species) %>% 
  mutate(descriptor = substr(index, 2, 2)) %>% 
  dplyr::select(complex, term, SpeciesLong, descriptor, estimate) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "TWI", `2` = "$NCI")) %>% 
  mutate(term = recode_factor(term, alpha = "$\\alpha$", 
                              beta = "$\\beta$", 
                              gamma = "$\\gamma$")) %>%
  reshape2::dcast(term + descriptor ~ SpeciesLong, value.var = "estimate") %>%
  kable(caption = "Model parameters for each complex with scaled descriptors.",
        escape = F, digits = 3, format = "pandoc") %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:3)
```

```{r posterior_species2, fig.cap="Model parameters posterior.", fig.width=10, fig.height=14}
lapply(lapply(fits, as.array, pars = "beta"), 
       bayesplot::mcmc_intervals_data) %>% 
  bind_rows(.id = "complex") %>% 
  mutate(descriptor = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(descriptor = gsub("([[:punct:]])", "", descriptor)) %>% 
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = as.numeric(substr(descriptor, 1, 1))) %>% 
  left_join(species) %>% 
  mutate(complex = gsub("E. ", "", complex)) %>% 
  mutate(descriptor = substr(descriptor, 2, 2)) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "TWI", `2` = "NCI")) %>% 
  dplyr::select(complex, parameter, descriptor, SpeciesLong, ll, l, m, h, hh) %>% 
  filter(parameter != "alpha") %>% 
  mutate(parameter = paste0(parameter, "[", descriptor, "]")) %>% 
  ggplot(aes(x = SpeciesLong,
             xend = SpeciesLong, 
             col = SpeciesLong, fill = SpeciesLong)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_grid(complex ~ parameter, labeller = label_parsed, 
             scales = "free") +
  xlab("") + ylab("") + 
  scale_fill_discrete(guide = "none") +
  scale_color_discrete(guide = "none") 
```

```{r SpeciesDistribution2, fig.cap="Relative abundance of species within species complexes. Predicted relative abundance of species within species complexes with dashed line indicating observed relative abundance of species within species complex in Paracou, and solid line and area representing respectively the mean and the 95% confidence interval of projected relative abundance of species complexes depending on descriptors. The color indicates the species complex.", fig.height=10}
predSpecies %>%
    group_by(complex) %>% 
  left_join(data %>% 
              group_by(Complex) %>% 
              mutate(N_complex = n()) %>% 
              group_by(Complex, N_complex, Genus, Species) %>% 
              summarise(N_species = n()) %>% 
              mutate(RelAbund = N_species / N_complex),
            by = c("complex" = "Complex", "species" = "Species")) %>% 
  mutate(species = ifelse(species == "grandiflora_form2", "grandiflora", species)) %>%
  mutate(Species = paste0(substr(Genus, 1,1), ". ", species)) %>% 
  ggplot(data = ., aes(x = value, col = Species)) + 
  geom_ribbon(aes(ymin = Y5, ymax = Y95), alpha = 0.2) +
  geom_line(aes(y = Y), lwd = 1.3) +
  geom_hline(aes(yintercept = RelAbund, col = Species), linetype = "dashed") +
  facet_grid(complex ~ variable, scales = "free") +
  guides(col = guide_legend(nrow=2)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  ylab("Species relative abundance") + xlab("Standardized value") +
  scale_y_sqrt() +
  scale_color_manual(values = c(RColorBrewer::brewer.pal(6, "Dark2")[1:6], 
                       RColorBrewer::brewer.pal(3, "Dark2")[1:2],
                       RColorBrewer::brewer.pal(8, "Dark2")[1:8],
                       RColorBrewer::brewer.pal(3, "Dark2")[1:2],
                       RColorBrewer::brewer.pal(3, "Dark2")[1:3]))
```
