---
title: "Model form"
date: '`r Sys.Date()`'
author: Sylvain Schmitt
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
vertical_layout: scroll
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
```

Intro
==================

Issue {data-width=200}
-------------------------------------

The aim of this document is to find the best general form of model to fit a distribution model. Especially we are thinking for a model that can deal with three specific cases :

* The environmental variable has no effect
* The niche optimum is in a intermediate value of the environmental variable range
* The niche optimum is at a limit of the environmental variable range

For this we will build three fake data sets distributed along our environmental variable as follow :

#### Models summary

```{r ModelsTable, render="pandoc"}
data.frame(
  Type = c("Bernoulli"),
  Name = c("$B_0$", "$B_{\\alpha}$", "$B_{\\alpha, \\alpha_2}$", "$B_{\\alpha, \\beta}$"),
  Formula = c("$Presence \\sim \\mathcal{Bernoulli}(\\theta)$",
              "$Presence \\sim \\mathcal{Bernoulli}(logit(\\alpha_0+\\alpha*Environment))$",
              "$Presence \\sim \\mathcal{Bernoulli}(logit(\\alpha*Environment+\\alpha_2*Environment^2))$",
              "$Presence \\sim \\mathcal{Bernoulli}(logit(\\alpha*Environment+Environment^{\\beta}))$")  
) %>% 
  kable(caption = "Models summary.",
        format = "html", escape = F) %>% 
  kable_styling("striped", full_width = F) %>% 
  collapse_rows(1)
```


Data {data-width=200}
-------------------------------------

#### Fake data

```{r fakeData}
n <- 100
data <- list(
  no = data.frame(Environment = seq(0, 1, length.out = n),
                  Presence = c(rep(0:1, n))),
  intermediate = data.frame(Environment = seq(0, 1, length.out = n),
                            Presence = c(rep(0,round((n-round(n/3))/2)), 
                                         rep(1,round(n/3)), 
                                         rep(0,(n-round((n-round(n/3))/2)-round(n/3))))),
  limit = data.frame(Environment = seq(0, 1, length.out = n), 
                     Presence = c(rep(0,(n-round(n/3))), rep(1,round(n/3))))
)
mdata <- lapply(data, function(x) list(N = nrow(x),
                                       Presence = x$Presence,
                                       Environment = x$Environment,
                                       weights = ifelse(x$Presence == 0,
                                                        1/(2*sum(x$Presence == 1)),
                                                        1/(2*sum(x$Presence == 0)))))
bind_rows(data, .id = "id") %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  facet_wrap(~id, nrow = 3)
```

$B_0$
==================

Column 1
-------------------------------------

#### Model

The first family of model are Bernoulli law to model the probability of presence. $B_0$ is the null model estimating a single theta along the whole environmental gradient.

$$Presence \sim \mathcal{Bernoulli}(\theta)$$

#### Code

```{stan , output.var="Model", echo=T, eval=F}
  data {
    int<lower=1> N ; // # obs
    int<lower=0, upper=1> Presence[N] ;
    vector<lower=0, upper=1>[N] weights ;
  }
  parameters {
    real<lower=0, upper=1> theta ;
  }
  model {
    target += weights*bernoulli_lpmf(Presence | theta) ;
  }
```

#### Summary

```{r }
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- names(mdata)
# save(fits, file = "./distribution_save/B0.Rdata")
load("./distribution_save/B0.Rdata")
pars <- c("theta")
lapply(as.list(names(mdata)), function(type)
  broom::tidyMCMC(fits[[type]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
    mutate(Type = type)) %>%
  bind_rows() %>%
  kable()
```

#### Posteriors

```{r , fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(names(data), function(type)
  mcmc_intervals(as.array(fits[[type]]), pars = pars)),
  labels = names(data), nrow = 2)
```

Column 2
-------------------------------------

#### Chains

```{r , fig.cap="Markov chains trace plot after warmup for intermediate data."}
mcmc_trace(as.array(fits$intermediate), pars = c(pars, "lp__"),
             facet_args = list(labeller = label_parsed))
```

#### Pairs

```{r , fig.cap="Markov chains pairs plot after warmup  for intermediate data."}
mcmc_pairs(as.array(fits$intermediate), pars = c(pars, "lp__"))
```

#### Predictions

```{r , fig.cap="Predictions."}
lapply(as.list(names(data)), function(type)
  cbind(type = type, data[[type]],
        mu = apply(as.matrix(fits[[type]], pars = "theta"), 2, mean),
        t(apply(as.matrix(fits[[type]], pars = "theta"), 2, quantile, probs = c(0.05, 0.95))))) %>% 
  bind_rows() %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'red', alpha = 0.2) +
  geom_line(aes(y = `5%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  facet_wrap(~ type, scales = "free", nrow = 3)
```

$B_{\alpha}$
==================

Column 1
-------------------------------------

#### Model

$B_{\alpha}$ is a Bernoulli logistic regression estimating a sigmoidal incresing or decreasing $\alpha$ along the whole environmental gradient in addition to the intercept $\alpha_0$.

$$Presence \sim \mathcal{Bernoulli}(logit(\alpha_0+\alpha*Environment))$$

#### Code

```{stan , output.var="Model", echo=T, eval=F}
  data {
    int<lower=1> N ; // # obs
    int<lower=0, upper=1> Presence[N] ;
    vector[N] Environment ;
    vector<lower=0, upper=1>[N] weights ;
  }
  parameters {
    real alpha_0 ; // Intercept
    real alpha ; // sigmoidal slope
  }
  model {
    target += weights*bernoulli_logit_lpmf(Presence | alpha_0 + alpha*Environment) ;
  }
  generated quantities {
    vector<lower=0, upper=1>[N] mu ;
    mu = inv_logit(alpha_0 + alpha*Environment) ;
  }
```

#### Summary

```{r }
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- names(mdata)
# save(fits, file = "./distribution_save/Balpha.Rdata")
load("./distribution_save/Balpha.Rdata")
pars <- c("alpha_0", "alpha")
lapply(as.list(names(mdata)), function(type)
  broom::tidyMCMC(fits[[type]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
    mutate(Type = type)) %>%
  bind_rows() %>%
  kable()
```

#### Posteriors

```{r , fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(names(data), function(type)
  mcmc_intervals(as.array(fits[[type]]), pars = pars)),
  labels = names(data), nrow = 2)
```

Column 2
-------------------------------------

#### Chains

```{r , fig.cap="Markov chains trace plot after warmup for intermediate data."}
mcmc_trace(as.array(fits$intermediate), pars = c(pars, "lp__"),
             facet_args = list(labeller = label_parsed))
```

#### Pairs

```{r , fig.cap="Markov chains pairs plot after warmup  for intermediate data."}
mcmc_pairs(as.array(fits$intermediate), pars = c(pars, "lp__"))
```

#### Predictions

```{r , fig.cap="Predictions."}
lapply(as.list(names(data)), function(type)
  cbind(type = type, data[[type]],
        mu = apply(as.matrix(fits[[type]], pars = "mu"), 2, mean),
        t(apply(as.matrix(fits[[type]], pars = "mu"), 2, quantile, probs = c(0.05, 0.95))))) %>% 
  bind_rows() %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'red', alpha = 0.2) +
  geom_line(aes(y = `5%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  facet_wrap(~ type, scales = "free", nrow = 3)
```

$B_{\alpha, \alpha_2}$
==================

Column 1
-------------------------------------

#### Model

$B_{\alpha, \alpha_2}$is a Bernoulli logistic regression allowing a quadratic form in addition to sigmoids with the $\alpha_2$ parameter.

$$Presence \sim \mathcal{Bernoulli}(logit(\alpha*Environment+\alpha_2*Environment^2))$$

#### Code

```{stan , output.var="Model", echo=T, eval=F}
  data {
    int<lower=1> N ; // # obs
    int<lower=0, upper=1> Presence[N] ;
    vector[N] Environment ;
    vector<lower=0, upper=1>[N] weights ;
  }
  parameters {
    real alpha ; // sigmoidal slope
    real alpha_2 ; // quadratic form
  }
  model {
    target += weights*bernoulli_logit_lpmf(Presence | alpha*Environment + alpha_2 * Environment .* Environment) ;
  }
  generated quantities {
    vector<lower=0, upper=1>[N] theta ;
    theta = inv_logit(alpha * Environment  + alpha_2 * Environment .* Environment) ;
  }
```

#### Summary

```{r }
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- names(mdata)
# save(fits, file = "./distribution_save/Balpha2.Rdata")
load("./distribution_save/Balpha2.Rdata")
pars <- c("alpha", "alpha_2")
lapply(as.list(names(mdata)), function(type)
  broom::tidyMCMC(fits[[type]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
    mutate(Type = type)) %>%
  bind_rows() %>%
  kable()
```

#### Posteriors

```{r , fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(names(data), function(type)
  mcmc_intervals(as.array(fits[[type]]), pars = pars)),
  labels = names(data), nrow = 2)
```

Column 2
-------------------------------------

#### Chains

```{r , fig.cap="Markov chains trace plot after warmup for intermediate data."}
mcmc_trace(as.array(fits$intermediate), pars = c(pars, "lp__"),
             facet_args = list(labeller = label_parsed))
```

#### Pairs

```{r , fig.cap="Markov chains pairs plot after warmup  for intermediate data."}
mcmc_pairs(as.array(fits$intermediate), pars = c(pars, "lp__"))
```

#### Predictions

```{r , fig.cap="Predictions."}
lapply(as.list(names(data)), function(type)
  cbind(type = type, data[[type]],
        mu = apply(as.matrix(fits[[type]], pars = "theta"), 2, mean),
        t(apply(as.matrix(fits[[type]], pars = "theta"), 2, quantile, probs = c(0.05, 0.95))))) %>% 
  bind_rows() %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'red', alpha = 0.2) +
  geom_line(aes(y = `5%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  facet_wrap(~ type, scales = "free", nrow = 3)
```

$B_{\alpha, \beta}$
==================

Column 1
-------------------------------------

#### Model

$B_{\alpha, \beta}$ is a Bernoulli logistic regression allowing a polynomial form in addition to sigmoids with the $\beta$ power parameter on the environmental variable.

$$Presence \sim \mathcal{Bernoulli}(logit(\alpha*Environment+Environment^{\beta}))$$

#### Code

```{stan , output.var="Model", echo=T, eval=F}
  data {
    int<lower=1> N ; // # obs
    int<lower=0, upper=1> Presence[N] ;
    vector[N] Environment ;
    vector<lower=0, upper=1>[N] weights ;
  }
  parameters {
    real alpha_0 ; // Intercept
    real alpha ; // sigmoidal slope
    real<lower=0, upper=4> beta ; // power form
  }
  model {
    for(n in 1:N)
      target += weights*bernoulli_logit_lpmf(Presence[n] | alpha*Environment[n] + pow(Environment[n], beta)) ;
  }
  generated quantities {
    vector<lower=0, upper=1>[N] theta ;
    for(n in 1:N)
      theta[n] = inv_logit(alpha * Environment[n]  + pow(Environment[n], beta)) ;
  }
```

#### Summary

```{r }
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- names(mdata)
# save(fits, file = "./distribution_save/Bbeta.Rdata")
load("./distribution_save/Bbeta.Rdata")
pars <- c("alpha", "beta")
lapply(as.list(names(mdata)), function(type)
  broom::tidyMCMC(fits[[type]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
    mutate(Type = type)) %>%
  bind_rows() %>%
  kable()
```

#### Posteriors

```{r , fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(names(data), function(type)
  mcmc_intervals(as.array(fits[[type]]), pars = pars)),
  labels = names(data), nrow = 2)
```

Column 2
-------------------------------------

#### Chains

```{r , fig.cap="Markov chains trace plot after warmup for limit data."}
mcmc_trace(as.array(fits$limit), pars = c(pars, "lp__"),
             facet_args = list(labeller = label_parsed))
```

#### Pairs

```{r , fig.cap="Markov chains pairs plot after warmup  for limit data."}
mcmc_pairs(as.array(fits$limit), pars = c(pars, "lp__"))
```

#### Predictions

```{r , fig.cap="Predictions."}
lapply(as.list(names(data)), function(type)
  cbind(type = type, data[[type]],
        mu = apply(as.matrix(fits[[type]], pars = "theta"), 2, mean),
        t(apply(as.matrix(fits[[type]], pars = "theta"), 2, quantile, probs = c(0.05, 0.95))))) %>% 
  bind_rows() %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'red', alpha = 0.2) +
  geom_line(aes(y = `5%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  facet_wrap(~ type, scales = "free", nrow = 3)
```

$B_{\alpha, \beta}2$
==================

Column 1
-------------------------------------

#### Model

$B_{\alpha, \beta}2$ is a Bernoulli logistic regression allowing a polynomial form in addition to sigmoids with the $\beta$ power parameter on the environmental variable.

$$Presence \sim \mathcal{Bernoulli}(logit(\alpha*(Environment+Environment^{\beta})))$$

#### Code

```{stan , output.var="Model", echo=T, eval=F}
  data {
    int<lower=1> N ; // # obs
    int<lower=0, upper=1> Presence[N] ;
    vector[N] Environment ;
    vector<lower=0, upper=1>[N] weights ;
  }
  parameters {
    real alpha_0 ; // Intercept
    real alpha ; // sigmoidal slope
    real<lower=0, upper=4> beta ; // power form
  }
  model {
    for(n in 1:N)
      target += weights*bernoulli_logit_lpmf(Presence[n] | alpha*(Environment[n] + pow(Environment[n], beta))) ;
  }
  generated quantities {
    vector<lower=0, upper=1>[N] theta ;
    for(n in 1:N)
      theta[n] = inv_logit(alpha*(Environment[n]  + pow(Environment[n], beta))) ;
  }
```

#### Summary

```{r }
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- names(mdata)
# save(fits, file = "./distribution_save/Bbeta2.Rdata")
load("./distribution_save/Bbeta2.Rdata")
pars <- c("alpha", "beta")
lapply(as.list(names(mdata)[-1]), function(type)
  broom::tidyMCMC(fits[[type]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
    mutate(Type = type)) %>%
  bind_rows() %>%
  kable()
```

#### Posteriors

```{r , fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(names(data)[-1], function(type)
  mcmc_intervals(as.array(fits[[type]]), pars = pars)),
  labels = names(data), nrow = 2)
```

Column 2
-------------------------------------

#### Chains

```{r , fig.cap="Markov chains trace plot after warmup for limit data."}
mcmc_trace(as.array(fits$limit), pars = c(pars, "lp__"),
             facet_args = list(labeller = label_parsed))
```

#### Pairs

```{r , fig.cap="Markov chains pairs plot after warmup  for limit data."}
mcmc_pairs(as.array(fits$limit), pars = c(pars, "lp__"))
```

#### Predictions

```{r , fig.cap="Predictions."}
lapply(as.list(names(data)[-1]), function(type)
  cbind(type = type, data[[type]],
        mu = apply(as.matrix(fits[[type]], pars = "theta"), 2, mean),
        t(apply(as.matrix(fits[[type]], pars = "theta"), 2, quantile, probs = c(0.05, 0.95))))) %>% 
  bind_rows() %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'red', alpha = 0.2) +
  geom_line(aes(y = `5%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  facet_wrap(~ type, scales = "free", nrow = 3)
```

$B_{\alpha, \beta}3$
==================

Column 1
-------------------------------------

#### Model

$B_{\alpha, \beta}3$ is a Bernoulli logistic regression allowing a polynomial form in addition to sigmoids with the $\beta$ power parameter on the environmental variable.

$$Presence \sim \mathcal{Bernoulli}(logit(Environment^{\alpha}+Environment^{\beta})))~|~\alpha<\beta$$

#### Code

```{stan , output.var="Model", echo=T, eval=F}
  data {
    int<lower=1> N ; // # obs
    int<lower=0, upper=1> Presence[N] ;
    vector[N] Environment ;
    vector<lower=0, upper=1>[N] weights ;
  }
  parameters {
    real alpha_0 ; // Intercept
    real<lower=0, upper=beta> alpha ; // first power form
    real<lower=0, upper=4> beta ; // second power form
  }
  model {
    for(n in 1:N)
      target += weights*bernoulli_logit_lpmf(Presence[n] | pow(Environment[n], alpha) + pow(Environment[n], beta)) ;
  }
  generated quantities {
    vector<lower=0, upper=1>[N] theta ;
    for(n in 1:N)
      theta[n] = inv_logit(pow(Environment[n], alpha) + pow(Environment[n], beta)) ;
  }
```

#### Summary

```{r }
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- names(mdata)
# save(fits, file = "./distribution_save/Bbeta3.Rdata")
load("./distribution_save/Bbeta3.Rdata")
pars <- c("alpha", "beta")
lapply(as.list(names(mdata)[-1]), function(type)
  broom::tidyMCMC(fits[[type]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
    mutate(Type = type)) %>%
  bind_rows() %>%
  kable()
```

#### Posteriors

```{r , fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(names(data)[-1], function(type)
  mcmc_intervals(as.array(fits[[type]]), pars = pars)),
  labels = names(data), nrow = 2)
```

Column 2
-------------------------------------

#### Chains

```{r , fig.cap="Markov chains trace plot after warmup for limit data."}
mcmc_trace(as.array(fits$limit), pars = c(pars, "lp__"),
             facet_args = list(labeller = label_parsed))
```

#### Pairs

```{r , fig.cap="Markov chains pairs plot after warmup  for limit data."}
mcmc_pairs(as.array(fits$limit), pars = c(pars, "lp__"))
```

#### Predictions

```{r , fig.cap="Predictions."}
lapply(as.list(names(data)[-1]), function(type)
  cbind(type = type, data[[type]],
        mu = apply(as.matrix(fits[[type]], pars = "theta"), 2, mean),
        t(apply(as.matrix(fits[[type]], pars = "theta"), 2, quantile, probs = c(0.05, 0.95))))) %>% 
  bind_rows() %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'red', alpha = 0.2) +
  geom_line(aes(y = `5%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  facet_wrap(~ type, scales = "free", nrow = 3)
```

# Discussion

*Discuss and summarise all built models, if possible in regard of litterature...*

### Conclusion

*Models that we will test...*

### References