---
title: "A01a: Single taxon model"
date: '`r Sys.Date()`'
author: Sylvain Schmitt
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
    theme: flatly
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
```

```{r data}
n <- 100
data <- list(
  no = data.frame(Environment = seq(0, 1, length.out = 100),
                  Presence = c(rep(0:1, 50))),
  intermediate = data.frame(Environment = seq(0, 1, length.out = n),
                            Presence = c(rep(0, 20), 
                                         rep(0:1,10),
                                         rep(1,20),
                                         rep(0:1,10),
                                         rep(0,20))),
  limit = data.frame(Environment = seq(0, 1, length.out = n), 
                     Presence = c(rep(0,30), rep(0:1,20), rep(1,30)))
)
mdata <- lapply(data, function(x) list(N = nrow(x),
                                       Presence = x$Presence,
                                       Environment = x$Environment))
```

```{r fits}
# ModelB0 <- stan_model("./distribution_models/Bernoullis/B0.stan")
# fitsB0 <- lapply(mdata, function(x) sampling(ModelB0, chains = 2, data = x))
# names(fitsB0) <- names(mdata)
# ModelBalpha <- stan_model("./distribution_models/Bernoullis/Balpha.stan")
# fitsBalpha <- lapply(mdata, function(x) sampling(ModelBalpha, chains = 2, data = x))
# names(fitsBalpha) <- names(mdata)
# ModelBalpha2 <- stan_model("./distribution_models/Bernoullis/Balpha2.stan")
# fitsBalpha2 <- lapply(mdata, function(x) sampling(ModelBalpha2, chains = 2, data = x))
# names(fitsBalpha2) <- names(mdata)
# ModelBbeta <- stan_model("./distribution_models/Bernoullis/Bbeta.stan")
# fitsBbeta <- lapply(mdata, function(x) sampling(ModelBbeta, chains = 2, data = x))
# names(fitsBbeta) <- names(mdata)
# ModelBbeta2 <- stan_model("./distribution_models/Bernoullis/Bbeta2.stan")
# fitsBbeta2 <- lapply(mdata, function(x) sampling(ModelBbeta2, chains = 2, data = x))
# names(fitsBbeta2) <- names(mdata)
# ModelBbeta3  <- stan_model("./distribution_models/Bernoullis/Bbeta3.stan")
# fitsBbeta3 <- lapply(mdata, function(x) sampling(ModelBbeta3, chains = 2, data = x))
# names(fitsBbeta3) <- names(mdata)
# fits <- list(B0 = fitsB0, Balpha = fitsBalpha, Balpha2 = fitsBalpha2,
#              Bbeta = fitsBbeta, Bbeta2 = fitsBbeta2, Bbeta3 = fitsBbeta3)
# save(fits, file = "./distribution_save/SingleModel.Rdata")
load("./distribution_save/SingleModel.Rdata")
```

# Introduction

The aim of this document is to find the best general form of model to fit a distribution model. Especially we are seeking for a model that can deal with three specific cases :

* The environmental variable has no effect
* The niche optimum is a intermediate value of the environmental variable range
* The niche optimum is a limit of the environmental variable range

# Models

We used Bernoulli distribution to model species complex probability of presence. $B_0$ is the null model estimating a single theta along the whole environmental gradient. Thus $B_0$ cannot include any environmental effect. $B_{\alpha}$ is a Bernoulli logistic regression estimating a sigmoidal increasing or decreasing $\alpha$ along the whole environmental gradient in addition to the intercept $\alpha_0$. $B_{\alpha, \alpha_2}$ is a Bernoulli logistic regression allowing a quadratic form in addition to sigmoids with the $\alpha_2$ parameter. $B_{\alpha, \beta}$ is a Bernoulli logistic regression allowing a polynomial form in addition to sigmoids with the $\beta$ power parameter on the environmental variable. $B_{\alpha, \beta}2$ is a Bernoulli logistic regression allowing a polynomial form in addition to sigmoids with the $\beta$ power parameter on the environmental variable. $B_{\alpha, \beta}3$ is a Bernoulli logistic regression allowing a polynomial form in addition to sigmoids with the $\beta$ power parameter on the environmental variable.

```{r ModelsTable, render="pandoc"}
data.frame(
  Name = c("$B_0$", "$B_{\\alpha}$", "$B_{\\alpha, \\alpha_2}$", "$B_{\\alpha, \\beta}$",
           "$B_{\\alpha, \\beta}2$", "$B_{\\alpha, \\beta}3$"),
  Formula = c("$Presence \\sim \\mathcal{Bernoulli}(logit^{-1}(\\alpha_0))$",
              "$Presence \\sim \\mathcal{Bernoulli}(logit^{-1}(\\alpha_0+\\alpha*Environment))$",
              "$Presence \\sim \\mathcal{Bernoulli}(logit^{-1}(\\alpha_0+\\alpha*Environment+\\alpha_2*Environment^2))$",
              "$Presence \\sim \\mathcal{Bernoulli}(logit^{-1}(\\alpha_0+\\alpha*Environment+Environment^{\\beta}))$",
              "$Presence \\sim \\mathcal{Bernoulli}(logit^{-1}(\\alpha_0+\\alpha*(Environment+Environment^{\\beta})))$",
              "$Presence \\sim \\mathcal{Bernoulli}(logit^{-1}(\\alpha_0+Environment^{\\alpha}+Environment^{\\beta}))$")  
) %>% 
  kable(caption = "Models summary.",
        format = "html", escape = F) %>% 
  kable_styling("striped", full_width = F) %>% 
  collapse_rows(1)
```

# Results

$B_{\alpha}$ is responding well to the optimum in the limit of the environmental range and to no effect from the environmental variable. But $B_{\alpha}$ can't detect ecological optimum in the middle of the ecological range, besides it's the most commonly used model ! $B_{\alpha, \alpha_2}$ is giving the expected result in every case. $B_{\alpha, \beta}$, $B_{\alpha, \beta}2$, and $B_{\alpha, \beta}3$ are not dealing well with the intermediate niche optimum case. 

```{r Predictions, fig.cap="Predictions."}
lapply(names(fits), function(model){
  lapply(as.list(names(data)), function(type)
    cbind(type = type, data[[type]],
          mu = apply(as.matrix(fits[[model]][[type]], pars = "theta"), 2, mean),
          t(apply(as.matrix(fits[[model]][[type]], pars = "theta"), 2, 
                  quantile, probs = c(0.05, 0.95))))) %>% 
    bind_rows() %>% 
    mutate(model = model)
}) %>% bind_rows() %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'red', alpha = 0.2) +
  geom_line(aes(y = `5%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  facet_grid(model ~ type, scales = "free")
```

# Discussion

In conclusion, $B_{\alpha, \alpha_2}$ is the best model in terms of convergence, likelihood, number of parameters, and prediction quality. Consequently we will retain $B_{\alpha, \alpha_2}$ to model species complex against the whole forest. We now need to develop a multinomial Dirichlet model to model joint niches of several species within complexes. We will additionnally check theoritically the behavior of $B_{\alpha, \alpha_2}$ in case of unweighted sampling in another document.
