---
title: "A01b: Single taxon model weight"
date: '`r Sys.Date()`'
author: Sylvain Schmitt
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
    theme: flatly
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
```

```{r data}
absence <- data.frame(Environment = seq(0, 1, length.out = 1000), Presence = 0)
data <- list(
  no = rbind(absence, 
      data.frame(Environment = sample(absence$Environment, 100), Presence = 1)),
  intermediate = rbind(absence, 
      data.frame(Environment = sample(absence$Environment[400:600], 100), Presence = 1)),
  limit = rbind(absence, 
      data.frame(Environment = sample(absence$Environment[800:1000], 100), Presence = 1))
)
mdata <- lapply(data, function(x) list(N = nrow(x),
                                       Presence = x$Presence,
                                       Environment = x$Environment,
                                       weights = ifelse(x$Presence == 0,
                                                        sum(x$Presence)/length(x$Presence),
                                                        1-sum(x$Presence)/length(x$Presence))))
```

```{r fits}
# unweighted <- stan_model("./distribution_models/Bernoullis/Balpha2.stan")
# fitsUnweighted <- lapply(mdata, function(x) sampling(unweighted, chains = 2, data = x))
# names(fitsUnweighted) <- names(mdata)
# weighted <- stan_model("./distribution_models/Bernoullis/Balpha2weighted.stan")
# fitsWeighted <- lapply(mdata, function(x) sampling(weighted, chains = 2, data = x))
# names(fitsWeighted) <- names(mdata)
# fits <- list(unweighted = fitsUnweighted, weighted = fitsWeighted)
# save(fits, file = "./distribution_save/SingleModelWeight.Rdata")
load("./distribution_save/SingleModelWeight.Rdata")
```

# Introduction

The aim of this document is to find the best practice to deal with unbalance sampling between absences and presences (*e.g.* thousand absences observations against hundred presences).

# Models

We used the model $B_{\alpha, \alpha_2}$ is a Bernoulli logistic regression allowing a quadratic form in addition to sigmoids with the $\alpha_2$ parameter:
$$Presence \sim \mathcal{Bernoulli}(logit^{-1}(\alpha_0+\alpha*Environment+\alpha_2*Environment^2))$$
with unweigthed observations or with absences weight equal to 9% and presence wieght equalt to 91%.

# Results

The unweighted performed better with a better assessment of uncertainty. Both model are predicting species relative abundance and not species habitat suitability.

```{r Predictions, fig.cap="Predictions."}
lapply(names(fits), function(model){
  lapply(as.list(names(data)), function(type)
    cbind(type = type, data[[type]],
          mu = apply(as.matrix(fits[[model]][[type]], pars = "theta"), 2, mean),
          t(apply(as.matrix(fits[[model]][[type]], pars = "theta"), 2, 
                  quantile, probs = c(0.05, 0.95))))) %>% 
    bind_rows() %>% 
    mutate(model = model)
}) %>% bind_rows() %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'red', alpha = 0.2) +
  facet_grid(model ~ type, scales = "free")
```

# Discussion

We thus kept the unweighted model for further analysis keeping in mind that our model will predict species complex relative abundance depending on the environment gradient and not habitat suitability.
