---
title: "A02: Abiotic distributions"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  github_document: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(parallel)
library(tidyverse)
library(ggfortify)
library(raster)
library(leaflet)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
```

```{r trees}
trees <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(CensusYear == 2015) %>%
  filter(Species != "Indet.") %>% 
  mutate(DBH = CircCorr*pi) %>% 
  collect()
```

```{r env}
# env <- trees %>%
#   dplyr::select(idTree, Xutm, Yutm) %>%
#   unique()
# coordinates(env) <- ~Xutm + Yutm
# proj4string(env) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
# env <- spTransform(env, CRSobj = crs)
# topo <- stack(
#   raster(file.path(path, "topography", "DEM_1m_2015.tif")),
#   raster(file.path(path, "topography", "RelativeElevation_1m.tif")),
#   raster(file.path(path, "topography", "slope_1m.tif")),
#   raster(file.path(path, "topography", "curvature_1m.tif")),
#   raster(file.path(path, "topography", "TRI_1m.tif")),
#   raster(file.path(path, "topography", "TWI_1m.tif"))
# )
# names(topo) <- c("DEM", "RelativeElevation", "Slope", "Curvature", "TRI", "TWI")
# topo <- projectRaster(topo, crs = crs)
# env <- data.frame(cbind(env@data, raster::extract(topo, env)))
# rm(topo)
# save(env, file = "./distribution_save/env.Rdata")
load("./distribution_save/env.Rdata")
```

```{r data}
data <- trees %>% 
  left_join(env)
```

```{r complexes}
complexes <- bind_rows(
  data.frame(Complex = "Aspidosperma", Genus = "Aspidosperma",
             Species = c("carapanauba", "shultesii", "excelsum", "oblongum", 
                         "desmanthum", "album", "spruceanum", "sandwithianum", "helstonei", "sp.1CAY-ATDN")),
  data.frame(Complex = "Dyospiros", Genus = "Dyospiros",
             Species = c("capreifolia", "carbonaria", "martinii", "guianensis", "vestitia")),
  data.frame(Complex = "Eschweilera clade Chartacea",
             Genus = c("Eschweilera", "Lecythis", "Eschweilera", "Courataria", "Lecythis", "Lecythis"),
             Species = c("simiorum", "holocogyne", "congestiflora", "multiflora", "chartacea", "zabucajo")),
  data.frame(Complex = "Eschweilera clade Parvifolia", Genus = "Eschweilera",
             Species = c("pedicellata", "coriacea", "decolorans", "sagotiana", "parviflora",
                         "micrantha", "grandiflora", "chartaceifolia")),
  data.frame(Complex = "Lecythis", Genus = "Lecythis",
             Species = c("idatimon", "persistens", "persistenssubspaurantiaca", "corrugata")),
  data.frame(Complex = "Couratari", Genus = "Couratari",
             Species = c("calcynia", "oblongifolia", "guianensis")),
  data.frame(Complex = "Pourouma", Genus = "Pourouma",
             Species = c("guianensis", "saülensis", "sp.2CAY-ATDN", "minor", "melionii", "bicolor",
                         "villosa", "sp.5CAY-ATDN")),
  data.frame(Complex = "Cecropia", Genus = c("Pourouma", "Cecropia", "Cecropia"),
             Species = c("mollis", "obtusa", "sciadophylla")),
  data.frame(Complex = "Licania1", Genus = "Licania",
             Species = c("menbranacea", "ovalifolia", "micrantha", "canescens", "laxiflora",
                         "alba", "majuscula")),
  data.frame(Complex = "Licania2", Genus = "Licania",
             Species = c("octandra", "sprucei", "bicornis", "minutiflora")),
  data.frame(Complex = "Couepia", Genus = c(rep("Couepia", 7), rep("Licania", 2)),
             Species = c("obovata", "abrantha", "magnolifolia", "joaquinea", "caryophilloides",
                         "guianensis", "bracteosa", "heteromorpha", "latistipula")),
    data.frame(Complex = "Symphonia", Genus = "Symphonia",
             Species = c("globulifera", "sp.1"))
)
complexes <- complexes %>% 
  filter(Complex %in% c("Symphonia", "Eschweilera clade Parvifolia"))
data <- left_join(complexes, data) %>% 
  filter(!is.na(Plot))
```

# Introduction

The aim of this document is to study the distribution at micro-environmental scale with abiotic environment of two species complexes : *Symphonia globulifera* morphotypes and *Eschweilera Parvifolia clade* species. *Symphonia globulifera* includes two morphotypes, *S. globulifera* and *S. sp1*, and *Eschweilera Parvifolia clade* 11 species. We wish to use a bayseian approach including both biotic and abiotic environment and maybe the ontogoeny thourg diameter at breast hieght (DBH).

One of the first thing aiming this study as such fine micro-scale started from @Allie2015 figure (see figure \@ref(fig:Allie)). Effectivelly for the two species complexes studied she has shown a differentation of habitat (between *S. globulifera* and *S. sp1* and between, *E. coriacea* and *E. sagotiana*). But the ecological niche defined only included the centroid of the niche and we were wondering about the real niche taking into account intraspecific variability of habitat between sistser species within species complexes.

```{r Allie, fig.cap = "Environmental variables selected in @Allie2015 including intraspecific variability."}
data %>% 
  ggplot(aes(RelativeElevation, Slope, 
             col = Species, size = DBH)) +
  geom_point(alpha = 0.2) +
  facet_wrap(~ Complex, nrow = 2)
```
```{r Allie2, fig.cap = "Environmental variables selected in @Allie2015 including intraspecific variability."}
data %>% 
  filter(Species %in% c("coriacea", "sagotiana",
                        "globulifera", "sp.1")) %>%
  ggplot(aes(TWI, Slope, 
             col = Species, size = DBH)) +
  geom_point(alpha = 0.2) +
  facet_wrap(Complex ~ Species, nrow = 2)
```

# Graphical exploration

Base on abiotic variable PCA (see figure \@ref(fig:pcaABiot)), We selected less correlated variable: DEM, TWI, and slope.

```{r pcaABiot, fig.cap="Abiotic variable PCA"}
autoplot(princomp(~ DEM + RelativeElevation + Slope + Curvature + TRI + TWI, 
                    data = data, cor = T), data = data,
         colour = "Species", alpha = 0.1, size = "DBH",
         loadings.label.size = 6,
         loadings.label.colour = 'black', loadings.label.vjust = 1.1,
         loadings = T, loadings.label = T, loadings.colour = 'black') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted") +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1)) +
  scale_color_discrete(guide= "none") +
  facet_wrap(~ Complex)
```

```{r densityAbiot,fig.cap="Abiotic variable density plot for selected variables."}
data %>% 
  dplyr::select(Complex, Species, DEM, TWI, Slope) %>% 
  reshape2::melt(id.var = c("Complex", "Species")) %>% 
  group_by(variable) %>% 
  mutate(value = scale(value)) %>% 
  ungroup() %>% 
  ggplot(aes(value, fill = Species, col = Species)) + 
  geom_density(alpha = 0.3) +
  facet_grid(variable ~ Complex, scales = "free")
```

# Niche analysis

```{r dataReduction}
# data <- filter(data, Plot == 16) # for tests
# data <- filter(data, Genus == "Symphonia")
```

To build the niche model, we have several conceptual and technical challenges to solve. But the most important and first questions is how to consider absence data. Effectivelly, what we have for Paracou is the information but where we can fin species individuals, the presence information. But we have to decide how we consider absence from one habitat or putative absence called pseudo-absence. First of all, we can use recommandations of *Barbet Massin ... (to include in Mendely*): *"we recommend the use of a large number (e.g. 10 000) of pseudo‐absences with equal weighting for presences and absences when using regression techniques"*. Secondly, because the model is partly built on pseudo-absences, we will proceed in two steps: (i) first we will built the complex distribution using individual from all species within the complex as presence and all other trees as absence, (ii) secondly we will built species distribution inside the complex considering all individual from one species inside the complex as presence and all individuals from other species of the complex as absence.

## Model

 representing the habitat suitability for the taxon (the complex or the species). $\theta_{taxon}$  is an hyperparameter depending on abitoic environment represented here by the digital elevation model $DEM$, the topographic wetness index $TWI$ and the slope $Slope$.


```{stan Model, output.var="Model", echo=T, eval=F}
  data {
    int<lower=1> N ; // # obs
    int<lower=0, upper=1> Presence[N] ;
    vector<lower=0, upper=1>[N] DEM ;
    vector<lower=0, upper=1>[N] TWI ;
    vector<lower=0, upper=1>[N] Slope ;
    vector<lower=0, upper=1>[N] weights ;
  }
  parameters {
    real alpha_0 ; // intercept
    real alpha_TWI ; // sigmoidal slope
    real alpha_TWI2 ; // quadratic form
    real alpha_DEM ; // sigmoidal slope
    real alpha_DEM2 ; // quadratic form
    real alpha_Slope ; // sigmoidal slope
    real alpha_Slope2 ; // quadratic form
  }
  model {
    target += weights*bernoulli_logit_lpmf(Presence | alpha_0 + alpha_TWI*TWI + alpha_TWI2 * TWI .* TWI + alpha_DEM*DEM + alpha_DEM2 * DEM .* DEM + alpha_Slope*Slope + alpha_Slope2 * Slope .* Slope) ;
  }
  generated quantities {
    vector<lower=0, upper=1>[N] theta ;
    theta = inv_logit(alpha_0 + alpha_TWI*TWI + alpha_TWI2 * TWI .* TWI + alpha_DEM*DEM + alpha_DEM2 * DEM .* DEM + alpha_Slope*Slope + alpha_Slope2 * Slope .* Slope) ;
  }
```

## Symphonia

```{r fitMabiotic1Sympho}
datam <- dplyr::filter(data, Genus == "Symphonia")
species <- c("globulifera", "sp.1")
# fits <- lapply(as.list(species), function(sp)
#                sampling(Model, chains = 2,
#                         data = list(N = nrow(datam),
#                                     Presence = as.numeric(datam$Species == sp),
#                                     DEM = datam$DEM/max(datam$DEM),
#                                     TWI = datam$TWI/max(datam$TWI),
#                                     Slope = datam$Slope/max(datam$Slope),
#                                     weights = ifelse(datam$Species == sp,
#                                                      1/(2*sum(datam$Species == sp)),
#                                                      1/(2*sum(datam$Species != sp))))))
# names(fits) <- species
# save(fits, file = "./distribution_save/MSympho.Rdata")
load("./distribution_save/MSympho.Rdata")
pars <- c("alpha_0", "alpha_DEM", "alpha_DEM2", "alpha_TWI", "alpha_TWI2",
          "alpha_Slope", "alpha_Slope2")
lapply(as.list(species), function(sp)
  broom::tidyMCMC(fits[[sp]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
    mutate(Species = sp)) %>%
  bind_rows() %>%
  kable(caption = "Summary table of the model")
```

```{r Cabiotic1Sympho, fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(as.list(species), function(sp)
  mcmc_intervals(as.array(fits[[sp]]), pars = pars)),
  labels = species, nrow = 2)
```

```{r Tabiotic1Sympho,fig.cap="Markov chains trace plot after warmup."}
mcmc_trace(as.array(fits$globulifera), pars = c(pars, "lp__")) +
  ggtitle("Symphonnia globulifera")
```

```{r 2abiotic1Sympho, fig.cap="Markov chains pairs plot after warmup."}
mcmc_pairs(as.array(fits$globulifera), pars = c(pars, "lp__"))
```

```{r Abiotic1Sympho, fig.cap="Probability density (mu)."}
pars_val <- lapply(fits, function(fit) apply(as.matrix(fit, pars = c("alpha_0", "alpha_DEM", "alpha_DEM2", "alpha_TWI", "alpha_TWI2", "alpha_Slope", "alpha_Slope2")), 2, mean))

proj <- lapply(as.list(species), function(sp)
  data.frame(cbind(model = sp, species = datam$Species,
        Presence = as.numeric(datam$Species == sp),
        mu = apply(as.matrix(fits[[sp]], pars = "theta"), 2, mean),
        t(apply(as.matrix(fits[[sp]], pars = "theta"), 2, quantile, probs = c(0.05, 0.95))),
        DEM = paste0(psych::logistic(pars_val[[sp]]["alpha_0"] + 
                                   pars_val[[sp]]["alpha_DEM"]*datam$DEM/max(datam$DEM) +
                                   pars_val[[sp]]["alpha_DEM2"]*(datam$DEM/max(datam$DEM))^2 +  
                                   pars_val[[sp]]["alpha_TWI"]*mean(datam$TWI/max(datam$TWI)) + 
                                   pars_val[[sp]]["alpha_TWI"]*mean(datam$TWI/max(datam$TWI))^2 +
                                   pars_val[[sp]]["alpha_Slope"]*mean(datam$Slope/max(datam$Slope)) +
                                   pars_val[[sp]]["alpha_Slope2"]*mean(datam$Slope/max(datam$Slope))^2),
                     "/", datam$DEM),
        TWI = paste0(psych::logistic(pars_val[[sp]]["alpha_0"] + 
                                   pars_val[[sp]]["alpha_DEM"]*mean(datam$DEM/max(datam$DEM)) +
                                   pars_val[[sp]]["alpha_DEM2"]*mean(datam$DEM/max(datam$DEM))^2 +  
                                   pars_val[[sp]]["alpha_TWI"]*datam$TWI/max(datam$TWI) + 
                                   pars_val[[sp]]["alpha_TWI"]*datam$TWI/max(datam$TWI)^2 +
                                   pars_val[[sp]]["alpha_Slope"]*mean(datam$Slope/max(datam$Slope)) +
                                   pars_val[[sp]]["alpha_Slope2"]*mean(datam$Slope/max(datam$Slope))^2),
                     "/", datam$TWI),
        Slope = paste0(psych::logistic(pars_val[[sp]]["alpha_0"] + 
                                   pars_val[[sp]]["alpha_DEM"]*mean(datam$DEM/max(datam$DEM)) +
                                   pars_val[[sp]]["alpha_DEM2"]*mean(datam$DEM/max(datam$DEM))^2 +  
                                   pars_val[[sp]]["alpha_TWI"]*mean(datam$TWI/max(datam$TWI)) + 
                                   pars_val[[sp]]["alpha_TWI"]*mean(datam$TWI/max(datam$TWI))^2 +
                                   pars_val[[sp]]["alpha_Slope"]*datam$Slope/max(datam$Slope) +
                                   pars_val[[sp]]["alpha_Slope2"]*datam$Slope/max(datam$Slope)^2),
                       "/", datam$Slope))))
names(proj) <- species
bind_rows(proj) %>% 
  rename(`5%` = X5., `95%` = X95.) %>% 
  mutate(Presence = as.numeric(as.character(Presence))) %>% 
  mutate_at(vars("mu", "5%", "95%"), as.numeric) %>% 
  dplyr::select(model, species, Presence, mu, `5%`, `95%`, DEM, TWI, Slope) %>% 
  reshape2::melt(id.vars = c("model", "species", "Presence", "mu", "5%", "95%")) %>% 
  separate(value, c("mu_var", "value"), "/", convert = T) %>% 
  ggplot(aes(x = value)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  scale_color_discrete("Presence") +
  geom_point(aes(y = mu, alpha = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = NA, alpha = 0.2) +
  geom_line(aes(y = mu_var, col = "Prediciton"), lwd = 2) +
  facet_grid(model ~ variable, scales = "free") +
  ylab("Probability density")
```


```{r NicheAbiotic1Sympho, fig.cap="Niche (HabitatSuitability)."}
lapply(as.list(species), function(sp)
  cbind(Model = sp, datam,
        Presence = as.numeric(datam$Species == sp),
        mu = apply(as.matrix(fits[[sp]], pars = "theta"), 2, mean))) %>% 
  bind_rows() %>%
  filter(Model == Species) %>% 
  dplyr::select(Species, DBH, mu, TWI, DEM, Slope) %>% 
  reshape2::melt(id.vars = c("Species", "DBH", "mu", "TWI")) %>% 
  ggplot(aes(TWI, value, col = mu, size = DBH)) +
  geom_point() +
  facet_grid(variable~Species, scales = "free") +
  scale_color_gradient2("Habitat\nsuitability", low = scales::muted("blue"), mid = "white",
                        high = scales::muted("red"), midpoint = 0.5)
```

```{r NicheAbiotic2Sympho, fig.cap="Niche (HabitatSuitability)."}
pars_val <- apply(as.matrix(fits$globulifera, pars = c("alpha_0", "alpha_DEM", "alpha_DEM2", "alpha_TWI",
                                           "alpha_TWI2", "alpha_Slope", "alpha_Slope2")), 2, mean)
n <- 100
HS <- matrix(nrow = n, ncol = n)
DEM <- seq(0, 1, length.out = n)
TWI <- seq(0, 1, length.out = n)
for(i in 1:n)
  for(j in 1:n)
    HS[i, j] <- psych::logistic(pars_val["alpha_0"] + 
                                   pars_val["alpha_DEM"]*DEM[i] +
                                   pars_val["alpha_DEM2"]*DEM[i]^2 +
                                   pars_val["alpha_TWI"]*TWI[j] +
                                   pars_val["alpha_TWI"]*TWI[j]^2 +
                                   pars_val["alpha_Slope"]*0.5 +
                                   pars_val["alpha_Slope2"]*0.5^2)
reshape2::melt(HS, value.name = "HS") %>% 
  mutate(Var1 = DEM[Var1]*max(datam$DEM), Var2 = TWI[Var2]*max(datam$TWI)) %>% 
  rename(DEM = Var1, TWI = Var2) %>% 
  ggplot(aes(x = DEM, y = TWI, z = HS, fill = HS)) +
  geom_raster() +
  geom_contour() +
  scale_fill_distiller("S. globulifera\nHabitat\nsuitability (%)", palette = "Spectral")
```

```{r projection, fig.cap="Symphonia globulifera morphotype globulifera distribution in the Plot 16."}
limits <- subset(shapefile(file.path(path, "limits", "OverallPlots.shp")))
# env <- stack(raster(file.path(path, "topography", "DEM_1m_2015.tif")),
#              raster(file.path(path, "topography", "TWI_1m.tif")),
#              raster(file.path(path, "topography", "slope_1m.tif")))
# env <- crop(env, limits)
# names(env) <- c("DEM", "TWI", "Slope")
# pars_val <- apply(as.matrix(fits$globulifera, pars = c("alpha_0", "alpha_DEM", "alpha_DEM2", "alpha_TWI",
#                                            "alpha_TWI2", "alpha_Slope", "alpha_Slope2")), 2, mean)
# projection <- calc(env, function(x){psych::logistic(sum(pars_val["alpha_0"],
#                                           pars_val["alpha_DEM"]*x[1], pars_val["alpha_DEM2"]*x[1]*x[1],
#                                           pars_val["alpha_TWI"]*x[2], pars_val["alpha_TWI2"]*x[2]*x[2],
#                                           pars_val["alpha_Slope"]*x[3], pars_val["alpha_Slope2"]*x[3]*x[3]))})
# save(projection, file = "./distribution_save/SymphoProj.Rdata")
load("./distribution_save/SymphoProj.Rdata")
projection_df <- as(aggregate(projection, 4), "SpatialPixelsDataFrame")
projection_df <- as.data.frame(projection_df)
colnames(projection_df) <- c("HS", "x", "y")
ggplot() +  
  geom_tile(data=projection_df, aes(x=x, y=y, fill=(1-HS)*100)) +
  geom_polygon(data = fortify(limits),
                 aes(long, lat, group=group),
                 colour = "black", lwd = 1.1, alpha = 0) +
  coord_equal() +
  cowplot::theme_map() +
  scale_fill_distiller("S. globulifera\nHabitat\nsuitability (%)", palette = "Spectral")
```

`dismo::nicheOverlap`  functions as other compute niche overlap from distribution projected on rasters which miss the overlap computation of the theoritical niche in the full environmental space tested. We will build an overlap index as the intersection of habitat suitability distribution (product of distribution functions) divided by the total sum of habitat suitability distribution (sum of distribution functions) :

$$I_{overlap} = \frac{\cap niche}{\sum niche} = \frac{\prod_sdistribution_s}{\sum_sdistribution_s} =  \frac{\prod_s f_s(DEM, TWI, Slope)}{\sum_s f_s(DEM, TWI, Slope)}$$

```{r overlap, eval=F}
distribution <- function(DEM, TWI, Slope, pars)
  psych::logistic(sum(pars_val["alpha_0"],
                      pars_val["alpha_DEM"]*DEM, pars_val["alpha_DEM2"]*DEM,
                      pars_val["alpha_TWI"]*TWI, pars_val["alpha_TWI2"]*TWI*TWI,
                      pars_val["alpha_Slope"]*Slope, pars_val["alpha_Slope2"]*Slope*Slope))
pars_val <- lapply(fits, function(fit)
  apply(as.matrix(fit, pars = c("alpha_0", "alpha_DEM", "alpha_DEM2", "alpha_TWI",
                                "alpha_TWI2", "alpha_Slope", "alpha_Slope2")), 2, mean))
pars_val
```

## Eschweilera

```{r fitMabiotic1Maho, eval=F}
datam <- filter(data, Complex == "Eschweilera clade Parvifolia")
species <- c("coriacea", "sagotiana")
# fits <- lapply(as.list(species), function(sp)
#                sampling(Model, chains = 2,
#                         data = list(N = nrow(datam),
#                                     Presence = c(10^-6, 1-10^-6)[as.numeric(datam$Species == sp)+1],
#                                     DEM = datam$DEM/max(datam$DEM),
#                                     TWI = datam$TWI/max(datam$TWI),
#                                     Slope = datam$Slope/max(datam$Slope),
#                                     weights = ifelse(datam$Species == sp,
#                                                      1/(2*sum(datam$Species == sp)),
#                                                      1/(2*sum(datam$Species != sp))))))
# names(fits) <- species
# save(fits, file = "./distribution_save/MMaho1.Rdata")
load("./distribution_save/MMaho1.Rdata")
pars <- c("alpha_0", "alpha_DEM", "alpha_DEM2", "alpha_TWI", "alpha_TWI2",
          "alpha_Slope", "alpha_Slope2", "beta_0", "beta_DEM", "beta_DEM2",
          "beta_TWI", "beta_TWI2", "beta_Slope", "beta_Slope2")
lapply(as.list(species), function(sp)
  broom::tidyMCMC(fits[[sp]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
    mutate(Species = sp)) %>%
  bind_rows() %>%
  kable(caption = "Summary table of the model")
```

```{r Cabiotic1Maho, fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(as.list(species), function(sp)
  mcmc_intervals(as.array(fits[[sp]]), pars = pars)),
  labels = species, nrow = 2)
```

```{r Tabiotic1Maho,fig.cap="Markov chains trace plot after warmup.", eval=F}
mcmc_trace(as.array(fits$coriacea), pars = c(pars, "lp__")) +
  ggtitle("Symphonnia globulifera")
```

```{r 2abiotic1Maho, fig.cap="Markov chains pairs plot after warmup.", eval=F}
mcmc_pairs(as.array(fits$globulifera), pars = c(pars, "lp__"))
```

```{r Yabiotic1Maho, fig.cap="Predictions.", eval=F}
lapply(as.list(species), function(sp)
  cbind(species = sp, datam,
        Presence = as.numeric(datam$Species == sp),
        mu = apply(as.matrix(fits[[sp]], pars = "mu"), 2, mean),
        t(apply(as.matrix(fits[[sp]], pars = "mu"), 2, quantile, probs = c(0.05, 0.95))))) %>% 
  bind_rows() %>% 
  dplyr::select(species, Presence, mu, `5%`, `95%`, DEM, TWI, Slope) %>% 
  reshape2::melt(id.vars = c("species", "Presence", "mu", "5%", "95%")) %>% 
  ggplot(aes(x = value)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = NA, alpha = 0.2) +
  facet_grid(species ~ variable, scales = "free") +
  ylab("Habitat suitability") +
  scale_color_discrete("Presence")
```

```{r NicheAbiotic1Maho, fig.cap="Niche.", eval=F}
lapply(as.list(species), function(sp)
  cbind(species = sp, datam,
        Presence = as.numeric(datam$Species == sp),
        mu = apply(as.matrix(fits[[sp]], pars = "mu"), 2, mean),
        t(apply(as.matrix(fits[[sp]], pars = "mu"), 2, quantile, probs = c(0.05, 0.95))))) %>% 
  bind_rows() %>% 
  ggplot(aes(TWI, DEM, col = mu)) +
  geom_point() +
  facet_wrap(~species) +
  scale_color_continuous("Habitat\nsuitability")
```

# References

