---
title: "A02b: Joint taxon model weight"
date: '`r Sys.Date()`'
author: Sylvain Schmitt
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
    theme: flatly
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
```

```{r data}
data <- list(
  no = data.frame(Environment = c(sample(seq(0, 1, length.out = 1000), 850, replace = T),
                                  sample(seq(0, 1, length.out = 1000), 130, replace = T),
                                  sample(seq(0, 1, length.out = 1000), 20, replace = T)),
                  A = c(rep(1, 850), rep(0, 130), rep(0, 20)),
                  B = c(rep(0, 850), rep(1, 130), rep(0, 20)),
                  C = c(rep(0, 850), rep(0, 130), rep(1, 20))),
  intermediate = data.frame(Environment = c(sample(seq(0, 0.4, length.out = 1000), 850, replace = T),
                                  sample(seq(0.3, 0.7, length.out = 1000), 130, replace = T),
                                  sample(seq(0.6, 1, length.out = 1000), 20, replace = T)),
                            A = c(rep(1, 850), rep(0, 130), rep(0, 20)),
                            B = c(rep(0, 850), rep(1, 130), rep(0, 20)),
                            C = c(rep(0, 850), rep(0, 130), rep(1, 20))),
  limit = data.frame(Environment = c(sample(seq(0, 0.4, length.out = 1000), 
                                            850, replace = T, prob = 1 - seq(0, 0.4, length.out = 1000)),
                                  sample(seq(0.3, 0.7, length.out = 1000), 130, replace = T),
                                  sample(seq(0.6, 1, length.out = 1000), 
                                         20, replace = T, prob = seq(0.6, 1, length.out = 1000))),
                     A = c(rep(1, 850), rep(0, 130), rep(0, 20)),
                     B = c(rep(0, 850), rep(1, 130), rep(0, 20)),
                     C = c(rep(0, 850), rep(0, 130), rep(1, 20)))
)
data <- lapply(data, function(x) x %>% 
         reshape2::melt(id.vars = "Environment", 
                        variable.name = "Species", value.name = "Presence") %>% 
         filter(Presence != 0) %>% 
         dplyr::select(-Presence))
mdata <- lapply(data, function(x) 
  list(N = nrow(x),
       S = length(unique(x$Species)),
       K = 1,
       X = matrix(x$Environment),
       Y = sapply(levels(as.factor(x$Species)), 
                  function(sp) as.numeric(x$Species == sp))))
```

```{r fit}
# Model <- stan_model("./distribution_models/JointModel.stan")
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- names(mdata)
# save(fits, file = "./distribution_save/DirichletWeight.Rdata")
load("./distribution_save/DirichletWeight.Rdata")
pars <- c("alpha", "beta", "gamma")
```

# Introduction

The aim of this document is to find the best practice to deal with unbalance sampling between absences and presences (e.g. 850 individuals from species A, 130 from species B and 20 for species C).

# Models

The model used follow a Multinomial and Dirichlet conjugated distributions with a quadratic formed generalized through the softmax function. Each observation is defined by a simplex of size number of species $S$ coded in 0 for all species absent and 1 for species present. Consequently, $S$ values of $\alpha$, $\beta$ and $\gamma$ are inferred to compute the niche of each species along the environmental variable relativelly to other species distributions (sum being equal to 1).

$$
Presence_{species} \sim \mathcal{Dirichlet~Multinomial}(softmax(\alpha + \beta*Environment + \gamma*Environment^2)) \\
{Log~Probability~Mass~Function}_{Dirichlet~Multinomial}(y|\alpha) = \Gamma(\sum \alpha)  + \sum(\Gamma(\alpha + y)) - \Gamma(\sum \alpha+\sum y) - \sum\Gamma(\alpha)
$$

# Results

The unweighted performed better with a better assessment of uncertainty. Both model are predicting species relative abundance and not species habitat suitability.

```{r Predictions, fig.cap="Predictions."}
preds <- lapply(names(mdata), function(type){
  N <- 100
  Env <- seq(0, 1, length.out = 100)
  S <- 3
  alpha <- as.matrix(fits[[type]], pars = "alpha")
  beta <- as.matrix(fits[[type]], pars = "beta")
  gamma <- as.matrix(fits[[type]], pars = "gamma")
  Yp <- abind::abind(sapply(1:length(Env), function(i) 
    sapply(1: nrow(alpha), function(n)
      tsensembler:::softmax(alpha[n,] + beta[n,] * Env[i] + gamma[n,]*Env[i]^2)
      ), simplify = F), along = 3)
  lapply(1:S, function(sp)
    data.frame(
      type = type,
      species = sp,
      x = Env,
      Yobs = mdata[[type]]$Y[,sp],
      Y = apply(Yp[sp,,], 2, mean),
      Y5 = apply(Yp[sp,,], 2, quantile, probs = 0.05),
      Y95 = apply(Yp[sp,,], 2, quantile, probs = 0.95))) %>% 
    bind_rows()
}) %>% bind_rows()
preds %>% 
  group_by(type, species) %>% 
  mutate(SpRelAbund = sum(Yobs)/1000) %>% 
  ungroup() %>% 
  ggplot(aes(x = x, col = as.factor(species))) + 
  geom_ribbon(aes(ymin = Y5, ymax = Y95), alpha = 0.2) +
  geom_line(aes(y = Y), lwd = 1.3) +
  geom_hline(aes(yintercept =  SpRelAbund, col = as.factor(species)), 
             linetype = "dashed") +
  facet_wrap(~ type, nrow = 3) +
  theme(legend.position = "bottom") +
  scale_y_sqrt()
```

# Discussion

