---
title: "A03: Niche descriptors"
date: '`r Sys.Date()`'
author: Sylvain Schmitt
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
    theme: flatly
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---
```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
library(parallel)
library(tidyverse)
library(ggfortify)
library(raster)
library(leaflet)
library(rstan)
library(bayesplot)
library(factoextra)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
```

```{r trees}
d <- 20
trees <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(CensusYear == 2015) %>%
  filter(Species != "Indet.") %>% 
  mutate(DBH = CircCorr/pi) %>% 
  collect() %>% 
  filter(Xfield > d, Xfield < 250-d, Yfield > d, Yfield < 250-d)
```

```{r abiotic}
# env <- trees %>%
#   dplyr::select(idTree, Xutm, Yutm) %>%
#   unique()
# coordinates(env) <- ~Xutm + Yutm
# proj4string(env) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
# env <- spTransform(env, CRSobj = crs)
# topo <- stack(
#   raster(file.path(path, "topography", "DEM_1m_2015.tif")),
#   raster(file.path(path, "topography", "RelativeElevation_1m.tif")),
#   raster(file.path(path, "topography", "slope_1m.tif")),
#   raster(file.path(path, "topography", "curvature_1m.tif")),
#   raster(file.path(path, "topography", "TRI_1m.tif")),
#   raster(file.path(path, "topography", "TWI_1m.tif"))
# )
# names(topo) <- c("DEM", "RelativeElevation", "Slope", "Curvature", "TRI", "TWI")
# topo <- projectRaster(topo, crs = crs)
# env <- data.frame(cbind(env@data, raster::extract(topo, env)))
# rm(topo)
# save(env, file = "./distribution_save/env.Rdata")
# load("./distribution_save/env.Rdata")
```

```{r biotic}
# treesXY <- trees %>%
#   dplyr::select(idTree, Xutm, Yutm) %>%
#   unique()
# coordinates(treesXY) <- ~Xutm + Yutm
# proj4string(treesXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
# treesXY <- spTransform(treesXY, CRSobj = crs)
# MNC <- raster(file.path(path, "topography", "MNC_ParacouAvril2009_1m.tif"))
# names(MNC) <- c("MNC")
# crs(MNC) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
# MNC <- projectRaster(MNC, crs = crs)
# env$MNC <- raster::extract(MNC, treesXY)
# rm(treesXY, MNC)
# save(env, file = "./distribution_save/env.Rdata")
load("./distribution_save/env.Rdata")
```

```{r Competition}
# cl <- makeCluster(6)
# clusterExport(cl, list("data", "path", "d"))
# Competition <- clusterMap(cl,
#   function(id, plot, census, x, y, genus){
#   library(dplyr)
#   src_sqlite(file.path(path, "trees/Paracou.sqlite")) %>%
#     tbl("Paracou") %>%
#     filter(CensusYear == census) %>%
#     filter(Plot == plot) %>%
#     filter(idTree != id) %>%
#     mutate(DBHj = CircCorr/pi) %>%
#     filter(DBHj >= 10) %>% # != INRA plots
#     mutate(dij = sqrt((x - Xutm)^2+(y - Yutm)^2)) %>%
#     filter(dij < d) %>%
#     rename(idTreej = idTree, Genusj = Genus, Speciesj = Species) %>%
#     mutate(idTree = id) %>%
#     dplyr::select(idTree, idTreej, Genusj, Speciesj, DBHj, dij) %>%
#     collect()},
#   id = trees$idTree,
#   plot = trees$Plot,
#   x = trees$Xutm,
#   y = trees$Yutm,
#   census = trees$CensusYear,
#   genus = trees$Genus,
#   SIMPLIFY = F)
# stopCluster(cl)
# rm(cl)
# Competition <- bind_rows(Competition)
# save(Competition, file = "./distribution_save/Competition.Rdata")
load("./distribution_save/Competition.Rdata")
Competition <- Competition %>% 
  left_join(dplyr::select(trees, idTree, Genus, Species)) %>% 
  group_by(idTree) %>% 
  mutate(dij = ifelse(dij <1, 1, dij)) %>% 
  summarise(NCI = log(sum(DBHj^2*exp(-0.25*dij))+1),
            NCIgenus = log(sum(DBHj^2*exp(-0.25*dij)*as.numeric(Genus == Genusj))+1),
            NCIspecies = log(sum(DBHj^2*exp(-0.25*dij)*as.numeric(Species == Speciesj))+1))
```

```{r complexes}
complexes <- bind_rows(
  data.frame(Complex = "Chartacea", Genus = "Eschweilera",
             Species = c("simiorum", "congestiflora")),
  data.frame(Complex = "Parvifolia", Genus = "Eschweilera",
             Species = c("pedicellata", "coriacea", "decolorans", "sagotiana",
                         "wachenheimii", "grandiflora_form2")),
  data.frame(Complex = "Licania", Genus = "Licania",
             Species = c("alba", "membranacea", "canescens", "micrantha",
                         "ovalifolia", "sprucei", "densiflora",
                         "laxiflora", "parvifructa")),
  data.frame(Complex = "Iryanthera", Genus = "Iryanthera",
             Species = c("hostmannii", "sagotiana")),
  data.frame(Complex = "Talisia", Genus = "Talisia",
             Species = c("hexaphylla", "praealta", "simaboides")),
  data.frame(Complex = "Symphonia", Genus = "Symphonia",
             Species = c("globulifera", "sp.1")))
```

```{r data}
data <- trees %>% 
  left_join(env) %>% 
  left_join(complexes) %>% 
  left_join(Competition) %>% 
  filter(!is.na(Plot))
```

# Introduction

Subsequent analysis aimed to select descriptors of species complex distribution.

Species distribution, and therefore species complex distribution, will be shaped both by abiotic environment and biotic interactions. We will not explore tree interactions with their environment trough herbivory and pathogens, besides their huge impact on tree life history [but see @VanderPutten2001].

We selected descriptors of species distribution by studying descriptors co-variation and keeping only little correlated variables with the best ecological meaning to represent abiotic environment and biotic interactions.

# Material & Methods

Descriptors co-variation has been investigated tow-by-two by correlations and together by principal component analysis. Co-variations have been investigated together and separately for biotic and abiotic descriptors.

# Results

## Correlations

Globally, topographic variables derived from the digital elevation model showed high level of co-variation (correlation between 0.57 and 0.9) at the exception of curvature. Regarding biotic descriptors digital canopy model and neighbor crowding index showed a non negligible correlation of 0.39 and neighbor crowding index for genus and species were correlated to 0.76.

```{r correlations}
data %>% 
  dplyr::select(DBH, 
                DEM, RelativeElevation, Slope, Curvature, TRI, TWI,
                MNC, NCI, NCIgenus, NCIspecies) %>% 
  na.omit() %>% 
  cor(.) %>% 
  corrplot::corrplot.mixed()
```

```{r pca12}
fviz_pca_var(princomp(~ DBH + DEM + RelativeElevation +
                        Slope + Curvature + TRI + TWI +
                        MNC + NCI + NCIgenus + NCIspecies, data, cor = T),
             axes = c(1, 2), geom = c("arrow", "text"), col.var = "contrib")
```

```{r pca13}
fviz_pca_var(princomp(~ DBH + DEM + RelativeElevation +
                        Slope + Curvature + TRI + TWI +
                        MNC + NCI + NCIgenus + NCIspecies, data, cor = T),
             axes = c(1, 3), geom = c("arrow", "text"), col.var = "contrib")
```

## Biotic interactions


```{r pcaBiotic}
fviz_pca_var(princomp(~ MNC + NCI + NCIgenus + NCIspecies, data, cor = T),
             axes = c(1, 2), geom = c("arrow", "text"), col.var = "contrib")
```

## Abiotic environment

```{r pcaAbiotic}
fviz_pca_var(princomp(~ DEM + RelativeElevation +
                        Slope + Curvature + TRI + TWI, data, cor = T),
             axes = c(1, 2), geom = c("arrow", "text"), col.var = "contrib")
```

# Conclusion

We thus selected three little correlated variables ($r < 0.13$) to represent species biotic and abiotic environment: (i) topographic wetness index (TWI) representing  water accumulation areas, crucial to tropical ecosystems [@ferry2010higher], (ii) neighbor crowding index of heterospecific ($NCI_{hetero}$) representing hetersopecific interaction and (iii) neighbor crowding index of conspecific ($NCI_{con}$) representing conspecific interaction

# References
