---
title: "A05: Complexes distributions"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  github_document: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
linestretch: 1.5
---

```{r setup, include=FALSE}
# rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
library(parallel)
library(tidyverse)
library(ggfortify)
library(raster)
library(leaflet)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
```

```{r trees}
d <- 20
trees <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(CensusYear == 2015) %>%
  filter(Xfield > d, Xfield < 250-d, Yfield > d, Yfield < 250-d) %>% 
  filter(Species != "Indet.") %>% 
  mutate(DBH = CircCorr/pi) %>% 
  collect()
```

```{r env}
# env <- trees %>%
#   dplyr::select(idTree, Xutm, Yutm) %>%
#   unique()
# coordinates(env) <- ~Xutm + Yutm
# proj4string(env) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
# env <- spTransform(env, CRSobj = crs)
# topo <- stack(
#   raster(file.path(path, "topography", "DEM_1m_2015.tif")),
#   raster(file.path(path, "topography", "RelativeElevation_1m.tif")),
#   raster(file.path(path, "topography", "slope_1m.tif")),
#   raster(file.path(path, "topography", "curvature_1m.tif")),
#   raster(file.path(path, "topography", "TRI_1m.tif")),
#   raster(file.path(path, "topography", "TWI_1m.tif"))
# )
# names(topo) <- c("DEM", "RelativeElevation", "Slope", "Curvature", "TRI", "TWI")
# topo <- projectRaster(topo, crs = crs)
# env <- data.frame(cbind(env@data, raster::extract(topo, env)))
# rm(topo)
# save(env, file = "./distribution_save/env.Rdata")
load("./distribution_save/env.Rdata")
```


```{r Competition}
# cl <- makeCluster(4)
# clusterExport(cl, list("data", "path"))
# Competition <- clusterMap(cl,
#   function(id, plot, census, x, y, genus){
#   library(dplyr)
#   src_sqlite(file.path(path, "trees/Paracou.sqlite")) %>%
#     tbl("Paracou") %>%
#     filter(CensusYear == census) %>%
#     filter(Plot == plot) %>%
#     filter(idTree != id) %>%
#     mutate(DBH = CircCorr/pi) %>%
#     filter(DBH >= 10) %>% # != INRA plots
#     mutate(dij = sqrt((x - Xutm)^2+(y - Yutm)^2)) %>%
#     filter(dij < 10) %>%
#     summarise(BA = sum(0.25*pi*DBH^2),
#               BAgenus =  sum(0.25*pi*DBH^2*as.numeric(Genus == genus))) %>%
#     mutate(idTree = id) %>%
#     dplyr::select(idTree, BA, BAgenus) %>%
#     collect()},
#   id = trees$idTree,
#   plot = trees$Plot,
#   x = trees$Xutm,
#   y = trees$Yutm,
#   census = trees$CensusYear,
#   genus = trees$Genus,
#   SIMPLIFY = F)
# stopCluster(cl)
# rm(cl)
# Competition <- bind_rows(Competition)
# save(Competition, file = "./distribution_save/Competition.Rdata")
load("./distribution_save/Competition.Rdata")
```

```{r complexes}
Complexes <- bind_rows(
  data.frame(Complex = "E. Chartacea", Genus = "Eschweilera",
             Species = c("simiorum", "congestiflora")),
  data.frame(Complex = "E. Parvifolia", Genus = "Eschweilera",
             Species = c("pedicellata", "coriacea", "decolorans", "sagotiana",
                         "wachenheimii", "grandiflora_form2")),
  data.frame(Complex = "Licania", Genus = "Licania",
             Species = c("alba", "membranacea", "canescens", "micrantha",
                         "ovalifolia", "sprucei", "densiflora",
                         "laxiflora", "parvifructa")),
  data.frame(Complex = "Iryanthera", Genus = "Iryanthera",
             Species = c("hostmannii", "sagotiana")),
  data.frame(Complex = "Talisia", Genus = "Talisia",
             Species = c("hexaphylla", "praealta", "simaboides")),
  data.frame(Complex = "Symphonia", Genus = "Symphonia",
             Species = c("globulifera", "sp.1")))
```

```{r data}
data <- trees %>% 
  left_join(env) %>% 
  left_join(Complexes) %>% 
  filter(!is.na(Plot))
```

```{r mdata}
complexes <- unique(Complexes$Complex)
data <- mutate(data, Complex = ifelse(is.na(Complex), "no", Complex))
mdata <- lapply(complexes, function(complex) {
  data_complex <- left_join(data,
                            Competition %>% 
                              left_join(Complexes, by = c("Genusj" = "Genus", "Speciesj" = "Species")) %>%
                              mutate(Complex = ifelse(is.na(Complex), "no", Complex)) %>% 
                              left_join(dplyr::select(trees, idTree, Genus, Species)) %>% 
                              mutate(dij = ifelse(dij <1, 1, dij)) %>% 
                              mutate(other = as.numeric(Complex != complex)) %>% 
                              mutate(con = as.numeric(Genusj == Genus, Speciesj == Species)) %>% 
                              mutate(hetero = ifelse(Complex == complex & con == 0, 1, 0)) %>% 
                              group_by(idTree) %>% 
                              summarise(NCIother = log(sum(DBHj^2*exp(-0.25*dij)*other)+1),
                                        NCIhetero = log(sum(DBHj^2*exp(-0.25*dij)*hetero)+1),
                                        NCIcon = log(sum(DBHj^2*exp(-0.25*dij)*con)+1)))
  list(N = nrow(data_complex),
       K = 4,
       Y = as.numeric(data_complex$Complex == complex),
       X = dplyr::select(data_complex, TWI, NCIother, NCIhetero, NCIcon) %>%
         mutate_all(funs(scale)) %>%
         as.matrix(),
       w = ifelse(data_complex$Complex == complex,
                  1-sum(data_complex$Complex == complex)/length(data_complex$Complex),
                  sum(data_complex$Complex == complex)/length(data_complex$Complex)))
})
names(mdata) <- complexes
reshape2::melt(mdata$Symphonia$X) %>% 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~ Var2, scales = "free")
```

```{r fit}
Model <- stan_model("distribution_models/SingleModelUnweighted.stan")
fits <- lapply(mdata, function(data) sampling(Model, chains = 2, data))
names(fits) <- complexes
save(fits, file = "./distribution_save/complexes.Rdata")
load("./distribution_save/complexes.Rdata")
pars <- c("alpha", "beta", "gamma")
```

# Introduction

The aim of this document is to study the distribution at micro-environmental scale with biotic and abiotic environment of whole species complexes in Paracou (not species within complexes).

# Material and Methods

## Study site

Paracou

## Individuals

Ddefined in document [A04-Complexes](A04-Complexes.html):

* *E. charatcaea*
* *E. Parvifolia*
* *Licania*
* *Iryanthera*
* *Talisia*
* *Symphonia*

## Environmental variables

Defined in document [A03-Descriptors](A03-Descriptors.html):

* TWI
* $NCI_{other}$
* $NCI_{complex}$

## Analysis

We analysed species complex distribution by comparing species complex occurrence, considered as presence, against all occurrences of tree belonging to other species, considered as pseudo-absence. Species complex distribution has been investigating through logistic regression following a Bernoulli distribution (model with the best likelihood and results see document [A01-SingleModel](A01-SingleModel.html)):

$$Presence \sim \mathcal{Bernoulli}[logit^{-1}(\alpha + \beta*X+\gamma*X^2)]$$
with $X$ the matrix of environmental descriptors of size $[N,K]$, $\alpha$ the intercept, $\beta$ a vector of size $K$ representing the slope and $\gamma$ a vector of size $K$ representing the quadratic form.

Environmental descriptors were all scaled for the model inference, in order to ease model inference and later compare strength of effects between traits and between effects. A Bayesian method was used to infer parameters of the model regarding each leaf traits using `stan` language and `rstan` package in the R environment (see supplementary material [S1: Model] for model equations and `stan` code).

# Results

The model inference correctly converged without divergent transition and resulted in a correct predictions density distribution (see supplementary material [S2: Diagnostic]).

```{r summary}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = pars, droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "complex") %>% 
  mutate(descriptor = gsub("([[:alpha:]])", "", term)) %>% 
  dplyr::select(complex, term, descriptor, estimate) %>% 
  mutate(descriptor = gsub("([[:punct:]])", "", descriptor)) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "TWI", 
                                    `2` = "NCIother", `3` = "NCIhetero",
                                    `4` = "NCIcon")) %>% 
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(term = recode_factor(term, alpha = "$\\alpha$", 
                              beta = "$\\beta$", 
                              gamma = "$\\gamma$")) %>% 
  reshape2::dcast(term + descriptor ~ complex, value.var = "estimate") %>% 
  kable(caption = "Model parameters for each complex with scaled descriptors.",
        escape = F, digits = 3, format = "pandoc") %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1)
```

```{r posterior, fig.cap="Model parameters posterior."}
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit, pars = pars))) %>% 
  bind_rows(.id = "complex") %>% 
  mutate(descriptor = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(descriptor = gsub("([[:punct:]])", "", descriptor)) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "TWI", 
                                    `2` = "NCIother", `3` = "NCIhetero",
                                    `4` = "NCIcon")) %>% 
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>%
  filter(parameter != "alpha") %>% 
  dplyr::select(-outer_width, -inner_width, -point_est) %>% 
  dplyr::select(complex, parameter, descriptor, ll, l, m, h, hh) %>% 
  ggplot(aes(x = complex, xend = complex, col = descriptor, fill = descriptor)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_grid(descriptor ~ parameter, labeller = label_parsed, scales = "free_y") +
  xlab("") + ylab("") + 
  scale_fill_discrete(guide = "none") +
  scale_color_discrete(guide = "none")
```

```{r prediciton, fig.cap="Predictions."}
preds <- lapply(names(fits), function(complex){

  TWI <- seq(from = min(mdata[[complex]]$X[,1]), 
             to = max(mdata[[complex]]$X[,1]), length.out = 100)
  NCIother <- seq(from = min(mdata[[complex]]$X[,2]), 
             to = max(mdata[[complex]]$X[,2]), length.out = 100)
  NCIhetero <- seq(from = min(mdata[[complex]]$X[,3]), 
             to = max(mdata[[complex]]$X[,3]), length.out = 100)
  NCIcon <- seq(from = min(mdata[[complex]]$X[,4]), 
                   to = max(mdata[[complex]]$X[,4]), length.out = 100)

  alpha <- as.matrix(fits[[complex]], pars = "alpha")
  beta <- as.matrix(fits[[complex]], pars = "beta")
  gamma <- as.matrix(fits[[complex]], pars = "gamma")

  Ytwi <- do.call("rbind", sapply(1: nrow(alpha), function(n) (arm::invlogit(alpha[n] + beta[n,1]*TWI + beta[n,2]*mean(NCIother) + beta[n,3]*mean(NCIhetero) + beta[n,4]*mean(NCIcon) + gamma[n,1]*TWI^2 + gamma[n,2]*mean(NCIother)^2 + gamma[n,3]*mean(NCIhetero)^2 + gamma[n,4]*mean(NCIcon)^2)), simplify = F))
  Ynciother <- do.call("rbind", sapply(1: nrow(alpha), function(n) (arm::invlogit(alpha[n] + beta[n,1]*mean(TWI) + beta[n,2]*NCIother + beta[n,3]*mean(NCIhetero) + beta[n,4]*mean(NCIcon) + gamma[n,1]*mean(TWI)^2 + gamma[n,2]*NCIother^2 + gamma[n,3]*mean(NCIhetero)^2 + gamma[n,4]*mean(NCIcon)^2)), simplify = F))
  Yncihetero <- do.call("rbind", sapply(1: nrow(alpha), function(n) (arm::invlogit(alpha[n] + beta[n,1]*mean(TWI) + beta[n,2]*mean(NCIother) + beta[n,3]*NCIhetero + beta[n,4]*mean(NCIcon) + gamma[n,1]*mean(TWI)^2 + gamma[n,2]*mean(NCIother)^2 + gamma[n,3]*NCIhetero^2 + gamma[n,4]*mean(NCIcon)^2)), simplify = F))
  Yncicon <- do.call("rbind", sapply(1: nrow(alpha), function(n) (arm::invlogit(alpha[n] + beta[n,1]*mean(TWI) + beta[n,2]*mean(NCIother) + beta[n,3]*mean(NCIhetero) + beta[n,4]*NCIcon + gamma[n,1]*mean(TWI)^2 + gamma[n,2]*mean(NCIother)^2 + gamma[n,3]*mean(NCIhetero)^2 + gamma[n,4]*NCIcon^2)), simplify = F))
  
  bind_rows(
    data.frame(
      Complex = complex,
      variable = "TWI",
      value = TWI,
      Y = apply(Ytwi, 2, mean),
      Y5 = apply(Ytwi, 2, quantile, probs = 0.05),
      Y95 = apply(Ytwi, 2, quantile, probs = 0.95)),
    data.frame(
      Complex = complex,
      variable = "NCIother",
      value = NCIother,
      Y = apply(Ynciother, 2, mean),
      Y5 = apply(Ynciother, 2, quantile, probs = 0.05),
      Y95 = apply(Ynciother, 2, quantile, probs = 0.95)),
    data.frame(
      Complex = complex,
      variable = "NCIhtereo",
      value = NCIhetero,
      Y = apply(Yncihetero, 2, mean),
      Y5 = apply(Yncihetero, 2, quantile, probs = 0.05),
      Y95 = apply(Yncihetero, 2, quantile, probs = 0.95)),
    data.frame(
      Complex = complex,
      variable = "NCIcon",
      value = NCIcon,
      Y = apply(Yncicon, 2, mean),
      Y5 = apply(Yncicon, 2, quantile, probs = 0.05),
      Y95 = apply(Yncicon, 2, quantile, probs = 0.95))
  )}) %>% bind_rows()
preds %>% 
  left_join(group_by(data, Complex) %>% 
              summarise(Yo = n()/nrow(.))) %>% 
  ggplot(aes(x = value, col = Complex)) +
  geom_ribbon(aes(ymin = Y5, ymax = Y95), alpha = 0.2) +
  geom_line(aes(y = Y)) +
  geom_hline(aes(yintercept = Yo, col = Complex), lty = "dashed") +
  facet_wrap(~ variable, nrow = 1) + 
  xlab("") + ylab("Species complex relative abundance") +
  scale_y_sqrt() +
  theme(legend.position = "bottom")
```

```{r PosteriorNPrediction, fig.cap="Model parameters posterior and predictions."}
g1 <-lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit, pars = "beta"))) %>% 
  bind_rows(.id = "complex") %>% 
  mutate(descriptor = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(descriptor = gsub("([[:punct:]])", "", descriptor)) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "TWI", 
                                    `2` = "NCIother", `3` = "NCIhetero",
                                    `4` = "NCIcon")) %>% 
  mutate(descriptor = factor(descriptor, levels = c("NCIcon", "NCIhetero", 
                                                    "NCIother", "TWI"))) %>% 
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>%
  dplyr::select(-outer_width, -inner_width, -point_est) %>% 
  dplyr::select(complex, parameter, descriptor, ll, l, m, h, hh) %>% 
  ggplot(aes(x = complex, xend = complex, col = complex, fill = complex)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_wrap(~ descriptor, labeller = label_parsed, scales = "free_y", nrow = 1) +
  ylab("") + xlab("Species\ncomplex") + 
  scale_fill_discrete(guide = "none") +
  scale_color_discrete(guide = "none") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
g2 <-preds %>% 
  left_join(group_by(data, Complex) %>% 
              summarise(Yo = n()/nrow(.))) %>% 
  ggplot(aes(x = value, col = Complex)) +
  geom_ribbon(aes(ymin = Y5, ymax = Y95), alpha = 0.2) +
  geom_line(aes(y = Y)) +
  geom_hline(aes(yintercept = Yo, col = Complex), lty = "dashed") +
  facet_wrap(~ variable, nrow = 1, scales = "free") + 
  xlab("") + ylab("Species complex relative abundance") +
  scale_y_sqrt() +
  theme(legend.position = "bottom",
        strip.background = element_blank(),
        strip.text.x = element_blank())
cowplot::plot_grid(g1, g2, nrow = 2, rel_heights = c(1,2), 
                   labels = LETTERS[1:2], label_x = 0.1)
```

# Discussion

# Supplementray materials

## S1: Model

```{stan Model, output.var="Model", echo=T, eval=F, cache = F}
data {
  int<lower=1> N ;                // # of individuals
  int<lower=1> K ;                // # of environmental descriptors
  int<lower=0, upper=1> Y[N] ;    // individuals presence or absence (0-1) 
  matrix[N,K] X ;                 // environmental descriptors
  vector<lower=0, upper=1>[N] w ; // weights for pseudo-absence to be equal to presence
}
parameters {
  real alpha ;       // intercept
  vector[K] beta ;   // sigmoidal slope
  vector[K] gamma ;  // quadratic form
}
model {
  alpha ~ normal(0, 10^6) ; // priors
  beta ~ normal(0, 10^6) ;
  gamma ~ normal(0, 10^6) ;
  target += w*bernoulli_lpmf(Y | inv_logit(alpha + X * beta + X .* X * gamma)) ; // likelihood
}
```

## S2: Diagnostic

Model inference converged ($\hat R = 1$, figure \@ref(fig:mcmc)) for every complex without divergent transition, and relatively uncorrelated parameters.

```{r mcmc, fig.cap="Markov chains trace plot after warmup for leaf thickness model."}
# lapply(fits, check_divergences)
mcmc_trace(as.array(fits$Symphonia, pars = c(pars, "lp__")),
             facet_args = list(labeller = label_parsed))
# mcmc_pairs(as.array(fits$invSLA), regex_pars = pars)
```

# References