---
title: Pervasive local-scale species niche differentiation with topography within widespread Neotropical tree species complexes
author: Sylvain Schmitt$^{*,1}$, Niklas Tysklind$^2$, Géraldine Derroire$^2$, Myriam Heuertz$^1$, Bruno Herault$^{3,4}$
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
```

```{r complexes}
Complexes <- bind_rows(
  data.frame(Complex = "Chartacea", Genus = "Eschweilera",
             Species = c("simiorum", "congestiflora")),
  data.frame(Complex = "Parvifolia", Genus = "Eschweilera",
             Species = c("pedicellata", "coriacea", "decolorans", "sagotiana",
                         "wachenheimii", "grandiflora_form2")),
  data.frame(Complex = "Licania", Genus = "Licania",
             Species = c("alba", "membranacea", "canescens", "micrantha",
                         "ovalifolia", "sprucei", "densiflora",
                         "laxiflora", "parvifructa")),
  data.frame(Complex = "Iryanthera", Genus = "Iryanthera",
             Species = c("hostmannii", "sagotiana")),
  data.frame(Complex = "Talisia", Genus = "Talisia",
             Species = c("hexaphylla", "praealta", "simaboides")),
  data.frame(Complex = "Symphonia", Genus = "Symphonia",
             Species = c("globulifera", "sp.1")))
```

```{r data}
d <- 20
trees <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(CensusYear == 2015) %>%
  filter(Xfield > d, Xfield < 250-d, Yfield > d, Yfield < 250-d) %>% 
  filter(Species != "Indet.") %>% 
  mutate(DBH = CircCorr/pi) %>% 
  collect()
load("./distribution_save/env.Rdata")
```

```{r predComplex}
load("./distribution_save/Competition.Rdata")
data <- trees %>% 
  left_join(env) %>% 
  left_join(Complexes) %>% 
  filter(!is.na(Plot))
complexes <- unique(Complexes$Complex)
data <- mutate(data, Complex = ifelse(is.na(Complex), "no", Complex))
mdata <- lapply(complexes, function(complex) {
  data_complex <- left_join(data,
                            Competition %>% 
                              left_join(Complexes, by = c("Genusj" = "Genus", "Speciesj" = "Species")) %>% 
                              mutate(Complex = ifelse(is.na(Complex), "no", Complex)) %>% 
                              mutate(dij = ifelse(dij <1, 1, dij)) %>% 
                              mutate(strata = as.numeric(Complex == complex)) %>% 
                              group_by(idTree) %>% 
                              summarise(NCIother = log(sum(DBHj^2*exp(-0.25*dij)*abs(strata-1))+1),
                                        NCIcomplex = log(sum(DBHj^2*exp(-0.25*dij)*strata)+1)))
  list(N = nrow(data_complex),
       K = 3,
       Y = as.numeric(data_complex$Complex == complex),
       X = dplyr::select(data_complex, TWI, NCIother, NCIcomplex) %>%
         mutate_all(funs(scale)) %>%
         as.matrix(),
       w = ifelse(data_complex$Complex == complex,
                  1-sum(data_complex$Complex == complex)/length(data_complex$Complex),
                  sum(data_complex$Complex == complex)/length(data_complex$Complex)))
})
names(mdata) <- complexes
load("./distribution_save/complexes.Rdata")
names(fits)[1:2] <- c("Chartacea", "Parvifolia")
fitsComplex <- fits
predComplex <- lapply(names(fits), function(complex){

  TWI <- seq(from = min(mdata[[complex]]$X[,1]), 
             to = max(mdata[[complex]]$X[,1]), length.out = 100)
  NCIother <- seq(from = min(mdata[[complex]]$X[,2]), 
             to = max(mdata[[complex]]$X[,2]), length.out = 100)
  NCIcomplex <- seq(from = min(mdata[[complex]]$X[,3]), 
             to = max(mdata[[complex]]$X[,3]), length.out = 100)

  alpha <- as.matrix(fits[[complex]], pars = "alpha")
  beta <- as.matrix(fits[[complex]], pars = "beta")
  gamma <- as.matrix(fits[[complex]], pars = "gamma")

  Ytwi <- do.call("rbind", sapply(1: nrow(alpha), function(n) (arm::invlogit(alpha[n] + beta[n,1]*TWI + beta[n,2]*mean(NCIother) + beta[n,3]*mean(NCIcomplex) + gamma[n,1]*TWI^2 + gamma[n,2]*mean(NCIother)^2 + gamma[n,3]*mean(NCIcomplex)^2)), simplify = F))
  Ynciother <- do.call("rbind", sapply(1: nrow(alpha), function(n) (arm::invlogit(alpha[n] + beta[n,1]*mean(TWI) + beta[n,2]*NCIother + beta[n,3]*mean(NCIcomplex) + gamma[n,1]*mean(TWI)^2 + gamma[n,2]*NCIother^2 + gamma[n,3]*mean(NCIcomplex)^2)), simplify = F))
  Yncicomplex <- do.call("rbind", sapply(1: nrow(alpha), function(n) (arm::invlogit(alpha[n] + beta[n,1]*mean(TWI) + beta[n,2]*mean(NCIother) + beta[n,3]*NCIcomplex + gamma[n,1]*mean(TWI)^2 + gamma[n,2]*mean(NCIother)^2 + gamma[n,3]*NCIcomplex^2)), simplify = F))

  bind_rows(
    data.frame(
      Complex = complex,
      variable = "TWI",
      value = TWI,
      Y = apply(Ytwi, 2, mean),
      Y5 = apply(Ytwi, 2, quantile, probs = 0.05),
      Y95 = apply(Ytwi, 2, quantile, probs = 0.95)),
    data.frame(
      Complex = complex,
      variable = "NCIother",
      value = NCIother,
      Y = apply(Ynciother, 2, mean),
      Y5 = apply(Ynciother, 2, quantile, probs = 0.05),
      Y95 = apply(Ynciother, 2, quantile, probs = 0.95)),
    data.frame(
      Complex = complex,
      variable = "NCIcomplex",
      value = NCIcomplex,
      Y = apply(Yncicomplex, 2, mean),
      Y5 = apply(Yncicomplex, 2, quantile, probs = 0.05),
      Y95 = apply(Yncicomplex, 2, quantile, probs = 0.95)))
}) %>% bind_rows()
```

```{r predSpecies}
load("./distribution_save/Competition.Rdata")
Competition <- Competition %>% 
  mutate(dij = ifelse(dij <1, 1, dij)) %>% 
  left_join(dplyr::select(trees, idTree, Genus, Species)) %>% 
  mutate(strata = as.numeric(Genusj == Genus, Speciesj == Species)) %>% 
  group_by(idTree) %>% 
  summarise(NCIhetero = log(sum(DBHj^2*exp(-0.25*dij)*abs(strata-1))+1),
            NCIcon = log(sum(DBHj^2*exp(-0.25*dij)*strata)+1))
data <- trees %>% 
  left_join(env) %>% 
  left_join(Complexes) %>% 
  left_join(Competition) %>% 
  filter(!is.na(Plot)) %>% 
  filter(!is.na(Complex)) %>% 
  group_by(Complex, Species) %>% 
  filter(n() > 10) %>%
  ungroup()
complexes <- unique(Complexes$Complex)
mdata <- lapply(complexes, function(complex) {
  data<- filter(data, Complex == complex)
  list(N = nrow(data),
       S = length(unique(data$Species)),
       K = 3,
       Y = sapply(levels(as.factor(data$Species)),
                  function(sp) as.numeric(data$Species == sp)),
       X = dplyr::select(data, TWI, NCIhetero, NCIcon) %>%
         mutate_all(funs(scale)) %>%
         as.matrix())
})
names(mdata) <- complexes
load("./distribution_save/species.Rdata")
predSpecies <- lapply(complexes, function(complex){

  N <- 25
  TWI <- seq(from = min(mdata[[complex]]$X[,1]), 
             to = max(mdata[[complex]]$X[,1]), length.out = N)
  BA <- seq(from = min(mdata[[complex]]$X[,2]), 
             to = max(mdata[[complex]]$X[,2]), length.out = N)
  BAgenus <- seq(from = min(mdata[[complex]]$X[,3]), 
             to = max(mdata[[complex]]$X[,3]), length.out = N)
  
  S <- mdata[[complex]]$S
  K <- mdata[[complex]]$K
  alpha <- as.matrix(fits[[complex]], pars = "alpha")
  beta0 <- as.matrix(fits[[complex]], pars = "beta")
  gamma0 <- as.matrix(fits[[complex]], pars = "gamma")
  beta <- array(dim = c(nrow(alpha), S, K))
  gamma <- array(dim = c(nrow(alpha), S, K))
  for(k in 1:K){
    beta[,,k] <- beta0[,seq.int(1, 3*S, S)[k]:(seq.int(1, 3*S, S)+(S-1))[k]]
    gamma[,,k] <- gamma0[,seq.int(1, 3*S, S)[k]:(seq.int(1, 3*S, S)+(S-1))[k]]
  }
    
  Yp <- abind::abind(sapply(1:length(TWI), function(i) 
    sapply(1: nrow(alpha), function(n)
      tsensembler:::softmax(alpha[n,] + beta[n,,1] * TWI[i] + beta[n,,2]*BA[i] + beta[n,,3]*BAgenus[i] + gamma[n,,1]*TWI[i]^2 + gamma[n,,2]*BA[i]^2 + gamma[n,,3]*BAgenus[i]^2)
    ), simplify = F), along = 3)
  Ytwi <- abind::abind(sapply(1:length(TWI), function(i) 
    sapply(1: nrow(alpha), function(n)
      tsensembler:::softmax(alpha[n,] + beta[n,,1]*TWI[i] + beta[n,,2]*mean(BA) + beta[n,,3]*mean(BAgenus) + gamma[n,,1]*TWI[i]^2 + gamma[n,,2]*mean(BA)^2 + gamma[n,,3]*mean(BAgenus)^2)
    ), simplify = F), along = 3)
  Yba <- abind::abind(sapply(1:length(TWI), function(i) 
    sapply(1: nrow(alpha), function(n)
      tsensembler:::softmax(alpha[n,] + beta[n,,1]*mean(TWI) + beta[n,,2]*BA[i] + beta[n,,3]*mean(BAgenus) + gamma[n,,1]*mean(TWI)^2 + gamma[n,,2]*BA[i]^2 + gamma[n,,3]*mean(BAgenus)^2)
    ), simplify = F), along = 3)
  Ybagenus <- abind::abind(sapply(1:length(TWI), function(i) 
    sapply(1: nrow(alpha), function(n)
      tsensembler:::softmax(alpha[n,] + beta[n,,1]*mean(TWI) + beta[n,,2]*mean(BA) + beta[n,,3]*BAgenus[i] + gamma[n,,1]*mean(TWI)^2 + gamma[n,,2]*mean(BA)^2 + gamma[n,,3]*BAgenus[i]^2)
    ), simplify = F), along = 3)

  bind_rows(
    lapply(1:S, function(sp)
      data.frame(
        complex = complex,
        species = colnames(mdata[[complex]]$Y)[sp],
        variable = "TWI",
        value = TWI,
        Yp = apply(Yp[sp,,], 2, mean),
        Y = apply(Ytwi[sp,,], 2, mean),
        Y5 = apply(Ytwi[sp,,], 2, quantile, probs = 0.05),
        Y95 = apply(Ytwi[sp,,], 2, quantile, probs = 0.95))) %>% 
      bind_rows(),
    lapply(1:S, function(sp)
      data.frame(
        complex = complex,
        species = colnames(mdata[[complex]]$Y)[sp],
        variable = "NCIhetero",
        value = BA,
        Yp = apply(Yp[sp,,], 2, mean),
        Y = apply(Yba[sp,,], 2, mean),
        Y5 = apply(Yba[sp,,], 2, quantile, probs = 0.05),
        Y95 = apply(Yba[sp,,], 2, quantile, probs = 0.95))) %>% 
      bind_rows(),
    lapply(1:S, function(sp)
      data.frame(
        complex = complex,
        species = colnames(mdata[[complex]]$Y)[sp],
        variable = "NCIcon",
        value = BAgenus,
        Yp = apply(Yp[sp,,], 2, mean),
        Y = apply(Ybagenus[sp,,], 2, mean),
        Y5 = apply(Ybagenus[sp,,], 2, quantile, probs = 0.05),
        Y95 = apply(Ybagenus[sp,,], 2, quantile, probs = 0.95))) %>% 
      bind_rows())
}) %>% bind_rows()
```

$^1$ UMR BIOGECO, Université de Bordeaux - INRA, 69 route d’Arcachon, F-33612 Cestas cedex France

$^2$ UMR ECOFOG (Agroparistech, CNRS, INRA, Université des
Antilles, Université de la Guyane, Cirad), Campus Agronomique, 97310 Kourou, French Guiana

$^3$ Cirad, Univ Montpellier, UR Forests & Societies, Montpellier, France

$^4$ INP-HB, Institut National Polytechnique Félix Houphouët-Boigny, BP 1093, Yamoussoukro, Ivory Coast

__$^*$ Corresponding author:__ sylvain.m.schmitt@gmail.com

\newpage 

# Abstract

> To be written

## Keywords

species complex; species distribution model; species; ...

\newpage 

# Introduction 

Tropical forests and more especially Neotropics shelter the highest level of biodiversity worldwide [@Gaston2000], reconginzed more than 40 years ago [@connell_diversity_1978]. Lowland Amazonia include 4962 known tree species over 810 genera [@TerSteege2013]. Many Neotropical genera are thus species rich and present higher level of genetic polyphormisms than species-poor genera, likely related to introgressions between sister species [@Caron2019]. Among this species-rich genera, several are qualified of species complexes because they share a large ammount of genetic variation due to a recent common ancestor and/or hybridization [@baraloto_using_2012-1]. In addition, species linked by hybridization form syngameons, generally non-randomly structured among plants [@Boecklen2017].

<!-- **Species complexes importance in evolution ?** -->
Species complexes are key to understand and test hypothesis on species delimitation, and thus help understanding the processes that have promoted neotropical diversification [@Pinheiro2018].
Simulations suggest that low hybridization success promote species-coexistence in the community, resulting syngameaon may be selectively advantageous, and can be considred as genomic mutalism, decreasing risk of extinction while increasing local adaptation [@Cannon2015b].

<!-- **Other informations ?** -->
Despite the putative importance of species complexes in evolution, a little is known about the ecological distribution of the complex and the belonging species' niches.
Congeneric taxa have been suggested to present pervasive species-habitat associations with topography in French Guiana [@Allie2015].
Nutrient-ppor white-sand and clay forests specialist species are often found in the same genera in the Peruvian Amazon, and sister species edaphic specialization is promoted by herbivores [@Fine2004].
*Pagamea coriacea* (Rubiaceae) live in sympatry with two differentiated population living in more shaded habitats and over sandstone bedrocks opposed to deeper sandy soils and canopy individuals [@Esteves2013].

**Whay and What we did**
**Our questions**
**Our hypotheses**


List of articles that might be used:

- marques 2014
- Seehausen 2004
- Whitney et al., 2010

- Cannon Lerdeau 2019
- Levi 2019
- Berehegaray 2015 (minor)
- Cronk Suarez Gonzalez 2018
- Barluenga & Meyer 2004
- Barluenga 2006
- gene flow: Anderson, 1953;Morjan and Rieseberg, 2004; Arnold et al., 2008b; Givnish, 2010
- Advantageous alleles Fitzpatrick et al., 2010
- genetic rescue Baskett andGomulkiewicz, 2011

# Material and Methods

## Study site

The study was conducted in the northernmost part of the Guiana Plateau region, at the Paracou field station. The site was characterized by an average of 3041 mm annual rainfall and a mean air temperature of 25.71 °C. Old tropical forest with an exceptional richness (over 750 woody species) has developed among the succession of small hills of this area, rising to 10–40 m a.s.l. [@Gourlet-Fleury2004].

The site is made of 16 permanent plots (fifteen 6.25 ha plus one 25 ha) which have been censused (DBH>10) every 1-2 years for more than 35 years. Nine of the plots were logged and subjected to human-induced disturbance in 1986.

## Species complexes

Species complex were defined based on evidence for low phylogenetic resolution or plastid DNA sharing in the clade [@baraloto_using_2012-1; @Gonzalez2009]: *Eschweilera* with clades Parvifolia and Chartacea (Lecythidaceae; Chave *et al*, unpublished; Heuertz *et al*. unpublished); *Licania* [Chrysobalanaceae; @Bardon2016]; *Iryanthera* (Myristicaceae); *Symphonia* (Clusiaceae); *Talisia* (Sapindaceae). We further removed species with less than 10 individuals in Paracou in 2015, resulting in 6 species complexes including 2 to 9 species (Table \@ref(tab:ComplexTab)).

## Environmental variables

Three little correlated (Pearson's r < 0.13, see supplementary material S1) environmental descriptors were chosen to depict species and species complexes distribution. Topographic wetness index ($TWI$) was selected between various abiotic descriptors available and highlighted water accumulation areas, which is crucial information at local scale when it comes to tropical ecosystems [@ferry2010higher].

A neighbor crowding index [$NCI$; @Uriarte2004a] was used as a descriptor of biotic interactions through tree neighborhood, implying that other biotic effects, such as herbivory and pathogen effects, were not directly taken into account. The neighborhood crowding index $NCI_i$ from tree individual $i$ was calculated with the following formula:
$$NCI_i = \sum _{j|\delta_{i,j}<20m} ^{J_i} DBH_j ^2 e^{-\frac{1}{4}\delta_{i,j}}$$
where $DBH_j$ is the diameter of neighboring tree $j$ and $\delta_{i,j}$ its distance to individual tree $i$. $NCI_i$ is computed for all neighbors at a distance $\delta_{i,j}$ inferior to 20m because NCI showed very little effect beyond 20m in preliminary analysis. Size effect of neighbors through their diameter was set to 2 to represent neighbors through their basal area. Distance effect of neighbors was set to -0.25 corresponding to neighbors beyond 20m having less than 1% effect compared to the effect of neighbors at 0m. The neighbor crowding index was further split between neighbors belonging to the same taxa
(neighbors of the complex $NCI_{complex}$ at the species complex level and conspecific neighbors $NCI_{con}$ at the sepcies level) and neighbors of remaining taxa (neighbors not belonging to the complex $NCI_{other}$ at the species complex level and heterospecific neighbors $NCI_{hetero}$ at the sepcies level).

Descriptors used were derived from field censuses ($NCI_{other}$ and $NCI_{genus}$) or derived from a 1-m resolution digital elevation model built using LiDAR campaign done in 2015 ($TWI$) using SAGA-GIS [@Conrad2015].

## Analysis

### Complexes distribution

Species complex distribution was infered by comparing species complex occurrence, considered as presence, against all occurrences of tree belonging to other species, considered as pseudo-absence. Species complex presence $Presence$ was infered with a logistic regression following a Bernoulli distribution (best model form tested, see , see supplementary material S2 and S3):
$$Presence_{species~complex} \sim Bernoulli[logit^{-1}(\alpha + \beta*X+\gamma*X^2)]$$
were $X$ is the matrix of environmental descriptors, $\alpha$ is the intercept, $\beta$ is a vector representing the slope and $\gamma$ is a vector representing the quadratic form of environment variable effects.

### Species distribution

Species joint-distributions were infered within species complexes with softmax regression following a conjugated Dirchlet and Multinomial distribution:
$$Presence_{species|species~complex} \sim Dirichlet~Multinomial(softmax(\alpha + \beta*X + \gamma*X^2))$$
were $Presence_{species|species~complex}$ is the tree species with a vector of 0 and 1 (0 for all species of the species complex it belongs except the tree species equal to 1),  $X$ is the matrix of environmental descriptors, $\alpha$ is the vector of species intercepts, $\beta$ is a matrix representing the slope and $\gamma$ is a matrix representing the quadratic form of environment variable effects for each species.

Environmental descriptors were all scaled for the model inference, in order to ease model inference and later compare strength of effects between distributions. A Bayesian method was used to infer parameters of the model regarding each complex and species using `stan` language [@Carpenter2017] and `rstan` package [@Team2018] in the R environment [@RCoreTeam2018].

# Results

The topographic wetness index ($TWI$) and the neighbor crowding index ($NCI$) had weak effects on relative abundance of species complex (Fig. \@ref(fig:ComplexesDistribution)). Nevertheless, all relative abundances of species complexes increased with $NCI_{complex}$ (except for *Talisia*). And increased value of $TWI$ resulted in an increased relative abundances of *Symphonia* and *Iryanthera*.

On the other hand, the topographic wetness index ($TWI$) strongly influenced species distribution within complexes, with a gradient of response from positive to negative effect within each complex (Fig. \@ref(fig:betaTWI)). For instance, *S. globulifera*, *E. coriacea*, and *I. hostmanii* species relative abundance increased with water accumulation when it decreased for *S. sp1*, *E. sagotiana*, and *I. sagotiana* within *Symphonia*, *Parvifolia* and *Iryanthera* complexes respectively (Fig. \@ref(fig:SpeciesDistribution)). Still, the $TWI$ influence was low for *Chartacea* and *Talisia* complexes.

Finally, $NCI_{hetero}$ and $NCI_{con}$ had no or weak effects on species relative abundance within species complexes. Still conspecific interactions ($NCI_{con}$) had an effect on species from *Parvifolia* and *Symphonia* complexes with a change of dominance within the complex with increasing congeneric interactions (*E. sagotiana* and *E. decolorans* replacing *E. coriacea* and *E. wachenheimii*, *S. globulifera* replacing *S. sp1* within *Parvifolia* and  *Symphonia* complexes respectively). And heterospecific interactions ($NCI_{hetero}$) had an effect on species from *Symphonia* complexe with a change of dominance within the complex with increasing heterospecific interactions (*S. sp1* replacing *S. globulifera*).

# Discussion

> Species complexes of tropical tree are widely spread across habitats at local-scale, but their species show pervasive niche differentiation mainly with water accumulation and a little with congeneric interactions __[WIP]__.

## Species complexes are widely spread across habitats

Biotic and abiotic environment null or weak effect on species complexes distribution suggest that species complexes are widely spread across habitats at local-scale. Topographic wetness index (TWI) reflects topographic position where water accumulates, whereas neighbor crowding index of trees not belonging to the species complex (NCIother) represents competition with neighbors and thus access to water, nutrients and light. Consequently, TWI and NCIother represent the combined effects of abiotic and biotic environment in defining local scale habitats for species and species complexes distributions. In our study, TWI and NCIother had no effect on species complexes distribution, at the exception of a weak positive effect of TWI on *Symphonia* and *Iryanthera* species abundances, highlighting the widespread distribution of species complexes across habitats.

Several of this species complexes belongs to genera already highlighted to be widespread across Amazonia [@TerSteege2013] and known to integrate species with differentiated local habitats preferences [@Allie2015]. Furthermore, we highlighted species complexes distribution across all habitats of small hills from Paracou. *Symphonia* and *Iryanthera* have an advantage in wettest environments, already evidenced experimentally [@Ferry2007], but this result should be taken with caution as the inventory does not integrate palm trees, besides hyper abundant in wet seasonally flodded bottomlands.

On the other hand, neighbor crowding index of trees belonging to the species complex (NCIcomplex) had a positive effect on relative abundances of all species complexes, highlighting trees of the same species complexes to be aggregated in a radius of 20 meters. Species aggregated distribution is a widely observed phenomenon in forests pan-tropically [@Zhang2013; @PLOTKIN2000; @Wang2012; @condit_spatial_2000; @Forget1999]. Aggregation can be explained by dispersal limitation or habitat association. Consequently, due to the weak to null effect of abiotic and biotic environment variables on species complexes distribution, we assumed species complexes to be distributed in aggregates due to the sole effect of dispersal limitation of their species.

## Species show pervasive niche differentiation within species complexes

The opposed response of sister species relative distribution wihtin species complexes of *Iryanthera*, *Licania*, *Symphonia* and *Eschweilera* clade *Parvifolia* to biotic and abiotic environment highlight a pervasive niche differentiation wihtin species complexes among four clades abundant locally [citation] or at the regional scale [@TerSteege2013]. More specifically, sister species of *Iryanthera*, *Licania*, & *Symphonia* species complexes are segregated along a gradient of dry habitats with increased heterospecific tree interaction to wetter habitats with low heterospecific tree interactions. Similarly, sister species of *Eschweilera* clade *Parvifolia* species complex are segregated along a gradient of dry habitats with increased conspecific tree interactions to wetter habitats with low conspecific tree interactions. Moreover, because of the widespread distribution of every species complexes, the sister species relative distribution conditional to the complex distribution is thus conserved across habitats ($P(Species)=P(Complex).P(Species|Complex)$).

Increased dynamic of bottomlands of Paracou lowlands [@ferry2010higher] could explain sister species segregation along a gradient of dry habitats with increased heterospecific tree interaction to wetter habitats with low heterospecific tree interactions due to increased tree mortality and lower tree biomass. 

Janzen Connell for *Parvifolia* ?

Opposed species-habitat association with topography and seasonal flooding between sister species have already been highlighted in litterature across several genera at local [@Allie2015; @Ferry2007] and regional scale [@Guitet2015; @Fine2014]. The pervasive niche differentiation within species complexes with topographic position and seasonal flooding might related to repeated functional strategy adaptation [Schmitt *et al*., in prep; @Ferry2007] to local environment with several diversification events across taxa (*e.g.* @Fine2014 for Protiae).

Evolutionary benefits of habitat specialization ?

# Acknowledgements

We thanks the university of Bordeaux for a PhD grant to Sylvain Schmitt. We are grateful to Pascal Petronelli and the CIRAD inventory team for their work on tree inventories and botanical identification.

# Authors’ contributions

NT and SS conceived the ideas; SS, BH and NT designed methodology; XX analysed model outputs; SS, and XX led the writing of the manuscript. All authors contributed critically to the drafts and gave final approval for publication.

# References

<div id="refs"></div>

\newpage 

## Table captions

__Table \@ref(tab:ComplexTab)__ Identified species complexes with species including more than 10 individuals in Paracou.


## Figure captions

__Figure \@ref(fig:ComplexesDistribution)__ Relative abundance of species complexes. Subplot **A** represent $\beta$ parameter values for each descriptor and species complex, with the circle representing the mean value of the parameter posterior, thick lines the 50% confidence interval and thin lines the 95% confidence interval. Subplot **B**, represent species complex predicted relative abundance, with dashed line indicating observed relative abundance of species complex in Paracou, and solid line and area representing respectivelly the mean and the 95% confidence interval of projected relative abundance of species complexes depending on descriptors. The color indicates the species complex.

__Figure \@ref(fig:betaTWI)__ TWI and NCI heterospecific and conspecific effects on species relative distribution within species complexes. $\beta_{TWI}$, $\beta_{NCIhetero}$, $\beta_{TWIcon}$ posterior are represented for each species complex and species with mean (circle), 50% credibility interval (thick segment) and 95% confidence interval (thin segment).

__Figure \@ref(fig:SpeciesDistribution)__ Relative abundance of species wihtin species complexes. Predicted relative abundance of species within species complexes with dashed line indicating observed relative abundance of species within species complex in Paracou, and solid line and area representing respectivelly the mean and the 95% confidence interval of projected relative abundance of species complexes depending on descriptors. The color indicates the species complex.

\newpage 

<!-- tables -->

```{r ComplexTab}
trees %>% 
  left_join(Complexes, by = c("Genus", "Species")) %>% 
  mutate(Species = ifelse(Species == "grandiflora_form2", "grandiflora", Species)) %>% 
  filter(!is.na(Complex)) %>% 
  group_by(Complex, Genus, Species) %>% 
  summarise(N = n()) %>% 
  arrange(Complex, desc(N)) %>% 
  kable(caption = "Identified species complexes with species including more than 10 individuals in Paracou.", format = "pandoc") %>% 
  kable_styling(full_width = F)
```

<!-- figures -->

```{r ComplexesDistribution, fig.cap="Relative abundance of species complexes. Subplot **A** represent $\\beta$ parameter values for each descriptor and species complex, with the circle representing the mean value of the parameter posterior, thick lines the 50% confidence interval and thin lines the 95% confidence interval. Subplot **B**, represent species complex predicted relative abundance, with dashed line indicating observed relative abundance of species complex in Paracou, and solid line and area representing respectivelly the mean and the 95% confidence interval of projected relative abundance of species complexes depending on descriptors. The color indicates the species complex."}
g1 <-lapply(fitsComplex, function(fit) 
  bayesplot::mcmc_intervals_data(as.array(fit, pars = "beta"))) %>% 
  bind_rows(.id = "complex") %>% 
  mutate(descriptor = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(descriptor = gsub("([[:punct:]])", "", descriptor)) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "TWI", 
                                    `2` = "NCIother", `3` = "NCIcomplex")) %>% 
  mutate(descriptor = factor(descriptor, levels = c("NCIcomplex",
                                                    "NCIother", "TWI"))) %>% 
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>%
  dplyr::select(-outer_width, -inner_width, -point_est) %>% 
  dplyr::select(complex, parameter, descriptor, ll, l, m, h, hh) %>% 
  ggplot(aes(x = complex, xend = complex, col = complex, fill = complex)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_wrap(~ descriptor, labeller = label_parsed, scales = "free_y") +
  ylab("") + xlab("Species\ncomplex") + 
  scale_fill_discrete(guide = "none") +
  scale_color_discrete(guide = "none") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
g2 <-predComplex %>% 
  left_join(group_by(data, Complex) %>% 
              summarise(Yo = n()/nrow(.))) %>% 
  ggplot(aes(x = value, col = Complex)) +
  geom_ribbon(aes(ymin = Y5, ymax = Y95), alpha = 0.2) +
  geom_line(aes(y = Y)) +
  geom_hline(aes(yintercept = Yo, col = Complex), lty = "dashed") +
  facet_wrap(~ variable, nrow = 1, scales = "free") + 
  xlab("") + ylab("Species complex relative abundance") +
  scale_y_sqrt() +
  theme(legend.position = "bottom",
        strip.background = element_blank(),
        strip.text.x = element_blank())
cowplot::plot_grid(g1, g2, nrow = 2, rel_heights = c(1,2), 
                   labels = LETTERS[1:2], label_x = 0.1)
```

```{r betaTWI, fig.cap="TWI and NCI heterospecific and conspecific effects on species relative distribution within species complexes. $\\beta_{TWI}$, $\\beta_{NCIhetero}$, $\\beta_{TWIcon}$ posterior are represented for each species complex and species with mean (circle), 50% credibility interval (thick segment) and 95% confidence interval (thin segment).", fig.height=10} 
species <- data %>%
  dplyr::select(Complex, Genus, Species) %>% 
  unique() %>% 
  mutate(Species = ifelse(Species == "grandiflora_form2", "grandiflora", Species)) %>% 
  group_by(Complex) %>% 
  arrange(Complex, Species) %>% 
  mutate(species = row_number()) %>% 
  mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species))
lapply(fits, function(fit) 
  bayesplot::mcmc_intervals_data(as.array(fit, pars = "beta"))) %>% 
  bind_rows(.id = "Complex") %>% 
  mutate(descriptor = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(descriptor = gsub("([[:punct:]])", "", descriptor)) %>% 
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = as.numeric(substr(descriptor, 1, 1))) %>% 
  left_join(species) %>% 
  mutate(descriptor = substr(descriptor, 2, 2)) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "TWI", 
                                    `2` = "NCIhetero", `3` = "NCIcon")) %>% 
  dplyr::select(Complex, parameter, descriptor, SpeciesLong, ll, l, m, h, hh) %>% 
  mutate(parameter = ifelse(parameter != "alpha",
                            paste0(parameter, "[", descriptor, "]"),
                            parameter)) %>% 
  mutate(Complex = gsub("E. ", "", Complex)) %>% 
  ggplot(aes(x = datatools::reorder_within(SpeciesLong, m, Complex), 
             xend = datatools::reorder_within(SpeciesLong, m, Complex), 
             col = SpeciesLong, fill = SpeciesLong)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_grid(Complex ~ parameter, labeller = label_parsed, scales = "free") +
  datatools::scale_x_reordered() +
  xlab("") + ylab("") + 
  scale_fill_discrete(guide = "none") +
  scale_color_discrete(guide = "none")
```

```{r SpeciesDistribution, fig.cap="Relative abundance of species wihtin species complexes. Predicted relative abundance of species within species complexes with dashed line indicating observed relative abundance of species within species complex in Paracou, and solid line and area representing respectivelly the mean and the 95% confidence interval of projected relative abundance of species complexes depending on descriptors. The color indicates the species complex.", fig.height=10}
predSpecies %>%
    group_by(complex) %>% 
  left_join(data %>% 
              group_by(Complex) %>% 
              mutate(N_complex = n()) %>% 
              group_by(Complex, N_complex, Genus, Species) %>% 
              summarise(N_species = n()) %>% 
              mutate(RelAbund = N_species / N_complex),
            by = c("complex" = "Complex", "species" = "Species")) %>% 
  mutate(species = ifelse(species == "grandiflora_form2", "grandiflora", species)) %>%
  mutate(Species = paste0(substr(Genus, 1,1), ". ", species)) %>% 
  do(g = ggplot(data = ., aes(x = value, col = Species)) + 
       geom_ribbon(aes(ymin = Y5, ymax = Y95), alpha = 0.2) +
       geom_line(aes(y = Y), lwd = 1.3) +
       geom_hline(aes(yintercept = RelAbund, col = Species), linetype = "dashed") +
       facet_grid(complex ~ variable, scales = "free_x") +
       guides(col = guide_legend(nrow=3)) +
       theme(legend.position = "bottom") +
       ylab("Species relative abundance") + xlab("Standardized value") +
       scale_y_sqrt()) %>% 
  gridExtra::grid.arrange(grobs = .$g)
```
