```{r setup_interaction, include=FALSE}
#rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "../../data/Paracou/"
```

```{r data_interaction, eval=T}
# traitsSymphonia <- googlesheets::gs_title("Measures_Symphonia") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   mutate(SLA = as.numeric(SLA), LDMC = as.numeric(LDMC)) %>%
#   mutate(Genus = "Symphonia") %>%
#   rename(Species = Morphotype) %>%
#   mutate(Species = ifelse(Species == "Indet.",
#                           c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G",
#                                      "sp.1" =  "S")], Species))
# traitsEschweilera <- googlesheets::gs_title("Measures_Eschweilera") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   filter(!(Plot == 14 & SubPlot == 1 & TreeFieldNum == 760)) %>%  # outlier
#   filter(!(Species %in% c("congestiflora","simiorum","persistens")))
# traits <- bind_rows(traitsEschweilera, traitsSymphonia) %>%
#   mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>%
#   group_by(idTree, Plot, SubPlot, TreeFieldNum, Genus, Species,
#            SpeciesLong, Bark, Dawkins) %>%
#   summarise_at(vars("SLA", "LDMC", "LT", "LA", "CC",
#                     "brBT", "brBD", "brWD"), mean, na.rm = T) %>%
#   ungroup() %>%
#   mutate(invSLA=1/SLA) %>%
#   mutate(invLA=1/LA)
# load("./functional_save/env.Rdata")
# Individuals <-  left_join(traits, env, by = "idTree", suffix = c("", ".y"))
# rm(traitsEschweilera, traitsSymphonia, traits, paracou, env)
# save(Individuals, file = "./functional_save/Individuals.Rdata")
load("./functional_save/Individuals.Rdata")
load("./functional_save/CompetitionMatrix.Rdata")
sdDBH <- sd(Individuals$DBH)
Individuals <- Individuals %>% 
  rename(LMA = invSLA) %>% 
  group_by(Species) %>% 
  filter(n() > 20) %>% 
  ungroup() %>% 
  left_join(Competition %>% 
              group_by(idTree) %>% 
              summarise(NCI = (1/mean(1-AreaOutside20))*sum(DBHj^2*exp(-0.25*dij)))) %>%
  mutate(NCI = log(NCI)) %>% 
  mutate_at(vars("LMA", "LDMC", "LT", "invLA", "CC", "DBH", "TWI", "NCI"),
            funs(./sd(., na.rm = T)))
```

```{r mdata_interaction}
traits <- c("LMA", "LDMC", "LT", "invLA", "CC")
complexes <- c("Symphonia", "Eschweilera")
models <- sapply(complexes, function(complex) paste0(complex, "-", traits))
mdata <- lapply(models, function(model){
  trait <- strsplit(model, "-")[[1]][2]
  complex <- strsplit(model, "-")[[1]][1]
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    filter(Genus == complex)
  list(N = nrow(data_trait),
       S = length(unique(data_trait$Species)),
       Trait = unlist(data_trait[trait]),
       DBH = data_trait$DBH,
       TWI = data_trait$TWI,
       NCI = data_trait$NCI,
       species = as.numeric(as.factor(data_trait$Species)))
  })
names(mdata) <- models
```

```{r fits_interaction}
# fits <- lapply(mdata, function(x)
#   sampling(Model, chains = 2, data = x, save_warmup = F))
# names(fits) <- models
# save(fits, file = "./functional_save/InteractionReduced.Rdata")
load("./functional_save/InteractionReduced.Rdata")
S_sqrt_trans <- function() scales::trans_new("S_sqrt",
                                             function(x) sign(x)*sqrt(abs(x)),
                                             function(x) x^2*sign(x))
```

```{r species_interaction}
species <- lapply(models, function(model){
  trait <- strsplit(model, "-")[[1]][2]
  complex <- strsplit(model, "-")[[1]][1]
  Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    filter(Genus == complex) %>% 
    mutate(species = as.numeric(as.factor(Species))) %>% 
    mutate(SpeciesLong = paste(substr(complex, 1, 1), Species)) %>% 
    dplyr::select(Genus, Species, SpeciesLong, species) %>% 
    unique()
}) %>% bind_rows() %>% 
  unique() %>% 
  mutate(complex = Genus)
```

```{r predicitons_interaction}
traits <- c("LMA", "LDMC", "LT", "invLA", "CC")
complexes <- c("Symphonia", "Eschweilera")
models <- sapply(complexes, function(complex) paste0(complex, "-", traits))
preds <- lapply(models, function(model){
  trait <- strsplit(model, "-")[[1]][2]
  complex <- strsplit(model, "-")[[1]][1]
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    filter(Genus == complex)
  
  DBH <- data_trait$DBH
  TWI <- data_trait$TWI
  NCI <- data_trait$NCI
  species <- as.numeric(as.factor(data_trait$Species))
  
  alpha <- as.matrix(fits[[model]], pars = "alpha")
  betaDBH <- as.matrix(fits[[model]], pars = "betaDBH")
  betaTWI <- as.matrix(fits[[model]], pars = "betaTWI")
  betaNCI <- as.matrix(fits[[model]], pars = "betaNCI")
  
  Yp <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * TWI +  betaNCI[n,species] * NCI) * (DBH / (betaDBH[n,species] + DBH)), simplify = F))
  Yalpha <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * mean(TWI) +  betaNCI[n,species] * mean(NCI)) * (mean(DBH) / (betaDBH[n,species] + mean(DBH))), simplify = F))
  Ydbh <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * mean(TWI) +  betaNCI[n,species] * mean(NCI)) * (DBH / (betaDBH[n,species] + DBH)), simplify = F))
  Ytwi <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * TWI +  betaNCI[n,species] * mean(NCI)) * (mean(DBH) / (betaDBH[n,species] + mean(DBH))), simplify = F))
  Ynci <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * mean(TWI) +  betaNCI[n,species] * NCI) * (mean(DBH) / (betaDBH[n,species] + mean(DBH))), simplify = F))
  
  data.frame(
    Model = model,
    Trait = trait,
    Complex = complex,
    DBH = data_trait$DBH, 
    TWI = data_trait$TWI, 
    NCI = NCI,
    Genus = data_trait$Genus,
    Species = data_trait$Species,
    Y = unlist(data_trait[trait]),
    Yp = apply(Yp, 2, mean),
    Yp5 = apply(Yp, 2, quantile, probs = 0.05),
    Yp95 = apply(Yp, 2, quantile, probs = 0.95),
    Yalpha = apply(Yalpha, 2, mean),
    Yalpha5 = apply(Yalpha, 2, quantile, probs = 0.05),
    Yalpha95 = apply(Yalpha, 2, quantile, probs = 0.95),
    Ydbh = apply(Ydbh, 2, mean),
    Ydbh5 = apply(Ydbh, 2, quantile, probs = 0.05),
    Ydbh95 = apply(Ydbh, 2, quantile, probs = 0.95),
    Ytwi = apply(Ytwi, 2, mean),
    Ytwi5 = apply(Ytwi, 2, quantile, probs = 0.05),
    Ytwi95 = apply(Ytwi, 2, quantile, probs = 0.95),
    Ynci = apply(Ynci, 2, mean),
    Ynci5 = apply(Ynci, 2, quantile, probs = 0.05),
    Ynci95 = apply(Ynci, 2, quantile, probs = 0.95)
  )}) %>% bind_rows()
```

# Iteraction Model Reduced

## Results

### Summary

```{r summarySpecies}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = c("alpha", "betaDBH", "betaTWI",
                                "betaNCI", "sigma", "lp__"), 
                  droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-") %>% 
  mutate(species = gsub("([[:alpha:]])", "", term)) %>% 
  dplyr::select(complex, trait, term, species, estimate, std.error, rhat) %>% 
  mutate(species = gsub("([[:punct:]])", "", species)) %>% 
  mutate(species = as.numeric(species)) %>% 
  left_join(species) %>% 
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(term = recode_factor(term, 
                              alpha = "$\\alpha$",
                              betaDBH = "$\\beta_{DBH}$", 
                              betaTWI = "$\\beta_{TWI}$", 
                              betaNCI = "$\\beta_{Comp}$",
                              sigma = "$\\sigma^2$",
                              lp = "log-likelihood")) %>% 
  mutate(trait = recode_factor(trait, 
                              LMA = "$LMA$",
                              LDMC = "$LDMC$", 
                              LT = "$LT$", 
                              invLA = "$\\frac{1}{LA}$",
                              CC = "$CC$")) %>% 
  reshape2::dcast(complex + term + SpeciesLong ~ trait, value.var = "estimate") %>% 
  rename(Parameter = term, Species = SpeciesLong) %>% 
  kable(caption = "Model parameters for each species with reduced traits and descriptors.",
        escape = F, digits = 3, format = "pandoc") %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```

### Intercept

```{r speciesIntercept, fig.cap="Model species effect for Intercept."}
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = "alpha"))) %>% 
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-") %>%
  mutate(species = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = gsub("([[:alpha:]])", "", species)) %>% 
  mutate(species = as.numeric(species)) %>% 
  left_join(species) %>% 
  mutate(trait = recode_factor(trait, invLA = "frac(1, LA)")) %>% 
  ggplot(aes(x = datatools::reorder_within(SpeciesLong, m, trait), 
             xend = datatools::reorder_within(SpeciesLong, m, trait),
             color = SpeciesLong, fill = SpeciesLong)) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  datatools::scale_x_reordered() +
  facet_grid(trait ~ ., labeller = label_parsed, scales = "free_y") +
  xaxis_title(F) +
  ylab(expression(alpha)) +
  scale_y_continuous(trans="S_sqrt") +
  scale_color_brewer("", palette = "Set1", guide = "none") +
  scale_fill_brewer("", palette = "Set1", guide = "none")
```

### DBH

```{r speciesDBH, fig.cap="Model species effect for $DBH_{90}$."}
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = "betaDBH"))) %>% 
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-",remove = F) %>%
  mutate(species = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = gsub("([[:alpha:]])", "", species)) %>% 
   mutate(species = as.numeric(species)) %>% 
  left_join(species) %>% 
  mutate(trait = recode_factor(trait, invLA = "frac(1, LA)")) %>% 
  mutate_at(vars("ll", "l", "m", "h", "hh"), funs(.*sdDBH*9)) %>% 
  ggplot(aes(x = datatools::reorder_within(SpeciesLong, m, trait), 
             xend = datatools::reorder_within(SpeciesLong, m, trait),
             color = SpeciesLong, fill = SpeciesLong)) +
  geom_hline(yintercept = c(10, 57, 68), color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  datatools::scale_x_reordered() +
  facet_grid(trait ~ Genus, labeller = label_parsed, 
             scales = "free_y") +
  ylab(expression(DBH[90]~(cm))) +
  scale_y_log10() +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")
```

```{r speciesDBH2, fig.cap="Model species effect for DBH."}
preds %>% 
  mutate(Trait = recode_factor(Trait, invLA = "frac(1, LA)")) %>% 
  ggplot(aes(DBH*sdDBH, col = Species)) +
  geom_point(aes(y = Y), alpha = 0.2) +
  geom_ribbon(aes(ymin = Ydbh5, ymax = Ydbh95), alpha = 0.2) +
  geom_line(aes(y = Ydbh)) +
  facet_grid(Trait ~ Complex, scales = "free", labeller = label_parsed) +
  scale_y_log10()
```

### TWI

```{r speciesTWI, fig.cap="Model species effect for TWI."}
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = "betaTWI"))) %>% 
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-") %>%
  mutate(species = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = gsub("([[:alpha:]])", "", species)) %>% 
    mutate(species = as.numeric(species)) %>% 
  left_join(species) %>% 
  mutate(trait = recode_factor(trait, invLA = "frac(1, LA)")) %>% 
  ggplot(aes(x = datatools::reorder_within(SpeciesLong, m, trait), 
             xend = datatools::reorder_within(SpeciesLong, m, trait),
             color = SpeciesLong, fill = SpeciesLong)) +
  geom_hline(yintercept = 0, color = "gray90", size = 1) +
  geom_hline(yintercept = c(-0.25, 0.25), 
             color = "gray90", size = 0.5, lty = "dashed") +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_grid(trait ~ Genus, labeller = label_parsed, 
              scales = "free_y") +
  datatools::scale_x_reordered() +
  xaxis_title(F) +
  ylab(expression(beta[TWI])) +
  scale_y_continuous(trans="S_sqrt") +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none") +
  ylim(-0.5, 0.5)
```

```{r speciesTWI2, fig.cap="Model species effect for DBH."}
preds %>% 
  mutate(Trait = recode_factor(Trait, invLA = "frac(1, LA)")) %>% 
  ggplot(aes(TWI, col = Species)) +
  geom_point(aes(y = Y), alpha = 0.2) +
  geom_ribbon(aes(ymin = Ytwi5, ymax = Ytwi95), alpha = 0.2) +
  geom_line(aes(y = Ytwi)) +
  facet_grid(Trait ~ Complex, scales = "free", labeller = label_parsed)
```

### NCI

```{r speciesNCI, fig.cap="Model species effect for NCI."}
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = "betaNCI"))) %>% 
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-") %>%
  mutate(species = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = gsub("([[:alpha:]])", "", species)) %>% 
    mutate(species = as.numeric(species)) %>% 
  mutate(trait = recode_factor(trait, invLA = "frac(1, LA)")) %>%
  left_join(species) %>% 
  ggplot(aes(x = datatools::reorder_within(SpeciesLong, m, trait), 
             xend = datatools::reorder_within(SpeciesLong, m, trait),
             color = SpeciesLong, fill = SpeciesLong)) +
  geom_hline(yintercept = 0, color = "gray90", size = 1) +
  geom_hline(yintercept = c(-0.25, 0.25), 
             color = "gray90", size = 0.5, lty = "dashed") +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  datatools::scale_x_reordered() +
    facet_grid(trait ~ Genus, labeller = label_parsed, 
              scales = "free") +
  xaxis_title(F) +
  ylab(expression(beta[NCI])) +
  scale_y_continuous(trans="S_sqrt") +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none") +
  ylim(-0.5, 0.5)
```

```{r speciesNCI2, fig.cap="Model species effect for DBH."}
preds %>% 
  mutate(Trait = recode_factor(Trait, invLA = "frac(1, LA)")) %>% 
  ggplot(aes(NCI, col = Species)) +
  geom_point(aes(y = Y), alpha = 0.2) +
  geom_ribbon(aes(ymin = Ynci5, ymax = Ynci95), alpha = 0.2) +
  geom_line(aes(y = Ynci)) +
  facet_grid(Trait ~ Complex, scales = "free", labeller = label_parsed)
```

### Sensitivity

```{r sensitivity, fig.cap="Model sensitivity to variables. ...", eval=F}
preds %>%
  dplyr::select(Trait, Genus, Species, Yp, Y, Yalpha, Ydbh, Ytwi, Ynci) %>% 
  rename(Intercept = Yalpha, DBH = Ydbh, TWI = Ytwi, NCI = Ynci) %>% 
  reshape2::melt(id.vars = c("Trait", "Genus", "Species", "Yp", "Y"),
                 value.name = "Yvariable") %>% 
  group_by(Trait, Genus, Species, variable) %>%
  mutate(Dvariable = sqrt((Y-Yp)^2), Dmean = sqrt((Y - mean(Y))^2)) %>% 
  group_by(Trait, Genus, Dvariable > Dmean) %>% 
  summarise(n())
  
#   mutate(value = (Yp-Yvariable)/(Yp-mean(Y))*100) %>% 
#   group_by(Trait, Genus, variable) %>%
#   summarise(mean = mean(value), sd = sd(value)) %>% 
#   group_by(Trait, Genus) %>% summarise(sum(mean))
# 
#     mutate(variable = factor(variable, levels = c("DBH", "TWI", "NCI"))) %>%
#     ggplot(aes(variable, fill = variable)) +
#   geom_bar(aes(y = mean), stat = "identity") +
#   geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2) + 
#   facet_grid(Trait ~ Genus, scales = "free") +
#   scale_fill_discrete(guide = "none") +
#   xlab("") + ylab(expression(Delta(Prediction))) +
#   ylim(0,100)
#     
# preds %>% 
#   group_by(Trait, Complex, Species) %>% 
#   mutate(Ymean = mean(Y)) %>% 
#   ggplot(aes(Ymean, Yalpha, col = Species)) + geom_point() +   
#   facet_grid(Trait ~ Genus, scales = "free") +
#   geom_abline(slope = 1, intercept = 1)
```

```{r}
Individuals %>%
  dplyr::select(Genus, Species, DBH, TWI, NCI, LMA, LDMC, LT, invLA, CC) %>%
  reshape2::melt(id.vars = c("Genus", "Species", "DBH", "TWI", "NCI"),
                 variable.name = "Trait") %>%
  group_by(Trait, Genus) %>%
  do(model = lm(value ~ Species + DBH + DBH:TWI + DBH:NCI, data = .)) %>% 
  broom::tidy(model) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = gsub("Species", "", term)) %>% 
  mutate(`P-value` = ifelse(p.value < 0.001, "***",
                            ifelse(p.value < 0.01, "**",
                                   ifelse(p.value < 0.05, "*",
                                          ifelse(p.value < 0.1, ".", 
                                                 "n.s."))))) %>% 
  mutate(Coefficient = round(estimate,2)) %>% 
  mutate(value = paste(Coefficient, `P-value`)) %>%
  reshape2::dcast(term + Genus ~ Trait, value.var = "value") %>% 
  kable(escape = F, format = "pandoc") %>% 
  kable_styling(full_width = F)
```


## Supplementary materials

### S1: Model

We explained leaf trait $Trait$ variation with ontogeny ($DBH$), abiotic environment ($TWI$), biotic interactions ($NCI$) and taxonomic levels (species complex $C$ and species $S$) with a linear mixed model including interaction between ontogeny and environment (more likely than an additive form without interactions, see analysis [A07-Additive](A07-Additive.html)). We obtained the following model (see [Material and Methods] for more detail on the model form):

\begin{equation}
  T_{s,i} \sim \mathcal{N}(\frac{DBH_i}{{\beta_{DBH}}_{s} + DBH_i} \times (\alpha_{s} + {\beta_{TWI}}_{s}.TWI_i + {\beta_{Comp}}_{s}.\sum_{j=1}^{J} DBH_j^{2} . e^{-{\alpha_{NCI}}.\delta_{i,j}}), \sigma)  
\end{equation} 

Resulting in the following `stan` code with selected priors:

```{stan stanModel, output.var="Model", echo=T, eval=F, cache=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1> S ; // # species
  vector[N] Trait ; // Trait
  vector[N] DBH ; // Diameter at Breast Height
  vector[N] TWI ; // Topographic Wetness Index
  vector[N] NCI ; // DBH neighbor
  int<lower=1, upper=S> species[N] ; // species index
}
parameters {
  vector [S] alpha ; // Intercept
  vector<lower=0> [S]  betaDBH ; // DBH half-load
  vector[S] betaTWI ; // TWI slope
  vector[S] betaNCI ; // Competition slope
  real<lower=0> sigma ; // Variance
}
model {
  alpha ~ normal(0, 10^6) ; // Priors
  betaDBH ~ normal(0,1) ;
  betaTWI ~ normal(0,1) ;
  betaNCI ~ normal(0,1) ;
  sigma ~ cauchy(0, 5) ;
  Trait ~ normal((alpha[species] + betaTWI[species] .* TWI +  betaNCI[species] .* NCI) .* (DBH ./ (betaDBH[species] + DBH)), sigma) ; // Likelihood
}
```

### S2: Diagnostic

```{r mcmc, fig.cap="Markov chains trace plot after warmup for leaf thickness model."}
# lapply(fits, check_divergences)
mcmc_trace(as.array(fits$`Symphonia-LT`, pars = c("alpha", "betaDBH", "betaTWI",
                                                  "betaNCI", "sigma", "lp__")),
             facet_args = list(labeller = label_parsed))
```
