---
title: "A01: Traits covariation"
date: '`r Sys.Date()`'
author: Sylvain Schmitt
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(ggfortify)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
path <- "../../data/Paracou/"
```

```{r data}
traitsSymphonia <- googlesheets::gs_title("Measures_Symphonia") %>% 
  googlesheets::gs_read("AllTraits") %>% 
  mutate(SLA = as.numeric(SLA), LDMC = as.numeric(LDMC)) %>% 
  mutate(Genus = "Symphonia") %>% 
  rename(Species = Morphotype)
traitsEschweilera <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("AllTraits") %>% 
  dplyr::select(-DBH)
traits <- bind_rows(traitsEschweilera, traitsSymphonia) %>% 
  mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) ; rm(traitsEschweilera, traitsSymphonia)
paracou <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>%
  tbl("Paracou") %>%
  filter(idTree %in% traits$idTree) %>%
  filter(Genus %in% c("Lecythis", "Eschweilera") & Plot %in% c(1, 6, 11, 13:15) & CensusYear == 2017 |
         Genus %in% c("Lecythis", "Eschweilera") & Plot == 16 & CensusYear == 2015 |
         Genus == "Symphonia" & CensusYear == 2015) %>%
  mutate(DBH = CircCorr*pi) %>% 
  collect()
data <- left_join(traits, paracou) ; rm(traits, paracou)
data <- data %>% # outliers removal
  filter(Plot != 14 & SubPlot != 1 & TreeFieldNum != 760) # rejection individual
data <- list(
  leaf = data %>% 
  dplyr::select(Plot, SubPlot, TreeFieldNum, Leaf,
                Genus, Species, SpeciesLong, Bark, Dawkins, DBH,
                SLA, LDMC, LT, LA, CC) %>% 
  group_by(Plot, SubPlot, TreeFieldNum,
           Genus, Species, SpeciesLong, Bark, Dawkins, DBH) %>% 
  summarise_at(vars("SLA", "LDMC", "LT", "LA", "CC"), mean, na.rm = T) %>% 
  ungroup() %>% 
  drop_na(SLA, LDMC, LT, LA, CC),
  wood = data %>% 
    dplyr::select(Plot, SubPlot, TreeFieldNum, Leaf,
                Genus, Species, SpeciesLong, Bark, Dawkins, DBH,
                SLA, LDMC, LT, LA, CC, brBT, brBD, brWD) %>% 
  group_by(Plot, SubPlot, TreeFieldNum,
           Genus, Species, SpeciesLong, Bark, Dawkins, DBH) %>% 
  summarise_at(vars("SLA", "LDMC", "LT", "LA", "CC", "brBT", "brBD", "brWD"), 
               mean, na.rm = T) %>% 
  ungroup() %>% 
  drop_na(SLA, LDMC, LT, LA, CC, brBT, brBD, brWD)
)
```

# Introduction

*What and why...*
Dataset gathers measures of leaves and wood's trait over 838 trees. From those were calculated different functionnal traits. 
For leaf traits, each individual (although there are few exceptions) had 5 leaves sampled. Functionnal traits were averaged per individual for this part of the anylisis.

Starting with all those traits we want to understand how they are correlated between them, and how individuals shared out and how they distinguished according to those traits.

# Material & Methods

*Traits covariation by PCA...*

# Results

*Describe all observed relations in each case (\@ref(fig:PCAleafComplexes) to refer to a figure)...*

## Leaf traits

### Between complexes

```{r PCAleafComplexes, fig.cap='Principal Component Analysis (PCA) of leaf trait  between complexes.'}
data$leaf %>% 
  autoplot(princomp(~ SLA + LDMC + LT + LA + CC, data = ., cor = T), 
           data = .,
           colour = "Genus", alpha = 0.5, size = "DBH",
           loadings.label.size = 6,
           loadings.label.colour = 'black', loadings.label.vjust = 1.1,
           loadings = T, loadings.label = T, loadings.colour = 'black') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted") +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1))
# g <- ggExtra::ggMarginal(g, groupFill = T)
```
69,54% of total variance is conserved on the two first axis of PCA. On this representation
* CC and LT seem quite correlated
* while CC/LT and SLA are really anti-correlated
* LDMC and LA seem well correlated
* SLA and LA aren't correlated (which is a strange interpretation)
The two species complexes are well separated on this decomposition, what shows they react diferently regarding the axis of LDMC/LA.
Moreover for the two complexes we observed a general trend for DBH repartition along the CC/SLA axis. Apparently smaller DBH seems to be correlated with a higher SLA, which could reflect a strategy of high investment in leaves development for youngest tree (as far as ontogeny and DBH are well correlated)

### Within complexes

```{r PCAleafSymphonia, fig.cap='Principal Component Analysis (PCA) of leaf trait wihtin Symphonia complex.'}
data$leaf %>% 
  filter(Genus == "Symphonia") %>% 
  autoplot(princomp(~ SLA + LDMC + LT + LA + CC, data = ., cor = T), 
           data = .,
           colour = "SpeciesLong", alpha = 0.5, size = "DBH",
           loadings.label.size = 6,
           loadings.label.colour = 'black', loadings.label.vjust = 1.1,
           loadings = T, loadings.label = T, loadings.colour = 'black') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted") +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1))
```
75,12% of total variance is conserved on the two first axis of PCA. On this representation
* CC/LDMC and LT seem not  correlated
* aswell as SLA and LA aren't correlated (which is a strange interpretation)
* while LT and SLA are really anti-correlated
* LDMC and CC seem well correlated
Among Symphonia complex, the two morphotypes seem to well diferenciate along an LDMC/LA axis. As well as for the intercomplex analysis, DBH seems also to create a more obvious gradient along the same LT/SLA axis. 

```{r PCAleafEschweilera, fig.cap='Principal Component Analysis (PCA) of leaf trait wihtin Eschweilera complex.'}
data$leaf %>% 
  filter(Genus != "Symphonia") %>% 
  autoplot(princomp(~ SLA + LDMC + LT + LA + CC, data = ., cor = T), 
           data = .,
           colour = "SpeciesLong", alpha = 0.5, size = "DBH",
           loadings.label.size = 6,
           loadings.label.colour = 'black', loadings.label.vjust = 1.1,
           loadings = T, loadings.label = T, loadings.colour = 'black') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted")
```
65,86% of total variance is conserved on the two first axis of PCA. On this representation:
* CC and LT seem not  correlated
* aswell as SLA and LA aren't correlated (which is a strange interpretation)
* while LT and SLA are really anti-correlated
* LA and CC seem well correlated
First PCA axis is well described by LT/SLA, which draw an axis of conservation/acquisition strategy. The second axis is supported mostly by CC and LA. 
Species which were most sampled show a real diferenciation along axis. Sagotiana and Coriacea diferenciate really well along first axis.

## Within *Symphonia* species

```{r PCAleafSymphoniaSpecies, fig.cap='Principal Component Analysis (PCA) of leaf trait wihtin Symphonia species.'}
pcas <- data$leaf %>% 
  filter(Genus == "Symphonia") %>% 
  filter(Species != "Indet.") %>% 
  mutate(Dawkins = gsub('([[:alpha:]])', "", Dawkins)) %>% 
  group_by(SpeciesLong) %>% 
  do(pcas = autoplot(princomp(~ SLA + LDMC + LT + LA + CC, data = ., cor = T), 
           data = .,
           colour = "Dawkins", alpha = 0.5, size = "DBH",
           loadings.label.size = 6,
           loadings.label.colour = 'black', loadings.label.vjust = 1.1,
           loadings = T, loadings.label = T, loadings.colour = 'black') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted"))
cowplot::plot_grid(plotlist = pcas$pcas, labels = pcas$SpeciesLong)
```

## Within *Eschweilera* species

```{r PCAleafEschweileraSpecies, fig.cap='Principal Component Analysis (PCA) of leaf trait wihtin Eschweilera species.'}
pcas <- data$leaf %>% 
  filter(Genus != "Symphonia") %>% 
  mutate(Dawkins = gsub('([[:alpha:]])', "", Dawkins)) %>% 
  group_by(SpeciesLong) %>% 
  filter(n() > 10) %>% 
  do(pcas = autoplot(princomp(~ SLA + LDMC + LT + LA + CC, data = ., cor = T), 
           data = .,
           colour = "Dawkins", alpha = 0.5, size = "DBH",
           loadings.label.size = 6,
           loadings.label.colour = 'black', loadings.label.vjust = 1.1,
           loadings = T, loadings.label = T, loadings.colour = 'black') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted"))
cowplot::plot_grid(plotlist = pcas$pcas, labels = pcas$SpeciesLong)
```

## Wood traits

### Within complexes

```{r PCAwoodSymphonia, fig.cap='Principal Component Analysis (PCA) of wood trait wihtin Symphonia complex.'}
data$wood %>% 
  filter(Genus == "Symphonia") %>% 
  autoplot(princomp(~ SLA + LDMC + LT + LA + CC + brBT + brBD + brWD, data = ., cor = T), 
           data = .,
           colour = "SpeciesLong", alpha = 0.5, size = "DBH",
           loadings.label.size = 6,
           loadings.label.colour = 'black', loadings.label.vjust = 1.1,
           loadings = T, loadings.label = T, loadings.colour = 'black') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted") +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1))
```

# Discussion

*Discuss and summarise all observed relations in regard of litterature...*

## Conclusion

*Take home message...*

# References
