---
title: "A02: Environmental descriptors selection"
date: '`r Sys.Date()`'
author: Sylvain Schmitt
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(parallel)
library(tidyverse)
library(ggfortify)
library(raster)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "../../data/Paracou/"
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
```

```{r data}
traitsSymphonia <- googlesheets::gs_title("Measures_Symphonia") %>% 
  googlesheets::gs_read("AllTraits") %>% 
  mutate(SLA = as.numeric(SLA), LDMC = as.numeric(LDMC)) %>% 
  mutate(Genus = "Symphonia") %>% 
  rename(Species = Morphotype)
traitsEschweilera <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("AllTraits") %>% 
  dplyr::select(-DBH)
traits <- bind_rows(traitsEschweilera, traitsSymphonia) %>% 
  mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) ; rm(traitsEschweilera, traitsSymphonia)
paracou <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>%
  tbl("Paracou") %>%
  filter(idTree %in% traits$idTree) %>%
  filter(Genus %in% c("Lecythis", "Eschweilera") & Plot %in% c(1, 6, 11, 13:15) & CensusYear == 2017 |
         Genus %in% c("Lecythis", "Eschweilera") & Plot == 16 & CensusYear == 2015 |
         Genus == "Symphonia" & CensusYear == 2015) %>%
  mutate(DBH = CircCorr*pi) %>% 
  collect()
trees <- left_join(traits, paracou) ; rm(traits, paracou)
```

```{r env}
# env <- trees %>%
#   dplyr::select(idTree, Xutm, Yutm) %>%
#   unique()
# coordinates(env) <- ~Xutm + Yutm
# proj4string(env) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
# env <- spTransform(env, CRSobj = crs)
# topo <- stack(
#   raster(file.path(path, "topography", "DEM_1m_2015.tif")),
#   raster(file.path(path, "topography", "RelativeElevation_1m.tif")),
#   raster(file.path(path, "topography", "slope_1m.tif")),
#   raster(file.path(path, "topography", "curvature_1m.tif")),
#   raster(file.path(path, "topography", "aspect_1m.tif")),
#   raster(file.path(path, "topography", "TRI_1m.tif")),
#   raster(file.path(path, "topography", "TWI_1m.tif"))
# )
# names(topo) <- c("DEM", "RelativeElevation", "Slope", "Curvature",
#                  "Aspect", "TRI", "TWI")
# topo <- projectRaster(topo, crs = crs)
# env <- data.frame(cbind(env@data, raster::extract(topo, env)))
# rm(topo)
# save(env, file = "./functional_save/env.Rdata")
load("./functional_save/env.Rdata")
```

# Introduction

This document test for environmental descriptors correlation in order to help select best biotic and abiotic descriptors.

# Material & Methods

**We need to add biotic environment with local basal area ($BA10$ and $BA20$), DAwkins for light, and DBH for ontogeny !**

# Results

```{r pairs, fig.cap="Pairs plot"}
env %>% 
  dplyr::select(DEM, RelativeElevation, Slope, Curvature, 
                Aspect, TRI, TWI) %>% 
  GGally::ggpairs()
```

```{r corrplot, fig.cap="Correlation plot"}
env %>% 
  dplyr::select(DEM, RelativeElevation, Slope, Curvature, 
                Aspect, TRI, TWI) %>% 
  na.omit() %>% 
  cor(.) %>% 
  corrplot::corrplot.mixed()
```

```{r PCA, fig.cap="Principal compoonent analysis plot"}
pca <- princomp(~ DEM + RelativeElevation + Slope + Curvature + 
                Aspect + TRI + TWI, env, cor = T)
autoplot(pca, 
         colour = "lightgrey",
         loadings.label.size = 6,
         loadings.label.colour = 'black', loadings.label.vjust = 1.1,
         loadings = T, loadings.label = T, loadings.colour = 'black') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted") +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1))
```

# Conclusion

**Following conclusion is wrong, to be changed !**

We consequently decided to use 5 weakly ($r < 0.75$) correlated descriptors to model functional traits variation :

* 2 abiotic variables
     * Elevation, representing ...
     * Topographic wetness index, representing ...
* 2 biotic variables
     * The local competition index
     * A disturbance index
* 1 indvidual variable
     * DBH as the ontogenetic descriptor
     
# References

