---
title: "A06 : Biotic Model (NCI)"
date: '`r Sys.Date()`'
author: Sylvain Schmitt & Anne Baranger
output:
  bookdown::html_document2:
    number_sections: no
    toc: true
    toc_float: yes
    theme: flatly
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
  bookdown::word_document2: default
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()[ls() != "Model"]) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = 2)
rstan_options(auto_write = T)
path <- "../../data/Paracou/"
```

```{r data, eval=T}
# traitsSymphonia <- googlesheets::gs_title("Measures_Symphonia") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   mutate(SLA = as.numeric(SLA), LDMC = as.numeric(LDMC)) %>%
#   mutate(Genus = "Symphonia") %>%
#   rename(Species = Morphotype) %>%
#   mutate(Species = ifelse(Species == "Indet.",
#                           c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G",
#                                      "sp.1" =  "S")], Species))
# traitsEschweilera <- googlesheets::gs_title("Measures_Eschweilera") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   filter(!(Plot == 14 & SubPlot == 1 & TreeFieldNum == 760)) %>%  # outlier
#   filter(!(Species %in% c("congestiflora","simiorum","persistens")))
# traits <- bind_rows(traitsEschweilera, traitsSymphonia) %>%
#   mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>% 
#   group_by(idTree, Plot, SubPlot, TreeFieldNum, Genus, Species,
#            SpeciesLong, Bark, Dawkins) %>%
#   summarise_at(vars("SLA", "LDMC", "LT", "LA", "CC",
#                     "brBT", "brBD", "brWD"), mean, na.rm = T)  %>%
#   ungroup() %>%
#   mutate(invSLA=1/SLA) %>%
#   mutate(invLA=1/LA)
# load("./functional_save/env.Rdata")
# Individuals <-  left_join(traits, env, by = "idTree", suffix = c("", ".y"))
# rm(traitsEschweilera, traitsSymphonia, traits, paracou, env)
# save(Individuals, file = "./functional_save/Individuals.Rdata")
load("./functional_save/Individuals.Rdata")
load("./functional_save/CompetitionMatrix.Rdata")
```

# Introduction {-}

The aim of this document is to fit the biotic model (equation \@ref(eq:Competition)) with observed values of functional traits for our individuals. We prepared data excluding oultiers and scaling all continuous variable to ease model fit. We tested the abiotic model with mixed effects. We tested first 5 ways to deal with edge effects with a reduce dataset of 200 observations for $LA$ with neighbors until 20m (table \@ref(tab:EdgeEffecttable)). Finally, we built the full models for all traits and all observations for all neighbors in a radius of 20m with the method to deal with edge effect with the highest likelihood.

We explained leaf trait $Trait$ according to competition following a normal law with a linear model form around a Neighborhood Crowding Index (including both neighbor diameter and distance) :

\begin{equation} 
Trait_i \sim \mathcal{N}(\alpha + \beta_{comp}. \sum_{j~|~\delta_{i,j}<\delta_{max}}^{J_i} DBH_j^{\beta_{NCI}}.e^{-\alpha_{NCI}.\delta_{i,j}}, \sigma) \\
(\alpha, \beta_{comp}, \beta_{NCI}) \sim \mathcal{N}^{3}(0, + \infty) \\
(\alpha_{NCI}, \sigma) \sim \Gamma(- \infty, - \infty) 
(\#eq:Competition)
\end{equation}

where:

* $Trait_i$ is a functional trait of an individual i associated to a $DBH_i$ value ;
* $i \in [1,I]$ where I is the number of individuals, i.e. 838 in our dataset;
* $j \in [1,J_i]$ where $J_i$ is the number of neighbors within the radius $\delta_{max}$ surrounding the individual i, and DBH_j the value of DBH for neighbor j;
* $\delta_{i,j}$ is the distance between the individual i and one of its neighbors j;
* $\alpha_{NCI}$ represents the effect of neighbors distance to the individual, lower $\alpha_{NCI}$ enable to strengthen competition effects from trees farther to the individual $i$. $\alpha_{NCI}$ will be computed with positive values;
* $\beta_{NCI}$ represents the idea that bigger neighbor will increase competition on the focal individual, thus $\beta_{NCI}$ was chosen equal to 2 to represent neighbor surface.
* $\alpha$ and $\beta_{comp}$ are linear model parameters and could range in positive or negative value. Normal flat laws were chosen as non informative flat priors on on $\mathbb{R}$;
* $\sigma$ is normal law standard error, and will, therefore, be chosen with positive values;
* Gamma law was chosen as non informative flat prior on $\mathbb{R}^+$ for neighbors distance effect and Normal law standard error.


```{r EdgeEffecttable}
data.frame(Fit = 0:4,
           Include = c("No", rep("Yes", 4)),
           Correction = c("", rep("No", 2), rep("Yes", 2)),
           Weighing = c("", "No", "Yes", "No", "Yes")) %>% 
  kable(caption = "Edge effects test", 
        col.names = c("Test", "Include edge effect", "Correct edge effect", 
                      "Weight corresponding observations"))
```


# Edge effect test

## Test 0

> Edge effect excluded from data

```{stan, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  int<lower=1> S ; // # of species
  int<lower=1> C ; // # of complexes
  vector[N] Trait ; // Trait
  vector[J] DBHj ; // DBH neighbor
  vector[J] Deltaj ; // Distance neighbor
  int<lower=1, upper=N>  individual[J] ; // individuals index in neighbor matrix
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
}
parameters {
  vector[S] alpha_s ; // Intercept for species
  vector[C] alpha_c ; // Intercept for complexes
  vector[S] betaComp_s ; // Biotic slope for species
  vector[C] betaComp_c ; // Biotic slope for complexes
  real<lower=0> alphaNCI ;
  real<lower=0, upper=10> sigmaIntercept ; // Intercept variance for species
  real<lower=0, upper=10> sigmaComp ; // Biotic slope variance for species
  real<lower=0> sigma ;
}
transformed parameters {
  vector[J] NCIj ;
  vector[N] NCI ;
  NCIj = (DBHj .* DBHj + exp(-alphaNCI * Deltaj)) ;
  NCI = rep_vector(0.0, N) ;
  for(j in 1:J)
   NCI[individual[j]] += NCIj[j] ;
}
model {
  alphaNCI ~ lognormal(0, 1) ;
  alpha_s ~ normal(alpha_c[speciesincomplex], sigmaIntercept) ;
  betaComp_s ~ normal(betaComp_c[speciesincomplex], sigmaComp) ;
  Trait ~ normal(alpha_s[species] +  betaComp_s[species] .* NCI, sigma) ; // Likelihood
}
generated quantities {
  matrix[N,C] Trait_pred ;
  for(c in 1:C)
    Trait_pred[,c] = alpha_c[c] + betaComp_c[c] * NCI ; // Predictions
}
```

```{r }
data_trait <- Individuals %>% 
  filter(!is.na(invLA)) %>% 
  left_join(select(Competition, idTree, EdgeEffect20)) %>% 
  filter(EdgeEffect20 == 0) %>% 
  sample_n(200)
Competition_trait <- Competition
Competition_trait$idTree <- match(Competition_trait$idTree, data_trait$idTree)
Competition_trait <- filter(Competition_trait, !is.na(idTree))
mdata <- list(N = nrow(data_trait),
       J = nrow(Competition_trait),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = as.numeric(scale(data_trait$invLA, center = F)),
       individual = Competition_trait$idTree,
       DBHj = as.numeric(scale(Competition_trait$DBHj, center = F)),
       Deltaj = as.numeric(scale(Competition_trait$dij, center = F)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2],
       weights = weights,
       data = data_trait,
       competition = Competition_trait)
# fit <- sampling(Model, chains = 2, data = mdata)
# save(fit, file = "./functional_save/NCItest0.Rdata")
```

## Test 1

> Edge effect without weight nor correction

```{stan, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  int<lower=1> S ; // # of species
  int<lower=1> C ; // # of complexes
  vector[N] Trait ; // Trait
  vector[J] DBHj ; // DBH neighbor
  vector[J] Deltaj ; // Distance neighbor
  int<lower=1, upper=N>  individual[J] ; // individuals index in neighbor matrix
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
}
parameters {
  vector[S] alpha_s ; // Intercept for species
  vector[C] alpha_c ; // Intercept for complexes
  vector[S] betaComp_s ; // Biotic slope for species
  vector[C] betaComp_c ; // Biotic slope for complexes
  real<lower=0> alphaNCI ;
  real<lower=0, upper=10> sigmaIntercept ; // Intercept variance for species
  real<lower=0, upper=10> sigmaComp ; // Biotic slope variance for species
  real<lower=0> sigma ;
}
transformed parameters {
  vector[J] NCIj ;
  vector[N] NCI ;
  NCIj = (DBHj .* DBHj + exp(-alphaNCI * Deltaj)) ;
  NCI = rep_vector(0.0, N) ;
  for(j in 1:J)
   NCI[individual[j]] += NCIj[j] ;
}
model {
  alphaNCI ~ lognormal(0, 1) ;
  alpha_s ~ normal(alpha_c[speciesincomplex], sigmaIntercept) ;
  betaComp_s ~ normal(betaComp_c[speciesincomplex], sigmaComp) ;
  Trait ~ normal(alpha_s[species] +  betaComp_s[species] .* NCI, sigma) ; // Likelihood
}
generated quantities {
  matrix[N,C] Trait_pred ;
  for(c in 1:C)
    Trait_pred[,c] = alpha_c[c] + betaComp_c[c] * NCI ; // Predictions
}
```

```{r }
data_trait <- Individuals %>% 
  filter(!is.na(invLA)) %>% 
  sample_n(200)
Competition_trait <- Competition
Competition_trait$idTree <- match(Competition_trait$idTree, data_trait$idTree)
Competition_trait <- filter(Competition_trait, !is.na(idTree))
mdata <- list(N = nrow(data_trait),
       J = nrow(Competition_trait),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = as.numeric(scale(data_trait$invLA, center = F)),
       individual = Competition_trait$idTree,
       DBHj = as.numeric(scale(Competition_trait$DBHj, center = F)),
       Deltaj = as.numeric(scale(Competition_trait$dij, center = F)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2],
       weights = weights,
       data = data_trait,
       competition = Competition_trait)
# fit <- sampling(Model, chains = 2, data = mdata)
# save(fit, file = "./functional_save/NCItest1.Rdata")
```

## Test 2

> Edge effect with weighted observations

```{stan, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  int<lower=1> S ; // # of species
  int<lower=1> C ; // # of complexes
  vector[N] Trait ; // Trait
  vector[J] DBHj ; // DBH neighbor
  vector[J] Deltaj ; // Distance neighbor
  int<lower=1, upper=N>  individual[J] ; // individuals index in neighbor matrix
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
  vector[N] weights ;
}
parameters {
  vector[S] alpha_s ; // Intercept for species
  vector[C] alpha_c ; // Intercept for complexes
  vector[S] betaComp_s ; // Biotic slope for species
  vector[C] betaComp_c ; // Biotic slope for complexes
  real<lower=0> alphaNCI ;
  real<lower=0, upper=10> sigmaIntercept ; // Intercept variance for species
  real<lower=0, upper=10> sigmaComp ; // Biotic slope variance for species
  real<lower=0> sigma ;
}
transformed parameters {
  vector[J] NCIj ;
  vector[N] NCI ;
  NCIj = (DBHj .* DBHj + exp(-alphaNCI * Deltaj)) ;
  NCI = rep_vector(0.0, N) ;
  for(j in 1:J)
   NCI[individual[j]] += NCIj[j] ;
}
model {
  alphaNCI ~ lognormal(0, 1) ;
  alpha_s ~ normal(alpha_c[speciesincomplex], sigmaIntercept) ;
  betaComp_s ~ normal(betaComp_c[speciesincomplex], sigmaComp) ;
  for(n in 1:N)
    target += weights[n] * normal_lpdf(Trait[n] | alpha_s[species] +  betaComp_s[species]*NCI[n], sigma) ; // Likelihood
}
generated quantities {
  matrix[N,C] Trait_pred ;
  for(c in 1:C)
    Trait_pred[,c] = alpha_c[c] + betaComp_c[c] * NCI ; // Predictions
}
```

```{r }
data_trait <- Individuals %>% 
  filter(!is.na(invLA)) %>% 
  sample_n(200)
Competition_trait <- Competition
Competition_trait$idTree <- match(Competition_trait$idTree, data_trait$idTree)
Competition_trait <- filter(Competition_trait, !is.na(idTree))
data_trait <- data_trait %>% 
  left_join(select(Competition, idTree, AreaOutside20) 
            %>% unique())
mdata <- list(N = nrow(data_trait),
       J = nrow(Competition_trait),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = as.numeric(scale(data_trait$invLA, center = F)),
       individual = Competition_trait$idTree,
       DBHj = as.numeric(scale(Competition_trait$DBHj, center = F)),
       Deltaj = as.numeric(scale(Competition_trait$dij, center = F)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2],
       weights = 1-data_trait$AreaOutside20,
       data = data_trait,
       competition = Competition_trait)
# fit <- sampling(Model, chains = 2, data = mdata)
# save(fit, file = "./functional_save/NCItest2.Rdata")
```

## Test 3

> Edge effect with corrected observations

```{stan, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  int<lower=1> S ; // # of species
  int<lower=1> C ; // # of complexes
  vector[N] Trait ; // Trait
  vector[J] DBHj ; // DBH neighbor
  vector[J] Deltaj ; // Distance neighbor
  int<lower=1, upper=N>  individual[J] ; // individuals index in neighbor matrix
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
  vector[N] weights ;
}
parameters {
  vector[S] alpha_s ; // Intercept for species
  vector[C] alpha_c ; // Intercept for complexes
  vector[S] betaComp_s ; // Biotic slope for species
  vector[C] betaComp_c ; // Biotic slope for complexes
  real<lower=0> alphaNCI ;
  real<lower=0, upper=10> sigmaIntercept ; // Intercept variance for species
  real<lower=0, upper=10> sigmaComp ; // Biotic slope variance for species
  real<lower=0> sigma ;
}
transformed parameters {
  vector[J] NCIj ;
  vector[N] NCI ;
  NCIj = (DBHj .* DBHj + exp(-alphaNCI * Deltaj)) ;
  NCI = rep_vector(0.0, N) ;
  for(j in 1:J)
   NCI[individual[j]] += NCIj[j] ;
}
model {
  alphaNCI ~ lognormal(0, 1) ;
  alpha_s ~ normal(alpha_c[speciesincomplex], sigmaIntercept) ;
  betaComp_s ~ normal(betaComp_c[speciesincomplex], sigmaComp) ;
  Trait ~ normal(alpha_s[species] +  betaComp_s[species] .* (1 ./ weights) .* NCI, sigma) ; // Likelihood
}
generated quantities {
  matrix[N,C] Trait_pred ;
  for(c in 1:C)
    Trait_pred[,c] = alpha_c[c] + betaComp_c[c] * (1 ./ weights) .* NCI ; // Predictions
}
```


```{r }
data_trait <- Individuals %>% 
  filter(!is.na(invLA)) %>% 
  sample_n(200)
Competition_trait <- Competition
Competition_trait$idTree <- match(Competition_trait$idTree, data_trait$idTree)
Competition_trait <- filter(Competition_trait, !is.na(idTree))
data_trait <- data_trait %>% 
  left_join(select(Competition, idTree, AreaOutside20) 
            %>% unique())
mdata <- list(N = nrow(data_trait),
       J = nrow(Competition_trait),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = as.numeric(scale(data_trait$invLA, center = F)),
       individual = Competition_trait$idTree,
       DBHj = as.numeric(scale(Competition_trait$DBHj, center = F)),
       Deltaj = as.numeric(scale(Competition_trait$dij, center = F)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2],
       weights = 1-data_trait$AreaOutside20,
       data = data_trait,
       competition = Competition_trait)
# fit <- sampling(Model, chains = 2, data = mdata)
# save(fit, file = "./functional_save/NCItest3.Rdata")
```

## Test 4

> Edge effect with weighted and corrected observations


```{stan, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  int<lower=1> S ; // # of species
  int<lower=1> C ; // # of complexes
  vector[N] Trait ; // Trait
  vector[J] DBHj ; // DBH neighbor
  vector[J] Deltaj ; // Distance neighbor
  int<lower=1, upper=N>  individual[J] ; // individuals index in neighbor matrix
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
  vector[N] weights ;
}
parameters {
  vector[S] alpha_s ; // Intercept for species
  vector[C] alpha_c ; // Intercept for complexes
  vector[S] betaComp_s ; // Biotic slope for species
  vector[C] betaComp_c ; // Biotic slope for complexes
  real<lower=0> alphaNCI ;
  real<lower=0, upper=10> sigmaIntercept ; // Intercept variance for species
  real<lower=0, upper=10> sigmaComp ; // Biotic slope variance for species
  real<lower=0> sigma ;
}
transformed parameters {
  vector[J] NCIj ;
  vector[N] NCI ;
  NCIj = (DBHj .* DBHj + exp(-alphaNCI * Deltaj)) ;
  NCI = rep_vector(0.0, N) ;
  for(j in 1:J)
   NCI[individual[j]] += NCIj[j] ;
}
model {
  alphaNCI ~ lognormal(0, 1) ;
  alpha_s ~ normal(alpha_c[speciesincomplex], sigmaIntercept) ;
  betaComp_s ~ normal(betaComp_c[speciesincomplex], sigmaComp) ;
  for(n in 1:N)
    target += weights[n] * normal_lpdf(Trait[n] | alpha_s[species] +  betaComp_s[species]*NCI[n], sigma) ; // Likelihood
}
generated quantities {
  matrix[N,C] Trait_pred ;
  for(c in 1:C)
    Trait_pred[,c] = alpha_c[c] + betaComp_c[c] * (1 ./ weights) .* NCI ; // Predictions
}
```

```{r }
data_trait <- Individuals %>% 
  filter(!is.na(invLA)) %>% 
  sample_n(200)
Competition_trait <- Competition
Competition_trait$idTree <- match(Competition_trait$idTree, data_trait$idTree)
Competition_trait <- filter(Competition_trait, !is.na(idTree))
data_trait <- data_trait %>% 
  left_join(select(Competition, idTree, AreaOutside20) 
            %>% unique())
mdata <- list(N = nrow(data_trait),
       J = nrow(Competition_trait),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = as.numeric(scale(data_trait$invLA, center = F)),
       individual = Competition_trait$idTree,
       DBHj = as.numeric(scale(Competition_trait$DBHj, center = F)),
       Deltaj = as.numeric(scale(Competition_trait$dij, center = F)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2],
       weights = 1-data_trait$AreaOutside20,
       data = data_trait,
       competition = Competition_trait)
# fit <- sampling(Model, chains = 2, data = mdata)
# save(fit, file = "./functional_save/NCItest4.Rdata")
```

## Summary

```{r }
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
tests <- paste0("test", 0:4)
fits <- lapply(tests, function(test){
  # cat(test, "\n")
  load(paste0("./functional_save/NCI", test, ".Rdata"))
  return(fit)
})
names(fits) <- tests
pars <- c("alpha_c", "betaComp_c", "alphaNCI", "sigma")
table <- lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = c(pars, "lp__"), droppars = NULL, rhat = T))
table %>% 
  bind_rows(.id = "Test") %>% 
  reshape2::melt(id.vars = c("Test", "term")) %>% 
  reshape2::dcast(term + variable ~ Test, value.var = "value") %>% 
  kable() %>%
  kable_styling() %>% 
  collapse_rows(1)
```

```{r EdgeEffecttable2}
data.frame(Fit = 0:4,
           Include = c("No", rep("Yes", 4)),
           Correction = c("", rep("No", 2), rep("Yes", 2)),
           Weighing = c("", "No", "Yes", "No", "Yes")) %>% 
  kable(caption = "Edge effects test", 
        col.names = c("Test", "Include edge effect", "Correct edge effect", 
                      "Weight corresponding observations"))
```

## Posteriors

```{r , fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(tests, function(test)
  mcmc_intervals(as.array(fits[[test]]), regex_pars = pars)),
  labels = tests, nrow = 2)
```

```{r , fig.cap="Model parameters posterior.", fig.width=12, fig.height=12}
cowplot::plot_grid(plotlist = lapply(tests, function(test)
  mcmc_intervals(as.array(fits[[test]]), regex_pars = c("Comp"))),
  labels = tests, nrow = 2)
```

```{r , fig.cap="Model parameters posterior."}
gdata <- lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit), regex_pars = c("lp__"))) %>% 
  bind_rows(.id = "test") 
ggplot(gdata, aes(x = test, xend = test, col = test, fill = test)) +
  geom_point(aes(y = m), shape = 21, size = 10) +
  geom_segment(aes_(y = ~ ll, yend = ~ hh), size = 1, show.legend = F) +
  geom_segment(aes_(y = ~ l, yend = ~ h), size = 2) +
  coord_flip() +
  scale_y_sqrt() +
  ggtitle("Log likelihood")
```

## Chains

```{r , fig.cap="Model parameters posterior.", fig.width=12, fig.height=12}
cowplot::plot_grid(plotlist = lapply(tests, function(test)
  mcmc_trace(as.array(fits[[test]]), pars = c("alphaNCI"))),
  labels = tests, nrow = 2)
```

## Pairs

```{r , fig.cap="Markov chains pairs plot after warmup  for intermediate data.", fig.height=12, fig.width=12}
mcmc_pairs(as.array(fits$test2), regex_pars = pars)
```

## Conclusion

We will use test 3, i.e. we will include individuals with edge effect with edge effet correction without weighing likelihood in the bayesian model.

# Full model

```{r mdata, fig.cap="prepared data"}
traits <- c("invSLA", "LDMC", "LT", "invLA", "CC")
mdata <- lapply(traits, function(trait){
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    left_join(select(Competition, idTree, AreaOutside20) 
              %>% unique())
  Competition_trait <- Competition
  Competition_trait$idTree <- match(Competition_trait$idTree, data_trait$idTree)
  Competition_trait <- filter(Competition_trait, !is.na(idTree))
  list(N = nrow(data_trait),
       J = nrow(Competition_trait),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = as.numeric(scale(data_trait[trait], center = F)),
       individual = Competition_trait$idTree,
       DBHj = as.numeric(scale(Competition_trait$DBHj, center = F)),
       Deltaj = as.numeric(scale(Competition_trait$dij, center = F)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2],
       weights = 1-data_trait$AreaOutside20,
       data = data_trait,
       competition = Competition_trait)})
names(mdata) <- traits
```

## Code

```{stan, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  int<lower=1> S ; // # of species
  int<lower=1> C ; // # of complexes
  vector[N] Trait ; // Trait
  vector[J] DBHj ; // DBH neighbor
  vector[J] Deltaj ; // Distance neighbor
  int<lower=1, upper=N>  individual[J] ; // individuals index in neighbor matrix
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
  vector[N] weights ;
}
parameters {
  vector[S] alpha_s ; // Intercept for species
  vector[C] alpha_c ; // Intercept for complexes
  vector[S] betaComp_s ; // Biotic slope for species
  vector[C] betaComp_c ; // Biotic slope for complexes
  real<lower=0> alphaNCI ;
  real<lower=0, upper=10> sigmaIntercept ; // Intercept variance for species
  real<lower=0, upper=10> sigmaComp ; // Biotic slope variance for species
  real<lower=0> sigma ;
}
transformed parameters {
  vector[J] NCIj ;
  vector[N] NCI ;
  NCIj = (DBHj .* DBHj + exp(-alphaNCI * Deltaj)) ;
  NCI = rep_vector(0.0, N) ;
  for(j in 1:J)
   NCI[individual[j]] += NCIj[j] ;
}
model {
  alphaNCI ~ lognormal(0, 1) ;
  alpha_s ~ normal(alpha_c[speciesincomplex], sigmaIntercept) ;
  betaComp_s ~ normal(betaComp_c[speciesincomplex], sigmaComp) ;
  Trait ~ normal(alpha_s[species] +  betaComp_s[species] .* (1 ./ weights) .* NCI, sigma) ; // Likelihood
}
generated quantities {
  matrix[N,C] Trait_pred ;
  for(c in 1:C)
    Trait_pred[,c] = alpha_c[c] + betaComp_c[c] * (1 ./ weights) .* NCI ; // Predictions
}
```

## Summary

```{r }
for(trait in traits){
  cat("### ", trait, " ### \n\n")
  fit <- sampling(Model, chains = 2, data = mdata[[trait]])
  save(fits, file = paste0("./functional_save/NCI_", trait, ".Rdata"))
  rm(fit) ; gc()
}
```

```{r }
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- traits
# save(fits, file = "./functional_save/NCI.Rdata")
# load("./functional_save/NCI.Rdata")
# pars <- c("alpha_c", "betaComp_c", "alphaNCI", "sigma")
# lapply(as.list(traits), function(trait)
#   broom::tidyMCMC(fits[[trait]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
#    mutate(Trait = trait)) %>%
#   bind_rows() %>%
#   kable() %>%
#   kable_styling()
```

## Posteriors

```{r , fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(traits, function(trait)
  mcmc_intervals(as.array(fits[[trait]]), regex_pars = pars)),
  labels = names(data), nrow = 2)
```

```{r , fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(traits, function(trait)
  mcmc_intervals(as.array(fits[[trait]]), regex_pars = "Comp")),
  labels = names(data), nrow = 2)
```

## Chains

```{r , fig.cap="Markov chains trace plot after warmup for intermediate data."}
mcmc_trace(as.array(fits$LDMC), regex_pars = c(pars, "lp__"),
             facet_args = list(labeller = label_parsed))
```

## Pairs

```{r , fig.cap="Markov chains pairs plot after warmup  for intermediate data.", fig.height=12, fig.width=12}
mcmc_pairs(as.array(fits$invSLA), regex_pars = pars)
```

## Predictions

```{r , ev, fig.cap="Predictions.", fig.width=12, fig.height=12}
lapply(as.list(names(mdata)), function(type)
  cbind(type = type, 
        Y = mdata[[type]]$Trait,
        NCI = apply(as.matrix(fits[[type]], pars = "NCI"), 2, mean),
        Individuals[!is.na(unlist(Individuals[,type])),],
        muEschweilera = apply(as.matrix(fits[[type]], 
                                        pars = paste0("Trait_pred[",
                                                      1:mdata[[type]]$N,",1]")), 
                              2, mean),
        muSymphonia = apply(as.matrix(fits[[type]], 
                                      pars = paste0("Trait_pred[",
                                                    1:mdata[[type]]$N,",2]")), 
                            2, mean),
        muEschweilera5 = apply(as.matrix(fits[[type]], 
                                         pars = paste0("Trait_pred[",
                                                       1:mdata[[type]]$N,",1]")), 
                               2, quantile, probs = 0.05),
        muSymphonia5 = apply(as.matrix(fits[[type]], 
                                       pars = paste0("Trait_pred[",
                                                      1:mdata[[type]]$N,",2]")), 
                             2, quantile, probs = 0.05),
        muEschweilera95 = apply(as.matrix(fits[[type]], 
                                          pars = paste0("Trait_pred[",
                                                      1:mdata[[type]]$N,",1]")), 
                                2, quantile, probs = 0.95),
        muSymphonia95 = apply(as.matrix(fits[[type]], 
                                        pars = paste0("Trait_pred[",
                                                      1:mdata[[type]]$N,",2]")), 
                              2, quantile, probs = 0.95)
  )) %>% 
  bind_rows() %>% 
  ggplot(aes(x = NCI)) +
  geom_point(aes(y = Y, col = as.factor(Genus))) +
  geom_ribbon(aes(ymin = muEschweilera5, ymax = muEschweilera95,
                  col = "Eschweilera"), alpha = 0.2) +
  geom_ribbon(aes(ymin = muSymphonia5, ymax = muSymphonia95,
                  col = "Symphonia"), alpha = 0.2) +
  geom_line(aes(y = muEschweilera, col = "Eschweilera")) +
  geom_line(aes(y = muSymphonia, col = "Symphonia")) +
  facet_wrap(~ type, scales = "free", nrow = 3)
```

# Conclusion

# References
