---
title: "A07 : Full Model"
date: '`r Sys.Date()`'
author: Sylvain Schmitt & Anne Baranger
output:
  bookdown::html_document2:
    number_sections: no
    toc: true
    toc_float: yes
    theme: flatly
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
  bookdown::word_document2: default
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
#rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "../../data/Paracou/"
```

```{r data, eval=T}
traitsSymphonia <- googlesheets::gs_title("Measures_Symphonia") %>% 
  googlesheets::gs_read("AllTraits") %>% 
  mutate(SLA = as.numeric(SLA), LDMC = as.numeric(LDMC)) %>% 
  mutate(Genus = "Symphonia") %>% 
  rename(Species = Morphotype) %>% 
  mutate(Species = ifelse(Species == "Indet.", 
                          c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G", 
                                     "sp.1" =  "S")], Species))
traitsEschweilera <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("AllTraits") %>% 
  filter(!(Plot == 14 & SubPlot == 1 & TreeFieldNum == 760)) %>%  # outlier
  filter(!(Species %in% c("congestiflora","simiorum","persistens")))
traits <- bind_rows(traitsEschweilera, traitsSymphonia) %>% 
  mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) 
paracou <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>%
  tbl("Paracou") %>%
  filter(idTree %in% traits$idTree) %>%
  filter(Genus %in% c("Lecythis", "Eschweilera") & Plot %in% c(1, 6, 11, 13:15) & CensusYear == 2017 |
         Genus %in% c("Lecythis", "Eschweilera") & Plot == 16 & CensusYear == 2015 |
         Genus == "Symphonia" & CensusYear == 2015) %>%
  mutate(DBH = CircCorr/pi) %>% 
  collect()
load("./functional_save/env.Rdata")
Individuals <- left_join(traits, paracou) %>% 
  left_join(env, by = "idTree", suffix = c("", ".y")) %>% 
  group_by(Plot, SubPlot, TreeFieldNum, Genus, Species,
           SpeciesLong, Bark, Dawkins, DBH, TWI, NCI) %>%
  summarise_at(vars("SLA", "LDMC", "LT", "LA", "CC",
                    "brBT", "brBD", "brWD"), mean, na.rm = T)  %>%
  ungroup() %>% 
  mutate(invSLA=1/SLA) %>% 
  mutate(invLA=1/LA)
load("./functional_save/CompetitionMatrix.Rdata")
rm(traitsEschweilera, traitsSymphonia, traits, paracou, env)
```

# Introduction {-}

The aim of this document is to fit the full model with observed values of functional traits for our individuals. We prepared data excluding oultiers and reducing variance of all continuous variable to ease model fit (see figure \@ref(fig:mdata)). We tested the full model with additive (1) and hierarchical (2) forms.

```{r mdata, fig.cap="prepared data"}
traits <- c("invSLA", "LDMC", "LT", "invLA", "CC")
mdata <- lapply(traits, function(trait){
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),]
  list(N = nrow(data_trait),
       J = nrow(Competition),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = as.numeric(scale(data_trait[trait], center = F)),
       DBH = as.numeric(scale(data_trait$DBH, center = F)),
       TWI = as.numeric(scale(data_trait$TWI, center = F)),
       individual = as.numeric(as.factor(Competition$idTree)),
       DBHj = as.numeric(scale(Competition$DBHj, center = F)),
       Deltaj = as.numeric(scale(Competition$dij, center = F)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2])})
names(mdata) <- traits
lapply(as.list(traits), function(trait)
  cbind(Model = trait, 
        Trait = mdata[[trait]]$Trait, 
        Individuals[unlist(!is.na(Individuals[trait])),])) %>% 
  bind_rows() %>% 
  select(Model, Trait, Genus, DBH, TWI, NCI) %>% 
  reshape2::melt(id.vars = c("Model", "Trait", "Genus")) %>% 
  ggplot(aes(x = value, y = Trait, col = Genus)) +
  geom_point() +
  geom_smooth() +
  facet_grid(Model ~ variable, scales = "free")
```

# Additive model

We explained leaf trait $Trait$ variation with ontogeny ($DBH$), abiotic environment ($TWI$), biotic interactions ($NCI$) and taxonomic levels ($C, S$ mixed effects, respectivelly fixed and random):

\begin{equation} 
  T_{c,s,i} \sim \mathcal{N}(\alpha_{c,s}.\frac{DBH_i}{{\beta_{DBH}}_{c,s} + DBH_i} + {\beta_{TWI}}_{c,s}.TWI_i + {\beta_{Comp}}_{c,s}.\sum_{j=1}^{J} DBH_j^{2} . e^{-{\alpha_{NCI}}_{c,s}.\delta_{i,j}}, \sigma)    \\
  \Theta_{c,s} \sim \mathcal{N}^5(\Theta_c,\sigma_{species}) ~|~ \Theta =
   \begin{bmatrix}
           \alpha \\
           \beta_{DBH} \\
           \beta_{TWI} \\
           \beta_{Comp} \\
           \alpha_{NCI}
  \end{bmatrix},~\sigma_{species}
  \begin{bmatrix}
           \sigma_{Intercept} \\
           \sigma_{DBH} \\
           \sigma_{TWI} \\
           \sigma_{Comp} \\
           \sigma_{NCI}
  \end{bmatrix}  \\
  \Theta_c \sim \mathcal{N}^{5}(0, + \infty) \\
  (\sigma, \sigma_{species}) \sim \Gamma^{6}(- \infty, + \infty)
(\#eq:final)
\end{equation}

* $Trait_{c,s,i}$ is a trait of an individual i, from the species s, and the complex c;
* $i \in [1,I_s]$ where $I_s$ is the number of individuals in species s;
* $s \in [1,S_c]$ where $S_c$ is the number of species in complex c;
* $c \in [1,C]$ where $C$ is the number of complexes.
* $i \in [1,I]$ where I is the number of individuals;
* $j \in [1,J_i]$ where $J_i$ is the number of neighbors within the area of diameter $\delta_{max}$ surrounding the individual i, and $DBH_j$ the value of $DBH$ for neighbor $j$;
* $\alpha$ represents both the plateau of the Michaelis Menten and the intercept of each submodel (ontogeny, biotic & abiotic);
* $\beta_{DBH}$  is the value of DBH for which the trait account for half of its saturation value, and thus, ranges in positive values;
* $\beta_{TWI}$ is linear abiotic model parameters. $\beta_{TWI}$ can range in positive or negative value;
* $\beta_{Comp}$ is linear biotic model parameters. $\beta_{Comp}$ can range in positive or negative value;
* $\alpha_{NCI}$ represents the effect of neighbors distance to the individual and will be chosen positive;
* all $\sigma_{species}$ represents parameters variance for species;
* $\sigma$ and $\sigma_{species}$ are standard errors, and thus will have positive values;
* therefore, gamma laws were chosen as non informative flat priors on $\mathbb{R}^+$ for $\sigma$  and $\sigma_{species}$. Normal flat law were chosen as non informative flat prior on $\mathbb{R}$ for all $\Theta_{c,s}$ and $\Theta_c$.

## Code

```{stan, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  vector[N] Trait ; // Trait
  vector[N] DBH ; // Diameter at Breast Height
  vector[N] TWI ; // Topographic Wetness Index
  vector[N] DBHj ; // DBH neighbor
  vector[N] Deltaj ; // Distance neighbor
  int<lower=1, upper=N>  individual ; // individuals index in neighbor matrix
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
}
parameters {
  vector[S] alpha_s ; // Intercept for species
  vector[C] alpha_c ; // Intercept for complexes
  vector[S] betaDBH_s ; // DBH half-load for species
  vector[C] betaDBH_c ; // DBH half-load for complexes
  vector[S] betaTWI_s ; // Abiotic slope for species
  vector[C] betaTWI_c ; // Abiotic slope for complexes
  real<lower=0, upper=10> sigmaIntercept ; // Intercept variance for species
  real<lower=0, upper=10> sigmaDBH ; // DBH half-load variance for species
  real<lower=0, upper=10> sigmaTWI ; // Abiotic slope variance for species
  real<lower=0, upper=10> sigma ; // Variance
}
model {
  alpha_s ~ normal(alpha_c[speciesincomplex], sigmaIntercept) ;
  betaDBH_s ~ normal(betaDBH_c[speciesincomplex], sigmaDBH) ;
  betaTWI_s ~ normal(betaTWI_c[speciesincomplex], sigmaTWI) ;
  Trait ~ normal(alpha_s[species] .* (DBH ./ (betaDBH_s[species] + DBH)) + betaTWI_s[species] .* TWI, sigma) ; // Likelihood
} 
generated quantities {
  matrix[N,C] Trait_pred ;
  for(c in 1:C)
    Trait_pred[,c] = alpha_c[c] + betaTWI_c[c] *TWI ; // Predictions
}
```


## Summary

```{r , eval=F}
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- traits
# save(fits, file = "./functional_save/NCI.Rdata")
load("./functional_save/NCI.Rdata")
pars <- c()
lapply(as.list(traits), function(trait)
  broom::tidyMCMC(fits[[trait]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
   mutate(Trait = trait)) %>%
  bind_rows() %>%
  kable() %>%
  kable_styling()
```

## Posteriors

```{r , eval=F, fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(traits, function(trait)
  mcmc_intervals(as.array(fits[[trait]]), pars = pars)),
  labels = names(data), nrow = 2)
```

## Chains

```{r , eval=F, fig.cap="Markov chains trace plot after warmup for intermediate data."}
mcmc_trace(as.array(fits$LA), pars = c(pars, "lp__"),
             facet_args = list(labeller = label_parsed))
```

## Pairs

```{r , eval=F, fig.cap="Markov chains pairs plot after warmup  for intermediate data."}
mcmc_pairs(as.array(fits$CC), pars = pars)
```

## Predictions

```{r , eval=F, fig.cap="Predictions."}
lapply(as.list(traits), function(trait)
  cbind(fit = trait, 
        Y = mdata[[trait]]$Trait,
        Individuals[unlist(!is.na(Individuals[trait])),],
        Traitpred = apply(as.matrix(fits[[trait]], pars = "Traitpred"), 2, mean),
        sigmaTWI = apply(as.matrix(fits[[trait]], pars = "sigma"), 2, mean),
        mu5 = unlist(apply(as.matrix(fits[[trait]], pars = "Traitpred"), 2, quantile, probs = 0.05)),
        mu95 = unlist(apply(as.matrix(fits[[trait]], pars = "Traitpred"), 2, quantile, probs = 0.95)),
        sigmaTWI95 = unlist(apply(as.matrix(fits[[trait]], pars = "sigma"), 2, quantile, probs = 0.95))
        )) %>% 
  bind_rows() %>% 
  mutate(Y5sigmafix = qnorm(0.05, mean = mu5, sd = sigmaTWI)) %>% 
  mutate(Y95sigmafix = qnorm(0.95, mean = mu95, sd = sigmaTWI)) %>% 
  mutate(Y5 = qnorm(0.05, mean = mu5, sd = sigmaTWI95)) %>% 
  mutate(Y95 = qnorm(0.95, mean = mu95, sd = sigmaTWI95)) %>% 
  ggplot(aes(x = TWI)) +
  geom_point(aes(y = Y)) +
  geom_line(aes(y = Traitpred)) +
  geom_ribbon(aes(ymin = Y5, ymax = Y95, color = 'Ypred'), alpha = 0.2) +
  geom_ribbon(aes(ymin = Y5sigmafix, ymax = Y95sigmafix, color = 'Ysigmafix'), alpha = 0.2) +
  geom_ribbon(aes(ymin = mu5, ymax = mu95, color = 'mu'), alpha = 0.2) +
  facet_wrap(~ fit, scales = "free", nrow = 3)
```

## Diagnostic


# Hierarchical model

We explained leaf trait $Trait$ variation with ontogeny ($DBH$), abiotic environment ($TWI$), biotic interactions ($NCI$) and taxonomic levels ($C, S$ mixed effects, respectivelly fixed and random):

\begin{equation} 
  T_{c,s,i} \sim \mathcal{N}(\alpha_{env}.\frac{DBH_i}{{\beta_{DBH}}_{c,s} + DBH_i}, \sigma)    \\
  \alpha_{env} \sim \mathcal{N}(\alpha_{c,s} + {\beta_{TWI}}_{c,s}.TWI_i + {\beta_{Comp}}_{c,s}.\sum_{j=1}^{J} DBH_j^{2} . e^{-{\alpha_{NCI}}_{c,s}.\delta_{i,j}}, \sigma_{env})    \\
  \Theta_{c,s} \sim \mathcal{N}^5(\Theta_c,\sigma_{species}) ~|~ \Theta =
   \begin{bmatrix}
           \alpha \\
           \beta_{DBH} \\
           \beta_{TWI} \\
           \beta_{Comp} \\
           \alpha_{NCI}
  \end{bmatrix},~\sigma_{species}
  \begin{bmatrix}
           \sigma_{Intercept} \\
           \sigma_{DBH} \\
           \sigma_{TWI} \\
           \sigma_{Comp} \\
           \sigma_{NCI}
  \end{bmatrix}  \\
  \Theta_c \sim \mathcal{N}^{5}(0, + \infty) \\
  (\sigma, \sigma_{env}, \sigma_{species}) \sim \Gamma^{7}(- \infty, + \infty)
(\#eq:final)
\end{equation}

* $Trait_{c,s,i}$ is a trait of an individual i, from the species s, and the complex c;
* $i \in [1,I_s]$ where $I_s$ is the number of individuals in species s;
* $s \in [1,S_c]$ where $S_c$ is the number of species in complex c;
* $c \in [1,C]$ where $C$ is the number of complexes.
* $i \in [1,I]$ where I is the number of individuals;
* $j \in [1,J_i]$ where $J_i$ is the number of neighbors within the area of diameter $\delta_{max}$ surrounding the individual i, and $DBH_j$ the value of $DBH$ for neighbor $j$;
* $\alpha$ represents both the plateau of the Michaelis Menten and the intercept of each submodel (ontogeny, biotic & abiotic);
* $\beta_{DBH}$  is the value of DBH for which the trait account for half of its saturation value, and thus, ranges in positive values;
* $\beta_{TWI}$ is linear abiotic model parameters. $\beta_{TWI}$ can range in positive or negative value;
* $\beta_{Comp}$ is linear biotic model parameters. $\beta_{Comp}$ can range in positive or negative value;
* $\alpha_{NCI}$ represents the effect of neighbors distance to the individual and will be chosen positive;
* all $\sigma_{species}$ represents parameters variance for species;
* $\sigma$ and $\sigma_{species}$ are standard errors, and thus will have positive values;
* therefore, gamma laws were chosen as non informative flat priors on $\mathbb{R}^+$ for $\sigma$  and $\sigma_{species}$. Normal flat law were chosen as non informative flat prior on $\mathbb{R}$ for all $\Theta_{c,s}$ and $\Theta_c$.

## Code

```{stan, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  vector[N] Trait ; // Trait
  vector[N] DBH ; // Diameter at Breast Height
  vector[N] TWI ; // Topographic Wetness Index
  vector[N] DBHj ; // DBH neighbor
  vector[N] Deltaj ; // Distance neighbor
  int<lower=1, upper=N>  individual ; // individuals index in neighbor matrix
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
}
parameters {
  vector[S] alpha_s ; // Intercept for species
  vector[C] alpha_c ; // Intercept for complexes
  vector[S] betaDBH_s ; // DBH half-load for species
  vector[C] betaDBH_c ; // DBH half-load for complexes
  vector[S] betaTWI_s ; // Abiotic slope for species
  vector[C] betaTWI_c ; // Abiotic slope for complexes
  real<lower=0, upper=10> sigmaIntercept ; // Intercept variance for species
  real<lower=0, upper=10> sigmaDBH ; // DBH half-load variance for species
  real<lower=0, upper=10> sigmaTWI ; // Abiotic slope variance for species
  real<lower=0, upper=10> sigma ; // Variance
}
model {
  alpha_s ~ normal(alpha_c[speciesincomplex], sigmaIntercept) ;
  betaDBH_s ~ normal(betaDBH_c[speciesincomplex], sigmaDBH) ;
  betaTWI_s ~ normal(betaTWI_c[speciesincomplex], sigmaTWI) ;
  Trait ~ normal(, sigma) ; // Likelihood
} 
generated quantities {
  matrix[N,C] Trait_pred ;
  for(c in 1:C)
    Trait_pred[,c] = alpha_c[c] + betaTWI_c[c] *TWI ; // Predictions
}
```

## Summary

```{r , eval=F}
# fits <- lapply(mdata, function(x) sampling(Model, chains = 2, data = x))
# names(fits) <- traits
# save(fits, file = "./functional_save/TWIRNE.Rdata")
load("./functional_save/TWIRNE.Rdata")
pars <- c( "alpha_c", "betaTWI_c", "sigmaIntercept", "sigmaTWI", "sigma" )
lapply(as.list(traits), function(trait)
  broom::tidyMCMC(fits[[trait]], pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>%
   mutate(Trait = trait)) %>%
  bind_rows() %>%
  kable() %>%
  kable_styling()
```

## Posteriors

```{r , eval=F, fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(traits, function(trait)
  mcmc_intervals(as.array(fits[[trait]]), regex_pars = pars)),
  labels = names(data), nrow = 2)
```

```{r , eval=F, fig.cap="Model parameters posterior."}
cowplot::plot_grid(plotlist = lapply(traits, function(trait)
  mcmc_intervals(as.array(fits[[trait]]), regex_pars = "betaTWI")),
  labels = names(data), nrow = 2)
```

## Chains

```{r , eval=F, fig.cap="Markov chains trace plot after warmup for intermediate data."}
mcmc_trace(as.array(fits$SLA), regex_pars = c(pars, "lp__"),
             facet_args = list(labeller = label_parsed))
```

## Pairs

```{r , eval=F, fig.cap="Markov chains pairs plot after warmup  for intermediate data."}
mcmc_pairs(as.array(fits$SLA), regex_pars = pars)
```

## Predictions

```{r , eval=F, fig.cap="Predictions.", fig.width=12, fig.height=12}
lapply(as.list(names(mdata)), function(type)
  cbind(type = type, 
        Y = mdata[[type]]$Trait,
        TWI = as.numeric(scale(Individuals[!is.na(unlist(Individuals[,type])),"TWI"])),
        Individuals[!is.na(unlist(Individuals[,type])),],
        muEschweilera = apply(as.matrix(fits[[type]], 
                                        pars = paste0("Trait_pred[",
                                                      1:mdata[[type]]$N,",1]")), 
                              2, mean),
        muSymphonia = apply(as.matrix(fits[[type]], 
                                      pars = paste0("Trait_pred[",
                                                    1:mdata[[type]]$N,",2]")), 
                            2, mean),
        muEschweilera5 = apply(as.matrix(fits[[type]], 
                                         pars = paste0("Trait_pred[",
                                                       1:mdata[[type]]$N,",1]")), 
                               2, quantile, probs = 0.05),
        muSymphonia5 = apply(as.matrix(fits[[type]], 
                                       pars = paste0("Trait_pred[",
                                                      1:mdata[[type]]$N,",2]")), 
                             2, quantile, probs = 0.05),
        muEschweilera95 = apply(as.matrix(fits[[type]], 
                                          pars = paste0("Trait_pred[",
                                                      1:mdata[[type]]$N,",1]")), 
                                2, quantile, probs = 0.95),
        muSymphonia95 = apply(as.matrix(fits[[type]], 
                                        pars = paste0("Trait_pred[",
                                                      1:mdata[[type]]$N,",2]")), 
                              2, quantile, probs = 0.95)
  )) %>% 
  bind_rows() %>% 
  ggplot(aes(x = TWI)) +
  geom_point(aes(y = Y, col = as.factor(Genus))) +
  geom_ribbon(aes(ymin = muEschweilera5, ymax = muEschweilera95,
                  col = "Eschweilera"), alpha = 0.2) +
  geom_ribbon(aes(ymin = muSymphonia5, ymax = muSymphonia95,
                  col = "Symphonia"), alpha = 0.2) +
  geom_line(aes(y = muEschweilera, col = "Eschweilera")) +
  geom_line(aes(y = muSymphonia, col = "Symphonia")) +
  facet_wrap(~ type, scales = "free", nrow = 3)
```

## Diagnostic

# Conclusion

# References
