---
title: "A07 : Additive Model"
date: '`r Sys.Date()`'
author: Sylvain Schmitt & Anne Baranger
output:
  bookdown::html_document2:
    number_sections: no
    toc: true
    toc_float: yes
    theme: flatly
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
  bookdown::word_document2: default
linestretch: 1.5
---

```{r setup, include=FALSE}
#rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "../../data/Paracou/"
```

```{r data, eval=T}
# traitsSymphonia <- googlesheets::gs_title("Measures_Symphonia") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   mutate(SLA = as.numeric(SLA), LDMC = as.numeric(LDMC)) %>%
#   mutate(Genus = "Symphonia") %>%
#   rename(Species = Morphotype) %>%
#   mutate(Species = ifelse(Species == "Indet.",
#                           c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G",
#                                      "sp.1" =  "S")], Species))
# traitsEschweilera <- googlesheets::gs_title("Measures_Eschweilera") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   filter(!(Plot == 14 & SubPlot == 1 & TreeFieldNum == 760)) %>%  # outlier
#   filter(!(Species %in% c("congestiflora","simiorum","persistens")))
# traits <- bind_rows(traitsEschweilera, traitsSymphonia) %>%
#   mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>%
#   group_by(idTree, Plot, SubPlot, TreeFieldNum, Genus, Species,
#            SpeciesLong, Bark, Dawkins) %>%
#   summarise_at(vars("SLA", "LDMC", "LT", "LA", "CC",
#                     "brBT", "brBD", "brWD"), mean, na.rm = T) %>%
#   ungroup() %>%
#   mutate(invSLA=1/SLA) %>%
#   mutate(invLA=1/LA)
# load("./functional_save/env.Rdata")
# Individuals <-  left_join(traits, env, by = "idTree", suffix = c("", ".y"))
# rm(traitsEschweilera, traitsSymphonia, traits, paracou, env)
# save(Individuals, file = "./functional_save/Individuals.Rdata")
load("./functional_save/Individuals.Rdata")
load("./functional_save/CompetitionMatrix.Rdata")
```

# Introduction {-}

The aim of this document is to fit the additive full model with observed values of functional traits for our individuals. We prepared data excluding oultiers and reducing variance of all continuous variable to ease model fit (see figure \@ref(fig:mdata)).


We explained leaf trait $Trait$ variation with ontogeny ($DBH$), abiotic environment ($TWI$), biotic interactions ($NCI$) and taxonomic levels ($C, S$ mixed effects, respectivelly fixed and random):

\begin{equation} 
$$  T_{c,s,i} \sim \mathcal{N}(\alpha_{c,s}.\frac{DBH_i}{{\beta_{DBH}}_{c,s} + DBH_i} + {\beta_{TWI}}_{c,s}.TWI_i + {\beta_{Comp}}_{c,s}.\sum_{j=1}^{J} DBH_j^{2} . e^{-{\alpha_{NCI}}_{c,s}.\delta_{i,j}}, \sigma)    \\
  \Theta_{c,s} \sim \mathcal{N}^5(\Theta_c,\sigma_{species}) ~|~ \Theta =
   \begin{bmatrix}
           \alpha \\
           \beta_{DBH} \\
           \beta_{TWI} \\
           \beta_{Comp} \\
           \alpha_{NCI}
  \end{bmatrix},~\sigma_{species}
  \begin{bmatrix}
           \sigma_{Intercept} \\
           \sigma_{DBH} \\
           \sigma_{TWI} \\
           \sigma_{Comp} \\
           \sigma_{NCI}
  \end{bmatrix}  \\
  \Theta_c \sim \mathcal{N}^{5}(0, + \infty) \\
  (\sigma, \sigma_{species}) \sim \Gamma^{6}(- \infty, + \infty)$$
(\#eq:final)
\end{equation}

* $Trait_{c,s,i}$ is a trait of an individual i, from the species s, and the complex c;
* $i \in [1,I_s]$ where $I_s$ is the number of individuals in species s;
* $s \in [1,S_c]$ where $S_c$ is the number of species in complex c;
* $c \in [1,C]$ where $C$ is the number of complexes.
* $i \in [1,I]$ where I is the number of individuals;
* $j \in [1,J_i]$ where $J_i$ is the number of neighbors within the area of diameter $\delta_{max}$ surrounding the individual i, and $DBH_j$ the value of $DBH$ for neighbor $j$;
* $\alpha$ represents both the plateau of the Michaelis Menten and the intercept of each submodel (ontogeny, biotic & abiotic);
* $\beta_{DBH}$  is the value of DBH for which the trait account for half of its saturation value, and thus, ranges in positive values;
* $\beta_{TWI}$ is linear abiotic model parameters. $\beta_{TWI}$ can range in positive or negative value;
* $\beta_{Comp}$ is linear biotic model parameters. $\beta_{Comp}$ can range in positive or negative value;
* $\alpha_{NCI}$ represents the effect of neighbors distance to the individual and will be chosen positive;
* all $\sigma_{species}$ represents parameters variance for species;
* $\sigma$ and $\sigma_{species}$ are standard errors, and thus will have positive values;
* therefore, gamma laws were chosen as non informative flat priors on $\mathbb{R}^+$ for $\sigma$  and $\sigma_{species}$. Normal flat law were chosen as non informative flat prior on $\mathbb{R}$ for all $\Theta_{c,s}$ and $\Theta_c$.

```{r Ndata, fig.cap="Species headcount",fig.width=12}
Individuals %>% 
  group_by(Genus, Species) %>% 
  summarise(N = n()) %>% 
  arrange(desc(Genus), desc(N)) %>% 
  kable(caption = "Number of individuals per species and genus.", escape = F) %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1)
```

```{r mdata, fig.cap="prepared data",fig.width=12}
traits <- c("invSLA", "LDMC", "LT", "invLA", "CC")
mdata <- lapply(traits, function(trait){
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),]
  list(trait = trait,
       N = nrow(data_trait),
       J = nrow(Competition),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = as.numeric(scale(data_trait[trait], center = F)),
       DBH = as.numeric(scale(data_trait$DBH, center = F)),
       TWI = as.numeric(scale(data_trait$TWI, center = F)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2],
       data = data_trait)
})
names(mdata) <- traits

lapply(as.list(traits), function(trait)
  cbind(Model = trait, 
        Trait = mdata[[trait]]$Trait, 
        Individuals[unlist(!is.na(Individuals[trait])),])) %>% 
  bind_rows() %>% 
  select(Model, Trait, Genus, DBH, TWI) %>% 
  reshape2::melt(id.vars = c("Model", "Trait", "Genus")) %>% 
  ggplot(aes(x = value, y = Trait, col = Genus)) +
  geom_point() +
  geom_smooth() +
  facet_grid(Model ~ variable, scales = "free")
```

# Sampling

## Code

```{stan, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  int<lower=1> S ; // # species
  int<lower=1> C ; // # complex
  vector[N] Trait ; // Trait
  vector[N] DBH ; // Diameter at Breast Height
  vector[N] TWI ; // Topographic Wetness Index
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
}
parameters {
  vector [C] alpha_c ; // Intercept for complexes
  vector<lower=0> [C]  betaDBH_c ; // DBH half-load for complexes
  vector[C] betaTWI_c ; // Abiotic slope for complexes
  vector<lower=0> [S] alpha_s_tilde ; // Intercept for species
  vector<lower=0> [S] betaDBH_s_tilde ; // DBH half-load for species
  vector[S] betaTWI_s_tilde ; // Abiotic slope for species
  real<lower=0> sigmaIntercept ; // Intercept variance for species
  real<lower=0> sigmaDBH ; // DBH half-load variance for species
  real<lower=0> sigmaTWI ; // Abiotic slope variance for species
  real<lower=0> sigma ; // Variance
}
transformed parameters {
  vector[S] alpha_s ;
  vector<lower=0> [S] betaDBH_s ;
  vector[S] betaTWI_s ;
  alpha_s = alpha_c[speciesincomplex] + sigmaIntercept*alpha_s_tilde ;
  betaDBH_s = betaDBH_c[speciesincomplex] + sigmaDBH*betaDBH_s_tilde ;
  betaTWI_s = betaTWI_c[speciesincomplex] + sigmaTWI*betaTWI_s_tilde ;
}
model {
  alpha_c ~ normal(0, 10^6) ;
  betaDBH_c ~ lognormal(0,1) ;
  betaTWI_c ~ normal(0, 10^6) ;
  alpha_s_tilde ~ normal(0, 1) ;
  betaDBH_s_tilde ~ normal(0, 1) ;
  betaTWI_s_tilde ~ normal(0, 1) ;
  sigmaIntercept ~ lognormal(0, 1) ;
  sigmaDBH ~ lognormal(0, 1) ;
  sigmaTWI ~ lognormal(0, 1) ;
  sigma ~ cauchy(0, 5) ;
  Trait ~ normal(alpha_s[species] .* (DBH ./ (betaDBH_s[species] + DBH)) + betaTWI_s[species] .* TWI, sigma) ; // Likelihood
}
generated quantities {
  vector[N] Trait_pred ;
  real Rsquared ;
  vector[N] Trait_predDBH ;
  vector[N] Trait_predTWI ;
  vector[N] Trait_predNCI ;
  Trait_pred = alpha_c[complex] .* (DBH ./ (betaDBH_c[complex] + DBH)) + betaTWI_c[complex] .* TWI ;
  Rsquared = 1 - dot_self(Trait - Trait_pred)/dot_self(Trait - mean(Trait)) ;
  Trait_predDBH = alpha_c[complex] .* (DBH ./ (betaDBH_c[complex] + DBH)) + betaTWI_c[complex] * mean(TWI) ;
  Trait_predTWI = alpha_c[complex] .* (mean(DBH) ./ (betaDBH_c[complex] + mean(DBH))) + betaTWI_c[complex] .* TWI ;
  Trait_predNCI = alpha_c[complex] .* (mean(DBH) ./ (betaDBH_c[complex] + mean(DBH))) + betaTWI_c[complex] * mean(TWI) ;
}
```

## Summary

```{r }
# fits <- lapply(mdata, function(x)
#   sampling(Model, chains = 2, data = x, save_warmup = F,
#            include = F, pars = c('NCIj', "alpha_s", "betaDBH_s", "betaTWI_s",
#                                  "alpha_s_tilde", "betaDBH_s_tilde", "betaTWI_s_tilde")))
# names(fits) <- traits
# save(fits, file = "./functional_save/AdditiveAnne.Rdata")
load("./functional_save/AdditiveAnne.Rdata")
pars <- c("alpha_c", "betaDBH_c", "betaTWI_c",
          "sigmaIntercept", "sigmaDBH", "sigmaTWI", "sigma")
```

```{r summary1}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = pars, droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "Trait") %>% 
  mutate(complex = gsub("([[:alpha:]])", "", term)) %>% 
  select(Trait, term, complex, estimate, std.error, rhat) %>% 
  mutate(complex = gsub("([[:punct:]])", "", complex)) %>% 
  mutate(complex = recode_factor(complex, `2` = "Symphonia", 
                                 `1` = "Eschweilera")) %>% 
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(term = as.factor(gsub("c", "", term))) %>% 
  mutate(term = recode_factor(term, 
                              alpha = "$\\alpha$",
                              betaDBH = "$\\beta_{DBH}$", 
                              betaTWI = "$\\beta_{TWI}$", 
                              sigmaInterept = "$\\sigma^2_{Intercept}$",
                              sigmaDBH = "$\\sigma^2_{DBH}$", 
                              sigmaTWI = "$\\sigma^2_{TWI}$",
                              sigma = "$\\sigma^2$")) %>% 
  kable(caption = "Model parameters posterior.",
        escape = F, col.names = c("Trait", "Parameter", "Complex",
                                  "Estimate", "Standard Error", "$\\hat R$")) %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```


```{r summary2}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = pars, droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "Trait") %>% 
  mutate(complex = gsub("([[:alpha:]])", "", term)) %>% 
  select(Trait, term, complex, estimate) %>% 
  mutate(complex = gsub("([[:punct:]])", "", complex)) %>% 
  mutate(complex = recode_factor(complex, `2` = "Symphonia", 
                                 `1` = "Eschweilera")) %>% 
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(term = as.factor(gsub("c", "", term))) %>% 
  mutate(term = recode_factor(term, 
                              alpha = "$\\alpha$",
                              betaDBH = "$\\beta_{DBH}$", 
                              betaTWI = "$\\beta_{TWI}$", 
                              sigmaInterept = "$\\sigma^2_{Intercept}$",
                              sigmaDBH = "$\\sigma^2_{DBH}$", 
                              sigmaTWI = "$\\sigma^2_{TWI}$",
                              sigma = "$\\sigma^2$")) %>% 
  reshape2::dcast(term + complex ~ Trait, value.var = "estimate") %>% 
  kable(caption = "Model parameters posterior.",
        escape = F, col.names = c("Parameter", "Complex",
                                  "$CC$", "$\\frac{1}{LA}$", "$\\frac{1}{SLA}$",
                                  "$LDMC$", "$LT$")) %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```

## Posteriors

```{r posteriorsLpSigma, fig.cap="Model quality."}
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit),
                      pars = c("lp__", "sigma"))) %>% 
  bind_rows(.id = "trait") %>% 
  mutate(parameter = recode_factor(parameter, lp__ = "log-likelihood", 
                                   sigma = "sigma^2")) %>% 
  ggplot(aes(x = trait, xend = trait)) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_wrap(~ parameter, labeller = label_parsed, scales = "free") +
  xaxis_title(F) +
  yaxis_title(F)
```

```{r Rsquared}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = "Rsquared", droppars = NULL)) %>%
  bind_rows(.id = "Trait") %>% 
  select(Trait, estimate) %>% 
  kable(caption = "Model $R^2$",
        escape = F, col.names = c("Trait", "$R^2$")) %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T)
```


```{r posteriorsBeta, fig.width=12, fig.height=6}
gdata <- lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit),
                      regex_pars = c("alpha_c", "betaDBH_c",
                                     "betaTWI_c"))) %>% 
  bind_rows(.id = "trait") %>% 
  mutate(complex = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(complex = gsub("([[:punct:]])", "", complex)) %>% 
  mutate(complex = ifelse(complex == 2, "Symphonia", "Eschweilera")) %>% 
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(parameter = as.factor(gsub("c", "", parameter))) %>% 
  mutate(parameter = recode_factor(parameter, betaDBH = "beta[DBH]", 
                                   betaTWI = "beta[TWI]")) %>% 
  mutate(parameter = factor(parameter, levels = c("alpha", "beta[DBH]", 
                                                  "beta[TWI]")))
student <- lapply(fits, function(fit) 
  as.data.frame(fit, pars = c("alpha_c", "betaDBH_c",
                                     "betaTWI_c"))) %>% 
  bind_rows(.id = "trait") %>%
  reshape2::melt(id.vars = "trait", variable.name = "parameter") %>% 
  mutate(complex = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(complex = gsub("([[:punct:]])", "", complex)) %>% 
  mutate(complex = ifelse(complex == 2, "Symphonia", "Eschweilera")) %>% 
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(parameter = as.factor(gsub("c", "", parameter))) %>% 
  mutate(parameter = recode_factor(parameter, betaDBH = "beta[DBH]", 
                                   betaTWI = "beta[TWI]")) %>% 
  group_by(trait, parameter) %>% 
  do(student = t.test(value ~ complex, data = .)) %>% 
  mutate(student = student$p.value) %>% 
  mutate(student = ifelse(student < 0.001, "***",
                          ifelse(student < 0.01, "**",
                                 ifelse(student < 0.05, "*",
                                        ifelse(student < 0.1, ".", 
                                               "n.s."))))) %>% 
  left_join(gdata %>% 
              group_by(trait, parameter) %>% 
              summarise(x = mean(m)))
ggplot(gdata, aes(x = trait, xend = trait, col = complex, fill = complex)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  geom_text(aes(x = trait, y = x, label = student,
                col = NA, fill = NA), position = position_nudge(x = 0.1),
            data = student) +
  facet_wrap(~ parameter, labeller = label_parsed, scales = "free") +
  xaxis_title(F) +
  yaxis_title(F)
```

```{r posteriorsSigma2, fig.width=12, fig.height=6}
gdata <- lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit),
                      regex_pars = c("sigma"))) %>% 
  bind_rows(.id = "trait") %>% 
  filter(parameter != "sigma") %>% 
  mutate(parameter = recode_factor(parameter, 
                                   sigmaIntercept = "sigma[Intercept]^2",
                                   sigmaDBH = "sigma[DBH]^2", 
                                   sigmaTWI = "sigma[TWI]^2",
                                   sigmaComp = "sigma[Comp]^2"))
ggplot(gdata, aes(x = trait, xend = trait)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_wrap(~ parameter, labeller = label_parsed, scales = "free") +
  xaxis_title(F) +
  yaxis_title(F)
```

## Predictions

```{r , fig.cap="Predictions.", fig.width=12, fig.height=12}
lapply(as.list(c("DBH", "TWI")), function(variable)
  lapply(as.list(traits), function(trait) {
    data.frame(
      trait = trait, 
      variable = variable,
      Y = mdata[[trait]]$Trait,
      value = mdata[[trait]][[variable]],
      Genus = mdata[[trait]]$complex,
      mu = apply(as.matrix(fits[[trait]], pars = paste0("Trait_pred", variable)), 
                 2, mean),
      mu5 = apply(as.matrix(fits[[trait]], pars = paste0("Trait_pred", variable)), 
                  2, quantile, probs = 0.05),
      mu95 = apply(as.matrix(fits[[trait]], pars = paste0("Trait_pred", variable)), 
                   2, quantile, probs = 0.95)
    )
  }) %>% bind_rows()) %>% 
  bind_rows() %>% 
  ggplot(aes(x = value, col = as.factor(Genus))) +
  geom_point(aes(y = Y)) +
  geom_ribbon(aes(ymin = mu5, ymax = mu95), alpha = 0.2) +
  geom_line(aes(y = mu)) +
  facet_grid(trait ~ variable, scales = "free") +
  scale_color_discrete("Complex", labels = c("Eschweilera", "Symphonia"))
```

# Conclusion

# References
