---
title: "A08 : Interaction Model only with Species 4"
date: '`r Sys.Date()`'
author: Sylvain Schmitt
output:
  bookdown::html_document2:
    number_sections: no
    toc: true
    toc_float: yes
    theme: flatly
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::word_document2: default
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
#rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "../../data/Paracou/"
```

```{r data, eval=T}
# traitsSymphonia <- googlesheets::gs_title("Measures_Symphonia") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   mutate(SLA = as.numeric(SLA), LDMC = as.numeric(LDMC)) %>%
#   mutate(Genus = "Symphonia") %>%
#   rename(Species = Morphotype) %>%
#   mutate(Species = ifelse(Species == "Indet.",
#                           c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G",
#                                      "sp.1" =  "S")], Species))
# traitsEschweilera <- googlesheets::gs_title("Measures_Eschweilera") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   filter(!(Plot == 14 & SubPlot == 1 & TreeFieldNum == 760)) %>%  # outlier
#   filter(!(Species %in% c("congestiflora","simiorum","persistens")))
# traits <- bind_rows(traitsEschweilera, traitsSymphonia) %>%
#   mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>%
#   group_by(idTree, Plot, SubPlot, TreeFieldNum, Genus, Species,
#            SpeciesLong, Bark, Dawkins) %>%
#   summarise_at(vars("SLA", "LDMC", "LT", "LA", "CC",
#                     "brBT", "brBD", "brWD"), mean, na.rm = T) %>%
#   ungroup() %>%
#   mutate(invSLA=1/SLA) %>%
#   mutate(invLA=1/LA)
# load("./functional_save/env.Rdata")
# Individuals <-  left_join(traits, env, by = "idTree", suffix = c("", ".y"))
# rm(traitsEschweilera, traitsSymphonia, traits, paracou, env)
# save(Individuals, file = "./functional_save/Individuals.Rdata")
load("./functional_save/Individuals.Rdata")
load("./functional_save/CompetitionMatrix.Rdata")
Individuals <- Individuals %>% 
  group_by(Species) %>% 
  filter(n() > 10) %>% 
  ungroup() %>% 
  left_join(Competition %>% 
              group_by(idTree) %>% 
              summarise(NCI = (1/mean(1-AreaOutside20))*sum(DBHj^2*exp(-0.25*dij))))
```

```{r mdata}
traits <- c("invSLA", "LDMC", "LT", "invLA", "CC")
complexes <- c("Symphonia", "Eschweilera")
models <- sapply(complexes, function(complex) paste0(complex, "-", traits))
mdata <- lapply(models, function(model){
  trait <- strsplit(model, "-")[[1]][2]
  complex <- strsplit(model, "-")[[1]][1]
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    filter(Genus == complex) %>% 
    left_join(select(Competition, idTree, AreaOutside20) 
              %>% unique())
  Competition_trait <- Competition
  Competition_trait$idTree <- match(Competition_trait$idTree, data_trait$idTree)
  Competition_trait <- filter(Competition_trait, !is.na(idTree))
  list(Model = model,
       Complex = complex,
       N = nrow(data_trait),
       J = nrow(Competition_trait),
       S = length(unique(data_trait$Species)),
       Trait = unlist(data_trait[trait])/sd(unlist(data_trait[trait])),
       DBH = data_trait$DBH/sd(data_trait$DBH),
       TWI = data_trait$TWI/sd(data_trait$TWI),
       NCI = data_trait$NCI/sd(data_trait$NCI),
       weights = 1-data_trait$AreaOutside20,
       individual = as.numeric(as.factor(Competition_trait$idTree)),
       species = as.numeric(as.factor(data_trait$Species)),
       data = data_trait,
       competition = Competition_trait,
       sd = c(Trait = sd(unlist(data_trait[trait])), 
              DBH = sd(data_trait$DBH), 
              TWI = sd(data_trait$TWI), 
              NCI = sd(data_trait$NCI)))
})
names(mdata) <- models
```

```{r fits}
# fits <- lapply(mdata, function(x)
#   sampling(Model, chains = 2, data = x, save_warmup = F,
#            include = F, pars = c('NCIj')))
# names(fits) <- models
# save(fits, file = "./functional_save/InteractionSpecies4.Rdata")
load("./functional_save/InteractionSpecies4.Rdata")
pars <- c("alpha", "betaDBH", "betaTWI", "betaComp", "sigma")
```

```{r species}
species <- lapply(models, function(model){
  trait <- strsplit(model, "-")[[1]][2]
  complex <- strsplit(model, "-")[[1]][1]
  Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    filter(Genus == complex) %>% 
    mutate(species = as.numeric(as.factor(Species))) %>% 
    mutate(SpeciesLong = paste(substr(complex, 1, 1), Species)) %>% 
    select(Genus, Species, SpeciesLong, species) %>% 
    unique()
}) %>% bind_rows() %>% 
  unique() %>% 
  mutate(complex = Genus)
```

# Introduction {-}

# Material and Methods

# Results

```{r summarySpecies}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = c("alpha", "betaDBH", "betaTWI",
                                "betaComp", "sigma"), 
                  droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-") %>% 
  mutate(species = gsub("([[:alpha:]])", "", term)) %>% 
  select(complex, trait, term, species, estimate, std.error, rhat) %>% 
  mutate(species = gsub("([[:punct:]])", "", species)) %>% 
  mutate(species = as.numeric(species)) %>% 
  left_join(species) %>% 
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(term = recode_factor(term, 
                              alphas = "$\\alpha$",
                              betasDBH = "$\\beta_{DBH}$", 
                              betasTWI = "$\\beta_{TWI}$", 
                              betasComp = "$\\beta_{Comp}$", 
                              alphaNCI = "$\\alpha_{NCI}$",
                              sigma = "$\\sigma^2$")) %>% 
  mutate(trait = recode_factor(trait, 
                              invSLA = "$\\frac{1}{SLA}$",
                              LDMC = "$LDMC$", 
                              LT = "$LT$", 
                              invLA = "$\\frac{1}{LA}$",
                              CC = "$CC$")) %>% 
  reshape2::dcast(complex + term + SpeciesLong ~ trait, value.var = "estimate") %>% 
  rename(Parameter = term, Species = SpeciesLong) %>% 
  kable(caption = "Model parameters for each species with reduced traits and descriptors.",
        escape = F, digits = 3, format = "pandoc") %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```

```{r speciesIntercept, fig.cap="Model species effect for Intercept."}
S_sqrt_trans <- function() scales::trans_new("S_sqrt",
                                             function(x) sign(x)*sqrt(abs(x)),
                                             function(x) x^2*sign(x))
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = "alpha"))) %>% 
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-") %>%
  mutate(species = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = gsub("([[:alpha:]])", "", species)) %>% 
  mutate(species = as.numeric(species)) %>% 
  left_join(species) %>% 
  ggplot(aes(x = datatools::reorder_within(SpeciesLong, m, trait), 
             xend = datatools::reorder_within(SpeciesLong, m, trait),
             color = SpeciesLong, fill = SpeciesLong)) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  datatools::scale_x_reordered() +
  facet_grid(trait ~ Genus, labeller = label_parsed, 
             scales = "free") +
  xaxis_title(F) +
  ylab(expression(alpha)) +
  scale_y_continuous(trans="S_sqrt") +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")
```

```{r speciesDBH, fig.cap="Model species effect for $DBH_{90}$."}
S_sqrt_trans <- function() scales::trans_new("S_sqrt",
                                             function(x) sign(x)*sqrt(abs(x)),
                                             function(x) x^2*sign(x))
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = "betaDBH"))) %>% 
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-",remove = F) %>%
  mutate(species = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = gsub("([[:alpha:]])", "", species)) %>% 
   mutate(species = as.numeric(species)) %>% 
  left_join(species) %>% 
  left_join(lapply(mdata, function(data) 
    data.frame(variable = names(data$sd), sd = data$sd)) %>% 
      bind_rows(.id = "model") %>% 
      filter(variable == "DBH") %>% 
      reshape2::dcast(model ~ variable) %>% 
      rename(sdDBH = DBH)) %>% 
  mutate_at(vars("ll", "l", "m", "h", "hh"), funs(.*sdDBH*9)) %>% 
  ggplot(aes(x = datatools::reorder_within(SpeciesLong, m, trait), 
             xend = datatools::reorder_within(SpeciesLong, m, trait),
             color = SpeciesLong, fill = SpeciesLong)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  datatools::scale_x_reordered() +
  facet_grid(trait ~ Genus, labeller = label_parsed, 
             scales = "free") +
  ylab(expression(DBH[90]~(cm))) +
  scale_y_continuous(trans="S_sqrt") +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")
```

```{r speciesTWI, fig.cap="Model species effect for TWI."}
S_sqrt_trans <- function() scales::trans_new("S_sqrt",
                                             function(x) sign(x)*sqrt(abs(x)),
                                             function(x) x^2*sign(x))
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = "betaTWI"))) %>% 
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-") %>%
  mutate(species = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = gsub("([[:alpha:]])", "", species)) %>% 
    mutate(species = as.numeric(species)) %>% 
  left_join(species) %>% 
  ggplot(aes(x = datatools::reorder_within(SpeciesLong, m, trait), 
             xend = datatools::reorder_within(SpeciesLong, m, trait),
             color = SpeciesLong, fill = SpeciesLong)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_grid(trait ~ Genus, labeller = label_parsed, 
              scales = "free") +
  datatools::scale_x_reordered() +
  xaxis_title(F) +
  ylab(expression(beta[TWI])) +
  scale_y_continuous(trans="S_sqrt") +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")
```

```{r speciesNCI, fig.cap="Model species effect for NCI."}
S_sqrt_trans <- function() scales::trans_new("S_sqrt",
                                             function(x) sign(x)*sqrt(abs(x)),
                                             function(x) x^2*sign(x))
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = "betaComp"))) %>% 
  bind_rows(.id = "model") %>% 
  separate(model, c("complex", "trait"), "-") %>%
  mutate(species = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(species = gsub("([[:alpha:]])", "", species)) %>% 
    mutate(species = as.numeric(species)) %>% 
  left_join(species) %>% 
  ggplot(aes(x = datatools::reorder_within(SpeciesLong, m, trait), 
             xend = datatools::reorder_within(SpeciesLong, m, trait),
             color = SpeciesLong, fill = SpeciesLong)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  datatools::scale_x_reordered() +
    facet_grid(trait ~ Genus, labeller = label_parsed, 
              scales = "free") +
  xaxis_title(F) +
  ylab(expression(beta[Comp])) +
  scale_y_continuous(trans="S_sqrt") +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")
```

```{r predicitons}
traits <- c("invSLA", "LDMC", "LT", "invLA", "CC")
complexes <- c("Symphonia", "Eschweilera")
models <- sapply(complexes, function(complex) paste0(complex, "-", traits))
preds <- lapply(models, function(model){
  trait <- strsplit(model, "-")[[1]][2]
  complex <- strsplit(model, "-")[[1]][1]
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    filter(Genus == complex) %>% 
    left_join(select(Competition, idTree, AreaOutside20) 
              %>% unique())
  
  DBH <- data_trait$DBH/sd(data_trait$DBH)
  TWI <- data_trait$TWI/sd(data_trait$TWI)
  NCI <- data_trait$NCI/sd(data_trait$NCI)
  species <- as.numeric(as.factor(data_trait$Species))
  
  alpha <- as.matrix(fits[[model]], pars = "alpha")
  betaDBH <- as.matrix(fits[[model]], pars = "betaDBH")
  betaTWI <- as.matrix(fits[[model]], pars = "betaTWI")
  betaComp <- as.matrix(fits[[model]], pars = "betaComp")
  
  Y <- unlist(data_trait[trait])/sd(unlist(data_trait[trait]))
  Yp <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * TWI +  betaComp[n,species] * NCI) * (DBH / (betaDBH[n,species] + DBH)), simplify = F))
  Ycomplex <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * TWI +  betaComp[n,species] * NCI) * (DBH / (betaDBH[n,species] + DBH)), simplify = F))
  Ydbh <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * mean(TWI) +  betaComp[n,species] * mean(NCI)) * (DBH / (betaDBH[n,species] + DBH)), simplify = F))
  Ytwi <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * TWI +  betaComp[n,species] * mean(NCI)) * (mean(DBH) / (betaDBH[n,species] + mean(DBH))), simplify = F))
  Ycomp <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * mean(TWI) +  betaComp[n,species] * NCI) * (mean(DBH) / (betaDBH[n,species] + mean(DBH))), simplify = F))
  
  Ddbh <- sqrt((Yp - Ydbh)^2)
  Dtwi <- sqrt((Yp - Ytwi)^2)
  Dcomp <- sqrt((Yp - Ycomp)^2)
  
  data.frame(
    Model = model,
    Trait = trait,
    Complex = complex,
    DBH = data_trait$DBH, 
    TWI = data_trait$TWI, 
    NCI = NCI,
    Genus = data_trait$Genus,
    Species = data_trait$Species,
    Y = unlist(data_trait[trait])/sd(unlist(data_trait[trait])),
    Yp = apply(Yp, 2, mean),
    Yp5 = apply(Yp, 2, quantile, probs = 0.05),
    Yp95 = apply(Yp, 2, quantile, probs = 0.95),
    Ydbh = apply(Ydbh, 2, mean),
    Ydbh5 = apply(Ydbh, 2, quantile, probs = 0.05),
    Ydbh95 = apply(Ydbh, 2, quantile, probs = 0.95),
    Ytwi = apply(Ytwi, 2, mean),
    Ytwi5 = apply(Ytwi, 2, quantile, probs = 0.05),
    Ytwi95 = apply(Ytwi, 2, quantile, probs = 0.95),
    Ycomp = apply(Ycomp, 2, mean),
    Ycomp5 = apply(Ycomp, 2, quantile, probs = 0.05),
    Ycomp95 = apply(Ycomp, 2, quantile, probs = 0.95),
    Ddbh = apply(Ddbh, 2, mean),
    Ddbh5 = apply(Ddbh, 2, quantile, probs = 0.05),
    Ddbh95 = apply(Ddbh, 2, quantile, probs = 0.95),
    Dtwi = apply(Dtwi, 2, mean),
    Dtwi5 = apply(Dtwi, 2, quantile, probs = 0.05),
    Dtwi95 = apply(Dtwi, 2, quantile, probs = 0.95),
    Dcomp = apply(Dcomp, 2, mean),
    Dcomp5 = apply(Dcomp, 2, quantile, probs = 0.05),
    Dcomp95 = apply(Dcomp, 2, quantile, probs = 0.95)
  )}) %>% bind_rows()
```

```{r sensitivity, fig.cap="Model sensitivity to variables. ..."}
preds %>%
  select(Trait, Genus, Ddbh, Dtwi, Dcomp) %>% 
  rename(DBH = Ddbh, TWI = Dtwi, Competition = Dcomp) %>% 
  reshape2::melt(id.vars = c("Trait", "Genus")) %>% 
  mutate(variable = factor(variable, levels = c("Competition", "TWI", "DBH"))) %>%
  ggplot(aes(variable, value, fill = variable)) +
  geom_boxplot() +
  facet_grid(Trait ~ Genus, scales = "free") +
  coord_flip() +
  scale_y_sqrt() +
  scale_fill_discrete(guide = "none") +
  xlab("") + ylab(expression(Delta(Prediction)))
```

# Discussion

# Supplementary materials

## S1: Model

We explained leaf trait $Trait$ variation with ontogeny ($DBH$), abiotic environment ($TWI$), biotic interactions ($NCI$) and taxonomic levels (species complex $C$ and species $S$) with a linear mixed model including interaction between ontogeny and environment (more likely than an additive form without interactions, see analysis [A07-Additive](A07-Additive.html)). We obtained the following model (see [Material and Methods] for more detail on the model form):

\begin{equation}
  T_{s,i} \sim \mathcal{N}(\frac{DBH_i}{{\beta_{DBH}}_{s} + DBH_i} \times (\alpha_{s} + {\beta_{TWI}}_{s}.TWI_i + {\beta_{Comp}}_{s}.\sum_{j=1}^{J} DBH_j^{2} . e^{-{\alpha_{NCI}}.\delta_{i,j}}), \sigma)  
\end{equation} 

Resulting in the following `stan` code with selected priors:

```{stan stanModel, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1> S ; // # species
  vector[N] Trait ; // Trait
  vector[N] DBH ; // Diameter at Breast Height
  vector[N] TWI ; // Topographic Wetness Index
  vector[N] NCI ; // DBH neighbor
  int<lower=1, upper=S> species[N] ; // species index
}
parameters {
  vector [S] alpha ; // Intercept
  vector<lower=0> [S]  betaDBH ; // DBH half-load
  vector[S] betaTWI ; // TWI slope
  vector[S] betaComp ; // Competition slope
  real<lower=0> sigma ; // Variance
}
model {
  alpha ~ normal(0, 10^6) ; // Priors
  betaDBH ~ normal(0,1) ;
  betaTWI ~ normal(0,1) ;
  betaComp ~ lognormal(0,1) ;
  sigma ~ cauchy(0, 5) ;
  Trait ~ normal((alpha[species] + betaTWI[species] .* TWI +  betaComp[species] .* NCI) .* (DBH ./ (betaDBH[species] + DBH)), sigma) ; // Likelihood
}
```

## S2: Diagnostic

```{r mcmc, fig.cap="Markov chains trace plot after warmup for leaf thickness model."}
# lapply(fits, check_divergences)
mcmc_trace(as.array(fits$`Eschweilera-LT`, pars = c(pars, "lp__")),
             facet_args = list(labeller = label_parsed))
# mcmc_pairs(as.array(fits$invSLA), regex_pars = pars)
```

# References
