---
title: "A08 : Interaction Model"
date: '`r Sys.Date()`'
author: Sylvain Schmitt & Anne Baranger
output:
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
  bookdown::word_document2: default
  bookdown::html_document2:
    number_sections: no
    toc: true
    toc_float: yes
    theme: flatly
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
#rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "../../data/Paracou/"
```

```{r data, eval=T}
# traitsSymphonia <- googlesheets::gs_title("Measures_Symphonia") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   mutate(SLA = as.numeric(SLA), LDMC = as.numeric(LDMC)) %>%
#   mutate(Genus = "Symphonia") %>%
#   rename(Species = Morphotype) %>%
#   mutate(Species = ifelse(Species == "Indet.",
#                           c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G",
#                                      "sp.1" =  "S")], Species))
# traitsEschweilera <- googlesheets::gs_title("Measures_Eschweilera") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   filter(!(Plot == 14 & SubPlot == 1 & TreeFieldNum == 760)) %>%  # outlier
#   filter(!(Species %in% c("congestiflora","simiorum","persistens")))
# traits <- bind_rows(traitsEschweilera, traitsSymphonia) %>%
#   mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>%
#   group_by(idTree, Plot, SubPlot, TreeFieldNum, Genus, Species,
#            SpeciesLong, Bark, Dawkins) %>%
#   summarise_at(vars("SLA", "LDMC", "LT", "LA", "CC",
#                     "brBT", "brBD", "brWD"), mean, na.rm = T) %>%
#   ungroup() %>%
#   mutate(invSLA=1/SLA) %>%
#   mutate(invLA=1/LA)
# load("./functional_save/env.Rdata")
# Individuals <-  left_join(traits, env, by = "idTree", suffix = c("", ".y"))
# rm(traitsEschweilera, traitsSymphonia, traits, paracou, env)
# save(Individuals, file = "./functional_save/Individuals.Rdata")
load("./functional_save/Individuals.Rdata")
load("./functional_save/CompetitionMatrix.Rdata")
```

```{r mdata}
traits <- c("invSLA", "LDMC", "LT", "invLA", "CC")
mdata <- lapply(traits, function(trait){
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    left_join(select(Competition, idTree, AreaOutside20) 
              %>% unique())
  Competition_trait <- Competition
  Competition_trait$idTree <- match(Competition_trait$idTree, data_trait$idTree)
  Competition_trait <- filter(Competition_trait, !is.na(idTree))
  list(Model = trait,
       N = nrow(data_trait),
       J = nrow(Competition_trait),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = unlist(data_trait[trait])/sd(unlist(data_trait[trait])),
       DBH = data_trait$DBH/sd(data_trait$DBH),
       TWI = data_trait$TWI/sd(data_trait$TWI),
       DBHj = Competition_trait$DBHj/sd(Competition_trait$DBHj),
       Deltaj = Competition_trait$dij/sd(Competition_trait$dij),
       weights = 1-data_trait$AreaOutside20,
       individual = as.numeric(as.factor(Competition_trait$idTree)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2],
       data = data_trait,
       competition = Competition_trait,
       sd = c(Trait = sd(unlist(data_trait[trait])), 
              DBH = sd(data_trait$DBH), 
              TWI = sd(data_trait$TWI), 
              DBHj = sd(Competition_trait$DBHj), 
              Deltaj = sd(Competition_trait$dij)))
})
names(mdata) <- traits
```

```{r fits}
# fits <- lapply(mdata, function(x)
#   sampling(Model, chains = 2, data = x, save_warmup = F,
#            include = F, pars = c('NCIj', "alpha_s", "betaDBH_s", "betaTWI_s",
#                                  "betaComp_s", "alpha_s_tilde",
#                                  "betaDBH_s_tilde", "betaTWI_s_tilde", "betaComp_s_tilde")))
# names(fits) <- traits
# save(fits, file = "./functional_save/Interaction.Rdata")
load("./functional_save/Interaction.Rdata")
pars <- c("alpha", "betaDBH", "betaTWI", "betaComp", "alphaNCI",
          "sigmaIntercept", "sigmaDBH", "sigmaTWI", "sigmaComp", "sigma")
```

# Introduction {-}

Subsequent analysis aimed to explore determinant of individual leaf functional traits from Symphonia globulifera and Eschweilera clade Parvifolia species complexes. More especially, we investigated how individual ontogeny, abiotic environment, and biotic interactions shaped its leaf phenotype at several taxonomic scale, from individual to complex including species. Determinant of the functional traits have been studied with the interaction model developed in document [A03-Models](A03-Models.html) and are further detailed.

**Real introduction to be developped (maybe only in the first draft and not directly in this analysis).**

# Material and Methods

## Study site

The study was conducted in the northernmost part of the Guiana Plateau region, at the Paracou field station. The site was characterized by an average of 3041 mm annual rainfall and a mean air temperature of 25.71 °C. Old tropical forest with an exceptional richness (over 750 woody species) has developed among the succession of small hills of this area, rising to 10–40 m a.s.l. [@Gourlet-Fleury2004].

The site is made of 16 permanent plots (fifteen 6.25 ha plus one 25 ha) which have been censused (DBH>10) every 1-2 years for more than 35 years. Nine of the plots were logged and subjected to human-induced disturbance in 1986.

## Individuals sampling

Two sampling campaigns were led in 2017 and 2018 during dry season (between September and December). Sampling was made for respectively 14 and 2 species from *Eschweilera* (Lecythidaceae) and *Symphonia* (Clusiaceae) genera present in the Paracou station. Species from *Eschweilera* genus were all included in the *Parvifolia* clade [@Mori2016; Heuertz et al, unpublished]. Both *Eschweilera* clade *Parvifolia* and *Symphonia* species are Amazonian hyperdominant tree species [@TerSteege2013] and part of the most abundant trees within Paracou station (*Eschweilera sagotiana* from the clade *Parvifolia* is the most abundant tree species in 2017 for Paracou).

*Symphonia* genus only include the *Symphonia globulifera* L.f (Clusiaceae) species in French Guiana but presents two morphotypes, *S. globulifera* and *S. sp.1*, living in sympatry but apparently in differentiated habitats, with *S. globulifera* preferentially growing in valley bottoms and *S. sp1* preferentially exploiting a variety of drier habitats. Similarly, *Eschweilera* clade *parvifolia* include numerous species, with *E. sagotiana* and *E. coriacea* which have been shown to follow a niche differentiation similar to *S. sp1* and *S. globulifera* respectively [@Allie2015].

**We need to add that they are species complex.**

Between 232 and 1 individual have been sampled per species, depending on their natural abundance within the Paracou station, for a total of 823 individuals (Table \@ref(tab:headcount)). *Symphonia globulifera* were collected over all plots of the study site to reach targeted number, whereas *Eschweilera* clade *Parvifolia* were sampled only within non-disturbed plots. For each individual, 5 mature and healthy leaves in consistent light conditions were collected, and kept in humidified ziplock with enriched CO2 air in dark until measurement within the next 6 hours following standard protocol [@Perez-Harguindeguy2013]. In addition, light conditions with Dawkins index [@Dawkins1958] and tree, bark and leaf morphology were assessed for every individuals.

```{r headcount}
Individuals %>% 
  mutate(Species = gsub("_", " ", Species)) %>% 
  group_by(Genus, Species) %>% 
  summarise(N = n()) %>% 
  arrange(desc(N)) %>% 
  kable(caption = "Number of individuals sampled per genus and species.",
        escape = F, format = "pandoc") %>% 
  kable_styling(full_width = F)
```

## Trait measurements

We calculated 5 functional traits from leaf samples already highlighted in previous studies as phenotypic descriptor of major ecological strategies [@diaz_plant_1998; @wright_worldwide_2004] (Table \@ref(tab:Trait)): specific leaf area (leaf area divided by the fresh weight, $SLA$, $cm^2.g^{-1}$), leaf dry matter content (leaf dry weight divided by the fresh weight, $LDMC$, $g.g^{-1}$), leaf thickness ($LT$, $\mu m$), leaf area ($LA$, $cm^2$), leaf chlorophyll content ($CC$, $g.cm^{-2}$). Leaf petiole was removed for all measurements. Leaf fresh weight was measured with an analytic balance [brand], leaf fresh thickness with a micrometer [brand], leaf area through the ImageJ software [@Schneider2012] using scanned images of fresh leaves, and leaf chlorophyll content with a SPAD measurement [brand] converted with an allometric model [@Coste2010]. Leaf thickness and chlorophyll content measurement was repeated 3 times per leaf to reduce errors due to intra-leaf variation. Leaves were then kept in herbarium and oven-dried for at least 48 hours at 80°C before measurement of dry weight. 

```{r Trait}
Individuals %>% 
  select(SLA, LDMC, LT, LA, CC) %>% 
  reshape2::melt(variable.name = "trait") %>% 
  filter(!is.na(value)) %>% 
  group_by(trait) %>% 
  summarise(N = n(),
            Value = paste0("$",round(mean(value), 2), "\\pm", 
                           round(sd(value), 2), "$")) %>% 
  mutate(Trait = c("Specific Leaf Area", "Leaf Dry Matter Content",
                    "leaf Thickness", "Leaf Area", "Chlorophyll Content")) %>% 
  mutate(Unit = c("$cm^2.g^{-1}$", "$g.g^{-1}$", "$\\mu m$", "$cm^2$", "$g.cm^{-2}$")) %>%
  select(Trait, Unit, trait, N, Value) %>% 
  kable(col.names = c("Trait", "Unit", "Abbreviation", "N", "Mean (sd)"), format = "pandoc",
    caption = "Functional traits measured, with trait unit, abbreviation, headcount, mean and standard error.", escape = F) %>% 
  kable_styling(full_width = F)
```

## Environmental variables

Three little correlated (Pearson's r < 0.14, see analysis [A02-Environment](A02-Environment.html)) ontogenetic and environmental descriptors were chosen to depict individual leaf traits. Diameter at breast height ($DBH$, $cm$) was chosen as an indirect proxy for ontogeny effect [but see @roggy_links_2005] but was shown to be confounded with light access through Dawkins index (see analysis [A02-Environment](A02-Environment.html)).

A neighbor crowding index [$NCI$; @Uriarte2004a] was used as a descriptor of biotic asymmetric competition through tree neighborhood, implying that other biotic effects, as herbivory and pathogens effects, were not taken into account. The neighborhood crowding index $NCI_i$ from tree individual $i$ was calculated with the following formula:
$$NCI_i = \sum _{j|\delta_{i,j}<20m} ^{J_i} DBH_j ^2 e^{-\alpha.\delta_{i,j}}$$
with $DBH_j$ the diameter from neighboring tree $j$ and $\delta_{i,j}$ its distance to individual tree $i$. $NCI_i$ is computed for all neighbors at a distance $\delta_{i,j}$ inferior to 20m as often observed in literature [ref]. $\alpha$ represents the decrease of neighbors $DBH_j$ effect with distance (note that  set to 0 represent local basal area). 

Finally, topographic wetness index ($TWI$) was selected between various abiotic descriptors available (see analysis [A02-Environment](A02-Environment.html)) and highlighted water accumulation areas, which is crucial information at local scale when it comes to tropical ecosystems [@ferry2010higher]. Descriptors used were derived from field censuses ($DBH$ and $NCI$) or derived from a 1-m resolution digital elevation model built using LiDAR campaign done in 2015 ($TWI$) using SAGA-GIS [ref].


## Analysis

We explained leaf trait $Trait$ variation with ontogeny ($DBH$), abiotic environment ($TWI$), biotic interactions ($NCI$) and taxonomic levels (species complex $C$ and species $S$) with a linear mixed model including interaction between ontogeny and environment (more likely than an additive form without interactions, see analysis [A07-Additive](A07-Additive.html)).

Individual leaf trait $Trait_{c,s,i}$ for species complex $c \in [1;C]$ and species $s \in [1;S_c]$ is first explained by its ontogeny confounded with light though individual DBH $DBH_i$ with a Michaelis Menten form (more likely than a linear form, see analysis [A05-Ontogeny](A05-Ontogeny.html)):
$$T_{c,s,i} \sim \mathcal{N}(\alpha_{env}.\frac{DBH_i}{{\beta_{DBH}}_{c,s} + DBH_i}, \sigma)$$
Because of the Michaelis Menten form $\beta_{DBH}$ is the value of DBH for which the trait account for half of its mature value (when DBH tend to its maximum value). With higher $\beta_{DBH}$, traits evolves slower towards its plateau, consequently, ontogeny affects more traits values. We will further use $DBH_{90}$ equal to $9.\beta_{DBH}$ being the DBH value for which the trait account for 90% of its maximum value that we can assume as the DBH of trait maturity.

Finally $\alpha_{env}$ is the trait maximum value in the Michaelis Menten and will be represented by an additive linear form of environmental effects represented by $TWI$ for abiotic environment and by $NCI$ for biotic interactions:
$$alpha_{env} = \alpha_{c,s} + {\beta_{TWI}}_{c,s}.TWI_i + {\beta_{Comp}}_{c,s}.NCI_i$$
where $\alpha_{c,s}$ is the intercept, $\beta_{TWI}$ is the slope of TWI effect, e.g. positive $\beta_{TWI}$ represents an increase of the trait value with water accumulation; whereas $\beta_{Comp}$ is the slope of competition effect, e.g. negative $\beta_{Comp}$ highlight a decrease of trait value with competition.

Species complex effect was integrated as a fixed effect on each parameter of the model and was later tested by Student T test between the two complex values for each parameter, a non significative value indicating no differentiation of the complex for the tested effect. Species effect was integrated as a nested random effect within each parameter and evaluated through species variance for each parameter ($\sigma_{Intercept}$, $\sigma_{DBH}$, $\sigma_{TWI}$, $\sigma_{Comp}$), a high variance indicating a strong diversity of species answers within complex for the tested effect.

Traits, DBH and TWI were all reduced for the model inference, in order to ease model inference and later compare strength of effects between traits and between effects. A Bayesian method was used to infer parameters of the model regarding each leaf traits using `stan` language [ref] and `rstan` package [ref] in the R environment [ref] (see supplementary material [S1: Model] for model equations and `stan` code).

# Results

The model inference correctly converged without divergent transition and resulted in a correct predictions density distribution (see supplementary material [S2: Diagnostic]). Goodness of fit showed globally strong differences between traits (Table \@ref(tab:summary), and supplementary material [S4: Predictions]). SLA, LT displayed a low likelihood in association with a strong residual variation ($\sigma^2 = 0.763$), when CC had an even lower likelihood and increased residual variation ($\sigma^2 = 0.948$). On the other hand, LDMC and LA displayed relatively high likelihood and low residual variation ($\sigma^2 < 0.6$).

$\alpha$ parameter represents the intercept and $\sigma_{Intercept}$ the variation of species intercepts within complex. They varied from very high and stable value of $6.872 \pm 0.418$ for $CC_{Symphonia}$ to really low and variable value $0.415 \pm 1.294$ for $\frac{1}{LA}_{Eschweilera}$. They globally highlighted smaller, lighter and thicker, leaves with decreased SLA but increased chlorophyll for *Symphonia* compared to *Eschweilera*. More interestingly, LDMC and CC varied two fold less from one species to the other within complex compared to SLA, LT and LA.

$DBH_{90}$, the DBH for which the trait reach 90% of its mature value, varied from 5.22 to 117.62 cm with increasing uncertainty with increased value; whereas it showed a variation from 0.90 to 11 cm between species within complex (Figure \@ref(fig:ontogeny)). Globally, we can consider the effect of ontogeny stable within complex (low $\sigma_{DBH}$), except for $\frac{1}{LA}$ were it stays relatively low. In addition, considering that tree are censused starting from 10 cm DBH, we can consider $DBH_{90}$ under or close to 10 to be negligible. Consequently, we can consider CC and LDMC has not affected by ontogeny for both complex; whereas LT and SLA are moderately affected by ontogeny (maturity between 15 and 20 cm for *Escweilera* and around 58 cm for *Symphonia*); and LA is strongly affected by ontogeny (maturity at 51 and 117 cm for *Symphonia* and *Escweilera* respectively, close to the maximum sampled diameter for each). Note that patterns are similar between complex for each trait but with *Eschweilera* reaching maturity first except for LA, certainly due to a smaller sampled maximum DBH for *Eschweilera*.

Globally environmental variables have weak values of slope ($\beta_{TWI}$ and $\beta_{Comp}$ between -0.16 and 0.29) with high uncertainty but with a high variance between species within complex ($\sigma_{TWI}$ and $\sigma_{Comp}$ between 0.1 and 1.3, see Figure \@ref(fig:environment) and supplementary material [S3: Species parameters]). In detail, *Symphonia* globally answered to water accumulation, except for LT, with bigger less dense leaves with less chlorophyll but increased SLA in wet areas; whereas competition only seemed to affect LT and LA with smaller and thicker leaves with decreasing competition. On the other hand, *Escwheilera* answered to water accumulation mainly for CC and a little for LT with thicker leaves with less chlorophyll in dryer areas; whereas competition only seemed to affect LT and LA with once again smaller and thicker leaves with decreasing competition. Note that for described parameters species variance was lower than the complex parameter absolute value, indicating that the trend have low probability to be reversed for one species within the complex. In addition, response to competition were similar between complexes when response to water were different and sometimes opposed.

```{r summary}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = c(pars, "lp__"), droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "trait") %>% 
  mutate(complex = gsub("([[:alpha:]])", "", term)) %>% 
  select(trait, term, complex, estimate, std.error, rhat) %>% 
  mutate(complex = gsub("([[:punct:]])", "", complex)) %>% 
  mutate(complex = recode_factor(complex, `2` = "Symphonia", 
                                 `1` = "Eschweilera")) %>% 
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(term = as.factor(gsub("c", "", term))) %>% 
  mutate(term = recode_factor(term, 
                              alpha = "$\\alpha$",
                              betaDBH = "$\\beta_{DBH}$", 
                              betaTWI = "$\\beta_{TWI}$", 
                              betaComp = "$\\beta_{Comp}$", 
                              alphaNCI = "$\\alpha_{NCI}$",
                              sigmaInterept = "$\\sigma^2_{Intercept}$",
                              sigmaDBH = "$\\sigma^2_{DBH}$", 
                              sigmaTWI = "$\\sigma^2_{TWI}$",
                              sigmaComp = "$\\sigma^2_{Comp}$",
                              sigma = "$\\sigma^2$",
                              lp = "log-likelihood")) %>% 
  mutate(trait = recode_factor(trait, 
                              invSLA = "$\\frac{1}{SLA}$",
                              LDMC = "$LDMC$", 
                              LT = "$LT$", 
                              invLA = "$\\frac{1}{LA}$",
                              CC = "$CC$")) %>% 
  reshape2::dcast(term + complex ~ trait, value.var = "estimate") %>% 
  rename(Parameter = term, Complex = complex) %>% 
  kable(caption = "Model parameters for each complex with reduced traits and descriptors.",
        escape = F, digits = 3, format = "pandoc") %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```

```{r ontogeny, fig.cap="Ontogeny effect represented by the $DBH_{90}$ and associated species variance. Grey vertical line indicates the census recruitment DBH (10 cm), whereas red and blue vertical lines indicate maximum sampled diameter for respectively Symhponia and Eschweilera."}
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = c("betaDBH",
                                        "sigmaDBH")))) %>% 
  bind_rows(.id = "trait") %>% 
  mutate(complex = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(complex = gsub("([[:punct:]])", "", complex)) %>% 
  mutate(complex = ifelse(complex == "", NA, complex)) %>% 
  mutate(complex = recode_factor(complex, `2` = "Symphonia",
                                 `1` = "Eschweilera")) %>% 
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>%
  select(-outer_width, -inner_width, -point_est) %>% 
  select(trait, parameter, complex, ll, l, m, h, hh) %>% 
  left_join(lapply(mdata, function(data) 
    data.frame(variable = names(data$sd), sd = data$sd)) %>% 
      bind_rows(.id = "trait") %>% 
      filter(variable == "DBH") %>% 
      reshape2::dcast(trait ~ variable) %>% 
      rename(sdDBH = DBH)) %>% 
  mutate_at(vars("ll", "l", "m", "h", "hh"), funs(.*sdDBH)) %>% 
  mutate_at(vars("ll", "l", "m", "h", "hh"), funs(ifelse(parameter == "betaDBH",
                                                         .*9, .))) %>% 
  mutate(parameter = recode_factor(parameter, betaDBH = "DBH[90]",
                                   sigmaDBH = "sigma[DBH]^2")) %>%
  ggplot(aes(x = trait, xend = trait, col = complex, fill = complex)) +
  geom_hline(yintercept = 10, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_wrap(~ parameter, labeller = label_parsed, scales = "free",
             nrow = 2) +
  xaxis_title(F) + ylab("Diameter at Breast Hieght (DBH, cm)") +
  scale_color_discrete("Species complex", 
                       labels = c("Symphonia", "Eschweilera", "Both")) +
  scale_fill_discrete("Species complex",
                      labels = c("Symphonia", "Eschweilera", "Both")) +
  scale_y_sqrt(breaks = scales::pretty_breaks(10)) +
  theme(legend.position="bottom")
```

```{r environment, fig.cap="Environment effects represented by TWI and NCI (for reduced variables and traits)."}
S_sqrt_trans <- function() scales::trans_new("S_sqrt",
                                             function(x) sign(x)*sqrt(abs(x)),
                                             function(x) x^2*sign(x))
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = c("betaTWI",
                                        "sigmaTWI",
                                        "betaComp",
                                        "sigmaComp")))) %>% 
  bind_rows(.id = "trait") %>% 
  mutate(complex = gsub("([[:alpha:]])", "", parameter)) %>% 
  mutate(complex = gsub("([[:punct:]])", "", complex)) %>% 
  mutate(complex = ifelse(complex == "", NA, complex)) %>% 
  mutate(complex = recode_factor(complex, `2` = "Symphonia",
                                 `1` = "Eschweilera")) %>% 
  mutate(parameter = gsub("([[:digit:]])", "", parameter)) %>% 
  mutate(parameter = gsub("([[:punct:]])", "", parameter)) %>% 
  mutate(variable = ifelse(grepl("TWI", parameter), "TWI", "NCI")) %>% 
  mutate(parameter = ifelse(grepl("beta", parameter), "beta", "sigma^2")) %>% 
  mutate(parameter = factor(parameter, levels = c("beta", "sigma^2"))) %>%
  mutate(variable = factor(variable, levels = c("TWI", "NCI"))) %>%
  ggplot(aes(x = trait, xend = trait, col = complex, fill = complex)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_grid(variable ~ parameter, labeller = label_parsed, scales = "free") +
  xaxis_title(F) +
  yaxis_title(F) +
  scale_color_discrete("Species complex",
                       labels = c("Symphonia", "Eschweilera", "Both")) +
  scale_fill_discrete("Species complex",
                      labels = c("Symphonia", "Eschweilera", "Both")) +
  scale_y_continuous(trans="S_sqrt") +
  theme(legend.position="bottom")
```

# Discussion

> Discussion only include some ideas for the moment and has not been finished. First we are waiting the results for logged SLA and LA. Second the discussion need to be done with PCAs and sensitivity analysis results together to have a global overview of individual phenotype response to ontogeny and environment.

*Things to be discussed:*

- Model have a strong residual variation including intra-individual variation, measurement errors, and stochastic. Still LDMC and LA have relatively low value of $\sigma^2 < 0.6$, indicating that the individual level still explain a large part of leaf functional traits.
- The intercept varied strongly between species within complex for SLA, LT and LA $\sigma_{Intercept}^2 \sim 1$ but was low for LDMC and CC.
- Besides different between complexes (maybe because they do not have the same maximum DBH value), ontogeny effect seemed conserved between species within complexes (not expected !). Ontogeny effect was negligible for CC and LDMC, intermediate for LT and SLA (DBH of maturity 15-58 cm), and strong for LA (DBH of maturity 51-117 cm close to maximum observed DBH for each).
- Effect of environment (abiotic and biotic) is weak but highly variable between species within complex.
- Competition similarly decreased LT and LA for the two complex (with non-different parameters for LT).
- Water accumulation answer was different and sometimes opposed for the two complexes, with bigger less dense leaves with less chlorophyll but increased SLA in wet areas for *Symphonia* against less thick leaves with more chlorophyll for *Eschweilera* (interestingly though CC pattern completely change in PCAs for the two complex which might be linked to this change of response to water between the two complexes).

## Conclusion

*Take home messages…*

-

# Supplementary materials

## S1: Model

We explained leaf trait $Trait$ variation with ontogeny ($DBH$), abiotic environment ($TWI$), biotic interactions ($NCI$) and taxonomic levels (species complex $C$ and species $S$) with a linear mixed model including interaction between ontogeny and environment (more likely than an additive form without interactions, see analysis [A07-Additive](A07-Additive.html)). We obtained the following model (see [Material and Methods] for more detail on the model form):

\begin{equation} 
\begin{split}
  T_{c,s,i} \sim \mathcal{N}(\frac{DBH_i}{{\beta_{DBH}}_{c,s} + DBH_i} \times (\alpha_{c,s} + {\beta_{TWI}}_{c,s}.TWI_i + {\beta_{Comp}}_{c,s}.\sum_{j=1}^{J} DBH_j^{2} . e^{-{\alpha_{NCI}}_{c,s}.\delta_{i,j}}), \sigma)    \\
  \Theta_{c,s} \sim \mathcal{N}^5(\Theta_c,\sigma_{species}) ~|~ \Theta =
   \begin{bmatrix}
           \alpha \\
           \beta_{DBH} \\
           \beta_{TWI} \\
           \beta_{Comp}
  \end{bmatrix},~\sigma_{species}
  \begin{bmatrix}
           \sigma_{Intercept} \\
           \sigma_{DBH} \\
           \sigma_{TWI} \\
           \sigma_{Comp}
  \end{bmatrix}
\end{split}
\end{equation} 

Resulting in the following `stan` code with selected priors:

```{stan stanModel, output.var="Model", echo=T, eval=F}
data {
  int<lower=1>  N ; // # of observations
  int<lower=1>  J ; // # of neighbors
  int<lower=1> S ; // # species
  int<lower=1> C ; // # complex
  vector[N] Trait ; // Trait
  vector[N] DBH ; // Diameter at Breast Height
  vector[N] TWI ; // Topographic Wetness Index
  vector[J] DBHj ; // DBH neighbor
  vector[J] Deltaj ; // Distance neighbor
  vector[N] weights ;
  int<lower=1, upper=N>  individual[J] ; // individuals index in neighbor matrix
  int<lower=1, upper=S> species[N] ; // species index
  int<lower=1, upper=C> speciesincomplex[S] ; // species index wihtin complex
  int<lower=1, upper=C> complex[N] ; // species index
}
parameters {
  vector [C] alpha ; // Intercept for complexes
  vector<lower=0> [C]  betaDBH ; // DBH half-load for complexes
  vector[C] betaTWI ; // TWI slope for complexes
  vector[C] betaComp ; // Competition slope for complexes
  real<lower=0> alphaNCI ;
  vector<lower=0> [S] alpha_s_tilde ; // Intercept for species
  vector<lower=0> [S] betaDBH_s_tilde ; // DBH half-load for species
  vector[S] betaTWI_s_tilde ; // TWI slope for species
  vector[S] betaComp_s_tilde ; // Competition slope for species
  real<lower=0> sigmaIntercept ; // Intercept variance for species
  real<lower=0> sigmaDBH ; // DBH half-load variance for species
  real<lower=0> sigmaTWI ; // TWI slope variance for species
  real<lower=0> sigmaComp ; // Competition slope variance for species
  real<lower=0> sigma ; // Variance
}
transformed parameters {
  vector[S] alpha_s ; // Species uncentered effects
  vector<lower=0> [S] betaDBH_s ;
  vector[S] betaTWI_s ;
  vector[S] betaComp_s ;
  vector[J] NCIj ; // Crowding index
  vector[N] NCI ;
  alpha_s = alpha[speciesincomplex] + sigmaIntercept*alpha_s_tilde ;  // Species uncentered effects
  betaDBH_s = betaDBH[speciesincomplex] + sigmaDBH*betaDBH_s_tilde ;
  betaTWI_s = betaTWI[speciesincomplex] + sigmaTWI*betaTWI_s_tilde ;
  betaComp_s = betaComp[speciesincomplex] + sigmaComp*betaComp_s_tilde ;
  NCIj = DBHj .* DBHj .* exp(-alphaNCI * Deltaj) ;  // Crowding index
  NCI = rep_vector(0.0, N) ;
  for(j in 1:J)
   NCI[individual[j]] += NCIj[j] ;
  NCI = (1 ./ weights) .* NCI ;
  NCI = NCI ./ sd(NCI) ;
}
model {
  alpha ~ normal(0, 10^6) ; // Priors
  betaDBH ~ lognormal(0,1) ;
  betaTWI ~ normal(0, 10^6) ;
  betaComp ~ normal(0, 10^6) ;
  alphaNCI ~ lognormal(0, 1) ;
  alpha_s_tilde ~ normal(0, 1) ;
  betaDBH_s_tilde ~ normal(0, 1) ;
  betaTWI_s_tilde ~ normal(0, 1) ;
  betaComp_s_tilde ~ normal(0, 1) ;
  sigmaIntercept ~ lognormal(0, 1) ;
  sigmaDBH ~ lognormal(0, 1) ;
  sigmaTWI ~ lognormal(0, 1) ;
  sigmaComp ~ lognormal(0, 1) ;
  sigma ~ cauchy(0, 5) ;
  Trait ~ normal((alpha_s[species] + betaTWI_s[species] .* TWI +  betaComp_s[species] .* NCI) .* (DBH ./ (betaDBH_s[species] + DBH)), sigma) ; // Likelihood
} 
generated quantities {
  vector[N] Trait_pred ; // Predictions
  vector[N] Trait_predDBH ;
  vector[N] Trait_predTWI ;
  vector[N] Trait_predNCI ;
  Trait_pred = (alpha[complex] + betaTWI[complex] .* TWI + betaComp[complex] .* NCI) .* (DBH ./ (betaDBH[complex] + DBH));
  Trait_predDBH = (alpha[complex] + betaTWI[complex] * mean(TWI) + betaComp[complex] * mean(NCI)) .* (DBH ./ (betaDBH[complex] + DBH)) ;
  Trait_predTWI = (alpha[complex] + + betaTWI[complex] .* TWI + betaComp[complex] * mean(NCI)) .* (mean(DBH) ./ (betaDBH[complex] + mean(DBH))) ;
  Trait_predNCI = (alpha[complex] + betaTWI[complex] * mean(TWI) + betaComp[complex] .* NCI) .* (mean(DBH) ./ (betaDBH[complex] + mean(DBH))) ;
}
```

## S2: Diagnostic

Model inference converged ($\hat R = 1$, figure \@ref(fig:mcmc)) for every traits with relatively little or no divergent transition (< 0.1%), and relatively uncorrelated parameters. Post predictive check (PPC) with density overlay of observed trait values against predicted trait values showed relatively similar distribution taking into account that the model form was generic for every traits (figure \@ref(fig:ppc)). We can note that the model reproduced well multimodality for LDMC and invLA but had difficulty to represent high trait values for invSLA, LT and CC.

```{r mcmc, fig.cap="Markov chains trace plot after warmup for leaf thickness model."}
# lapply(fits, check_divergences)
mcmc_trace(as.array(fits$LT, pars = c(pars, "lp__")),
             facet_args = list(labeller = label_parsed))
# mcmc_pairs(as.array(fits$invSLA), regex_pars = pars)
```

```{r ppc, fig.cap="Post predictive check (PPC) with density overlay of observed trait values against predicted trait."}
cowplot::plot_grid(plotlist = mapply(function(fit, data) {
  yrep <- as.matrix(fit, pars = "Trait_pred")
  ppc_dens_overlay(data$Trait, yrep[sample(nrow(yrep), 25), ])
}, fit = fits, data = mdata, SIMPLIFY = F), labels = names(fits), nrow = 3)
```

## S3: Species parameters

Supplementary material S3 includes species parameters for every traits and for reduced TWI and NCI, and highlights the strong variance of slope between species within complex for the two descriptors.

```{r speciesPosteriorBeta, fig.cap="Model species effect for TWI and NCI (for reduced variables and traits)."}
S_sqrt_trans <- function() scales::trans_new("S_sqrt",
                                             function(x) sign(x)*sqrt(abs(x)),
                                             function(x) x^2*sign(x))
lapply(fits, function(fit) 
  mcmc_intervals_data(as.array(fit,
                               pars = c("betaTWI_s", 
                                        "betaComp_s")))) %>% 
  bind_rows(.id = "trait") %>% 
  separate(parameter, c("parameter", "species"), sep = "_") %>% 
  mutate(species = gsub("([[:punct:]])", "", species)) %>% 
  mutate(species = gsub("([[:alpha:]])", "", species)) %>% 
  mutate(species = levels(as.factor(Individuals$SpeciesLong))[as.numeric(species)]) %>% 
  mutate(complex = ifelse(substr(species, 1,1) == "E", "Eschweilera", "Symhponia")) %>% 
  mutate(parameter = recode_factor(parameter, betaTWI = "beta[TWI]",
                                   betaComp = "beta[Comp]")) %>%
  ggplot(aes(x = species, 
             xend = species, 
             color = species, fill = species)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_grid(parameter ~ trait, labeller = label_parsed, scales = "free") +
  xaxis_title(F) +
  yaxis_title(F) +
  scale_y_continuous(trans="S_sqrt") +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")
```

## S4: Predictions

Supplementary material S4 includes model predictions for every traits and descriptor (each descriptor is showed with other predictors set to their mean), and highlights the difference of goodness of fit among leaf functional traits.

```{r prediction, fig.cap="Predictions for every traits and descriptors (each dewscriptor is showed with other predictors set to their mean).", fig.width=12, fig.height=12}
lapply(as.list(c("DBH", "TWI", "NCI")), function(variable)
  lapply(as.list(traits), function(trait) {
    if(variable == "NCI") {
      value <- apply(as.matrix(fits[[trait]], pars = "NCI"), 2, mean)
    } else {
      value <- mdata[[trait]][[variable]]
    }
    data.frame(
      trait = trait, 
      variable = variable,
      Y = mdata[[trait]]$Trait,
      value = value,
      Genus = mdata[[trait]]$complex,
      mu = apply(as.matrix(fits[[trait]], pars = paste0("Trait_pred", variable)), 
                 2, mean),
      mu5 = apply(as.matrix(fits[[trait]], pars = paste0("Trait_pred", variable)), 
                  2, quantile, probs = 0.05),
      mu95 = apply(as.matrix(fits[[trait]], pars = paste0("Trait_pred", variable)), 
                   2, quantile, probs = 0.95)
    )
  }) %>% bind_rows()) %>% 
  bind_rows() %>% 
  ggplot(aes(x = value, col = as.factor(Genus))) +
  geom_point(aes(y = Y)) +
  geom_ribbon(aes(ymin = mu5, ymax = mu95), alpha = 0.2) +
  geom_line(aes(y = mu)) +
  facet_grid(trait ~ variable, scales = "free") +
  scale_color_discrete("Complex", labels = c("Eschweilera", "Symphonia"))
```

# References
