---
title: "A10 : Full models comparison"
date: '`r Sys.Date()`'
author: Sylvain Schmitt & Anne Baranger
output:
  bookdown::html_document2:
    number_sections: no
    toc: true
    toc_float: yes
    theme: flatly
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
  bookdown::word_document2: default
linestretch: 1.5
---

```{r setup, include=FALSE}
#rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
path <- "../../data/Paracou/"
```

```{r data, eval=T}
# traitsSymphonia <- googlesheets::gs_title("Measures_Symphonia") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   mutate(SLA = as.numeric(SLA), LDMC = as.numeric(LDMC)) %>%
#   mutate(Genus = "Symphonia") %>%
#   rename(Species = Morphotype) %>%
#   mutate(Species = ifelse(Species == "Indet.",
#                           c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G",
#                                      "sp.1" =  "S")], Species))
# traitsEschweilera <- googlesheets::gs_title("Measures_Eschweilera") %>%
#   googlesheets::gs_read("AllTraits") %>%
#   filter(!(Plot == 14 & SubPlot == 1 & TreeFieldNum == 760)) %>%  # outlier
#   filter(!(Species %in% c("congestiflora","simiorum","persistens")))
# traits <- bind_rows(traitsEschweilera, traitsSymphonia) %>%
#   mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>%
#   group_by(idTree, Plot, SubPlot, TreeFieldNum, Genus, Species,
#            SpeciesLong, Bark, Dawkins) %>%
#   summarise_at(vars("SLA", "LDMC", "LT", "LA", "CC",
#                     "brBT", "brBD", "brWD"), mean, na.rm = T) %>%
#   ungroup() %>%
#   mutate(invSLA=1/SLA) %>%
#   mutate(invLA=1/LA)
# load("./functional_save/env.Rdata")
# Individuals <-  left_join(traits, env, by = "idTree", suffix = c("", ".y"))
# rm(traitsEschweilera, traitsSymphonia, traits, paracou, env)
# save(Individuals, file = "./functional_save/Individuals.Rdata")
load("./functional_save/Individuals.Rdata")
load("./functional_save/CompetitionMatrix.Rdata")
```


```{r mdata, fig.cap="prepared data",fig.width=12}
traits <- c("invSLA", "LDMC", "LT", "invLA", "CC")
mdata <- lapply(traits, function(trait){
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    left_join(select(Competition, idTree, AreaOutside20) 
              %>% unique())
  Competition_trait <- Competition
  Competition_trait$idTree <- match(Competition_trait$idTree, data_trait$idTree)
  Competition_trait <- filter(Competition_trait, !is.na(idTree))
  list(trait = trait,
       N = nrow(data_trait),
       J = nrow(Competition_trait),
       C = length(unique(data_trait$Genus)),
       S = length(unique(data_trait$Species)),
       Trait = as.numeric(scale(data_trait[trait], center = F)),
       DBH = as.numeric(scale(data_trait$DBH, center = F)),
       TWI = as.numeric(scale(data_trait$TWI, center = F)),
       DBHj = as.numeric(scale(Competition_trait$DBHj, center = F)),
       Deltaj = as.numeric(scale(Competition_trait$dij, center = F)),
       weights = 1-data_trait$AreaOutside20,
       individual = as.numeric(as.factor(Competition_trait$idTree)),
       species = as.numeric(as.factor(data_trait$Species)),
       complex = as.numeric(as.factor(data_trait$Genus)),
       speciesincomplex = unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[order(unique(cbind(as.numeric(as.factor(data_trait$Species)), as.numeric(as.factor(data_trait$Genus))))[,1]),2],
       data = data_trait,
       competition = Competition_trait)
})
names(mdata) <- traits
```

```{r }
load("./functional_save/Additive.Rdata")
fits_additive <- fits
load("./functional_save/Interaction.Rdata")
fits_interaction <- fits
fits <- list(additive = fits_additive,
             interaction = fits_interaction)
pars <- c("alpha_c", "betaDBH_c", "betaTWI_c", "betaComp_c", "alphaNCI",
          "sigmaIntercept", "sigmaDBH", "sigmaTWI", "sigmaComp", "sigma")
summary <- lapply(fits, function(fits_model)
  lapply(fits_model, function(fit)
  broom::tidyMCMC(fit, pars = c(pars, "Rsquared", "lp__"), 
                  droppars = NULL, rhat = T)) %>% 
    bind_rows(.id = "Trait")) %>% 
  bind_rows(.id = "Model")
```

# Introduction

The aim of this document is to compare the full models with additive and interaction forms. We will use previously built and sampled models (see documents A03-Models, A07-Additive and A08-Interaction). Model will be compared based on likelihood difference, parameters posterior values and uncertainty, and predictions qualtiy.

# Results

## Likelihood

The full model wiht interaction is more likely in every case except for $\frac1{LA}$ but with a low value and it is way more likely for $LT$ (see table \@ref(tab:loglikelihood)).

```{r loglikelihood}
summary %>% 
  filter(term == "lp__") %>% 
  reshape2::dcast(Trait ~ Model, value.var = "estimate") %>% 
  mutate(Trait = recode_factor(Trait, 
                              CC = "$CC$",
                              invLA = "$\\frac{1}{LA}$",
                              invSLA = "$\\frac{1}{SLA}$",
                              LDMC = "$LDMC$",
                              LT = "$LT$")) %>%
  mutate(difference = interaction - additive) %>% 
  kable(caption = "Log-likelihood per Trait and Model form.",
        escape = F) %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T)
```

## Parameters

The two models parameters are almost always equal. For few of the $\beta$ the full model with interaction showed a lower value but with reduced uncertainty and an increased corresponding $\sigma_species$. 

```{r posteriorsBeta, fig.cap="$\\alpha$ and $\\beta$ posteriors between the two forms of full model,i.e. additive and interaction.", fig.width=12, fig.height=12}
lapply(fits, function(fits_model)
  lapply(fits_model, function(fit)
    mcmc_intervals_data(as.array(fit),
                        regex_pars = c("alpha_c", "betaDBH_c", "betaTWI_c"))) %>% 
    bind_rows(.id = "Trait")) %>% 
  bind_rows(.id = "Model") %>% 
  mutate(parameter = recode_factor(parameter, 
                                   `alpha_c[1]` = "alpha[Eschweilera]",
                                   `alpha_c[2]` = "alpha[Symphonia]",
                                   `betaDBH_c[1]` = "beta[DBH-Eschweilera]", 
                                   `betaDBH_c[2]` = "beta[DBH-Symphonia]", 
                                   `betaTWI_c[1]` = "beta[TWI-Eschweilera]", 
                                   `betaTWI_c[2]` = "beta[TWI-Symphonia]", 
                                   `betaComp_c[1]` = "beta[Comp-Eschweilera]", 
                                   `betaComp_c[2]` = "beta[Comp-Symphonia]")) %>% 
  ggplot(aes(x = Trait, xend = Trait, col = Model, fill = Model)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_wrap(~ parameter, labeller = label_parsed, scales = "free") +
  xaxis_title(F) +
  yaxis_title(F)
```

## Predictions

```{r ppcgraphs}
g <- lapply(fits, function(fits_model)
  mapply(function(fit, data) 
    ppc_dens_overlay(data$Trait, as.matrix(fit, pars = "Trait_pred")), 
         fit = fits_model, data = mdata, SIMPLIFY = F))
```

```{r ppcgraphsAdditive, fig.cap="Post predictive checks for additive model.", fig.width=12, fig.height=6}
cowplot::plot_grid(plotlist = g$additive, labels = names(g$additive))
```

```{r ppcgraphsInteraction, fig.cap="Post predictive checks for interaction model.", fig.width=12, fig.height=6}
cowplot::plot_grid(plotlist = g$interaction, labels = names(g$interaction))
```

```{r posteriorsSigma, fig.cap="$\\sigma$ and $\\sigma_{species}$ posteriors between the two forms of full model,i.e. additive and interaction.", fig.width=12, fig.height=12}
lapply(fits, function(fits_model)
  lapply(fits_model, function(fit)
    mcmc_intervals_data(as.array(fit),
                        regex_pars = c("sigma"))) %>% 
    bind_rows(.id = "Trait")) %>% 
  bind_rows(.id = "Model") %>% 
  mutate(parameter = recode_factor(parameter, 
                                   sigma = "sigma^2",
                                   sigmaIntercept = "sigma[Intercept]^2",
                                   sigmaDBH = "sigma[DBH]^2", 
                                   sigmaTWI = "sigma[TWI]^2",
                                   sigmaComp = "sigma[Comp]^2")) %>% 
  ggplot(aes(x = Trait, xend = Trait, col = Model, fill = Model)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  facet_wrap(~ parameter, labeller = label_parsed, scales = "free") +
  xaxis_title(F) +
  yaxis_title(F)
```

```{r , fig.cap="Predictions between the two forms of full model,i.e. additive and interaction.", fig.width=12, fig.height=24}
lapply(fits, function(fits_model)
  lapply(as.list(c("DBH", "TWI", "NCI")), function(variable)
    lapply(as.list(traits), function(trait) {
      if(variable == "NCI") {
        value <- apply(as.matrix(fits_model[[trait]], pars = "NCI"), 2, mean)
      } else {
        value <- mdata[[trait]][[variable]]
      }
      data.frame(
        trait = trait, 
        variable = variable,
        Y = mdata[[trait]]$Trait,
        value = value,
        Genus = mdata[[trait]]$complex,
        mu = apply(as.matrix(fits_model[[trait]], pars = paste0("Trait_pred", variable)), 
                   2, mean),
        mu5 = apply(as.matrix(fits_model[[trait]], pars = paste0("Trait_pred", variable)), 
                    2, quantile, probs = 0.05),
        mu95 = apply(as.matrix(fits_model[[trait]], pars = paste0("Trait_pred", variable)), 
                     2, quantile, probs = 0.95)
      )
    }) %>% bind_rows()) %>% 
    bind_rows()) %>% 
  bind_rows(.id = "Model") %>% 
  ggplot(aes(x = value, col = as.factor(Genus))) +
  geom_point(aes(y = Y)) +
  geom_ribbon(aes(ymin = mu5, ymax = mu95), alpha = 0.2) +
  geom_line(aes(y = mu)) +
  facet_grid(trait + Model ~ variable, scales = "free") +
  scale_color_discrete("Complex", labels = c("Eschweilera", "Symphonia"))
```

# Conclusion
