---
title: Title
author: Sylvain Schmitt$^{*,1}$, Bruno Herault$^{2,3}$, Emilie Ducouret$^4$, Anne Baranger$^5$, Niklas Tysklind$^6$,  Myriam Heuertz$^5$, Eric Marcon$^7$, Saint Omer Cazal $^6$, Geraldine Derroire$^{4}$ 
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(ggfortify)
library(broom)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
```

```{r data}
load("./functional_save/Individuals.Rdata")
Individuals0 <- Individuals
load("./functional_save/CompetitionMatrix.Rdata")
sdDBH <- sd(Individuals$DBH)
Individuals <- Individuals %>% 
  rename(LMA = invSLA) %>% 
  group_by(Species) %>% 
  filter(n() > 20) %>% 
  ungroup() %>% 
  left_join(Competition %>% 
              group_by(idTree) %>% 
              summarise(NCI = (1/mean(1-AreaOutside20))*sum(DBHj^2*exp(-0.25*dij)))) %>%
  mutate(NCI = log(NCI)) %>% 
  mutate_at(vars("LMA", "LDMC", "LT", "invLA", "CC", "DBH", "TWI", "NCI"),
            funs(./sd(., na.rm = T)))
```

```{r model}
load("./functional_save/InteractionReduced.Rdata")
traits <- c("LMA", "LDMC", "LT", "invLA", "CC")
complexes <- c("Symphonia", "Eschweilera")
models <- sapply(complexes, function(complex) paste0(complex, "-", traits))
species <- lapply(models, function(model){
  trait <- strsplit(model, "-")[[1]][2]
  complex <- strsplit(model, "-")[[1]][1]
  Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    filter(Genus == complex) %>% 
    mutate(species = as.numeric(as.factor(Species))) %>% 
    mutate(SpeciesLong = paste(substr(complex, 1, 1), Species)) %>% 
    select(Genus, Species, SpeciesLong, species) %>% 
    unique()
}) %>% bind_rows() %>% 
  unique() %>% 
  mutate(complex = Genus)
S_sqrt_trans <- function() scales::trans_new("S_sqrt",
                                             function(x) sign(x)*sqrt(abs(x)),
                                             function(x) x^2*sign(x))
```

```{r predicitons}
traits <- c("LMA", "LDMC", "LT", "invLA", "CC")
complexes <- c("Symphonia", "Eschweilera")
models <- sapply(complexes, function(complex) paste0(complex, "-", traits))
preds <- lapply(models, function(model){
  trait <- strsplit(model, "-")[[1]][2]
  complex <- strsplit(model, "-")[[1]][1]
  data_trait <- Individuals[!is.na(unlist(Individuals[,trait])),] %>% 
    filter(Genus == complex)
  
  DBH <- data_trait$DBH
  TWI <- data_trait$TWI
  NCI <- data_trait$NCI
  species <- as.numeric(as.factor(data_trait$Species))
  
  alpha <- as.matrix(fits[[model]], pars = "alpha")
  betaDBH <- as.matrix(fits[[model]], pars = "betaDBH")
  betaTWI <- as.matrix(fits[[model]], pars = "betaTWI")
  betaNCI <- as.matrix(fits[[model]], pars = "betaNCI")
  
  Yp <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * TWI +  betaNCI[n,species] * NCI) * (DBH / (betaDBH[n,species] + DBH)), simplify = F))
  Yalpha <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * mean(TWI) +  betaNCI[n,species] * mean(NCI)) * (mean(DBH) / (betaDBH[n,species] + mean(DBH))), simplify = F))
  Ydbh <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * mean(TWI) +  betaNCI[n,species] * mean(NCI)) * (DBH / (betaDBH[n,species] + DBH)), simplify = F))
  Ytwi <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * TWI +  betaNCI[n,species] * mean(NCI)) * (mean(DBH) / (betaDBH[n,species] + mean(DBH))), simplify = F))
  Ynci <- do.call("rbind", sapply(1: nrow(alpha), function(n) (alpha[n,species] + betaTWI[n,species] * mean(TWI) +  betaNCI[n,species] * NCI) * (mean(DBH) / (betaDBH[n,species] + mean(DBH))), simplify = F))
  
  data.frame(
    Model = model,
    Trait = trait,
    Complex = complex,
    DBH = data_trait$DBH, 
    TWI = data_trait$TWI, 
    NCI = NCI,
    Genus = data_trait$Genus,
    Species = data_trait$Species,
    Y = unlist(data_trait[trait]),
    Yp = apply(Yp, 2, mean),
    Yp5 = apply(Yp, 2, quantile, probs = 0.05),
    Yp95 = apply(Yp, 2, quantile, probs = 0.95),
    Yalpha = apply(Yalpha, 2, mean),
    Yalpha5 = apply(Yalpha, 2, quantile, probs = 0.05),
    Yalpha95 = apply(Yalpha, 2, quantile, probs = 0.95),
    Ydbh = apply(Ydbh, 2, mean),
    Ydbh5 = apply(Ydbh, 2, quantile, probs = 0.05),
    Ydbh95 = apply(Ydbh, 2, quantile, probs = 0.95),
    Ytwi = apply(Ytwi, 2, mean),
    Ytwi5 = apply(Ytwi, 2, quantile, probs = 0.05),
    Ytwi95 = apply(Ytwi, 2, quantile, probs = 0.95),
    Ynci = apply(Ynci, 2, mean),
    Ynci5 = apply(Ynci, 2, quantile, probs = 0.05),
    Ynci95 = apply(Ynci, 2, quantile, probs = 0.95)
  )}) %>% bind_rows()
```

$^1$ UMR BIOGECO, Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France

$^2$ Cirad, Univ Montpellier, UR Forests & Societies, Montpellier, France

$^3$ INP-HB, Institut National Polytechnique Félix Houphouët-Boigny, BP 1093, Yamoussoukro, Ivory Coast

$^4$ Cirad, UMR EcoFoG (Agroparistech, CNRS, Inra, Université des
Antilles, Université de la Guyane), Campus Agronomique, 97310 Kourou, French Guiana

$^5$ UMR BIOGECO, INRA, 69 route d’Arcachon, F-33612 Cestas cedex France

$^6$ INRA, UMR EcoFoG (Agroparistech, CNRS, Université des
Antilles, Université de la Guyane, Cirad), Campus Agronomique, 97310 Kourou, French Guiana

$^7$ Agroparistech, UMR EcoFoG (CNRS, Inra, Université des
Antilles, Université de la Guyane, Cirad), Campus Agronomique, 97310 Kourou, French Guiana

__$^*$ Corresponding author:__ sylvain.schmitt@agroparistech.fr

\newpage 
\newpage 

 # Abstract

> To be written

## Keywords

Intraspecific trait variation; species complexes; ...

\newpage 

# Introduction 

* Community trait-based study increase but species coexistence remains unclear
* What is a functional trait ?
    * Functional trait definition [@violle_let_2007]
    * Functional trait & ecology [@Wright2002]
    * Functional trait example - LMA [@Evans2001; @hodgson_is_2011]
* How do functional traits covary ?
    * Known level of covariation (community [@Bruelheide2018] and species [@diaz_global_2015])
    * Leaf Economics Spectrum [@wright_worldwide_2004; but see @lloyd_photosynthetically_2013; @osnas_global_2013]
    * Wood Economics Spectrum [@chave_towards_2009]
    * Plant Economics Spectrum [@reich_world-wide_2014; @freschet_evidence_2010]
* How do species coexist ?
    * Community > Species > Individual diversity
    * Community definition
    * Species definition [@Mayr1996]
    * Species coexistence
        * Niche [@Lortie2004a; @weiher_assembly_1995]
        * Neutral [@Hubbell2001]
        * Review [@wright2002plant]
* What are the limit of the species concept ?
    * debated concept [@Mayden1997]
    * numerous definitions [ @DeQueiroz2007a]
    * species complex & species coalescence [@Fujita2012] ? 
* How individuals vary within species ?
    * Individual vary in performance, phenotype and genotype
    * Represented by phenotypic variation / ITV in ecosystem
    * Factors : genetic, environment, stochastic [@Whitlock2007]
* Is individual variability important ?
    * relatively unexplored [@Hallgrimsson2005]
    * in tree allometric relations [@Vieilledent2010]
    * in leaf traits [@Hulshof2010a; @Messier2010]
    * within community itself [@Siefert2015a]
* Does individual variability play a role in community ?
    * environmental filtering [@Messier2010; @Paine2011]
    * raits shift along environmental and resources gradient [@Siefert2016; @jung_intraspecific_2010 ]
    * niche differentiation [@jung_intraspecific_2010; @Paine2011]
* Why individual variability could matter in community ?
    * ITV emphasized importance [@Albert2010a; @chave_neutral_2004; @Violle2012]
    * pass environmental filtering [@Clark2010]
    * exceeding negative interspecific interactions [@Chesson2000]
    * increase to answer to neighborhood [@Aarssen1983]
    * answer species coexistence paradox [@Laughlin2012]
* What did we do to explore individual variability origin ?
    * role of ontogeny, abiotic environment and tree neighborhood in shaping leaf functional trait at the individual level
    * 5 species and 2 genera from two Neotropical species complexes
    * 766 trees > 10 cm Paracou
    * Topography, water, nutriment, & light heterogeneity [@Allie2015; @ferry2010higher]
    * Census data + LiDAR data + Functional traits 
* Questions
    * Trait covariation between individuals of closely related tropical tree species
    * Ontogeny and environment impact on ITV of closely related tropical tree species
* Hypothesis
    * Covariation conserved in community & species so waited at between individuals [@Bruelheide2018; @diaz_global_2015; but see @messier_trait_2016]
    * Ontogeny & light strong role on individual leaf trait in interaction with abiotic and biotic environment

# Discussion

## Leaf traits covary from resources acquisition to conservation 

* PCA1 most of co-variations with LA, LT, LDMC, CC and LMA
* LMA, LDMC, and LT covariation has been linked to a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability
    * LDMC, 1/SLA (LMA), LTD (~LDMC) covariation in LES [@Baraloto2010]
    * SLA (1/LMA) and photosynthesis acquisition [@Evans2001]
    * 1/SLA (LMA) and LDMC covaraition with soil fertility [@hodgson_is_2011]
    * LDMC, SLA (1/LMA) and LT covariation [@Vile2005]
* LA covary with LDMC decoupled with LMA (opposed hypothesis)
    * LA participation to LES debated [@Ackerly2002]
    * Unclear role [@Nicotra2011]
    * LA negative ratio and leaf mass [@Yang2010] 
    * @Yang2010 hypothesized LA increase as an investment in photosynthesis or durability
    * the two  hyptohesis represent both side of the LES trade-off 
    * Does it explained contrasted co-variation of LA with LDMC and LMA ?
* CC oriented toward conservation co-varying with LMA and LT ?
    * already highlighted but unexplained in @Baraloto2010
    * why ?
* Sampled leaf traits seems to highlight a major covariation around the LES
* LES could be conserved at the individual level (too few traits to be affirmative)

## Ontogeny and light shape individual leaf traits

* DBH represent confounded effect of ontogeny and light
    * (SupplementaryMaterial ...)
    * light and ontogeny difficulty disentangled [@Lusk2007; @roggy_links_2005] but see [@Poorter2012]
    * light and ontogeny showed little to no interaction [@roggy_links_2005; @Sciences2009]
    * greater effect of ontogeny [@Rijkers2000; @Rozendaal2006b]
    * reduced effect of light on functional trait values [@Dang-Le2013; @Rozendaal2006b]
* DBH increased 1/LA, LT, and LMA during the whole tree radial growth with a decreasing intensity with increasing DBH toward a plateau for DBHmax
    * LMA, LT, LDMC (or similarly LTD) increase with ontogeny and light [@Lusk2007; @Spasojevic2014; @Niinemets2006; @Rozendaal2006b; @roggy_links_2005; @Sciences2009; @Dang-Le2013; @Rijkers2000]
    * LA decrease with ontogeny and light [@Dang-Le2013; @Rozendaal2006b]
* DBH had little or no effect on LDMC and CC
    * little or no effect of light and ontogeny on LDMC [@roggy_links_2005] 
    * no effect of light and ontogeny on CC [@Rijkers2000; @Rozendaal2006b]
* Ontogeny and light result in a shift from acquisitive to conservative functional strategy
    * following PCA result
* Height, light, herbivory and mechanical damage can explain conservative strategy with increasing ontogeny and light
    * Height effect on water potential due to gravity [@Koch2004; @WOODRUFF2007]
    * Light effect with thicker mesophyll with elongated and more layers of palisade parenchyma [@Chazdon1993]
    * Increased herbivory and mechanical damage in canopy [@Sterck1992]
* Shadow and development need can explain acquisitive strategy with decreasing ontogeny and light
    * Light effect with increase light capture need in shadower understory [@Chazdon1996]
    * Ontogeny effect with increased developement need for radial growth [@Herault2011]
* Ontogeny and light access cannot be dissociated *in natura* and their effect shape leaf traits all along tree growth/life
* Consequently shade and light leave concept should not be dissociated from ontogeny and ontogeny should not be discretized in large cohorts of age or diameter
    * @Wright2002a and @Coleman1994 advocate for a comprehensive assessment of ontogeny
    * Questionning the concept of sun-shade leaves [@Lusk2007; @Dang-Le2013]
    * Study showing light effect without controlling ontogeny [@Rozendaal2006b] ?
* Finally, we evidenced little effect of other environmental factors, advocating for the primordial role of light and ontogeny in individual *in natura* variation
    * answering @Dang-Le2013
    * (Fig. \@ref(fig:sensitivity))

## Leaf and individual sampling implications in trait-based studies

* Ontogeny and light have been showed to strongly shape individual variability within species
* Species differences within and among genus once this variation taken into account is low
    * (posteriors intercepts)
* Thus individual trait variation compared to species is strong
    * (supplementary material ..., Fig. \@ref(fig:varcomp)))
* We thus need of an increased sampling of individual controlling for light, habitat and ontogeny
    * (Fig. \@ref(fig:bootstrap))
* But regressions showed a strong residual variation linked to intra-individual variability
    * strong residual variation ($\sigma ^2 \in [0.5;1]$, Table \@ref(tab:parameters))
    * probably not measurement errors (low)
    * probably not climatic variation (controlled, but see Supplementary material..)
    * probably not intra-leaf variation (controlled with whole limb or repetitions)
    * probably intra-individual variation (supplementary material ..., Fig. \@ref(fig:varcomp)))
*  Moreover, functional traits intra-individual variation has already been highlighted in literature
    * leaf thickness, transfusion tissue, and mass per area increase with leaf height within the tree [@Koch2004; @Oldham2010]
    * leaves vein density increase and stomatal pore index and hydraulic conductance decrease with decreasing shoot length within _Ginkgo biloba_ individuals [@Leigh2011]
* We thus also need of an accurate sampling of individual leaf
    * control architectural development stage of branch
    * control branch position

## Intraspecific trait variability increase species tolerance to environmental filtering

* TWI represent low topographic position and water accumulation
* NCI represent neighbor competition and thus nutriment and light access
* Thus, TWI and NCI represent habitat differences with wetter and lighter habitats more dynamic and fertile from bottom lands to plateaus
    *  bottom lands with higher treefall, recruitment and growth rates inducing a strong light gradient combined with a higher fertility [@ferry2010higher]
    * topography correlated to nutrient [@Allie2015]
* TWI decreased 1/LA, LDMC, and LMA
* NCI increased 1/LA for *Symphonia*
* TWI had little effect on LT
* TWI had opposite effect on CC between complexes **??**
* Higher topographic position result in a shift from acquisitive to conservative strategy
* Tolerance to drier habitat can explain conservative strategy with increasing TWI **?**
    * tissues with smaller leaves allow more drought tolerance of individuals [@Mendez-Toribio2017]
* Increased dynamic and fertility from habitats can explain acquisitive strategy (both 1/LA, LDMC, and LMA with TWI and 1/LA increased with NCI)
* Decreased 1/LA, LDMC, and LMA from plateau to bottom lands have also been observed at the species level    
     * @kraft_functional_2008 decrease 1/LA and LMA in valley bottoms compared to ridge-tops at the species level
     * @Gotsch2010 found a decrease of SLA with decrease of water availability
     * @Cunningham2007  found a decrease of LMA towards both wetter sites with increased phosphorous
* Response to abiotic environment and biotic interaction is conserved from species to individuals broadening the niche of species
     * _S. globulifera_ and _E. coriacea_ have respectively decreased 1/LA, LDMC, and LMA than _E. sagotiana_ and _S. sp1_ (Posterior Intercepts)
     * _S. globulifera_ and _E. coriacea_ preferentially grow in flooded valley bottom than drier slopes and hilltops where grow _E. sagotiana_ and _S. sp1_ [@Allie2015; Schmitt et al. unpublished]
* ITV could increase species tolerance to environmental filtering [@Clark2010] and thus contribute 
