---
title: Installation instruction for "Introduction to bayesian modelling with WinBUGS"
author: "Sylvain Schmitt"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    theme: readable
    toc: true
    toc_float: yes
    number_sections: no
link-citations: yes
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = T, eval = F)
```

The aim of Wednesday lecture is to get familiar to Bayesian modelling and `WinBUGS` use through R with packages `R2WinBUGS` and `coda`. Consequently, you'll need to install everything and test installation before the presentation to try examples directly on your machine. Tools to be installed are listed below and detailed thereafter:

* __Mandatory tools__
    * `WinBUGS` _(software)_
    * `R2WinBUGS` _(R package)_
    * `coda` _(R package)_
    * `tydiverse` _(R package)_
* __Optional but recommended tools__
    * `rmarkdown` _(R package)_
    * `knitr` _(R package)_
* __Highly optional tools__
    * `reveal_js` _(R package)_

# Mandatory tools

## `WinBUGS`

`WinBUGS` will be the main tool used in Jean Jacques Boreux lecture to do Bayesian modelling, please install it following [MRC Biostatistics Unit instructions](https://www.mrc-bsu.cam.ac.uk/software/bugs/the-bugs-project-winbugs/).

## R packages associated to `WinBUGS` 

`R2WinBUGS` and `coda` are the two packages to run `WinBUGS` and analyse its output from R. Install them with the following command:

```{r}
install.packages(c("R2WinBUGS", "coda"))
```

We'll see during the lecture how to make the junction between `WinBUGS` and `R2WinBUGS`. Warning, Mac and Linux user have a specific way to make the junction because WinBUGS as been initially developed for Windows. 

## `tidyverse`

`tidyverse` is the only package used during the lecture to run code examples within R. Please install it with the following command to try and play with examples on your own machine:

```{r}
install.packages(c("tidyverse"))
```

# Optional tools

The presentation is an Rmarkdown document compiled into a presentation with `reveal_js` R package. To modify and compile the document on your own machine I recommend the installation of `rmarkdown` and `knitr`. You will be able to compile it as a pdf or an html. If you want specifically the presentation output, you will need to additionally install `reveal_js`.

```{r}
install.packages(c("rmarkdown", "knitr", "reveal_js"))
```

# Conclusion

If you have any trouble with the installation, don't hesitate to ask me.
