```{r setup_libraries, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(parallel)
library(tidyverse)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Symphonia_Paracou/"
```

# Libraries preparation protocol

The protocol is given per sample with the corresponding volume for a plate of 96 samples in bracket.

## Material preparation

* **0.5 $\mu L$ of 10mM TrisHCL + 10 mM NaCl** (48 $\mu L$ per plate)
* **46.875 $\mu L$ of 0.1X TE buffer** (2 136.875 $\mu L$ per plate) = 4.7 $\mu L$ of TE buffer (213.7 $\mu L$ per plate) + 42.3  $\mu L$ of water (1 923.3 $\mu L$ per plate)
* **400 $\mu L$ of fresh 80% Ethanol** (38 400  $\mu L$ per plate)

## Fragmentation

1. Prepare **6.5 $\mu L$ of samples** with *ca* **100 $ng$ of DNA** (see previous chapter)
2. **On ice**, pipette up and down **ULTRA II FS Reaction Buffer** 10X, vortex 5" and spin
3. **On ice**, vortex 5" **Ultra II FS Enzyme Mix** and spin
4. **On ice**, premix **0.5  $\mu L$ of Ultra II FS Enzyme Mix** (48 $\mu L$ per plate) with **1.75 $\mu L$ of Ultra II FS Reaction Buffer** (168 $\mu L$ per plate), vortex and spin
5. **On ice**, add **2.25  $\mu L$ of premix** to each sample, vortex 5" and spin:

```{r}
data.frame(
  Component = c("DNA", "Ultra II FS Enzyme Mix", "Ultra II FS Reaction Buffer", "Total"),
  V = c("6.5 $\\mu L$", "0.5 $\\mu L$", "1.75 $\\mu L$", "8.75 $\\mu L$"),
  Vplate = c("", "48 $\\mu L$", "168 $\\mu L$", "")
) %>% 
kable(col.names = c("Component", "Volume per library", "Volume per plate"))
```

6. Thermocycle with following programs depending on electrophoresis quality:
    * Good: 13'@37°C, 30'@65°C, Hold@-4°C
    * Medium: 9'@37°C, 30'@65°C, Hold@-4°C
    * Bad: 1'@37°C, 30'@65°C, Hold@-4°C
7. *Optionally, put the **NEBNext adaptor for Illumina** out of the freezer (long to melt)*
8. *Optionally, samples can be stored overnight at $-20^\circ$*

## Adaptor ligation

1. **On ice**, prepare diluted adaptor (1:5) with **0.125 $\mu L$ of NEBNext adaptor for Illumina** (12 $\mu L$ per plate) diluted into **0.5 $\mu L$ of 10mM TrisHCL + 10 mM NaCl** (48 $\mu L$ per plate)
2. **On ice**, premix **7.5  $\mu L$ of NEBNext Ultra II Ligation Master Mix** (720 $\mu L$ per plate) with **0.25 $\mu L$ of NEBNext ligation enhancer** (24 $\mu L$ per plate), vortex and spin
3. **On ice**, add **0.625 $\mu L$ of diluted adaptor** and **7.75 $\mu L$ of premix** to samples, mix and spin:

```{r}
data.frame(
  Component = c("DNA", "NEBNext Ultra II Ligation Master Mix", 
                "NEBNext ligation enhancer", "diluted NEBNext adaptor (1:5)", "Total"),
  V = c("8.75 $\\mu L$", "7.5 $\\mu L$", "0.25 $\\mu L$", "0.625 $\\mu L$", "17.25 $\\mu L$"),
  Vplate = c("", "720 $\\mu L$", "24 $\\mu L$", "50 $\\mu L$", "")
) %>% 
kable(col.names = c("Component", "Volume per library", "Volume per plate"))
```

4. Incubate 15'@20°C with **lid open**
5. **On ice**, add **0.75 $\mu L$ of USER Enzyme** (72 $\mu L$ per plate) to samples, mix and spin
6. Incubate 15'@37°C with lid hot (>47°C)
7. *Optionally, samples can be stored overnight at $-20^\circ$*

## Size selection

1. Bring the sample volume from **17.125 $\mu L$** to **27.125 $\mu L$** adding **10 $\mu L$ of 0.1X TE buffer** (1 716 $\mu L$ per plate)
1. Vortex **PGTB beads from batch E** at room temperature
1. Add **7 $\mu L$ (~0.28X) of beads** (672 $\mu L$ per plate) to each sample, mix, vortex 5" keeping beads, incubet 5', spin, place on magnet, wait 5', and transfer ~32 $\mu L$ of sample to a new plate
1. Add **3.5 $\mu L$ (~0.14X) of beads** (336 $\mu L$ per plate) to each sample, mix, wait 5' without the magnet
1. Place on magnet, wait 5" and discard ~35.5 $\mu L$ of supernatant
1. Add **100 $\mu L$ of fresh 80% Ethanol** (9 600 $\mu L$ per plate) to the beads on the magnet, wait 30' remove supernatant
1. Repeat, add **100 $\mu L$ of fresh 80% Ethanol** (9 600 $\mu L$ per plate) to the beads on the magnet, wait 30" remove supernatant
1. Air dry beads 3' on magnet
1. Remove magnet, elute into **12 $\mu L$ of hot 0.1X TE** (1 152 $\mu L$ per plate) (~40°C), mix, incubate 2', spin, put on magnet, wait 5'
1. Transfer 2 x **5 $\mu L$ of supernatant** to 2 new plates
1. *Optionally, samples can be stored overnight at $-20^\circ$*

## Enrichment and purification

*protocol given for delivered oligos at $100mM$, not NEBNext tag at $10mM$*

1. Prepare diluted index (1:10) with **0.16 $\mu L$ of Index Primer i5 and i7** diluted in **1.44 $\mu L$ of mQ $H_2O$** (1.92 $\mu L$ of i5 per row and 1.28 $\mu L$ of i7 per column)
2. Mix in each plate (2 for the 2 PCR), mix and spin :

```{r}
data.frame(
  Component = c("sample", "NEBNext Ultra II Q5 Master Mix", 
                "diluted Index Primer i5 (1:10)", "diluted Index Primer i7 (1:10)", "Total"),
  V = c("5 $\\mu L$", "8.3 $\\mu L$", "1.6 $\\mu L$", "1.6 $\\mu L$", "16.5 $\\mu L$"),
  Vplate = c("", "796.8 $\\mu L$", "19.2 $\\mu L$", "12.8 $\\mu L$", "")
) %>% 
kable(col.names = c("Component", "Volume per library", "Volume per plate/row/column"))
```

3. Thermocycle with following program
    * 30"@98°C
    * 8 cycles of 10"@98°C and 75"@65°C
    * 5'@65°C
    * Hold@4°C
4. *Optionally, amplify only the first plate, assess it with electrophoresis, and adjust cycles number for the second amplifcation depending on gel migration*
5. Pool PCR results (~16.5 $\mu L$ per sample) from the 2 plates into one (~33.3K $\mu L$ per sample)
6. Vortex **PGTB beads from batch E** at room temperature
7. Add **30 $\mu L$ (~0.9X) of beads** (2 880 $\mu L$ per plate), mix, vortex 5" keeping beads, spin, place on magnet, wait 5', and remove supernatant (~ 63 $\mu L$ per sample)
8. Add **100 $\mu L$ of fresh 80% Ethanol** (9 600 $\mu L$ per plate), wait 30' remove supernatant
9. Repeat, add **100 $\mu L$ of fresh 80% Ethanol** (9 600 $\mu L$ per plate), wait 30" remove supernatant
10. Air dry beads 3' on magnet
11. Remove magnet, elute into **22 $\mu L$ of hot 0.1X TE** (2 112 $\mu L$ per plate) (~40°C), mix, incubate 2', and spin
12. Place on magnet, wait 5', transfer **22 $\mu L$ of supernatant** to a new plate and store at $-20^\circ$
