```{r setup_lib_prep_res, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(parallel)
library(tidyverse)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Symphonia_Paracou/"
```

# Libraries preparation results

## Post-enrichment PCR1 quantification

After the enrichment and the purification of the first PCR (PCR1), we quantified double strand DNA in every plates in order to adjust the second PCR (PCR2), and more especially in order to increase the number of cycles in PCR2. We used both Quant-It and few samples on Qbit and transformed raw absorbance results into concentration with regressions.

```{r quantit}
quantit <- read_csv(file.path(path, "QuantIt", 
                   "SYMCAPTURE_Library_enriched_PCR1_Plate1_Plate2_QuantIt.csv"), skip = 53)[2:4] %>% mutate(QuantIt = 1) %>% 
  bind_rows(read_csv(file.path(path, "QuantIt", 
                   "SYMCAPTURE_Library_enriched_PCR1_Plate3_Plate4_Plate5_QuantIt.csv"), skip = 53)[2:4] %>% mutate(QuantIt = 2)) %>% 
  select(-`Well ID`) %>% 
  dplyr::rename(position = Well, fluorescence = `dosage_fluo_384_CB:501,527`) %>% 
  filter(!is.na(fluorescence)) %>%
  mutate(row = substr(position, 1, 1)) %>% 
  mutate(col = substr(position, 2, 3)) %>% 
  arrange(QuantIt, row, as.numeric(col)) %>%
  mutate(plate = NA) %>%
  mutate(plate = ifelse(QuantIt == 1 & row != "I" & (as.numeric(col) %% 2 != 0), 
                         1, plate)) %>% 
  mutate(plate = ifelse(QuantIt == 1 & row != "I" & (as.numeric(col) %% 2 == 0), 
                         2, plate)) %>% 
  mutate(plate = ifelse(QuantIt == 1 & row == "I", "ladder", plate)) %>% 
  filter(!(QuantIt == 1 & plate == "ladder" & as.numeric(col) > 12)) %>% 
  mutate(plate = ifelse(QuantIt == 2 & row %in% LETTERS[1:8] & (as.numeric(col) %% 2 != 0), 
                         3, plate)) %>% 
  mutate(plate = ifelse(QuantIt == 2 & row %in% LETTERS[1:8] & (as.numeric(col) %% 2 == 0), 
                         4, plate)) %>% 
  mutate(plate = ifelse(QuantIt == 2 & row %in% LETTERS[9:11], 5, plate)) %>% 
  mutate(plate = ifelse(QuantIt == 2 & row == "L", "ladder", plate)) %>% 
  filter(!(QuantIt == 2 & plate == "ladder" & as.numeric(col) > 12)) %>% 
  mutate(row_plate = NA) %>% 
  mutate(row_plate = ifelse(plate %in% as.character(1:4), row, row_plate)) %>% 
  mutate(col_plate = NA) %>% 
  mutate(col_plate = ifelse(plate %in% c(1,3), (as.numeric(col) + 1)/2, col_plate)) %>% 
  mutate(col_plate = ifelse(plate %in% c(2,4), as.numeric(col)/2, col_plate)) %>% 
  mutate(concentration = NA) %>% 
  mutate(concentration = ifelse(QuantIt == 1 & plate == "ladder", 
                                rep(c(0, 5, 10, 20, 40, 60), each = 2), concentration)) %>% 
  mutate(concentration = ifelse(QuantIt == 2 & plate == "ladder", 
                                rep(c(0, 5, 10, 20, 40, 60)*19/20, each = 2), concentration))
```

```{stan quantitModel, output.var="quantitModel", echo=F, eval=F}
  data {
    int N ;
    real fluorescence[N] ;
    real concentration[N] ;
  }
  parameters {
    real beta ;
    real sigma ;
  }
  model {
    for(n in 1:N)
      concentration[n] ~ normal(beta*fluorescence[n], sigma) ;
  }
```

```{r fitQuantit}
data1 <- quantit %>% filter(QuantIt == 1, plate == "ladder")
data2 <- quantit %>% filter(QuantIt == 2, plate == "ladder")
# fit1 <- sampling(quantitModel,
#                 data = list(N = nrow(data1),
#                             fluorescence = data1$fluorescence,
#                             concentration = data1$concentration))
# fit2 <- sampling(quantitModel,
#                 data = list(N = nrow(data2),
#                             fluorescence = data2$fluorescence,
#                             concentration = data2$concentration))
# save(fit1, fit2, file = "./probes_save/MQantIt.Rdata")
load("./symcapture_save/MQantIt.Rdata")
```

```{r quantitConcentration}
beta1 <- broom::tidyMCMC(fit1, droppars = NULL, rhat = T) %>% 
  filter(term == "beta") %>% select(estimate) %>% unlist()
beta2 <- broom::tidyMCMC(fit2, droppars = NULL, rhat = T) %>% 
  filter(term == "beta") %>% select(estimate) %>% unlist()
quantit <- quantit %>% 
  mutate(concentration = ifelse(QuantIt == 1 & plate != "ladder", 
                                fluorescence*beta1, concentration)) %>% 
  mutate(concentration = ifelse(QuantIt == 2 & plate != "ladder", 
                                fluorescence*beta2, concentration))
```

```{r qbit}
qbit <- data.frame(
  plate = as.character(1),
  row = "A",
  col = as.character(1:12),
  Qbit = c(0, 0.186, 2.02, 0.26, 5.6, 0.258, 23.8, 8.12, 0.238, 15.4, 12.8, 4.2) 
) %>% 
  left_join(quantit %>% 
              mutate(col_plate = as.character(col_plate)),
            by = c("plate" = "plate", "row" = "row_plate", "col" = "col_plate"))
```

```{stan qbitModelRes, output.var="qbitModel", echo=F, eval=F}
  data {
    int N ;
    real quantit[N] ;
    real qbit[N] ;
  }
  parameters {
    real beta ;
    real sigma ;
  }
  model {
    for(n in 1:N)
      qbit[n] ~ normal(beta*quantit[n], sigma) ;
  }
```

```{r fitQbit}
# fit <- sampling(qbitModel,
#                 data = list(N = nrow(qbit),
#                             quantit = qbit$concentration,
#                             qbit = qbit$Qbit))
# save(fit, file = "./probes_save/MQbit.Rdata")
load("./symcapture_save/MQbit.Rdata")
```

```{r quantitQbit}
beta <- broom::tidyMCMC(fit, droppars = NULL, rhat = T) %>% 
  filter(term == "beta") %>% select(estimate) %>% unlist()
quantit <- quantit %>% 
  mutate(qbit_est = beta*concentration)
```

```{r quantitConcentrationQbit1}
quantit %>% 
  filter(plate %in% 1:4) %>% 
  ggplot(aes(y = factor(row_plate,levels = rev(LETTERS[1:8])),
             x = factor(col_plate,levels = 1:12))) + 
  geom_point(aes(col = concentration), size =6)  +
  geom_text(aes(label = round(qbit_est)), size = 2.5, col = "white") +
  facet_wrap(~ plate, nrow = 2, labeller = "label_both") +
  theme_bw() +
  labs(x=NULL, y = NULL)
```

## Post-enrichment (PCR1 & PCR2) and amplification quantification

Library showing no band or light smir in the electrophoresis have been reamplified, resulting in amplified plates A1 and A2. We used the Quant-It to dose all samples (original libraries and reamplified libraries).

```{r quantit2}
quantit <- read_csv(file.path(path, "QuantIt", 
                   "SYMCAPTURE_P1_P2_P3.csv"), skip = 53)[2:4] %>% 
  mutate(QuantIt = 1) %>% 
  bind_rows(read_csv(file.path(path, "QuantIt", 
                   "SYMCAPTURE_A1_A2_P4_P5.csv"), skip = 53)[2:4] %>% 
              mutate(QuantIt = 2)) %>% 
  select(-`Well ID`) %>% 
  dplyr::rename(position = Well, fluorescence = `dosage_fluo_384_CB:501,527`) %>% 
  filter(!is.na(fluorescence)) %>%
  mutate(row = substr(position, 1, 1)) %>% 
  mutate(col = substr(position, 2, 3)) %>% 
  arrange(QuantIt, row, as.numeric(col)) %>%
  mutate(plate = NA) %>% 
  mutate(plate = ifelse(QuantIt == 1 & row %in% LETTERS[1:4], "P1", plate)) %>% 
  mutate(plate = ifelse(QuantIt == 1 & row %in% LETTERS[5:8], "P2", plate)) %>% 
  mutate(plate = ifelse(QuantIt == 1 & row %in% LETTERS[9:12], "P3", plate)) %>% 
  mutate(plate = ifelse(QuantIt == 1 & row == "M", "ladder", plate)) %>% 
  mutate(plate = ifelse(QuantIt == 2 & row %in% LETTERS[1:4], "A1", plate)) %>% 
  mutate(plate = ifelse(QuantIt == 2 & row %in% LETTERS[5:8], "A2", plate)) %>% 
  mutate(plate = ifelse(QuantIt == 2 & row %in% LETTERS[9:12], "P4", plate)) %>% 
  mutate(plate = ifelse(QuantIt == 2 & row %in% LETTERS[13:14], "P5", plate)) %>% 
  mutate(plate = ifelse(QuantIt == 2 & row == "O", "ladder", plate)) %>% 
  filter(!(plate == "ladder" & as.numeric(col) > 12)) %>% 
  mutate(row_plate = NA) %>% 
  group_by(plate, row, col) %>% 
  mutate(row_plate = ifelse(plate %in% c("P1", "A1") & (as.numeric(col) %% 2 != 0), 
                            LETTERS[which(LETTERS == row)*2-1], row_plate)) %>% 
  mutate(row_plate = ifelse(plate %in% c("P1", "A1") & (as.numeric(col) %% 2 == 0), 
                            LETTERS[which(LETTERS == row)*2], row_plate)) %>%
  mutate(row_plate = ifelse(plate %in% c("P2", "A2") & (as.numeric(col) %% 2 != 0), 
                            LETTERS[(which(LETTERS == row)-4)*2-1], row_plate)) %>% 
  mutate(row_plate = ifelse(plate %in% c("P2", "A2") & (as.numeric(col) %% 2 == 0), 
                            LETTERS[(which(LETTERS == row)-4)*2], row_plate)) %>% 
  mutate(row_plate = ifelse(plate %in% c("P3", "P4") & (as.numeric(col) %% 2 != 0), 
                            LETTERS[(which(LETTERS == row)-8)*2-1], row_plate)) %>% 
  mutate(row_plate = ifelse(plate %in% c("P3", "P4") & (as.numeric(col) %% 2 == 0), 
                            LETTERS[(which(LETTERS == row)-8)*2], row_plate)) %>%
  mutate(row_plate = ifelse(plate == "P5" & row == "M" & col %in% seq(1, 11, length.out = 6), "A", row_plate)) %>% 
  mutate(row_plate = ifelse(plate == "P5" & row == "M" & col %in% seq(2, 12, length.out = 6), "B", row_plate)) %>% 
  mutate(row_plate = ifelse(plate == "P5" & row == "M" & col %in% seq(13, 23, length.out = 6), "C", row_plate)) %>% 
  mutate(row_plate = ifelse(plate == "P5" & row == "M" & col %in% seq(14, 24, length.out = 6), "D", row_plate)) %>% 
    mutate(row_plate = ifelse(plate == "P5" & row == "N" & col %in% seq(1, 11, length.out = 6), "E", row_plate)) %>% 
  mutate(row_plate = ifelse(plate == "P5" & row == "N" & col %in% seq(2, 12, length.out = 6), "F", row_plate)) %>% 
  mutate(row_plate = ifelse(plate == "P5" & row == "N" & col %in% seq(13, 23, length.out = 6), "G", row_plate)) %>% 
  mutate(row_plate = ifelse(plate == "P5" & row == "N" & col %in% seq(14, 24, length.out = 6), "H", row_plate)) %>% 
  mutate(col_plate = NA) %>% 
  mutate(col_plate = ifelse(!(plate %in% c("P5", "ladder")) & (as.numeric(col) %% 2 != 0), 
                            (as.numeric(col)+1)/2, col_plate)) %>% 
  mutate(col_plate = ifelse(!(plate %in% c("P5", "ladder")) & (as.numeric(col) %% 2 == 0), 
                            as.numeric(col)/2, col_plate)) %>% 
  mutate(col_plate = ifelse(plate == "P5" & (as.numeric(col) %% 2 != 0) & as.numeric(col) < 13, (as.numeric(col)+1)/2, col_plate)) %>% 
  mutate(col_plate = ifelse(plate == "P5" & (as.numeric(col) %% 2 == 0) & as.numeric(col) < 13, as.numeric(col)/2, col_plate)) %>% 
  mutate(col_plate = ifelse(plate == "P5" & (as.numeric(col) %% 2 != 0) & as.numeric(col) > 12, (as.numeric(col)-11)/2, col_plate)) %>% 
  mutate(col_plate = ifelse(plate == "P5" & (as.numeric(col) %% 2 == 0) & as.numeric(col) > 12, (as.numeric(col)-12)/2, col_plate)) %>% 
  group_by(plate) %>% 
  mutate(concentration = NA) %>% 
  mutate(concentration = ifelse(plate == "ladder", 
                                rep(c(0, 5, 10, 20, 40, 60)*19/20, each = 2),
                                concentration)) %>% 
  ungroup()
```

```{r fitQuantit2}
data1 <- quantit %>% filter(QuantIt == 1, plate == "ladder")
data2 <- quantit %>% filter(QuantIt == 2, plate == "ladder")
# fit1 <- sampling(quantitModel,
#                 data = list(N = nrow(data1),
#                             fluorescence = data1$fluorescence,
#                             concentration = data1$concentration))
# fit2 <- sampling(quantitModel,
#                 data = list(N = nrow(data2),
#                             fluorescence = data2$fluorescence,
#                             concentration = data2$concentration))
# save(fit1, fit2, file = "./probes_save/MQantIt2.Rdata")
load("./symcapture_save//MQantIt2.Rdata")
```

```{r quantitConcentration2}
beta1 <- broom::tidyMCMC(fit1, droppars = NULL, rhat = T) %>% 
  filter(term == "beta") %>% select(estimate) %>% unlist()
beta2 <- broom::tidyMCMC(fit2, droppars = NULL, rhat = T) %>% 
  filter(term == "beta") %>% select(estimate) %>% unlist()
quantit <- quantit %>% 
  mutate(concentration = ifelse(QuantIt == 1 & plate != "ladder", 
                                fluorescence*beta1, concentration)) %>% 
  mutate(concentration = ifelse(QuantIt == 2 & plate != "ladder", 
                                fluorescence*beta2, concentration))
```

```{r quantitConcentrationPlot4}
quantit %>% 
  filter(plate != "ladder") %>% 
  ggplot(aes(y = factor(row_plate,levels = rev(LETTERS[1:8])),
             x = factor(col_plate,levels = 1:12))) + 
  geom_point(aes(col = concentration), size =6)  +
  geom_text(aes(label = round(concentration)), size = 2.5, col = "white") +
  facet_wrap(~ plate, ncol = 2, labeller = "label_both") +
  theme_bw() +
  labs(x=NULL, y = NULL) +
  scale_color_continuous(guide = F)
```

```{r}
googlesheets::gs_title("Symcapture") %>% 
  googlesheets::gs_read("Amplification") %>% 
  left_join(quantit %>% 
              filter(grepl("A", plate)) %>% 
              mutate(Plate_amplification = as.numeric(sub("A", "", plate))) %>% 
              mutate(Position_amplification = paste0(row_plate, col_plate)) %>% 
              mutate(Concentration_QuantIt = concentration) %>% 
              select(Plate_amplification, Position_amplification, Concentration_QuantIt)) %>% 
  write_csv(path = file.path(path, "Amplification.csv"))
```


## Amplification result

We used electrophoresis and noticed if re-amplified sampled had a band (B), a light smir (L) or nothing (A). Samples without anything (A) had their library remade.

```{r elecrtophoresisAmplified, fig.cap="Electrophoresis Amplified"}
googlesheets::gs_title("Symcapture") %>% 
  googlesheets::gs_read("Amplification") %>% 
  mutate(row = substr(Position_amplification, 1, 1)) %>% 
  mutate(col = substr(Position_amplification, 2, 3)) %>% 
  ggplot(aes(y = factor(row,levels = rev(LETTERS[1:8])),
             x = factor(col,levels = 1:12))) + 
  geom_point(aes(col = Quality_Electrophoresis), size =6)  +
  geom_text(aes(label = Quality_Electrophoresis), size = 2.5, col = "white") +
  facet_wrap(~ Plate_amplification, ncol = 2, labeller = "label_both") +
  theme_bw() +
  labs(x=NULL, y = NULL)
```

## Library & extraction repetition result

69 libraries were still not ok after re-amplification. Those library have been rebuilt from source DNA. Among them, 43 were still not good at the electrophoresis and have bee re-extracted. Electrophoresis of re-extracted samples show a high heterogeneity of size, and around 8 of them seem to have not worked. **All new samples need to be quantified.**.


