```{r setup_variantFiltering, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Symphonia_Paracou/"
```

```{bash variantFiltering, eval=F, echo=T}
vcftools --gzvcf reference.sequences_00.vcf.gz --missing-indv -c
vcftools --gzvcf reference.sequences_00.vcf.gz --missing-site -c > missing.txt
vcftools --gzvcf reference.sequences_00.vcf.gz --site-quality -c > QUAL.txt
vcftools --gzvcf reference.sequences_00.vcf.gz \
  --get-INFO AC \
  --get-INFO AF \
  --get-INFO QD \
  --get-INFO FS \
  --get-INFO SOR \
  -c > INFO.txt
```

```{r variantFilteringR}
filtering <- left_join(read_tsv(file.path(path, "Sequences", "variantCalling", "symcapture.vcf.stats", "missing.txt")),
          read_tsv(file.path(path, "Sequences", "variantCalling", "symcapture.vcf.stats", "QUAL.txt"))) %>% 
  left_join(read_tsv(file.path(path, "Sequences", "variantCalling", "symcapture.vcf.stats", "INFO.txt"))) %>% 
  mutate(QD = as.numeric(QD)) %>% 
  mutate(AF = as.numeric(AF)) %>% 
  sample_n(1000) # for tests
```

# Variant filtering

We will filter the final raw VCF with following filters (name, filter, description) using next histograms to set and test parameters values :

* __Missing__ `missing > 0.15`: filter out variant if they have more than 15% missing data (\@ref(fig:missing))
* __Allele count__ `AC > 3`: filter out variant if there are more than 2 allelic form (restricting to biallelic variants \@ref(fig:AC))
* __Quality ([QUAL](https://software.broadinstitute.org/gatk/documentation/article?id=7258))__ `QUAL < 30`: represents the likelihood of the site to be homozygote across all samples, we filter out variants having a low quality score (\@ref(fig:QUAL))
* __Quality depth ([QD](https://software.broadinstitute.org/gatk/documentation/tooldocs/3.8-0/org_broadinstitute_gatk_tools_walkers_annotator_QualByDepth.php))__ `QD < 2`: filter out variants with low variant confidence (\@ref(fig:QD))
* __Fisher strand bias ([FS](https://software.broadinstitute.org/gatk/documentation/tooldocs/3.8-0/org_broadinstitute_gatk_tools_walkers_annotator_FisherStrand.php))__ `FS > 60`: filter out variants based on Phred-scaled p-value using Fisher's exact test to detect strand bias (\@ref(fig:FS))
* __Strand odd ratio ([SOR](https://software.broadinstitute.org/gatk/documentation/tooldocs/3.8-0/org_broadinstitute_gatk_tools_walkers_annotator_StrandOddsRatio.php))__ `SOR < 3`: filter out variants based on Phred-scaled p-value used to detect strand bias (\@ref(fig:SOR))

For this we assessed different metrics using [`vcftools`](http://vcftools.sourceforge.net/man_latest.html), proceeded to the fltering, and quickly examined the result with genomic principal component analysis (gPCA):

* __Missing__ missing individuals histogram to set the threshold
* __AC__ allele count histogram to set the threshold
* __QUAL__ quality histogram to set the threshold
* __QD__ quality by depth histogram to set the threshold
* __FS__ Fisher strand ratio histogram to set the threshold
* __SOR__ Strand odd ratio histogram to set the threshold
* __Filtering__ filtering step with `gatk VariantFiltration`, `gatk SelectVariants` and `vcftools`
* __gPCA check__ quick assesment of the produced filtered VCF

<!-- recommandations https://www.france-bioinformatique.fr/sites/default/files/V04_FiltrageVariantNLapaluRoscoff2016_0.pdf -->
<!-- https://software.broadinstitute.org/gatk/documentation/article.php?id=6925 -->

## Missing

```{r missing, fig.cap="Missing data frequency per variant."}
Missingthresh <- 0.15
ggplot(filtering, aes(F_MISS)) + 
  geom_rect(xmin = Missingthresh, xmax = 100, ymin = 0, ymax = 100, col = "grey", alpha = 0.4) +
  geom_density(fill = "blue", alpha = 0.4, col = NA) +
  geom_vline(xintercept = Missingthresh, col = "red", linetype = "dashed") +
  xlab("Missing individuals percentage") +
  ylab("variant count") +
  ggtitle(paste0("Missing: ", round(nrow(filter(filtering, F_MISS < Missingthresh))/nrow(filtering)*100, 2),
                 "% (", nrow(filter(filtering, F_MISS < Missingthresh)), 
                 " / ", nrow(filtering), ") variant below ", Missingthresh))
```

## AC

```{r AC, fig.cap="Allelic count per variant."}
ACthresh <- 3
ggplot(filtering, aes(AC)) + 
  geom_rect(xmin = ACthresh, xmax = 100, ymin = 0, ymax = 100, col = "grey", alpha = 0.4) +
  geom_density(fill = "blue", alpha = 0.4, col = NA) +
  geom_vline(xintercept = ACthresh, col = "red", linetype = "dashed") +
  xlab("Allelic count") +
  ylab("variant count") +
  xlim(0, 100) +
  ggtitle(paste0("AC: ", round(nrow(filter(filtering, AC < ACthresh))/nrow(filtering)*100, 2),
                 "% (", nrow(filter(filtering, AC < ACthresh)), 
                 " / ", nrow(filtering), ") variant below ", ACthresh))
```

## QUAL

```{r QUAL, fig.cap="Variants quality."}
QUALthresh <- 30
ggplot(filtering, aes(QUAL)) + 
  geom_rect(xmin = 0, xmax = log10(QUALthresh), ymin = 0, ymax = 100, col = "grey", alpha = 0.4) +
  geom_density(fill = "blue", alpha = 0.4, col = NA) +
  geom_vline(xintercept = QUALthresh, col = "red", linetype = "dashed") +
  xlab("Quality depth") +
  ylab("variant count") +
  scale_x_log10() +
  ggtitle(paste0("QUAL: ", round(nrow(filter(filtering, QUAL > QUALthresh))/nrow(filtering)*100, 2),
                 "% (", nrow(filter(filtering, QUAL > QUALthresh)), 
                 " / ", nrow(filtering), ") variant above ", QUALthresh))
```

## QD

```{r QD, fig.cap="Variants quality by depth."}
QDthresh <- 2
ggplot(filtering, aes(QD)) + 
  geom_rect(xmin = 0, xmax = QDthresh, ymin = 0, ymax = 100, col = "grey", alpha = 0.4) +
  geom_density(fill = "blue", alpha = 0.4, col = NA) +
  geom_vline(xintercept = QDthresh, col = "red", linetype = "dashed") +
  xlab("Quality depth") +
  ylab("variant count") +
  ggtitle(paste0("QD: ", round(nrow(filter(filtering, QD > QDthresh))/nrow(filtering)*100, 2),
                 "% (", nrow(filter(filtering, QD > QDthresh)), 
                 " / ", nrow(filtering), ") variant above ", QDthresh))
```

## FS

```{r FS, fig.cap="Fisher strand ratio per variant."}
FSthresh <- 60
ggplot(filtering, aes(FS)) + 
  geom_rect(xmin = FSthresh, xmax = 1000, ymin = 0, ymax = 100, col = "grey", alpha = 0.4) +
  geom_density(fill = "blue", alpha = 0.4, col = NA) +
  geom_vline(xintercept = FSthresh, col = "red", linetype = "dashed") +
  xlab("Quality depth") +
  ylab("variant count") +
  ggtitle(paste0("FS: ", round(nrow(filter(filtering, FS < FSthresh))/nrow(filtering)*100, 2),
                 "% (", nrow(filter(filtering, FS < FSthresh)), 
                 " / ", nrow(filtering), ") variant above ", FSthresh))
```

## SOR

```{r SOR, fig.cap="Strand odd ratio per variant."}
SORthresh <- 3
ggplot(filtering, aes(SOR)) + 
  geom_rect(xmin = SORthresh, xmax = 100, ymin = 0, ymax = 100, col = "grey", alpha = 0.4) +
  geom_density(fill = "blue", alpha = 0.4, col = NA) +
  geom_vline(xintercept = SORthresh, col = "red", linetype = "dashed") +
  xlab("Quality depth") +
  ylab("variant count") +
  ggtitle(paste0("SOR: ", round(nrow(filter(filtering, SOR < SORthresh))/nrow(filtering)*100, 2),
                 "% (", nrow(filter(filtering, SOR < SORthresh)), 
                 " / ", nrow(filtering), ") variant below ", SORthresh))
```

## Filtering

Finally with all filtered combined we expect `r format(round(nrow(filter(filtering, F_MISS < Missingthresh, AC < ACthresh, QUAL > QUALthresh, QD > QDthresh, FS < FSthresh, SOR < SORthresh)) / 1000 * 26813513), big.mark = " ")` variants. We will first filter at 50% of missing data and will create an extra filtered VCF at 15% of missing. 42 sequences subset lost individuals in genotype call (over 878, 4%), blocking the functionning of `gatk CombineVariants`, so in a first time we removed the variants associated to those subest of reference sequences. We also extracted the Paracou individual from the final VCF for population genomics analysis.

```{r, eval=F}
format(round(nrow(filter(filtering, F_MISS < Missingthresh, AC < ACthresh, QUAL > QUALthresh, QD > QDthresh, FS < FSthresh, SOR < SORthresh)) / 1000 * 26813513), big.mark = " ")

filter(filtering, F_MISS < Missingthresh, AC < ACthresh, QUAL > QUALthresh, QD > QDthresh, FS < FSthresh, SOR < SORthresh) %>% 
    filter(AF < 0.5) %>% 
  ggplot(aes(AF)) +
  geom_histogram()
```

```{bash missingidv, eval=F, echo=T}
mkdir out
mkdir missing_ind
for file in $(ls symcapture.filtered.vcf/*.vcf.gz) ; do file=$(basename $file) ; file=${file%.*} ; echo "module load bioinfo/tabix-0.2.5 ; module load bioinfo/vcftools-0.1.15 ; vcftools --gzvcf symcapture.filtered.vcf/$file.gz --missing-indv -c > missing_ind/$file.missing.txt"; done > missingInd.sh
sarray -J missingInd -o out/%j.missingInd.out -e out/%j.missingInd.err -t 1:00:00 --mail-type=BEGIN,END,FAIL missingInd.sh
for file in $(ls missing_ind/*.missing.txt) ; do awk '{{if (NR!=1) print FILENAME"\t"$0}}' $file ; done > missingInd.txt
rm - r out
rm -r missing_ind
```

```{r missingindR}
missing <- read_tsv(file.path(path, "Sequences", "variantCalling", "missingInd.txt"), 
         col_names = c("ref", "sample", "ndata", "ngenotypes", "nmiss", "freq")) %>% 
  mutate(ref = gsub("missing_ind/reference.sequences_", "", ref)) %>% 
  mutate(ref = gsub(".vcf.missing.txt", "", ref)) %>% 
  mutate(sample = gsub(".g.vcf", "", sample)) %>% 
  reshape2::dcast(ref ~ sample, value.var = "freq") %>% 
  reshape2::melt(id.vars = "ref", variable.name = "sample", value.name = "missing") %>% 
  mutate(missing = ifelse(is.na(missing) | is.nan(missing), 1, missing))
read_tsv(file.path(path, "Sequences", "variantCalling", "missingInd.txt"), 
         col_names = c("ref", "sample", "ndata", "ngenotypes", "nmiss", "freq")) %>% 
  mutate(ref = gsub("missing_ind/reference.sequences_", "", ref)) %>% 
  mutate(ref = gsub(".vcf.missing.txt", "", ref)) %>% 
  mutate(sample = gsub(".g.vcf", "", sample)) %>% 
  group_by(ref) %>% 
  summarise(N = n()) %>% 
  filter(N == 432) %>% 
  select(ref) %>% 
  mutate(ref = paste0("symcapture.filtered.vcf/reference.sequences_", ref, ".vcf.gz")) %>% 
  write_tsv(file.path(path, "Sequences", "variantCalling", "nonmissing.list"), col_names = F)
googlesheets::gs_title("Symcapture") %>% 
              googlesheets::gs_read("Pop") %>% 
  mutate(Ind = paste0(Ind, ".g.vcf", "")) %>% 
  filter(Pop2 != "O") %>% 
  select(Ind) %>% 
  write_tsv(file.path(path, "Sequences", "variantCalling", "Paracou.ind"), col_names = F)
```


```{bash genotype, eval=F, echo=T}
echo -e '#!/bin/bash\n#SBATCH --time=48:00:00\n#SBATCH -J gather\n#SBATCH -o gather.out\n#SBATCH -e gather.err\n#SBATCH --mem=20G\n#SBATCH --cpus-per-task=1\n#SBATCH --mail-type=BEGIN,END,FAIL\nmodule load bioinfo/picard-2.14.1\njava -Xmx20g -jar $PICARD GatherVcfs \' > gather.sh
for file in $(cat nonmissing.list)
do
	echo -e '\tI=symcapture.raw.vcf/'$(basename $file)' \' >> gather.sh
done
echo -e '\tO=symcapture.all.raw.vcf.gz\n' >> gather.sh

srun --mem=80G --cpus-per-task=8 --pty bash

module load bioinfo/bcftools-1.8
bcftools view --max-alleles 2 --threads 8 symcapture.all.raw.vcf.gz | bgzip -c --threads 8 > symcapture.all.biallelic.vcf.gz

module load bioinfo/gatk-4.1.2.0
gatk IndexFeatureFile \
  -F symcapture.all.biallelic.vcf.gz
gatk SelectVariants \
  -V symcapture.all.biallelic.vcf.gz \
  -select-type SNP \
  -O symcapture.all.biallelic.snp.vcf.gz
gatk IndexFeatureFile \
  -F symcapture.all.biallelic.snp.vcf.gz

# https://hannahvmeyer.github.io/plinkQC/articles/plinkQC.html
module load bioinfo/plink2-v2.0_alpha2
mkdir test
plink2 \
  --threads 8 \
  --memory 80000 \
  --vcf symcapture.all.biallelic.snp.vcf.gz \
  --allow-extra-chr \
  --var-min-qual 30 \
  --mind 0.15 \
  --geno 0.15 \
  --maf 0.01 \
  --hwe 0.0001 \
  --make-bed --out test/symcapture.all.test
  
gatk VariantFiltration \
  -V input.vcf.gz \
  # --filter-expression "QUAL < 30.0 || QD < 2.0 || FS > 60.0 || SOR > 3.0" \
  --filter-expression "QD < 2.0 || FS > 60.0 || SOR > 3.0" \
  --filter-name "FAIL" \
  -O intermediate.vcf.gz
gatk SelectVariants \
  -V intermediate.vcf.gz \
  --exclude-filtered \
  -O final.vcf.gz

bcftools view -S Paracou.ind --threads 8 final.vcf.gz | bgzip --threads 8 -c > final.Paracou.vcf.gz
```
