```{r setup_growthanimal, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(raster)
library(bayesplot)
library(kinship2)
library(rstan)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = T, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
load("./symcapture_save/Env2.Rdata")
trees <- trees %>% 
  na.omit(pop) %>% # remove admixed
  group_by(Ind) %>% 
  mutate(Y0 = first(CensusYear), DBH0 = first(DBH), DBHtoday = last(DBH)) %>% 
  ungroup() %>% 
  dplyr::select(Ind, IID, pop, Plot, SubPlot, TreeFieldNum, Y0, DBH0, DBHtoday)
```

# Growth & Animal models

## Introduction

The aim of this model is to explore animal and growth model building theoritically in order to validate its behavior and use it on *Symphonia* and *Eschweilera* data. 

## Tes 1 - Simulated Kinship

Let's consider a set of $P$ populations including each $Fam$ families composed of $I = 14$ individuals with arbitrar relationships.

```{r kinship1, fig.cap="Kinship matrix"}
kin.df <- function(P= 3, Fam = 6){
  I <- 14 # Individuals
  data.frame(pop = rep(1:P, each = Fam*I),
                       fam = rep(1:Fam, each = I),
                       ind = rep(1:I, Fam*P),
                       father = rep(c(NA, NA, 1, 1, 1, 1, NA, 6, 6, 6, NA, 11, 11, 11), Fam*P),
                       mother = rep(c(NA, NA, 2, 2, 2, 2, NA, 7, 7, 7, NA, 3, 3, 3), Fam*P), 
                       sex = rep(c(1, 2, 2, 3, 3, 1, 2, 3, 3, 3, 1, 3, 3, 3), Fam*P)) %>% 
    mutate_at(c("father", "mother"), funs(ifelse(!is.na(.), paste0(pop,fam,.), NA))) %>% 
    mutate(ind = paste0(pop,fam,ind))
}
kin.mat <- function(P = 3, Fam = 6){
  ped.df <- kin.df(P, Fam)
  ped.ped <- pedigree(id = ped.df$ind, dadid = ped.df$father, momid = ped.df$mother, sex = ped.df$sex, famid = paste0(ped.df$pop, ped.df$fam))
  K <- as.matrix(kinship(ped.ped))
  K <- 2*K
}
heatmap(kin.mat(3, 6))
```

```{r gdata}
gdata <- function(
  I = 20, # Nb ind
  Y = 33, # Nb years
  P = 3, # Nb pops
  Gmax = c(0.54, 0.53, 0.44), # model parameters
  Dopt =  c(27, 31, 20),
  Ks = c(0.67, 0.73, 0.64),
  sigmaR = c(0.16, 9, 0.22),
  sigma = 0.14
){
  Gmax <- Gmax[1:P]
  Dopt <- Dopt[1:P]
  Ks <- Ks[1:P]
  years <- 1985:(1985+Y-1)
  pop <- sample(1:P, I, replace = T)
  DBH0 <- rlnorm(I, meanlog = mean(log(trees$DBH0)), 
                 sdlog = sd(log(trees$DBH0))) 
  Y0 <- sample(trees$Y0, I)
  Gmaxi <- rlnorm(I, mean = log(Gmax[pop]), sd = sqrt(sigmaR[1]))
  Dopti <- rlnorm(I, mean = log(Dopt[pop]), sd = sqrt(sigmaR[2]))
  Ksi <- rlnorm(I, mean = log(Ks[pop]), sd = sqrt(sigmaR[3]))
  DBH <- rep(NA, I)
  for (t in 1:(Y-1)) {
    for(i in 1:I){
      if(years[t] == Y0[i]) 
        DBH[i] <- DBH0[i] 
    }
    DBH <- DBH + Gmaxi * exp(-0.5*(log(DBH / Dopti) / Ksi) * (log(DBH / Dopti) / Ksi))
  }
  DBHtoday <- rnorm(I, DBH, sigma)
  list(params = list(Gmax = Gmax, Dopt = Dopt, Ks = Ks, 
                     sigmaR = sigmaR, sigma = sigma),
       data = data.frame(ind = 1:I, pop = pop, 
                         DBH0 = DBH0, Y0 = Y0, Y = Y, DBHtoday = DBHtoday),
       mdata = list(I = I, Y = Y, P = P, years = years, DBH0 = DBH0,
                    Y0 = Y0, DBHtoday = DBHtoday, pop = pop))
}
```

```{r gkdata}
gkdata <- function(
  P = 3, # Nb pops
  Fam = 6, # Nb family
  Y = 33, # Nb years
  Gmax = c(0.54, 0.53, 0.44), # model parameters
  Dopt =  c(27, 31, 20),
  Ks = c(0.67, 0.73, 0.64),
  sigmaR = c(0.16, 9, 0.22),
  sigmaG = c(0.37, 25, 0.38),
  sigma = 0.14
){
  K <- kin.mat(P, Fam)
  Kdf <- kin.df(P, fam)
  I <- nrow(K)
  Gmax <- Gmax[1:P]
  Dopt <- Dopt[1:P]
  Ks <- Ks[1:P]
  years <- 1985:(1985+Y-1)
  pop <- sample(1:P, I, replace = T)
  DBH0 <- rlnorm(I, meanlog = mean(log(trees$DBH0)), 
                 sdlog = sd(log(trees$DBH0))) 
  Y0 <- sample(trees$Y0, I)
  uG <- rnorm(I, sd = sqrt(sigmaG))
  uD <- rnorm(I, sd = sqrt(sigmaG))
  uK <- rnorm(I, sd = sqrt(sigmaG))
  Gmaxi <- rlnorm(I, mean = log(Gmax[Kdf$pop] + as.vector(t(chol(K)) %*% uG)), 
                  sd = sqrt(sigmaR[1]))
  Dopti <- rlnorm(I, mean = log(Dopt[Kdf$pop] + as.vector(t(chol(K)) %*% uD)), 
                  sd = sqrt(sigmaR[2]))
  Ksi <- rlnorm(I, mean = log(Ks[Kdf$pop] + as.vector(t(chol(K)) %*% uK)), 
                sd = sqrt(sigmaR[3]))
  DBH <- rep(NA, I)
  for (t in 1:(Y-1)) {
    for(i in 1:I){
      if(years[t] == Y0[i]) 
        DBH[i] <- DBH0[i] 
    }
    DBH <- DBH + Gmaxi * exp(-0.5*(log(DBH / Dopti) / Ksi) * (log(DBH / Dopti) / Ksi))
  }
  DBHtoday <- rnorm(I, DBH, sigma)
  list(params = list(Gmax = Gmax, Dopt = Dopt, Ks = Ks, 
                     sigmaR = sigmaR, sigmaG = sigmaG, sigma = sigma),
       data = data.frame(ind = 1:I, pop = pop, 
                         DBH0 = DBH0, Y0 = Y0, Y = Y, DBHtoday = DBHtoday),
       mdata = list(I = I, Y = Y, P = P, years = years, DBH0 = DBH0,
                    Y0 = Y0, DBHtoday = DBHtoday, pop = pop, K = K))
}
```

### Genotype

We will use the following animal model to estimate variance associated to the genotype without population variance at first:

$$y \sim \mathcal N(\mu + A.a,\sigma_R)~|~a \sim \mathcal N(0,\sigma_G)$$

```{r genotype1}
K <- kin.mat(1, 6)
sigmaY <- 1
sigmaG <- 0.3
sigmaR <- sigmaY - sigmaG
N <- nrow(K)
mu <- rnorm(1)
u <- rnorm(N, sd = sqrt(sigmaG))
r <- rnorm(N, sd = sqrt(sigmaR))
y <- mu + as.vector(t(chol(K)) %*% u) + r
ggplot(data.frame(y = y, u = u), aes(u, y)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  ggtitle(paste("Vy =", round(var(y), 2), 
                "mu =", round(mu, 2),
                "Vg=", round(var(as.vector((t(chol(K)) %*% u))), 2), 
                "Vr=", round(var(r), 2)))
animal1 <- stan_model("symcapture_models/Animal1.stan")
fitAnimal1 <- sampling(animal1, chains = 2, save_warmup = F, data = list(N = N, Y = y, K = K))
broom::tidyMCMC(fitAnimal1, pars = c("mu", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitAnimal1, pars = c("mu", "sigmaG", "sigmaR")))
```

### Population & Genotype

We will use the following animal model to estimate variance associated to the genotype with population variance:

$$y \sim \mathcal N(\mu_p + A.a,\sigma_R)~|~a \sim \mathcal N(o,\sigma_G)$$

```{r pop1}
P = 3
Fam = 6
K <- kin.mat(P, Fam)
ped.df <- kin.df(P, Fam)
sigmaP <- 0.8
sigmaG <- 0.8
sigmaR <- 0.2
N <- nrow(K)
p <- rnorm(P, sd = sqrt(sigmaP))
u <- rnorm(N, sd = sqrt(sigmaG))
r <- rnorm(N, sd = sqrt(sigmaR))
y <- p[ped.df$pop] + as.vector(t(chol(K)) %*% u) + r
ggplot(data.frame(y = y, u = u, pop = as.factor(ped.df$pop)),
       aes(u, y, col = pop)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  ggtitle(paste("Vy =", round(var(y), 2), 
                "Vp =", round(var(p), 2), 
                "Vg=", round(var(as.vector((t(chol(K)) %*% u))), 2), 
                "Vr=", round(var(r), 2)),
          paste("p =", paste0(round(p, 2), collapse = " ")))
animal2 <- stan_model("symcapture_models/Animal2.stan")
fitAnimal2 <- sampling(animal2, chains = 2, save_warmup = F, 
                       data = list(N = N, Y = y, K = K, P = P, population = ped.df$pop))
broom::tidyMCMC(fitAnimal2, pars = c("mu", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitAnimal2, pars = c("mu", "sigmaG", "sigmaR")))
```

### Growth - DBH

```{r growth1, eval=F}
mdata <- gdata1(I = 20, P = 1)
ggplot(mdata$data, aes(x = DBH0, y = DBHtoday, col = Y0, shape = as.factor(pop))) + 
  geom_point() + geom_abline()
growth1 <- stan_model("./symcapture_models/Growth5.stan")
fitGrowth1 <- sampling(growth1, chains = 2, data = mdata$mdata, save_warmup = F,
                      include = F, pars = c("DBH", "Gmaxi", "Dopti", "Ksi"))
unlist(mdata$params)
broom::tidyMCMC(fitGrowth1, droppars = NULL, rhat = T) %>% kable()
mcmc_trace(fitGrowth1) + 
  ggtitle(paste(unlist(mdata$params), collapse = " "),
          paste(names(unlist(mdata$params)), collapse = " "))
# mcmc_pairs(fitGrowth1, regex_pars = c("Gmax", "Dopt", "Ks"))
```

### Growth - DBH bis

```{r growth1b, eval=F}
mdata <- gdata1(I = 50, P = 1)
ggplot(mdata$data, aes(x = DBH0, y = DBHtoday, col = Y0, shape = as.factor(pop))) + 
  geom_point() + geom_abline()
growth1 <- stan_model("./symcapture_models/Growth5b.stan")
fitGrowth1 <- sampling(growth1, chains = 2, data = mdata$mdata, save_warmup = F,
                      include = F, pars = c("DBH", "Gmaxi"))
unlist(mdata$params)
broom::tidyMCMC(fitGrowth1, droppars = NULL, rhat = T) %>% kable()
mcmc_trace(fitGrowth1) + 
  ggtitle(paste(unlist(mdata$params), collapse = " "),
          paste(names(unlist(mdata$params)), collapse = " "))
```

### Growth - Gmax

```{r growth2, eval=F}
mdata <- gdata1(I = 20, P = 1)
ggplot(mdata$data, aes(x = DBH0, y = DBHtoday, col = Y0, shape = as.factor(pop))) + 
  geom_point() + geom_abline()
growth2 <- stan_model("./symcapture_models/Growth6.stan")
fitGrowth2 <- sampling(growth2, chains = 2, data = mdata$mdata, save_warmup = F,
                      include = F, pars = c("DBH", "Gmax"))
unlist(mdata$params)
broom::tidyMCMC(fitGrowth2, droppars = NULL, rhat = T, 
                estimate.method = "median") %>% kable()
mcmc_trace(fitGrowth2) + 
  ggtitle(paste(unlist(mdata$params), collapse = " "),
          paste(names(unlist(mdata$params)), collapse = " "))
```

### Growth & Genotype

```{r growthK1, eval=F}
growthanimal1 <- stan_model("symcapture_models/GrowthAnimalTest1.stan")
fitGrowthAnimal1 <- sampling(growthanimal1, chains = 2, save_warmup = F, 
                             data = list(N = nrow(data), I = max(data$i), y = data$y, 
                                         x = data $x, ind = data$i, K = K, L = c(0.1, 0, 0.1), U = c(2, 200, 3)))
broom::tidyMCMC(fitGrowthAnimal1, pars = c("theta", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitGrowthAnimal1, pars = c("theta", "sigmaG", "sigmaR")))
mcmc_pairs(as.array(fitGrowthAnimal1, pars = c("sigmaG", "sigmaR")))
```

### Growth & Gmax only

```{r growthK2, eval=F}
P = 1
Fam = 5
K <- kin.mat(P, Fam)
ped.df <- kin.df(P, Fam)
sigmaG <- 0.4
sigmaRt <- c(0.1, 10, 0.2)
sigmaR <- 0.1
N <- nrow(K)
Time = 3
u <- rnorm(N, sd = sqrt(sigmaG))
theta <- c(1, 30, 0.8)
r <- rnorm(N, sd = sqrt(sigmaR))
rt <- sapply(sigmaRt, function(sigma) rnorm(N, sd = sqrt(sigma)))
thetai <- matrix(nrow = N, ncol = 3)
thetai[,1] <- theta[1] + t(chol(K)) %*% u + rt[,1]
for(i in 2:3)
  thetai[,i] <- theta[i] + rt[,i]
data <- sapply(1:Time, function(t)
  data.frame(x = rlnorm(N, meanlog = 3.2, sdlog = 0.5),
             i = rownames(K)) %>% 
    mutate(y = thetai[,1]*exp(-0.5*(log(x/thetai[,2])/thetai[,3])^2) + r),
  simplify = F
) %>% bind_rows(.id = "t") %>% 
  mutate(i = as.numeric(as.factor(i)))
ggplot(data, aes(x, y, col = as.factor(i))) +
  geom_point() +
  stat_function(fun = function(x)
    theta[1]*exp(-0.5*(log(x/theta[2])/theta[3])^2), col = "red") +
  scale_color_discrete(guide = "none") +
  ggtitle(paste("Vg=", paste0(round(var(u), 2), collapse = " "),
                "Vr=", paste0(round(apply(rt, 2, var), 2), collapse = " ")),
          paste("theta =", paste0(round(theta, 2), collapse = " ")))
growthanimal2 <- stan_model("symcapture_models/GrowthAnimalTest2.stan")
fitGrowthAnimal2 <- sampling(growthanimal2, chains = 2, save_warmup = F,
                             data = list(N = nrow(data), I = max(data$i), y = data$y,
                                         x = data $x, ind = data$i, K = K, L = c(0.1, 0, 0.1), U = c(2, 200, 3)))
broom::tidyMCMC(fitGrowthAnimal2, pars = c("theta", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitGrowthAnimal2, pars = c("theta", "sigmaG", "sigmaR")))
mcmc_pairs(as.array(fitGrowthAnimal2, pars = c("theta", "sigmaG", "sigmaR")))
```

## Test2 - Symphonia Kinship

```{r kinship2}
path <- "~/Documents/BIOGECO/PhD/data/Symphonia_Paracou/Sequences/populationGenomics/"
fam <- read_delim(file.path(path, "gemma", "Gmax0", "paracou3pop.fam"), delim = " ",
                  col_names = c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE"))
inds <- unique(trees$IID)
K <- read_tsv(file.path(path, "..", "variantCalling", "paracou", "out.relatedness2"))  %>% 
  reshape2::dcast(INDV1 ~ INDV2, value.var = "RELATEDNESS_PHI") %>% 
  column_to_rownames("INDV1") %>% 
  as.matrix()
K <- K[inds, inds]
K[K < 0] <- 0
K <- K*2
K <- as.matrix(nearPD(K)$mat)
heatmap(K)
```

### Genotype

We will use the following animal model to estimate variance associated to the genotype without population variance at first:

$$y \sim \mathcal N(\mu + A.a,\sigma_R)~|~a \sim \mathcal N(0,\sigma_G)$$

```{r genotype2}
sigmaY <- 1
sigmaG <- 0.3
sigmaR <- sigmaY - sigmaG
N <- nrow(K)
mu <- rnorm(1)
u <- rnorm(N, sd = sqrt(sigmaG))
r <- rnorm(N, sd = sqrt(sigmaR))
y <- mu + as.vector(t(chol(K)) %*% u) + r
ggplot(data.frame(y = y, u = u), aes(u, y)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  ggtitle(paste("Vy =", round(var(y), 2), 
                "mu =", round(mu, 2),
                "Vg=", round(var(as.vector((t(chol(K)) %*% u))), 2), 
                "Vr=", round(var(r), 2)))
animal1 <- stan_model("symcapture_models/Animal1.stan")
fitAnimal1 <- sampling(animal1, chains = 2, save_warmup = F, data = list(N = N, Y = y, K = K))
broom::tidyMCMC(fitAnimal1, pars = c("mu", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitAnimal1, pars = c("mu", "sigmaG", "sigmaR")))
```

### Population & Genotype

We will use the following animal model to estimate variance associated to the genotype with population variance:

$$y \sim \mathcal N(\mu_p + A.a,\sigma_R)~|~a \sim \mathcal N(o,\sigma_G)$$

```{r pop2}
pop <- dplyr::select(trees, IID, pop) %>% unique() %>% 
  mutate(popNum = as.numeric(as.factor(pop))) %>% dplyr::select(popNum) %>% unlist()
P <- max(pop)
sigmaP <- 0.5
sigmaG <- 0.3
sigmaR <- 0.2
N <- nrow(K)
p <- rnorm(P, sd = sqrt(sigmaP))
u <- rnorm(N, sd = sqrt(sigmaG))
r <- rnorm(N, sd = sqrt(sigmaR))
y <- p[pop] + as.vector(t(chol(K)) %*% u) + r
animal2 <- stan_model("symcapture_models/Animal2.stan")
fitAnimal2 <- sampling(animal2, chains = 2, save_warmup = F, 
                       data = list(N = N, Y = y, K = K, P = P, population = pop))
ggplot(data.frame(y = y, u = u, pop = as.factor(pop)),
       aes(u, y, col = pop)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  ggtitle(paste("Vy =", round(var(y), 2), 
                "Vp =", round(var(p), 2), 
                "Vg=", round(var(as.vector((t(chol(K)) %*% u))), 2), 
                "Vr=", round(var(r), 2)),
          paste("p =", paste0(round(p, 2), collapse = " ")))
broom::tidyMCMC(fitAnimal2, pars = c("mu", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitAnimal2, pars = c("mu", "sigmaG", "sigmaR")))
```

