```{r setup_growthanimal, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(raster)
library(bayesplot)
library(kinship2)
library(rstan)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = T, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
load("./symcapture_save/Env2.Rdata")
trees <- trees %>% 
  na.omit(pop) %>% # remove admixed
  dplyr::select(Ind, IID, pop, Plot, SubPlot, TreeFieldNum, Y0, DBH0, DBHtoday)
pathCluster <- "~/Remotes/genotoul/work/PhD/documents/genomics/symcapture"
```

# Growth & Animal models

## Introduction

The aim of this model is to explore animal and growth model building theoritically in order to validate its behavior and use it on *Symphonia* and *Eschweilera* data. 

## Tes 1 - Simulated Kinship

Let's consider a set of $P$ populations including each $Fam$ families composed of $I = 14$ individuals with arbitrar relationships.

```{r kinship1, fig.cap="Kinship matrix"}
kin.df <- function(P= 3, Fam = 6){
  I <- 14 # Individuals
  data.frame(pop = rep(1:P, each = Fam*I),
                       fam = rep(1:Fam, each = I),
                       ind = rep(1:I, Fam*P),
                       father = rep(c(NA, NA, 1, 1, 1, 1, NA, 6, 6, 6, NA, 11, 11, 11), Fam*P),
                       mother = rep(c(NA, NA, 2, 2, 2, 2, NA, 7, 7, 7, NA, 3, 3, 3), Fam*P), 
                       sex = rep(c(1, 2, 2, 3, 3, 1, 2, 3, 3, 3, 1, 3, 3, 3), Fam*P)) %>% 
    mutate_at(c("father", "mother"), funs(ifelse(!is.na(.), paste0(pop,fam,.), NA))) %>% 
    mutate(ind = paste0(pop,fam,ind))
}
kin.mat <- function(P = 3, Fam = 6){
  ped.df <- kin.df(P, Fam)
  ped.ped <- pedigree(id = ped.df$ind, dadid = ped.df$father, momid = ped.df$mother, sex = ped.df$sex, famid = paste0(ped.df$pop, ped.df$fam))
  K <- as.matrix(kinship(ped.ped))
  K <- 2*K
}
heatmap(kin.mat(3, 6))
```

```{r gdata}
gdata <- function(
  I = 20, # Nb ind
  Y = 33, # Nb years
  P = 3, # Nb pops
  Gmax = c(0.54, 0.53, 0.44), # model parameters
  Dopt =  c(27, 31, 20),
  Ks = c(0.67, 0.73, 0.64),
  sigmaR = c(0.16, 9, 0.22),
  sigma = 0.14
){
  Gmax <- Gmax[1:P]
  Dopt <- Dopt[1:P]
  Ks <- Ks[1:P]
  years <- 1985:(1985+Y-1)
  pop <- sample(1:P, I, replace = T)
  DBH0 <- rlnorm(I, meanlog = mean(log(trees$DBH0)), 
                 sdlog = sd(log(trees$DBH0))) 
  Y0 <- sample(trees$Y0, I)
  # Gmaxi <- rlnorm(I, mean = log(Gmax[pop]), sd = sqrt(sigmaR[1]))
  # Dopti <- rnorm(I, mean = Dopt[pop], sd = sqrt(sigmaR[2]))
  # Ksi <- rlnorm(I, mean = log(Ks[pop]), sd = sqrt(sigmaR[3]))
  Gmaxi <- truncnorm::rtruncnorm(I, mean = Gmax[pop], sd = sqrt(sigmaR[1]), a = 0)
  Dopti <- truncnorm::rtruncnorm(I, mean = Dopt[pop], sd = sqrt(sigmaR[2]), a = 0)
  Ksi <- truncnorm::rtruncnorm(I, mean = Ks[pop], sd = sqrt(sigmaR[3]), a = 0)
  DBH <- rep(NA, I)
  for (t in 1:(Y-1)) {
    for(i in 1:I){
      if(years[t] == Y0[i]) 
        DBH[i] <- DBH0[i] 
    }
    DBH <- DBH + Gmaxi * exp(-0.5*(log(DBH / Dopti) / Ksi) * (log(DBH / Dopti) / Ksi))
  }
  DBHtoday <- rlnorm(I, log(DBH), sdlog = sigma)
  list(params = list(Gmax = Gmax, Dopt = Dopt, Ks = Ks, 
                     sigmaR = sigmaR, sigma = sigma),
       data = data.frame(ind = 1:I, pop = pop, 
                         DBH0 = DBH0, Y0 = Y0, Y = Y, DBHtoday = DBHtoday),
       mdata = list(I = I, Y = Y, P = P, years = years, DBH0 = DBH0,
                    Y0 = Y0, DBHtoday = DBHtoday, pop = pop))
}
```

```{r gkdata}
gkdata <- function(
  P = 3, # Nb pops
  Fam = 6, # Nb family
  Y = 33, # Nb years
  Gmax = c(0.54, 0.53, 0.44), # model parameters
  Dopt =  c(27, 31, 20),
  Ks = c(0.67, 0.73, 0.64),
  sigmaR = c(0.16, 9, 0.22),
  sigmaG = c(0.37, 25, 0.38),
  sigma = 0.14
){
  K <- kin.mat(P, Fam)
  Kdf <- kin.df(P, Fam)
  I <- nrow(K)
  Gmax <- Gmax[1:P]
  Dopt <- Dopt[1:P]
  Ks <- Ks[1:P]
  years <- 1984:(1984+Y-1)
  pop <- sample(1:P, I, replace = T)
  DBH0 <- rlnorm(I, meanlog = mean(log(trees$DBH0)), 
                 sdlog = sd(log(trees$DBH0))) 
  Y0 <- sample(trees$Y0, I)
  uG <- truncnorm::rtruncnorm(I, sd = sqrt(sigmaG[1]), a = -Gmax[Kdf$pop])
  uD <- truncnorm::rtruncnorm(I, sd = sqrt(sigmaG[2]), a = -Dopt[Kdf$pop])
  uK <- truncnorm::rtruncnorm(I, sd = sqrt(sigmaG[3]), a = -Ks[Kdf$pop])
  Gmaxi <- truncnorm::rtruncnorm(I, mean = Gmax[Kdf$pop] + as.vector(t(chol(K)) %*% uG),
                                 sd = sqrt(sigmaR[1]), a = 0)
  Dopti <- truncnorm::rtruncnorm(I, mean = Dopt[Kdf$pop] + as.vector(t(chol(K)) %*% uD), 
                                 sd = sqrt(sigmaR[2]), a = 0)
  Ksi <- truncnorm::rtruncnorm(I, mean = Ks[Kdf$pop] + as.vector(t(chol(K)) %*% uK), 
                               sd = sqrt(sigmaR[3]), a = 0)
  DBH <- rep(NA, I)
  for (t in 1:(Y-1)) {
    for(i in 1:I){
      if(years[t] == Y0[i]) 
        DBH[i] <- DBH0[i] 
    }
    DBH <- DBH + Gmaxi * exp(-0.5*(log(DBH / Dopti) / Ksi) * (log(DBH / Dopti) / Ksi))
  }
  DBHtoday <- rlnorm(I, log(DBH), sdlog = sigma)
  list(params = list(Gmax = Gmax, Dopt = Dopt, Ks = Ks, 
                     sigmaR = sigmaR, sigmaG = sigmaG, sigma = sigma),
       data = data.frame(ind = 1:I, pop = pop, 
                         DBH0 = DBH0, Y0 = Y0, Y = Y, DBHtoday = DBHtoday),
       mdata = list(I = I, Y = Y, P = P, years = years, DBH0 = DBH0,
                    Y0 = Y0, DBHtoday = DBHtoday, pop = pop, K = K))
}
```

### Genotype

We will use the following animal model to estimate variance associated to the genotype without population variance at first:

$$y \sim \mathcal N(\mu + A.a,\sigma_R)~|~a \sim \mathcal N(0,\sigma_G)$$

```{r genotype1}
K <- kin.mat(1, 6)
sigmaY <- 1
sigmaG <- 0.3
sigmaR <- sigmaY - sigmaG
N <- nrow(K)
mu <- rnorm(1)
u <- rnorm(N, sd = sqrt(sigmaG))
r <- rnorm(N, sd = sqrt(sigmaR))
y <- mu + as.vector(t(chol(K)) %*% u) + r
ggplot(data.frame(y = y, u = u), aes(u, y)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  ggtitle(paste("Vy =", round(var(y), 2), 
                "mu =", round(mu, 2),
                "Vg=", round(var(as.vector((t(chol(K)) %*% u))), 2), 
                "Vr=", round(var(r), 2)))
animal1 <- stan_model("symcapture_models/Animal1.stan")
fitAnimal1 <- sampling(animal1, chains = 2, save_warmup = F, data = list(N = N, Y = y, K = K))
broom::tidyMCMC(fitAnimal1, pars = c("mu", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitAnimal1, pars = c("mu", "sigmaG", "sigmaR")))
```

### Population & Genotype

We will use the following animal model to estimate variance associated to the genotype with population variance:

$$y \sim \mathcal N(\mu_p + A.a,\sigma_R)~|~a \sim \mathcal N(o,\sigma_G)$$

```{r pop1}
P = 3
Fam = 6
K <- kin.mat(P, Fam)
ped.df <- kin.df(P, Fam)
sigmaP <- 0.8
sigmaG <- 0.8
sigmaR <- 0.2
N <- nrow(K)
p <- rnorm(P, sd = sqrt(sigmaP))
u <- rnorm(N, sd = sqrt(sigmaG))
r <- rnorm(N, sd = sqrt(sigmaR))
y <- p[ped.df$pop] + as.vector(t(chol(K)) %*% u) + r
ggplot(data.frame(y = y, u = u, pop = as.factor(ped.df$pop)),
       aes(u, y, col = pop)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  ggtitle(paste("Vy =", round(var(y), 2), 
                "Vp =", round(var(p), 2), 
                "Vg=", round(var(as.vector((t(chol(K)) %*% u))), 2), 
                "Vr=", round(var(r), 2)),
          paste("p =", paste0(round(p, 2), collapse = " ")))
animal2 <- stan_model("symcapture_models/Animal2.stan")
fitAnimal2 <- sampling(animal2, chains = 2, save_warmup = F, 
                       data = list(N = N, Y = y, K = K, P = P, population = ped.df$pop))
broom::tidyMCMC(fitAnimal2, pars = c("mu", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitAnimal2, pars = c("mu", "sigmaG", "sigmaR")))
```

### Growth

```{r growth1}
# mdata <- gdata(I = 10*7, P = 1)
# save(mdata, file = file.path("symcapture_save", "growth5.Rdata"))
# file.copy("./symcapture_save/growth5.Rdata",
#           file.path(pathCluster, "symcapture_save"), overwrite = T)
# file.copy("./symcapture_models/Growth5.stan",
#           file.path(pathCluster, "symcapture_models"), overwrite = T)
# load(file = file.path("symcapture_save", "growth5.Rdata"))  # Begin cluster
# growth1 <- stan_model("./symcapture_models/Growth5.stan")
# fitGrowth1 <- sampling(growth1, chains = 2, data = mdata$mdata, save_warmup = F,
#                       include = F, pars = c("DBH", "Gmaxi", "Dopti", "Ksi"))
# save(mdata, fitGrowth1, file = file.path("symcapture_save", "growth5.Rdata"))  # End cluster
# file.copy(file.path(pathCluster, "symcapture_save"),
#           "./symcapture_save/growth5.Rdata", overwrite = T)
load(file = file.path("symcapture_save", "growth5.Rdata"))
ggplot(mdata$data, aes(x = DBH0, y = DBHtoday, col = Y0, shape = as.factor(pop))) + 
  geom_point() + geom_abline()
broom::tidyMCMC(fitGrowth1, droppars = NULL, rhat = T) %>% kable()
mcmc_trace(fitGrowth1) + 
  ggtitle(paste(unlist(mdata$params), collapse = " "),
          paste(names(unlist(mdata$params)), collapse = " "))
```

### Growth & Genotype

```{r growthK1, eval=F}
mdata <- gkdata(Fam = 3, P = 1)
ggplot(mdata$data, aes(x = DBH0, y = DBHtoday, col = Y0, shape = as.factor(pop))) + 
  geom_point() + geom_abline()
save(mdata, file = file.path("symcapture_save", "growthgeno1.Rdata"))
file.copy("./symcapture_save/growthgeno1.Rdata",
          file.path(pathCluster, "symcapture_save"), overwrite = T)
file.copy("./symcapture_models/GrowthGeno1.stan",
          file.path(pathCluster, "symcapture_models"), overwrite = T)
load(file = file.path("symcapture_save", "growthgeno1.Rdata")) # Begin cluster
growthgeno1 <- stan_model("./symcapture_models/GrowthGeno1.stan")
fitGrowthGeno1 <- sampling(growthgeno1, chains = 2, data = mdata$mdata, save_warmup = F,
                           include = F, pars = c("DBH", "Gmaxi", "Dopti", "Ksi",
                                                 "aG", "aK", "aD"))
save(mdata, fitGrowthGeno1, file = file.path("symcapture_save", "GrowthGeno1.stan")) # End cluster
file.copy(file.path(pathCluster, "symcapture_save"),
          "./symcapture_save/growthgeno1.Rdata", overwrite = T)
load(file = file.path("symcapture_save", "growth5.Rdata"))
unlist(mdata$params)
broom::tidyMCMC(fitGrowthGeno1, droppars = NULL, rhat = T) %>% kable()
mcmc_trace(fitGrowthGeno1) + 
  ggtitle(paste(unlist(mdata$params), collapse = " "),
          paste(names(unlist(mdata$params)), collapse = " "))
```

## Test2 - Symphonia Kinship

```{r kinship2}
path <- "~/Documents/BIOGECO/PhD/data/Symphonia_Paracou/Sequences/populationGenomics/"
fam <- read_delim(file.path(path, "gemma", "Gmax0", "paracou3pop.fam"), delim = " ",
                  col_names = c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE"))
inds <- unique(trees$IID)
K <- read_tsv(file.path(path, "..", "variantCalling", "paracou", "out.relatedness2"))  %>% 
  reshape2::dcast(INDV1 ~ INDV2, value.var = "RELATEDNESS_PHI") %>% 
  column_to_rownames("INDV1") %>% 
  as.matrix()
K <- K[inds, inds]
K[K < 0] <- 0
K <- K*2
K <- as.matrix(nearPD(K)$mat)
heatmap(K)
```

### Genotype

We will use the following animal model to estimate variance associated to the genotype without population variance at first:

$$y \sim \mathcal N(\mu + A.a,\sigma_R)~|~a \sim \mathcal N(0,\sigma_G)$$

```{r genotype2}
sigmaY <- 1
sigmaG <- 0.3
sigmaR <- sigmaY - sigmaG
N <- nrow(K)
mu <- rnorm(1)
u <- rnorm(N, sd = sqrt(sigmaG))
r <- rnorm(N, sd = sqrt(sigmaR))
y <- mu + as.vector(t(chol(K)) %*% u) + r
ggplot(data.frame(y = y, u = u), aes(u, y)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  ggtitle(paste("Vy =", round(var(y), 2), 
                "mu =", round(mu, 2),
                "Vg=", round(var(as.vector((t(chol(K)) %*% u))), 2), 
                "Vr=", round(var(r), 2)))
animal3 <- stan_model("symcapture_models/Animal1.stan")
fitAnimal1 <- sampling(animal1, chains = 2, save_warmup = F, data = list(N = N, Y = y, K = K))
broom::tidyMCMC(fitAnimal1, pars = c("mu", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitAnimal1, pars = c("mu", "sigmaG", "sigmaR")))
```

### Population & Genotype

We will use the following animal model to estimate variance associated to the genotype with population variance:

$$y \sim \mathcal N(\mu_p + A.a,\sigma_R)~|~a \sim \mathcal N(o,\sigma_G)$$

```{r pop2}
pop <- dplyr::select(trees, IID, pop) %>% unique() %>% 
  mutate(popNum = as.numeric(as.factor(pop))) %>% dplyr::select(popNum) %>% unlist()
P <- max(pop)
sigmaP <- 0.5
sigmaG <- 0.3
sigmaR <- 0.2
N <- nrow(K)
p <- rnorm(P, sd = sqrt(sigmaP))
u <- rnorm(N, sd = sqrt(sigmaG))
r <- rnorm(N, sd = sqrt(sigmaR))
y <- p[pop] + as.vector(t(chol(K)) %*% u) + r
animal2 <- stan_model("symcapture_models/Animal2.stan")
fitAnimal2 <- sampling(animal2, chains = 2, save_warmup = F, 
                       data = list(N = N, Y = y, K = K, P = P, population = pop))
ggplot(data.frame(y = y, u = u, pop = as.factor(pop)),
       aes(u, y, col = pop)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  ggtitle(paste("Vy =", round(var(y), 2), 
                "Vp =", round(var(p), 2), 
                "Vg=", round(var(as.vector((t(chol(K)) %*% u))), 2), 
                "Vr=", round(var(r), 2)),
          paste("p =", paste0(round(p, 2), collapse = " ")))
broom::tidyMCMC(fitAnimal2, pars = c("sigmaP", "sigmaG", "sigmaR"), droppars = NULL, rhat = T) %>% kable()
mcmc_trace(as.array(fitAnimal2, pars = c("sigmaP", "sigmaG", "sigmaR")))
```
