---
title: Installation for stan presentation
author: "Sylvain Schmitt"
output:
  bookdown::html_document2:
    theme: readable
    toc: true
    toc_float: yes
    number_sections: no
link-citations: yes
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = T, eval = F)
```


The aim of Thursday presentation is to get familiar to `stan` use through R with package `rstan` and associated packages. Consequently, you'll need to install everything and test installation before the presentation to try examples directly on your machine.

# `rstan`

Apparently depending on your operating system (Windows, Mac or Linux) the installation can be more or less complicated. Every explications are available on [`stan` wiki](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started). Anyway the final command should be this one:

```{r}
install.packages("rstan", repos = "https://cloud.r-project.org/", dependencies=TRUE)
```

> NB: omit the 's' in 'https' if you cannot handle https downloads

To test for `rstan` installation, use following commands:

```{r}
example("stan_model", package = "rstan", run.dontrun = TRUE)
```

# `rstan` associated packages

I will also introduce following packages that you can simply directly install through R:

* `bayesplot` for Bayesian output visualization using `ggplot2`
* `shinystan` for interactive `stan` output visualization
* `rstanarm` for Bayesian automatic regression modelling using `stan`
* `brms` Bayesian generalized multivariate non-linear multilevel models using `stan`

```{r}
install.packages(c("bayesplot", "shinystan", "rstanarm", "brms"))
```

# `greta`

I might do a short opening on `greta`. So if you want you can also install it.

```{r}
install.packages("greta")
```

# Conclusion

If you have any trouble with the installation, don't hesitate to ask me.
