```{r setupdata, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(lubridate)
library(ggfortify)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
data <- read_tsv("thickness_save/data.tsv")
```

```{r data, eval=F}
path <- "../../../data/Paracou/weather/"
# LT 
traits <- readxl::read_xlsx(path = file.path(path, "Measures_Symphonia.xlsx"), sheet = "AllTraits") %>% 
  dplyr::select(idTree, Plot, SubPlot, TreeFieldNum, Leaf, Morphotype, Bark, Dawkins, LT) %>% 
  na.omit()
individuals <- read_tsv(file.path(path, "Measures_Symphonia - Individuals.tsv")) %>% 
  mutate(date = dmy(paste(Date))) %>% 
  mutate(time = dmy_hm(paste(Date, Time))) %>% 
  dplyr::select(date, time, Plot, SubPlot, TreeFieldNum, Bark, Dawkins, Position)
data <- left_join(individuals, traits, by = c("Plot", "SubPlot", "TreeFieldNum", "Dawkins", "Bark"))
rm(individuals, traits)
## Paracou
paracou <- src_sqlite(file.path(path, "..", "trees", "Paracou.sqlite")) %>%
  tbl("Paracou") %>%
  filter(idTree %in% local(data$idTree)) %>%
  filter(Genus %in% c("Lecythis", "Eschweilera") & Plot %in% c(1, 6, 11, 13:15) & CensusYear == 2017 |
         Genus %in% c("Lecythis", "Eschweilera") & Plot == 16 & CensusYear == 2015 |
         Genus == "Symphonia" & CensusYear == 2015) %>%
  mutate(DBH = CircCorr/pi) %>% 
  collect()
data <- left_join(data, paracou)
rm(paracou)
# Weather
weather <- readxl::read_xlsx(path = file.path(path, "GX-METEO-2004 - 2018E.xlsx"), sheet = "Gxmeteo E") %>% 
  filter(Year == 2017, Month %in% 10:12)
weather <- weather %>% 
   mutate(hour = round(`heure+min`/100)) %>% 
  mutate(minutes = `heure+min`-100*hour) %>% 
  mutate(date = as_date(jour, origin = "2016-12-31")) %>% 
  mutate(hour = round(`heure+min`/100)) %>% 
  mutate(minutes = `heure+min`-100*hour) %>% 
  mutate(time = as_datetime(paste(date, hour, minutes, 0))) %>% 
  dplyr::rename(T55 = "Temp(55)", T30 = "Temp(30)", Hr55 = "Hr(55)", Hr30 = "Hr(30)", 
                windaspect = "Dir vent", windspeed = "Vit vent", rain = "Pluie", PET = "ETP penman (mm h-1) - OK") %>% 
  group_by(date) %>%
  mutate(rainday = sum(rain)) %>% 
  ungroup() %>% 
  mutate(rainfivedays = zoo::rollapplyr(rain, width = 240, FUN = sum, partial = T)) %>% 
  dplyr::select(date, time, T55, T30, Hr55, Hr30, Patm, windaspect, windspeed, rain, rainday, rainfivedays, vpd55, PET)
data <- left_join(data, weather, by = "date", suffix = c("tree", "weather")) %>% 
  mutate(dt = abs(timetree - timeweather)) %>% 
  group_by(idTree, Leaf) %>% 
  filter(dt == min(dt)) %>% 
  sample_n(1) %>% 
  ungroup()
rm(weather)
# REW
rew <- read_csv(file.path(path, "REW_1978-2017_Fabien.csv")) %>% 
  mutate(date = dmy(paste(day, month, year))) %>% 
  rename(rainfall = precipitation) %>% 
  dplyr::select(date, rainfall, REW)
data <- left_join(data, rew, by = "date")
rm(rew)
# save
write_tsv(data, path = "thickness_save/data.tsv")
```

# Weather data

This chapter introduce and explore environmental variables correlations and relations with leaf thickness. Leaf thickness is used per leaf (and not per tree or per plot).

## Origin

* `Measures_Symphonia.xlsx`: leaf thikcness and corresponding individuals data (My PhD)
* `GX-METEO-2004 - 2018E.xlsx`: Weather database from ECOFOG (Guyaflux eddy tower, Damien Bonal et Benoit Burban) 
* `REW_1978-2017_Fabien.csv`: Relative extractable water content computed with Fabien Wagner model

Tables have been joined with dates and shortest time difference into the `data.tsv` table.

## Variables

* **date**: tree sampling date YYYY-MM-DD
* **timetree**: tree sampling time YYYY-MM-DD hh:mm:ss
* **Plot**: Pracou plot id
* **SubPlot**: Paracou subplot id
* **TreeFieldNum**: tree field number in Paracou
* **Bark**: field identification of *Symphonia* morhpotype based on the bark during the first sampling campaign
* **Dawkins**: Dawkins index of tree exposure to light
* **Position**: sampled branched position in the tree
* **idTree**: tree id in the Paracou database
* **Leaf**: leaf id in the PhD database
* **Morphotype**: Pracou database identification of *Symphonia* morphotype
* **LT**: leaf thickness in $\mu m$
* **timeweather**: weather sampling time YYYY-MM-DD hh:mm:ss
* **T55**: Air temperature at 55 m in degree celsius
* **T30**: Air temperature at 30 m in degree celsius
* **Hr55**: Air humidity at 55 m in percent
* **Hr30**: Air humidity at 30 m in percent
* **Patm**: Atmospheric pressure in mbar
* **windaspect**: wind aspect in degree
* **windspeed**: wind speed in $m.s^{-1}$
* **rain**: rainfall in mm in the last half hour
* **rainday**: rainfall in mm during the day (current day)
* **rainfivedays**: rainfall in mm for the last five days (120 hours)
* **vpd55**: Vapour Pressure deficit in kPa
* **PET**: Potential evapotranspiration (Penman) in mm
* **dt**: time difference between tree and weather sampling in seconds
* **rainfall**: rainfall in mm from REW database
* **REW**: relative extractable water computed with Fabein Wagner model

## Covariations

Figures \@ref(fig:cor) and \@ref(fig:pca) showed variables covariations among them and with leaf thickness. There are a lot of positive and negative covariations, as expected. And leaf thickness (LT) is mainly correlated to relative extractible water (REW), atmospheric pressure (Patm) and cummulated rain over five days (rainfivedays).

```{r cor, fig.cap="Variables correlations. Pearson's r below the diagonal and corresponding color above the diagonal."}
dplyr::select(data, LT, DBH, T55, T30, Hr55, Hr30, Patm, windaspect, windspeed, rain, rainday, rainfivedays, vpd55, PET, rainfall, REW) %>% 
  na.omit() %>%
  cor() %>% 
  corrplot::corrplot.mixed()
```

```{r pca, fig.cap="Principal component analysis of variables."}
autoplot(princomp(~ LT + DBH + T55 + T30 + Hr55 + Hr30 + Patm + windaspect + windspeed + rain + rainday + rainfivedays + vpd55 +  PET + REW + rainfall,
                  data = data, cor = T),
           loadings.label.size = 6,
           loadings.label.colour = 'black', loadings.label.vjust = 1.1,
           loadings = T, loadings.label = T, loadings.colour = 'black') +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted")
```

## Relations

Figures \@ref(fig:lmplot) and \@ref(fig:oob) and table \@ref(tab:lmtab) showed variables single relation with leaf thickness (LT).
There are a lot of significant positive and negative relations. 
But relative extractible water (REW) show the strongest relation with leaf thickness (LT) with the highest variable importance in the random forest model (Fig. \@ref(fig:oob)) and the highest $R^2_{adj}$ (0.292, Tab. \@ref(tab:lmtab)).
Beware, the $R^2$ is much smaller than the 73% of variance explained by Émilie. 
But she worked with traits aggregated by trees and per day, whereas here data are used at the leaf level.
I thus suggest to build a hierarchical/mixed model with weather fixed effect though REW combined tree and species random effects.

```{r lmplot, fig.cap="Linear regressions per explanatory variables."}
dplyr::select(data,
              date, timetree, idTree, Leaf, Plot, SubPlot, TreeFieldNum, Morphotype, LT, 
              DBH, T55, T30, Hr55, Hr30, Patm, windaspect, windspeed, rain, rainday, rainfivedays, vpd55, PET, rainfall, REW) %>% 
  reshape2::melt(id.vars = c("date", "timetree", "idTree", "Leaf", "Plot", "SubPlot", "TreeFieldNum", "Morphotype", "LT")) %>% 
  ggplot(aes(x = value, y = LT)) +
  geom_point(alpha = 0.3) +
  facet_wrap(~ variable, scales = "free") +
  geom_smooth(method = "lm")
```

```{r lmtab}
a <- dplyr::select(data,
              date, timetree, idTree, Leaf, Plot, SubPlot, TreeFieldNum, Morphotype, LT,
              DBH, T55, T30, Hr55, Hr30, Patm, windaspect, windspeed, rain, rainday, rainfivedays, vpd55, PET, rainfall, REW) %>% 
  reshape2::melt(id.vars = c("date", "timetree", "idTree", "Leaf", "Plot", "SubPlot", "TreeFieldNum", "Morphotype", "LT")) %>% 
  group_by(variable) %>% 
  do(lm = lm(LT ~ value, data = .)) %>% 
  broom::tidy(lm) %>% 
  filter(term != "(Intercept)") %>% 
  dplyr::select(variable, p.value) %>% 
  arrange(p.value)
b <- dplyr::select(data,
              date, timetree, idTree, Leaf, Plot, SubPlot, TreeFieldNum, Morphotype, LT, 
              DBH, T55, T30, Hr55, Hr30, Patm, windaspect, windspeed, rain, rainday, rainfivedays, vpd55, PET, rainfall, REW) %>% 
  reshape2::melt(id.vars = c("date", "timetree", "idTree", "Leaf", "Plot", "SubPlot", "TreeFieldNum", "Morphotype", "LT")) %>% 
  group_by(variable) %>% 
  do(lm = lm(LT ~ value, data = .)) %>% 
  broom::glance(lm) %>% 
  dplyr::select(variable, r.squared, adj.r.squared)
left_join(a, b) %>% 
  arrange(desc(adj.r.squared)) %>% 
  dplyr::mutate_if(is.numeric, funs(as.character(signif(., 3)))) %>%
  kable(col.names = c("Variable", "$p$", "$R^2$", "$R^2_{adj}$"),
        caption = "Linear regressions per explanatory variables.")
```


```{r oob, fig.cap="Variable importance calculated with an out-of-bag method applied on a random forest. Variable importance is represented by the mean increase in mean square error (MSE)."}
randomForest::randomForest(LT ~ DBH + T55 + T30 + Hr55 + Hr30 + Patm + windaspect + windspeed + 
                             rain + rainday + rainfivedays + vpd55 +  PET + REW + rainfall,
                           data = data,  na.action = na.omit,
                           ntree = 1000, replace = F , importance = T,
                           do.trace = F, keep.forest = T, keep.inbag = T)$importance %>% 
  as.data.frame() %>% 
  rownames_to_column("variable") %>% 
  dplyr::rename(MSEinc = "%IncMSE") %>% 
  arrange(desc(MSEinc)) %>% 
  ggplot(aes(variable, MSEinc)) + 
  geom_col() +
  coord_flip() +
  xlab("") + ylab("Mean increase in mean square error.")
```
