```{r setupmodels, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
options(mc.cores = 4)
rstan_options(auto_write = TRUE)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
```


```{r mdata}
data <- read_tsv("thickness_save/data.tsv") %>% 
  mutate(Species = ifelse(Species == "Indet.",
                          c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G",
                                                                      "sp.1" =  "S")], Species)) %>% 
  filter(!is.na(Species)) %>% 
  mutate(SpeciesNum = as.numeric(as.factor(Species))) %>% 
  mutate(IndNum = as.numeric(as.factor(idTree))) %>%
  mutate(Species = recode(Species,
                        "globulifera"= "S. globulifera",
                        "sp.1" = "S. sp1"))
mdata <- list(
  L = nrow(data),
  S = max(data$SpeciesNum),
  I = max(data$IndNum),
  LT = data$LT,
  REW = as.vector(scale(data$REW)),
  logDBH = as.vector(scale(log(data$DBH))),
  species = data$SpeciesNum,
  individual = data$IndNum
)
```

# Models

This chapter develop all the models used to infer leaf thickness variation with relative extractible water.

## Summary

Summary of all models and results (list of variables and parameters, and details below). 
Without random effect the model with REW, DBH and species is the most likely and the likelihood increase a lot while adding individual random effects (Fig. \@ref(fig:lp)).
We will thus use the model with REW, DBH, species and individuals for variance partitionning (we don't seek an predictive model).
With few uncertainty, after individual random effect, REW is the main factor explainging leaf trait variation (Fig. \@raf(fig:varpart) & Fir. \@ref(fig:R2)).

```{r lp, fig.cap="Models log likelhood."}
load("thickness_save/fits.Rdata")
lapply(fits, mcmc_intervals_data, pars = "lp__") %>% 
  bind_rows(.id = "model") %>% 
  filter(!(model %in% c("RxSDSI", "LogRxSDSI"))) %>% 
  mutate(model = recode(model,
                        "R"= "REW",
                        "RS" = "REW + Species",
                        "RxS" = "REW.Species",
                        "RxSS" = "REW.Species + Species",
                        "D"= "DBH",
                        "DS" = "DBH + Species",
                        "DxS" = "DBH.Species",
                        "DxSS" = "DBH.Species + Species",
                        "RxSDS" = "REW.Species + DBH + Species"
  )) %>% 
  mutate(low = ifelse(m < 0, "low", "up")) %>% 
  ggplot(aes(x = model, 
             xend = model,
             color = model, 
             fill = model)) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  datatools::scale_x_reordered() +
  xaxis_title(F) +
  ylab("Log-likelihood") +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")
```

```{r Predictions, message=FALSE, fig.cap="Predictions for the selected model."}
load("thickness_save/fits.Rdata")
as.data.frame(fits$LogRxSDSI, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, REW)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = REW, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95, group = Species), fill = "lightgrey", col = NA) +
  geom_line(aes(y = LTm)) +
  scale_color_manual("", values = c("#0000EE", "#CD2626")) +
  theme(legend.position = "bottom", legend.text = element_text(face = "italic")) +
  xlab("Relative extractable water (REW)") +
  ylab("Leaf thickness (LT)")
```

```{r varpart, fig.cap="Variance partitionning for the selected model."}
load("thickness_save/fits.Rdata")
mcmc_intervals_data(fits$RxSDSI, regex_pars = c("V")) %>%
  mutate(variance = recode_factor(parameter,
                           "Vrew" = "REW", "Vdbh" = "DBH", "Vspecies" = "Species",
                           "Vindividual" = "Individual", "Vresidual" = "Residual")) %>%
  mutate(pct = paste0(round(m / sum(m) * 100), "%")) %>%
  ggplot(aes(x = "LT", fill = variance)) +
  geom_col(aes(y = m)) +
  geom_text(aes(y = m, label = pct), col = "white",
            position = position_stack(vjust = .5)) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.line.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_discrete(expression(sigma^2))
```

## Variables & parameters

Summary of all variables, laws, and parameters choice and used in subsequent models.

### Variables

* $l$ leaf or observation ($L$)
* $i$ individual ($I$)
* $s$ species ($S$)
* $LT$ leaf thickness (vary with $l$)
* $REW$ relative extractible water content (vary with $i$)
* $DBH$ diameter at breast height (vary with $i$)

### Parameters

* $\alpha$ leaf thickness origin
* $\alpha$ leaf thickness origin with species fixed effect
* $\beta_{REW}$ slope of REW effect
* $\beta_{DBH}$ slope of DBH effect
* $\sigma$ residual variation
* $\gamma_i$ individual random effect 
* $\sigma_I$ individual random effect variation

## Models

### $R$ Linear variation with REW

$$LT_{l,i,s} \sim \mathcal{N}(\alpha+\beta_{REW}.REW_i,\sigma)$$

```{r RSampling, include=FALSE}
modelR <- stan_model("thickness_models/R.stan")
fitR <- sampling(modelR, mdata)
```

```{r RTable}
broom::tidyMCMC(fitR, pars = c("alpha", "beta_rew", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r RTrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitR), pars = c("alpha", "beta_rew", "sigma"), np = nuts_params(fitR)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r RPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitR), pars = c("alpha", "beta_rew", "sigma"), np = nuts_params(fitR))
```

```{r RPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitR), regex_pars =  c("alpha", "beta"), np = nuts_params(fitR))
```

```{r RPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitR, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, REW)) %>%  
  unique() %>% 
  ggplot(aes(x = REW)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

### $R,S$ Linear variation with REW & species fixed effect on interecept

$$LT_{l,i,s} \sim \mathcal{N}(\alpha_s+\beta_{REW}.REW_i,\sigma)$$

```{r RSSampling, include=FALSE}
modelRS <- stan_model("thickness_models/RS.stan")
fitRS <- sampling(modelRS, mdata)
```

```{r RSTable}
broom::tidyMCMC(fitRS, pars = c("alpha", "beta_rew", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r RSTrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitRS), regex_pars = c("alpha", "beta_rew", "sigma"), np = nuts_params(fitRS)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r RSPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitRS), regex_pars = c("alpha", "beta_rew", "sigma"), np = nuts_params(fitRS))
```

```{r RSPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitRS), regex_pars =  c("alpha", "beta"), np = nuts_params(fitRS))
```

```{r RSPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitRS, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, REW)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = REW, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

### $RxS$ Linear variation with REW & species fixed effect on REW slope

$$LT_{l,i,s} \sim \mathcal{N}(\alpha + {\beta_{REW}}_s.REW_i,\sigma)$$

```{r RxSSampling, include=FALSE}
modelRxS <- stan_model("thickness_models/RxS.stan")
fitRxS <- sampling(modelRxS, mdata)
```

```{r RxSTable}
broom::tidyMCMC(fitRxS, pars = c("alpha", "beta_rew", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r RxSTrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitRxS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitRxS)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r RxSPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitRxS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitRxS))
```

```{r RxSPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitRxS), regex_pars =  c("alpha", "beta"), np = nuts_params(fitRxS))
```

```{r RxSPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitRxS, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, REW)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = REW, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

### $RxS,S$ Linear variation with REW & species fixed effect on intercept and REW slope

$$LT_{l,i,s} \sim \mathcal{N}(\alpha_s + {\beta_{REW}}_s.REW_i,\sigma)$$

```{r RxSSSampling, include=FALSE}
modelRxSS <- stan_model("thickness_models/RxSS.stan")
fitRxSS <- sampling(modelRxSS, mdata)
```

```{r RxSSTable}
broom::tidyMCMC(fitRxSS, pars = c("alpha", "beta_rew", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r RxSSTrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitRxSS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitRxSS)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r RxSSPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitRxSS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitRxSS))
```

```{r RxSSsPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitRxSS), regex_pars =  c("alpha", "beta"), np = nuts_params(fitRxSS))
```

```{r RxSSPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitRxSS, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, REW)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = REW, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

### $D$ Linear variation with logarithm of DBH

$$LT_{l,i,s} \sim \mathcal{N}(\alpha+\beta_{DBH}.log(DBH_i),\sigma)$$

```{r DSampling, include=FALSE}
modelD <- stan_model("thickness_models/D.stan")
fitD <- sampling(modelD, mdata)
```

```{r DTable}
broom::tidyMCMC(fitD, pars = c("alpha", "beta_dbh", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r DTrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitD), regex_pars =  c("alpha", "beta"), np = nuts_params(fitD)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r DPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitD), regex_pars =  c("alpha", "beta") , np = nuts_params(fitD))
```


```{r DPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitD), regex_pars =  c("alpha", "beta"), np = nuts_params(fitD))
```

```{r DPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitD, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, DBH)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = DBH, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

### $D,S$ Linear variation with DBH & species fixed effect on interecept

$$LT_{l,i,s} \sim \mathcal{N}(\alpha_s+\beta_{DBH}.log(DBH_i),\sigma)$$

```{r DSSampling, include=FALSE}
modelDS <- stan_model("thickness_models/DS.stan")
fitDS <- sampling(modelDS, mdata)
```

```{r DSTable}
broom::tidyMCMC(fitDS, pars = c("alpha", "beta_dbh", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r DSTrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitDS), regex_pars = c("alpha", "beta", "sigma"), np = nuts_params(fitDS)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r DSPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitDS), regex_pars = c("alpha", "beta", "sigma"), np = nuts_params(fitDS))
```

```{r DSPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitDS), regex_pars =  c("alpha", "beta"), np = nuts_params(fitDS))
```

```{r DSPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitDS, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, DBH)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = DBH, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

### $DxS$ Linear variation with DBH & species fixed effect on DBH slope

$$LT_{l,i,s} \sim \mathcal{N}(\alpha + {\beta_{DBH}}_s.log(DBH_i),\sigma)$$

```{r DxSSampling, include=FALSE}
modelDxS <- stan_model("thickness_models/DxS.stan")
fitDxS <- sampling(modelDxS, mdata)
```

```{r DxSTable}
broom::tidyMCMC(fitDxS, pars = c("alpha", "beta_dbh", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r DxSTrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitDxS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitDxS)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r DxSPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitDxS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitDxS))
```

```{r DxSPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitDxS), regex_pars =  c("alpha", "beta"), np = nuts_params(fitDxS))
```

```{r DxSPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitDxS, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, DBH)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = DBH, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

### $DxS,S$ Linear variation with DBH & species fixed effect on intercept and DBH slope

$$LT_{l,i,s} \sim \mathcal{N}(\alpha_s + {\beta_{DBH}}_s.log(DBH_i),\sigma)$$

```{r DxSSSampling, include=FALSE}
modelDxSS <- stan_model("thickness_models/DxSS.stan")
fitDxSS <- sampling(modelDxSS, mdata)
```

```{r DxSSTable}
broom::tidyMCMC(fitDxSS, pars = c("alpha", "beta_dbh", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r DxSSTrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitDxSS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitDxSS)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r DxSSPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitDxSS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitDxSS))
```

```{r DxSSsPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitDxSS), regex_pars =  c("alpha", "beta"), np = nuts_params(fitDxSS))
```

```{r DxSSPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitDxSS, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, DBH)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = DBH, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

### $RxS,D,S$ Linear variation with REW and DBH & species fixed effect on intercept and REW slope

$$LT_{l,i,s} \sim \mathcal{N}(\alpha_s + {\beta_{REW}}_s.REW_i + \beta_{DBH}.log(DBH_i),\sigma)$$

```{r RxSDSSampling, include=FALSE}
modelRxSDS <- stan_model("thickness_models/RxSDS.stan")
fitRxSDS <- sampling(modelRxSDS, mdata)
```

```{r RxSDSTable}
broom::tidyMCMC(fitRxSDS, pars = c("alpha", "beta_rew", "beta_dbh", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r RxSDSTrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitRxSDS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitRxSDS)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r RxSDSPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitRxSDS), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitRxSDS))
```

```{r RxSDSPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitRxSDS), regex_pars =  c("alpha", "beta"), np = nuts_params(fitRxSDS))
```

```{r RxSDSPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitRxSDS, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, REW)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = REW, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

### $RxS,D,S,I$ Linear variation with REW and DBH & species fixed effect on intercept and REW slope & individual random effect

$$LT_{l,i,s} \sim \mathcal{N}(\alpha_s + \gamma_i + {\beta_{REW}}_s.REW_i + \beta_{DBH}.log(DBH_i),\sigma)$$
$$\gamma_i \sim \mathcal{N}(0,\sigma_I)$$

```{r RxSDSISampling, include=FALSE}
modelRxSDSI <- stan_model("thickness_models/RxSDSI.stan")
fitRxSDSI <- sampling(modelRxSDSI, mdata)
```

```{r RxSDSITable}
broom::tidyMCMC(fitRxSDSI, pars = c("alpha", "beta_rew", "beta_dbh", "sigmaI", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r RxSDSITrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitRxSDSI), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitRxSDSI)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r RxSDSIPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitRxSDSI), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitRxSDSI))
```

```{r RxSDSIPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitRxSDSI), regex_pars =  c("alpha", "beta"), np = nuts_params(fitRxSDSI))
```

```{r RxSDSIPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitRxSDSI, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, REW)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = REW, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

```{r  RxSDSIVariance, fig.cap="Model variance partitionning."}
mcmc_intervals_data(fitRxSDSI, regex_pars = c("V")) %>%
  mutate(variance = recode_factor(parameter,
                           "Vrew" = "REW", "Vdbh" = "DBH", "Vspecies" = "Species",
                           "Vindividual" = "Individual", "Vresidual" = "Residual")) %>%
  mutate(pct = paste0(round(m / sum(m) * 100), "%")) %>%
  ggplot(aes(x = "LT", fill = variance)) +
  geom_col(aes(y = m)) +
  geom_text(aes(y = m, label = pct), col = "white",
            position = position_stack(vjust = .5)) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.line.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_discrete(expression(sigma^2))
```

### $log(RxS,D,S,I)$ Linear variation with REW and DBH & species fixed effect on intercept and REW slope & individual random effect

$$LT_{l,i,s} \sim \mathcal{logN}(log(\alpha_s + \gamma_i + {\beta_{REW}}_s.REW_i + \beta_{DBH}.log(DBH_i)),\sigma)$$
$$\gamma_i \sim \mathcal{N}(0,\sigma_I)$$

```{r LogRxSDSISampling, include=FALSE}
# modeLogRxSDSI <- stan_model("thickness_models/LogRxSDSI.stan")
# fitLogRxSDSI <- sampling(modeLogRxSDSI, mdata) 
load("thickness_save/fits.Rdata")
fitLogRxSDSI <- fits$LogRxSDSI
```

```{r LogRxSDSITable}
broom::tidyMCMC(fitLogRxSDSI, pars = c("alpha", "beta_rew", "beta_dbh", "sigmaI", "sigma", "lp__"),
                droppars = NULL, ess = T, rhat = T, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error, rhat, ess) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error", "$\\hat R$", "$N_{eff}$")) 
```

```{r LogRxSDSITrace, fig.cap="Traceplot of model parameters."}
mcmc_trace(as.array(fitLogRxSDSI), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitLogRxSDSI)) +
  viridis::scale_color_viridis(discrete = T)
```

```{r LogRxSDSIPairs, message=FALSE, fig.cap="Pairs of model parameters."}
mcmc_pairs(as.array(fitLogRxSDSI), regex_pars =  c("alpha", "beta", "sigma"), np = nuts_params(fitLogRxSDSI))
```

```{r LogRxSDSIPosteriors, message=FALSE, fig.cap="Posteriors of model parameters."}
mcmc_intervals(as.array(fitLogRxSDSI), regex_pars =  c("alpha", "beta"), np = nuts_params(fitLogRxSDSI))
```

```{r LogRxSDSIPredictions, message=FALSE, fig.cap="Model predictions."}
as.data.frame(fitLogRxSDSI, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), "[:punct:]", convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, REW)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = REW, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95), fill = "lightgrey") +
  geom_line(aes(y = LTm))
```

```{r  LogRxSDSIVariance, fig.cap="Model variance partitionning."}
mcmc_intervals_data(fitLogRxSDSI, regex_pars = c("V")) %>%
  mutate(variance = recode_factor(parameter,
                           "Vrew" = "REW", "Vdbh" = "DBH", "Vspecies" = "Species",
                           "Vindividual" = "Individual", "Vresidual" = "Residual")) %>%
  mutate(pct = paste0(round(m / sum(m) * 100), "%")) %>%
  ggplot(aes(x = "LT", fill = variance)) +
  geom_col(aes(y = m)) +
  geom_text(aes(y = m, label = pct), col = "white",
            position = position_stack(vjust = .5)) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.line.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_discrete(expression(sigma^2))
```

<!-- saving -->

```{r fits, eval=F}
fits <- list(
  R = fitR,
  RS = fitRS,
  RxS = fitRxS,
  RxSS = fitRxSS,
  D = fitD,
  DS = fitDS,
  DxS = fitDxS,
  DxSS = fitDxSS,
  RxSDS = fitRxSDS,
  RxSDSI = fitRxSDSI,
  LogRxSDSI = fitLogRxSDSI
)
# fits <- list(
#   R = fits$R,
#   RS = fits$RS,
#   RxS = fits$RxS,
#   RxSS = fits$RxSS,
#   D = fits$D,
#   DS = fits$DS,
#   DxS = fits$DxS,
#   DxSS = fits$DxSS,
#   RxSDS = fits$RxSDS,
#   RxSDSI = fits$RxSDSI,
#   LogRxSDSI = fitLogRxSDSI
# )
save(fits, file = "thickness_save/fits.Rdata")
```

