---
title: "Environment Symphonia"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(raster)
library(leaflet)
library(tidyverse)
library(ggfortify)
# library(ggplot.acm)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Paracou"
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
```

# Intro

## intro

This document is an introduction to Paracou environmental data to be used in the PhD. We will first focus on *Symphonia globulifera* species (with morphotypes *globulifera* and *sp1*). Please see environmental paragraph of the PhD subject document first. We still introduce a brief summary of environmental data types.

* __Biotic interactions__ affect all resources availabilty through facilitation and competition among living organisms. We will focus on other tree individuals as interacting organisms, besides the large amount of living organisms also interacting with trees. Biotic interactions affect (i) light interception through shading between individual trees, responsible for the asymetric competition [e.g. @Ruger2011], (ii) water availability with water table movements [including both competition and facilitation, e.g. @Pretzsch2014], (iii) and nutrients availability and access, with both nutrients competition and facilitation (e.g. nitrogen with *Fabaceae* family). Biotic interactions are uneasy to approach but can be approximated through neighbouring indices [@Canham]. In addition shading is partly included in direct light measurement with spatial measurement such as Lidar.
* __Hydrology__, by definition, affect individuals access to water. Hydrology includes previously introduced direct measurement of water availability [water table depth and water logging , e.g. @ferry2010higher]. But hydrology also include indirect factors, such as river and other hydrographic networks. Finally, hydrology is intimately related to topography, and thus can be derived from topographic variables (see below).
* __Pedology__ affect both individuals access to water and nutrients [@Levins1971]. Pedology variables can be direct measurement of nutrients or soil property, which will affect trees access to nutrient and water. Consequently, soil properties will be proxies for both water and nutrient access.
* __Topography__ affect both individuals access to water and nutrients [e.g. @gunatilleke_specieshabitat_2006]. Topography first encompass the elevation of the studied area and lot of topographic variables and indices can be derived from elevation (slope, aspect, curvature, roughness, ruggedness, position, water flow direction, water accumulation). Derived information will be proxies for water and nutrient access depending on the indice.

## variables

```{r env}
rbind(
  c('NCI', 'Neighborhood competition index', 'biotic', 'continuous'),
  c('DCM', 'Digital canopy model', 'biotic', 'continuous'),
  c('gap', 'Treefall gaps', 'biotic', 'discrete'),
  c('log gap', 'Logging gaps', 'biotic', 'discrete'),
  c('Idist', 'Disturbance index', 'biotic', 'continuous'),
  c('WTD', 'Water table depth', 'hydrology', 'discrete'),
  c('dist creek', 'Distance to nearest creek', 'hydrology', 'continuous'),
  c('drainages', 'Drainages type', 'hydrology', 'discrete'),
  c('soil', 'Soil chemical survey', 'pedology', 'continuous' ),
  c('hydromorphy', 'Hydromorphy', 'pedology', 'discrete'),
  c('waterlog', 'Water logging', 'pedology', 'discrete'),
  c('DEM', 'Digital elevation model', 'topography', 'continuous'),
  c('slope', 'Slope', 'topography', 'continuous'),
  c('aspect', 'Aspect', 'topography', 'continuous'),
  c('curvature', 'Curvature', 'topography', 'continuous'),
  c('TRI', 'Topographic roughness index', 'topography', 'continuous'),
  c('FlowDir', 'Water flow direction', 'topography', 'continuous'),
  c('wetness', 'Topographic wetness index', 'topography', 'continuous')
) %>% 
  data.frame() %>% 
  rename(code = X1, name = X2, type = X3, variable = X4) %>% 
  kable(caption = "Available environmental variables for Paracou.")
```

```{r sympho}
sympho <- src_sqlite(file.path(path, "trees/Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(Genus == "Symphonia") %>% 
  filter(CensusYear == 2015) %>% 
  filter(CodeAlive == 1) %>% 
  filter(CodeMeas == 0) %>% 
  filter(Plot != 18) %>% 
  mutate(morphotype = Species) %>% 
  collect()
symphoXY <- sympho
coordinates(symphoXY) <- ~Xutm + Yutm
proj4string(symphoXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
symphoXY <- spTransform(symphoXY, CRSobj = crs)
symphoPal <- colorFactor(c('darkgreen', 'black', 'firebrick'), as.factor(symphoXY$morphotype))
```

```{r basemap}
limits <- shapefile(file.path(path, "limits", "OverallPlots.shp"))
limits <- spTransform(limits, CRSobj = crs)
basemap <- leaflet() %>%
  addPolylines(data = limits, color = 'black', group = "limits") %>%
  addCircles(data = symphoXY, radius = ~Circ/pi/10, 
             color = ~symphoPal(morphotype), group = "Symphonia") %>% 
  addLegend(pal = symphoPal, title = 'Morphotype', 
            values = levels(as.factor(symphoXY$morphotype)),
            position = "bottomright") %>% 
  addLayersControl(overlayGroups = c('limits', 'Symphonia'),
                   options = layersControlOptions(collapsed = T))
```

# Biotic

## nci

> Neighborhood competition index 
$$NCI_i = \sum_j (\frac{DBH_j}{DBH_i})^\beta*e^{-\alpha*d_{i,j}} $$

### NCI

```{r nci map}

```

### Morphotype distribution with NCI

```{r nci comp}

```

## dcm

> Digital canopy model

### DCM

```{r dcm map}
dcm <- raster(file.path(path, "biotic", 
                        "MNC_ParacouAvril2009_1m.tif"))
dem <- raster(file.path(path, "topography", "DEM_1m_2015.tif")) # for CRS
projection(dcm) <- projection(dem)
dcm <- projectRaster(dcm, crs = crs)
basemap %>% 
  addRasterImage(aggregate(dcm, 4))
```

### Morphotype distribution with DCM

```{r dcm comp}
sympho$dcm <- raster::extract(dcm, symphoXY)
sympho %>% 
  ggplot(aes(dcm, col = morphotype, fill = morphotype)) +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  xlab('Digital canopy model')
```

## gap

> Treefall gaps

### gap

```{r gap map, eval=FALSE}
# gap <- dcm < 10
# gap <- crop(gap, extent(symphoXY))
# gap <- rasterToPolygons(gap, dissolve = T)
# save(gap, file = "./environment_save/gap.Rdata")
load("./environment_save/gap.Rdata")
gap <- gap[2,]
gap <- disaggregate(gap)
gap$area <- area(gap)
gap <- subset(gap, area > 4) # Value to be fixed
d_gap <- raster(extent(symphoXY), resolution = 10^-4, crs = crs)
d_gap <- setValues(d_gap, 0)
d_gap <-  mask(d_gap, gap)
d_gap <- distance(d_gap)
basemap %>% 
  addRasterImage(log(d_gap+1))
```

### Morphotype distribution with gap

```{r gap comp, eval=FALSE}
sympho$d_gap <- raster::extract(d_gap, symphoXY)
sympho %>% 
  ggplot(aes(d_gap, col = morphotype, fill = morphotype)) +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  scale_x_log10() +
  xlab('Distance to the closest treefall gap')
```

## log gap

> Logging gaps

### log gap

```{r log map}
log_gap <- shapefile(file.path(path, "logging", "Gaps.shp"))
log_gap <- spTransform(log_gap, CRSobj = crs)
d_log_gap <- raster(extent(symphoXY), resolution = 10^-4, crs = crs)
d_log_gap <- setValues(d_log_gap, 0)
d_log_gap <-  mask(d_log_gap, log_gap)
d_log_gap <- distance(d_log_gap)
basemap %>% 
  addRasterImage(log(d_log_gap+1))
```

### Morphotype distribution with log gap

```{r log comp}
sympho$d_log_gap <- raster::extract(d_log_gap, symphoXY)
sympho %>% 
  ggplot(aes(d_log_gap, col = morphotype, fill = morphotype)) +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  scale_x_log10() +
  xlab('Distance to the closest logging gap')
```

## Idist

> Disturbance index
$$Idist_i = \sum_j e^{-\alpha*d_{i,j}} * S^\ beta $$

### Idist

```{r Idist map}

```

### Morphotype distribution with Idist

```{r Idist comp}

```

# Hydrology

## wtd

> Water table depth

### WTD

```{r wtd map}
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
wtd <- spTransform(wtd, CRSobj = crs)
wtdPal <- colorFactor('Blues', as.factor(wtd$TypeWatEN), reverse = T)
basemap %>%
  addPolygons(data = wtd, opacity = 0, fillOpacity = 0.5, fill = T,
              fillColor = ~wtdPal(TypeWatEN), label = ~TypeWatEN)
```

### Morphotype distribution with log gap

```{r wtd comp}
sympho$wtd <- raster::extract(wtd, symphoXY)$TypeWatEN
sympho %>% 
  ggplot(aes(wtd, fill = morphotype)) +
  geom_bar(position = "dodge") +
  scale_y_log10() +
  xlab('Water table depth') +
  coord_flip()
```

## dist creek

> Distance to nearest creek

### dist creek

```{r dist creek map}
creek <- shapefile(file.path(path, "hydrology", "Creeks.shp"))
creek <- spTransform(creek, CRSobj = crs)
d_creek <- raster(extent(symphoXY), resolution = 10^-4, crs = crs)
d_creek <- setValues(d_creek, 0)
d_creek <-  mask(d_creek, creek)
d_creek <- distance(d_creek)
basemap %>% 
  addRasterImage(log(d_creek+1))
```

### Morphotype distribution with dist creek

```{r dist creek comp}
sympho$d_creek <- raster::extract(d_creek, symphoXY)
sympho %>% 
  ggplot(aes(d_creek, col = morphotype, fill = morphotype)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.05,
                 alpha = 0.1, position = "dodge") +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  scale_x_log10() +
  xlab('Distance to the closest creek')
```

## drainages

> Drainage type

### drainages

```{r drainages map}
drainages <- shapefile(file.path(path, "pedology", "Drainages.shp"))
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
proj4string(drainages) <- proj4string(wtd)
drainages <- spTransform(drainages, CRSobj = crs)
drainagesPal <- colorFactor('Blues', 
                            as.factor(drainages$TypDrainEN), 
                            reverse = T)
basemap %>%
  addPolygons(data = drainages, opacity = 0, fillOpacity = 0.5, 
              fill = T,
              fillColor = ~drainagesPal(TypDrainEN), 
              label = ~TypDrainEN)
```

### Morphotype distribution with drainages

```{r drainages comp}
sympho$drainages <- raster::extract(drainages, symphoXY)$TypDrainEN
sympho %>% 
  ggplot(aes(drainages, fill = morphotype)) +
  geom_bar(position = "dodge") +
  xlab('Drainage type') +
  coord_flip()
```

# Pedology

## soil

> Soil survey

### soil

```{r soil map}
soil <- shapefile(file.path(path, "pedology", "SoilSurvey.shp"))
soil@data %>% 
  dplyr::select(Refus, MO, C, N, C_N, Polsen, PBray2, Al_KCl, Mg_ech, K_ech, Na_ech, S, CEC, TS) %>% 
  prcomp(scale. = T) %>%
  autoplot(loadings = T, loadings.label = T)
```

### Morphotype distribution with soil

```{r soil comp}

```

## hydromorphy

> Hydromorphy

### hydromorphy

```{r hydromorphy map}
hydromorphy <- shapefile(file.path(path, "pedology", "Hydromorphy.shp"))
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
proj4string(hydromorphy) <- proj4string(wtd)
hydromorphy <- spTransform(hydromorphy, CRSobj = crs)
hydromorphyPal <- colorFactor('Blues', 
                            as.factor(hydromorphy$Hydro), 
                            reverse = T)
basemap %>%
  addPolygons(data = hydromorphy, opacity = 0, fillOpacity = 0.5, 
              fill = T,
              fillColor = ~hydromorphyPal(Hydro), 
              label = ~Hydro)
```

### Morphotype distribution with hydromorphy

```{r hydromorphy comp}
sympho$hydromorphy <- raster::extract(hydromorphy, symphoXY)$Hydro
sympho %>% 
  ggplot(aes(hydromorphy, fill = morphotype)) +
  geom_bar(position = "dodge") +
  xlab('Hydromorphy presence')
```

## waterlog

> Water logging

### waterlog

```{r waterlog map}
waterlog <- shapefile(file.path(path, "pedology", "Waterlogging.shp"))
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
proj4string(waterlog) <- proj4string(wtd)
waterlog <- spTransform(waterlog, CRSobj = crs)
waterlogPal <- colorFactor('Blues', 
                            as.factor(waterlog$LevelWater), 
                            reverse = T)
basemap %>%
  addPolygons(data = waterlog, opacity = 0, fillOpacity = 0.5, 
              fill = T,
              fillColor = ~waterlogPal(LevelWater), 
              label = ~LevelWater)
```

### Morphotype distribution with waterlog

```{r waterlog comp}
sympho$waterlog <- raster::extract(waterlog, symphoXY)$LevelWater
sympho %>% 
  ggplot(aes(waterlog, fill = morphotype)) +
  geom_bar(position = "dodge") +
  xlab('Water level from water logging') +
  coord_flip()
```

# Topography

## wetness

> Topographic wetness index

### Wetness

```{r wetness map}
wetness <- raster(file.path(path, "topography", "TWI_1m.tif"))
dem <- raster(file.path(path, "topography", "DEM_1m_2015.tif")) # for CRS
projection(wetness) <- projection(dem)
wetness <- projectRaster(wetness, crs = crs)
basemap %>% 
  addRasterImage(aggregate(wetness, 4))
```

### Morphotype distributions with wetness

```{r wetness comp}
sympho$wetness <- raster::extract(wetness, symphoXY)
sympho %>% 
  ggplot(aes(wetness, col = morphotype, fill = morphotype)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.05,
                 alpha = 0.1, position = "dodge") +
  geom_density(alpha = 0.2) +
  xlab('Tropographic wetness index')
```

## dem

> Digital elevation model

### DEM

```{r dem map}
dem <- raster(file.path(path, "topography", "DEM_1m_2015.tif"))
dem <- projectRaster(dem, crs = crs)
basemap %>% 
  addRasterImage(aggregate(dem, 4))
```

### Morphotype distribution with DEM

```{r dem comp}
sympho$dem <- raster::extract(dem, symphoXY)
sympho %>% 
  ggplot(aes(dem, col = morphotype, fill = morphotype)) +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  xlab('Digital elevation model')
```

## slope

> Digital slope model

### slope

```{r slope map}
slope <- terrain(dem, 'slope')
basemap %>% 
  addRasterImage(aggregate(slope, 4))
```

### Morphotype distribution with slope

```{r slope comp}
sympho$slope <- raster::extract(slope, symphoXY)
sympho %>% 
  ggplot(aes(slope, col = morphotype, fill = morphotype)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.01,
                 alpha = 0.1, position = "dodge") +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  xlab('Digital slope model')
```

## curvature

> Digital curvature model

### curvature

```{r curvature map}
curvature <- raster(file.path(path, "topography", "curvature_1m.tif"))
dem <- raster(file.path(path, "topography", "DEM_1m_2015.tif")) # for CRS
projection(curvature) <- projection(dem)
curvature <- projectRaster(curvature, crs = crs)
curvature <- crop(curvature, extent(symphoXY))
basemap %>% 
  addRasterImage(aggregate(curvature, 4))
```

### Morphotype distributions with curvature

```{r curvature comp}
sympho$curvature <- raster::extract(curvature, symphoXY)
sympho %>% 
  ggplot(aes(curvature, col = morphotype, fill = morphotype)) +
  geom_density(alpha = 0.2) +
  xlab('Digital curvature model')
```

## aspect

> Digital aspect model

### aspect

```{r aspect map}
aspect <- terrain(dem, 'aspect')
basemap %>% 
  addRasterImage(aggregate(aspect, 4))
```

### Morphotype distribution with aspect

```{r aspect comp}
sympho$aspect <- raster::extract(aspect, symphoXY)
sympho %>% 
  ggplot(aes(aspect, col = morphotype, fill = morphotype)) +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  xlab('Digital aspect model')
```

## TRI

> Topographic ruggedness index

### TRI

```{r TRI map}
TRI <- terrain(dem, 'TRI')
basemap %>% 
  addRasterImage(aggregate(TRI, 4))
```

### Morphotype distribution with TRI

```{r TRI comp}
sympho$TRI <- raster::extract(TRI, symphoXY)
sympho %>% 
  ggplot(aes(TRI, col = morphotype, fill = morphotype)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.05,
                 alpha = 0.1, position = "dodge") +
  ylim(0, 1) +
  xlab('Topographic ruggedness index')
```

# Correlations

```{r cor}
library(ade4)
ord <- sympho %>% 
  select(d_log_gap, wtd, d_creek, drainages, hydromorphy,
         waterlog, wetness, dem, slope, curvature, aspect, TRI) %>% 
  mutate(wtd = factor(wtd)) %>% 
  mutate(drainages = factor(drainages)) %>% 
  na.omit() %>% 
  dudi.mix(scannf = F, nf = 3)
data(deug)
deug.dudi <- dudi.pca(deug$tab, center = deug$cent, scale = FALSE, scan = FALSE)
class(ord) <- class(deug.dudi)
factoextra::fviz_pca_var(ord, col.var = "contrib", repel = TRUE,
                         gradient.cols = c("#00AFBB", "#E7B800",
                                           "#FC4E07"), c(1,2))
factoextra::fviz_pca_var(ord, col.var = "contrib", repel = TRUE,
                         gradient.cols = c("#00AFBB", "#E7B800",
                                           "#FC4E07"), c(1,3))
factoextra::fviz_pca_var(ord, col.var = "contrib", repel = TRUE,
                         gradient.cols = c("#00AFBB", "#E7B800",
                                           "#FC4E07"), c(2,3))
sympho %>% 
  model.matrix(~ d_log_gap + wtd + d_creek + drainages + hydromorphy + waterlog + wetness + dem + slope + curvature + aspect + TRI, 
               data = .) %>% 
  cor() %>% 
  corrplot::corrplot()
```

# References
