---
title: "Environment Eschweilera"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(raster)
library(leaflet)
library(tidyverse)
library(ggfortify)
# library(ggplot.acm)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Paracou"
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
```

# Intro

## intro

This document focus on *Eschweilera* genus relation to environment. Please see environmental paragraph of the PhD subject document first. We still introduce a brief summary of environmental data types.

* __Biotic interactions__ affect all resources availabilty through facilitation and competition among living organisms. We will focus on other tree individuals as interacting organisms, besides the large amount of living organisms also interacting with trees. Biotic interactions affect (i) light interception through shading between individual trees, responsible for the asymetric competition [e.g. @Ruger2011], (ii) water availability with water table movements [including both competition and facilitation, e.g. @Pretzsch2014], (iii) and nutrients availability and access, with both nutrients competition and facilitation (e.g. nitrogen with *Fabaceae* family). Biotic interactions are uneasy to approach but can be approximated through neighbouring indices [@Canham]. In addition shading is partly included in direct light measurement with spatial measurement such as Lidar.
* __Hydrology__, by definition, affect individuals access to water. Hydrology includes previously introduced direct measurement of water availability [water table depth and water logging , e.g. @ferry2010higher]. But hydrology also include indirect factors, such as river and other hydrographic networks. Finally, hydrology is intimately related to topography, and thus can be derived from topographic variables (see below).
* __Pedology__ affect both individuals access to water and nutrients [@Levins1971]. Pedology variables can be direct measurement of nutrients or soil property, which will affect trees access to nutrient and water. Consequently, soil properties will be proxies for both water and nutrient access.
* __Topography__ affect both individuals access to water and nutrients [e.g. @gunatilleke_specieshabitat_2006]. Topography first encompass the elevation of the studied area and lot of topographic variables and indices can be derived from elevation (slope, aspect, curvature, roughness, ruggedness, position, water flow direction, water accumulation). Derived information will be proxies for water and nutrient access depending on the indice.

## variables

```{r env}
rbind(
  c('NCI', 'Neighborhood competition index', 'biotic', 'continuous'),
  c('DCM', 'Digital canopy model', 'biotic', 'continuous'),
  c('gap', 'Treefall gaps', 'biotic', 'discrete'),
  c('log gap', 'Logging gaps', 'biotic', 'discrete'),
  c('Idist', 'Disturbance index', 'biotic', 'continuous'),
  c('WTD', 'Water table depth', 'hydrology', 'discrete'),
  c('dist creek', 'Distance to nearest creek', 'hydrology', 'continuous'),
  c('drainages', 'Drainages type', 'hydrology', 'discrete'),
  c('soil', 'Soil chemical survey', 'pedology', 'continuous' ),
  c('hydromorphy', 'Hydromorphy', 'pedology', 'discrete'),
  c('waterlog', 'Water logging', 'pedology', 'discrete'),
  c('DEM', 'Digital elevation model', 'topography', 'continuous'),
  c('relele', 'Relative elevation', 'topography', 'continuous'),
  c('slope', 'Slope', 'topography', 'continuous'),
  c('aspect', 'Aspect', 'topography', 'continuous'),
  c('curvature', 'Curvature', 'topography', 'continuous'),
  c('TRI', 'Topographic roughness index', 'topography', 'continuous'),
  c('FlowDir', 'Water flow direction', 'topography', 'continuous'),
  c('wetness', 'Topographic wetness index', 'topography', 'continuous')
) %>% 
  data.frame() %>% 
  rename(code = X1, name = X2, type = X3, variable = X4) %>% 
  kable(caption = "Available environmental variables for Paracou.")
```

```{r eschweilera}
eschweilera <- src_sqlite(file.path(path, "trees/Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(Genus == "Eschweilera") %>% 
  filter(CensusYear == 2015) %>% 
  filter(CodeAlive == 1) %>% 
  filter(CodeMeas == 0) %>% 
  filter(Plot != 18) %>%  
  filter(Species %in% c("sagotiana", "coriacea", "decolorans")) %>% 
  collect()
coordinates(eschweilera) <- ~Xutm + Yutm
proj4string(eschweilera) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
eschweilera <- spTransform(eschweilera, CRSobj = crs)
eschweileraPal <- colorFactor(RColorBrewer::brewer.pal(length(unique(eschweilera$Species)), "Accent"),
                         as.factor(eschweilera$Species))
```

```{r basemap}
limits <- shapefile(file.path(path, "limits", "OverallPlots.shp"))
limits <- spTransform(limits, CRSobj = crs)
basemap <- leaflet() %>%
  addPolylines(data = limits, color = 'black', group = "limits") %>%
    addCircles(data = eschweilera, radius = ~CircCorr/pi/10,  label = ~Species,
             color = ~eschweileraPal(Species), group = ~Species) %>% 
  addLegend(pal = eschweileraPal, title = 'Eschweilera', 
            values = levels(as.factor(eschweilera$Species)),
            position = "bottomright") %>% 
  addLayersControl(overlayGroups = c('limits', eschweilera$Species),
                   options = layersControlOptions(collapsed = T))
```

# Biotic

## nci

> Neighborhood competition index 
$$NCI_i = \sum_j (\frac{DBH_j}{DBH_i})^\beta*e^{-\alpha*d_{i,j}} $$

### NCI

```{r nci map}

```

### Species distribution with NCI

```{r nci comp}

```

## dcm

> Digital canopy model

### DCM

```{r dcm map}
dcm <- raster(file.path(path, "biotic", 
                        "MNC_ParacouAvril2009_1m.tif"))
dem <- raster(file.path(path, "topography", "DEM_1m_2015.tif")) # for CRS
projection(dcm) <- projection(dem)
dcm <- projectRaster(dcm, crs = crs)
basemap %>% 
  addRasterImage(aggregate(dcm, 4))
```

### Species distribution with DCM

```{r dcm comp}
eschweilera$dcm <- raster::extract(dcm, eschweilera)
eschweilera %>% 
  ggplot(aes(dcm, col = Species, fill = Species)) +
  geom_density(alpha = 0.2) +
  xlab('Digital canopy model')
```

## gap

> Treefall gaps

### gap

```{r gap map, eval=FALSE}
# gap <- dcm < 10
# gap <- crop(gap, extent(eschweilera))
# gap <- rasterToPolygons(gap, dissolve = T)
# save(gap, file = "./environment_save/gap.Rdata")
load("../Symphonia/environment_save/gap2.Rdata")
gap <- gap[2,]
gap <- disaggregate(gap)
gap$area <- area(gap)
gap <- subset(gap, area > 4) # Value to be fixed
d_gap <- raster(extent(eschweilera), resolution = 10^-4, crs = crs)
d_gap <- setValues(d_gap, 0)
d_gap <-  mask(d_gap, gap)
d_gap <- distance(d_gap)
basemap %>% 
  addRasterImage(log(d_gap+1))
```

### Species distribution with gap

```{r gap comp, eval=FALSE}
eschweilera$d_gap <- raster::extract(d_gap, eschweilera)
eschweilera %>% 
  ggplot(aes(d_gap, col = Species, fill = Species)) +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  scale_x_log10() +
  xlab('Distance to the closest treefall gap')
```

## log gap

> Logging gaps

### log gap

```{r log map}
log_gap <- shapefile(file.path(path, "logging", "Gaps.shp"))
log_gap <- spTransform(log_gap, CRSobj = crs)
d_log_gap <- raster(extent(eschweilera), resolution = 10^-4, crs = crs)
d_log_gap <- setValues(d_log_gap, 0)
d_log_gap <-  mask(d_log_gap, log_gap)
d_log_gap <- distance(d_log_gap)
basemap %>% 
  addRasterImage(log(d_log_gap+1))
```

### Species distribution with log gap

```{r log comp}
eschweilera$d_log_gap <- raster::extract(d_log_gap, eschweilera)
eschweilera %>% 
  ggplot(aes(d_log_gap, col = Species, fill = Species)) +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  scale_x_log10() +
  xlab('Distance to the closest logging gap')
```

## Idist

> Disturbance index
$$Idist_i = \sum_j e^{-\alpha*d_{i,j}} * S^\ beta $$

### Idist

```{r Idist map}

```

### Species distribution with Idist

```{r Idist comp}

```

# Hydrology

## wtd

> Water table depth

### WTD

```{r wtd map}
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
wtd <- spTransform(wtd, CRSobj = crs)
wtdPal <- colorFactor('Blues', as.factor(wtd$TypeWatEN), reverse = T)
basemap %>%
  addPolygons(data = wtd, opacity = 0, fillOpacity = 0.5, fill = T,
              fillColor = ~wtdPal(TypeWatEN), label = ~TypeWatEN)
```

### Species distribution with log gap

```{r wtd comp}
eschweilera$wtd <- raster::extract(wtd, eschweilera)$TypeWatEN
eschweilera %>% 
  ggplot(aes(wtd, fill = Species)) +
  geom_bar(position = "dodge") +
  scale_y_log10() +
  xlab('Water table depth') +
  coord_flip()
```

## dist creek

> Distance to nearest creek

### dist creek

```{r dist creek map}
creek <- shapefile(file.path(path, "hydrology", "Creeks.shp"))
creek <- spTransform(creek, CRSobj = crs)
d_creek <- raster(extent(eschweilera), resolution = 10^-4, crs = crs)
d_creek <- setValues(d_creek, 0)
d_creek <-  mask(d_creek, creek)
d_creek <- distance(d_creek)
basemap %>% 
  addRasterImage(log(d_creek+1))
```

### Species distribution with dist creek

```{r dist creek comp}
eschweilera$d_creek <- raster::extract(d_creek, eschweilera)
eschweilera %>% 
  ggplot(aes(d_creek, col = Species, fill = Species)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.05,
                 alpha = 0.1, position = "dodge") +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  scale_x_log10() +
  xlab('Distance to the closest creek')
```

## drainages

> Drainage type

### drainages

```{r drainages map}
drainages <- shapefile(file.path(path, "pedology", "Drainages.shp"))
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
proj4string(drainages) <- proj4string(wtd)
drainages <- spTransform(drainages, CRSobj = crs)
drainagesPal <- colorFactor('Blues', 
                            as.factor(drainages$TypDrainEN), 
                            reverse = T)
basemap %>%
  addPolygons(data = drainages, opacity = 0, fillOpacity = 0.5, 
              fill = T,
              fillColor = ~drainagesPal(TypDrainEN), 
              label = ~TypDrainEN)
```

### Species distribution with drainages

```{r drainages comp}
eschweilera$drainages <- raster::extract(drainages, eschweilera)$TypDrainEN[!duplicated(raster::extract(drainages, eschweilera)$point.ID)]
eschweilera %>% 
  ggplot(aes(drainages, fill = Species)) +
  geom_bar(position = "dodge") +
  xlab('Drainage type') +
  coord_flip() +
  scale_y_log10()
```

# Pedology

## soil

> Soil survey

### soil

```{r soil map}
soil <- shapefile(file.path(path, "pedology", "SoilSurvey.shp"))
soil@data %>% 
  select(Refus, MO, C, N, C_N, Polsen, PBray2, Al_KCl, Mg_ech, K_ech, Na_ech, S, CEC, TS) %>% 
  prcomp(scale. = T) %>%
  autoplot(loadings = T, loadings.label = T)
```

### Species distribution with soil

```{r soil comp}

```

## hydromorphy

> Hydromorphy

### hydromorphy

```{r hydromorphy map}
hydromorphy <- shapefile(file.path(path, "pedology", "Hydromorphy.shp"))
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
proj4string(hydromorphy) <- proj4string(wtd)
hydromorphy <- spTransform(hydromorphy, CRSobj = crs)
hydromorphyPal <- colorFactor('Blues', 
                            as.factor(hydromorphy$Hydro), 
                            reverse = T)
basemap %>%
  addPolygons(data = hydromorphy, opacity = 0, fillOpacity = 0.5, 
              fill = T,
              fillColor = ~hydromorphyPal(Hydro), 
              label = ~Hydro)
```

### Species distribution with hydromorphy

```{r hydromorphy comp}
eschweilera$hydromorphy <- raster::extract(hydromorphy, eschweilera)$Hydro
eschweilera %>% 
  ggplot(aes(hydromorphy, fill = Species)) +
  geom_bar(position = "dodge") +
  xlab('Hydromorphy presence') +
  scale_y_log10()
```

## waterlog

> Water logging

### waterlog

```{r waterlog map}
waterlog <- shapefile(file.path(path, "pedology", "Waterlogging.shp"))
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
proj4string(waterlog) <- proj4string(wtd)
waterlog <- spTransform(waterlog, CRSobj = crs)
waterlogPal <- colorFactor('Blues', 
                            as.factor(waterlog$LevelWater), 
                            reverse = T)
basemap %>%
  addPolygons(data = waterlog, opacity = 0, fillOpacity = 0.5, 
              fill = T,
              fillColor = ~waterlogPal(LevelWater), 
              label = ~LevelWater)
```

### Species distribution with waterlog

```{r waterlog comp}
eschweilera$waterlog <- raster::extract(waterlog, eschweilera)$LevelWater
eschweilera %>% 
  ggplot(aes(waterlog, fill = Species)) +
  geom_bar(position = "dodge") +
  xlab('Water level from water logging') +
  scale_y_log10() +
  coord_flip()
```

# Topography

## relele

> Relative elevation

### Relative elevation

```{r relele map}
relele <- raster(file.path(path, "topography", "RelativeElevation_1m.tif"))
dem <- raster(file.path(path, "topography", "DEM_1m_2015.tif"))
projection(relele) <- projection(dem)
relele <- projectRaster(relele, crs = crs)
basemap %>% 
  addRasterImage(aggregate(relele, 4))
```

### Species distributions with relative elevation

```{r relele comp}
eschweilera$relele <- raster::extract(relele, eschweilera)
eschweilera@data %>% 
  ggplot(aes(relele, col = Species, fill = Species)) +
  geom_density(alpha = 0.2) +
  xlab('Relative elevation')
```

## dem

> Digital elevation model

### DEM

```{r dem map}
dem <- raster(file.path(path, "topography", "DEM_1m_2015.tif"))
dem <- projectRaster(dem, crs = crs)
basemap %>% 
  addRasterImage(aggregate(dem, 4))
```

### Species distribution with DEM

```{r dem comp}
eschweilera$dem <- raster::extract(dem, eschweilera)
eschweilera %>% 
  ggplot(aes(dem, col = Species, fill = Species)) +
  geom_density(alpha = 0.2) +
  xlab('Digital elevation model')
```

## wetness

> Topographic wetness index

### Wetness

```{r wetness map}
wetness <- raster(file.path(path, "topography", "TWI_1m.tif"))
dem <- raster(file.path(path, "topography", "DEM_1m_2015.tif")) # for CRS
projection(wetness) <- projection(dem)
wetness <- projectRaster(wetness, crs = crs)
basemap %>% 
  addRasterImage(aggregate(wetness, 4))
```

### Species distributions with wetness

```{r wetness comp}
eschweilera$wetness <- raster::extract(wetness, eschweilera)
eschweilera %>% 
  ggplot(aes(wetness, col = Species, fill = Species)) +
  # geom_histogram(aes(y = ..density..), binwidth = 0.05,
  #                alpha = 0.1, position = "dodge") +
  geom_density(alpha = 0.2) +
  xlab('Tropographic wetness index')
```

## slope

> Digital slope model

### slope

```{r slope map}
slope <- terrain(dem, 'slope')
basemap %>% 
  addRasterImage(aggregate(slope, 4))
```

### Species distribution with slope

```{r slope comp}
eschweilera$slope <- raster::extract(slope, eschweilera)
eschweilera %>% 
  ggplot(aes(slope, col = Species, fill = Species)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.01,
                 alpha = 0.1, position = "dodge") +
  geom_density(alpha = 0.2) +
  ylim(0, 1) +
  xlab('Digital slope model')
```

## curvature

> Digital curvature model

### curvature

```{r curvature map}
curvature <- raster(file.path(path, "topography", "curvature_1m.tif"))
dem <- raster(file.path(path, "topography", "DEM_1m_2015.tif")) # for CRS
projection(curvature) <- projection(dem)
curvature <- projectRaster(curvature, crs = crs)
curvature <- crop(curvature, extent(eschweilera))
basemap %>% 
  addRasterImage(aggregate(curvature, 4))
```

### Species distributions with curvature

```{r curvature comp}
eschweilera$curvature <- raster::extract(curvature, eschweilera)
eschweilera %>% 
  ggplot(aes(curvature, col = Species, fill = Species)) +
  geom_density(alpha = 0.2) +
  xlab('Digital curvature model')
```

## aspect

> Digital aspect model

### aspect

```{r aspect map}
aspect <- terrain(dem, 'aspect')
basemap %>% 
  addRasterImage(aggregate(aspect, 4))
```

### Species distribution with aspect

```{r aspect comp}
eschweilera$aspect <- raster::extract(aspect, eschweilera)
eschweilera %>% 
  ggplot(aes(aspect, col = Species, fill = Species)) +
  geom_density(alpha = 0.2) +
  xlab('Digital aspect model')
```

## TRI

> Topographic ruggedness index

### TRI

```{r TRI map}
TRI <- terrain(dem, 'TRI')
basemap %>% 
  addRasterImage(aggregate(TRI, 4))
```

### Species distribution with TRI

```{r TRI comp}
eschweilera$TRI <- raster::extract(TRI, eschweilera)
eschweilera %>% 
  ggplot(aes(TRI, col = Species, fill = Species)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.05,
                 alpha = 0.1, position = "dodge") +
  ylim(0, 1) +
  xlab('Topographic ruggedness index')
```

# TWI-RE

## Correlation

### Correlation Gaph

```{r corTWIREGraph}
ggplot(eschweilera@data, aes(wetness, relele)) +
  geom_point() +
  geom_smooth() +
  theme_bw()
```

### Correlation Stat

```{r corTWIREStat}
cor.test(eschweilera$wetness, eschweilera$relele)
```

## Topography

### Topography Graph

```{r topoTWIREgraph}
dplyr::select(eschweilera@data, wetness, relele, dem, d_creek, slope, curvature) %>% 
  reshape2::melt(c("wetness", "relele")) %>% 
  reshape2::melt(c("variable", "value"), variable.name = "variable2", value.name = "value2") %>% 
  ggplot(aes(value2, value)) +
  geom_point() +
  facet_grid(variable ~ variable2, scales = "free") +
  geom_smooth(method = "lm") +
  theme_bw()
```

### Topography Stat

```{r topoTWIREstat}
dplyr::select(eschweilera@data, wetness, relele, dem, d_creek, slope, curvature) %>% 
  reshape2::melt(c("wetness", "relele")) %>% 
  group_by(variable) %>% 
  do(m= aov(lm(value ~ wetness + relele, data = .))) %>% 
  broom::tidy(m) %>% 
  mutate(R2 = round(sumsq/sum(sumsq)*100)) %>% 
  ggplot(aes(x = variable, fill = term)) +
  geom_col(aes(y = R2)) +
  geom_text(aes(y = R2, label = R2), col = "white", position = position_stack(vjust = .5)) +
  theme_bw()
```

## Soil

### Soil Graph

```{r soilTWIREgraph}
proj4string(soil) <- crs(dem)
soil$wetness <- raster::extract(wetness, soil)
soil$relele <- raster::extract(relele, soil)
dplyr::select(soil@data, wetness, relele, Refus, MO, 
              C, N, C_N, Polsen, PBray2, Al_KCl, Mg_ech, K_ech, Na_ech, S, CEC, TS) %>% 
  reshape2::melt(c("wetness", "relele")) %>% 
  reshape2::melt(c("variable", "value"), variable.name = "variable2", value.name = "value2") %>% 
  ggplot(aes(value2, value)) +
  geom_point() +
  facet_wrap(~ variable2 + variable, scales = "free") +
  geom_smooth(method = "lm") +
  theme_bw()
```

### Soil Statistics

```{r soilTWIREstat}
dplyr::select(soil@data, wetness, relele, Refus, MO, 
              C, N, C_N, Polsen, PBray2, Al_KCl, Mg_ech, K_ech, Na_ech, S, CEC, TS) %>% 
  reshape2::melt(c("wetness", "relele")) %>% 
  group_by(variable) %>% 
  do(m= aov(lm(value ~ wetness + relele, data = .))) %>% 
  broom::tidy(m) %>% 
  mutate(R2 = round(sumsq/sum(sumsq)*100)) %>% 
  ggplot(aes(x = variable, fill = term)) +
  geom_col(aes(y = R2)) +
  geom_text(aes(y = R2, label = R2), col = "white", position = position_stack(vjust = .5)) +
  theme_bw()
```

## Hydrology

### Hydrology Graph

```{r hydroTWIREgraph}
dplyr::select(eschweilera@data, wetness, relele, wtd, hydromorphy, waterlog) %>% 
  reshape2::melt(c("wetness", "relele")) %>% 
  reshape2::melt(c("variable", "value"), variable.name = "variable2", value.name = "value2") %>% 
  na.omit() %>% 
  ggplot(aes(value, value2)) +
  geom_boxplot() +
  coord_flip() +
  facet_grid(variable ~ variable2, scales = "free") +
  theme_bw()
```

### Hydrology Statistics

```{r hydroTWIREstat}
```


# References
