---
title: Supplementary Materials
author: Sylvain Schmitt, Niklas Tysklind, Géraldine Derroire, Myriam Heuertz, Bruno Herault
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: no
    toc_float: no
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::word_document2: default
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

\makeatletter
\renewcommand*{\thetable}{\arabic{table}}
\renewcommand*{\thefigure}{\arabic{figure}}
\let\c@table\c@figure
\makeatother 

\renewcommand{\figurename}{Supplementary Material S.}
\renewcommand{\tablename}{Supplementary Material S.}

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(ggfortify)
library(broom)
library(rstan)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
```

```{r trees}
d <- 20
trees <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(CensusYear == 2015) %>%
  filter(Species != "Indet.") %>% 
  mutate(DBH = CircCorr/pi) %>% 
  collect() %>% 
  filter(Xfield > d, Xfield < 250-d, Yfield > d, Yfield < 250-d)
load("../distribution_save/env.Rdata")
load("../distribution_save/Competition.Rdata")
Competition <- Competition %>% 
  left_join(dplyr::select(trees, idTree, Genus, Species)) %>% 
  group_by(idTree) %>% 
  mutate(dij = ifelse(dij <1, 1, dij)) %>% 
  summarise(NCIhetero = log(sum(DBHj^2*exp(-0.25*dij)*(1-as.numeric(Species == Speciesj)))+1),
            NCIcon = log(sum(DBHj^2*exp(-0.25*dij)*as.numeric(Species == Speciesj))+1))
complexes <- bind_rows(
  data.frame(Complex = "Chartacea", Genus = "Eschweilera",
             Species = c("simiorum", "congestiflora")),
  data.frame(Complex = "Parvifolia", Genus = "Eschweilera",
             Species = c("pedicellata", "coriacea", "decolorans", "sagotiana",
                         "wachenheimii", "grandiflora_form2")),
  data.frame(Complex = "Licania", Genus = "Licania",
             Species = c("alba", "membranacea", "canescens", "micrantha",
                         "ovalifolia", "sprucei", "densiflora",
                         "laxiflora", "parvifructa")),
  data.frame(Complex = "Iryanthera", Genus = "Iryanthera",
             Species = c("hostmannii", "sagotiana")),
  data.frame(Complex = "Talisia", Genus = "Talisia",
             Species = c("hexaphylla", "praealta", "simaboides")),
  data.frame(Complex = "Symphonia", Genus = "Symphonia",
             Species = c("globulifera", "sp.1")))
data <- trees %>% 
  left_join(env) %>% 
  left_join(complexes) %>% 
  left_join(Competition) %>% 
  filter(!is.na(Plot))
```

```{r correlations, fig.cap="Correlation of abiotic topographic variables and neighbor crowding indexes (NCI). All variables, relative elevation (RelEl), slope, curvature (Curv), aspect, topographic ruggedness index (TRI), and topographic wetness index (TWI) are derived from the digital elevation model (DEM) obtained through LiDAR campaign in 2015, the digital canopy model (DCM) has been obtained in the same campaign, and neighbor crowding index (NCI) of heterospecific (NCIh) or conspecific (NCIc) are derived from Paracou census campaign of 2015."}
data %>% 
  dplyr::select(DEM, RelativeElevation, Slope, Curvature, TRI, TWI,
                MNC, NCIhetero, NCIcon) %>% 
  rename(DCM = MNC, NCIh = NCIhetero, NCIc = NCIcon, 
         RelEl = RelativeElevation, Curv = Curvature) %>% 
  na.omit() %>% 
  cor(.) %>% 
  corrplot::corrplot.mixed()
```

```{r ModelsTable, render="pandoc"}
data.frame(
  Name = c("$B_0$", "$B_{\\alpha}$", "$B_{\\alpha, \\alpha_2}$", "$B_{\\alpha, \\beta}$",
           "$B_{\\alpha, \\beta}2$", "$B_{\\alpha, \\beta}3$"),
  Formula = c("$Presence \\sim \\mathcal{B}ernoulli(logit^{-1}(\\alpha_0))$",
              "$Presence \\sim \\mathcal{B}ernoulli(logit^{-1}(\\alpha_0+\\alpha*Environment))$",
              "$Presence \\sim \\mathcal{B}ernoulli(logit^{-1}(\\alpha_0+\\alpha*Environment+\\alpha_2*Environment^2))$",
              "$Presence \\sim \\mathcal{B}ernoulli(logit^{-1}(\\alpha_0+\\alpha*Environment+Environment^{\\beta}))$",
              "$Presence \\sim \\mathcal{B}ernoulli(logit^{-1}(\\alpha_0+\\alpha*(Environment+Environment^{\\beta})))$",
              "$Presence \\sim \\mathcal{B}ernoulli(logit^{-1}(\\alpha_0+Environment^{\\alpha}+Environment^{\\beta}))$")  
) %>% 
  kable(caption = "Model tested for species complex distribution.", format = "pandoc", escape = F) %>% 
  kable_styling(full_width = F)
```

```{r Predictions, fig.cap="Habitat suitability of single taxa model for every model form (see Supplementary Material S2) and three theoretical species distribution cases. Three species distribution cases have been simulated: (i) the environmental variable has no effect, (ii) the niche optimum is an intermediate value of the environmental variable range, and (iii) the niche optimum is a limit of the environmental variable range. Blue and red dots represent simulated presences and absences respectively, whereas the black line represents the fitted corresponding model with its credibility interval in grey. The model B,2shows the best behavior with the three distribution cases."}
n <- 100
data <- list(
  no = data.frame(Environment = seq(0, 1, length.out = 100),
                  Presence = c(rep(0:1, 50))),
  intermediate = data.frame(Environment = seq(0, 1, length.out = n),
                            Presence = c(rep(0, 20), 
                                         rep(0:1,10),
                                         rep(1,20),
                                         rep(0:1,10),
                                         rep(0,20))),
  limit = data.frame(Environment = seq(0, 1, length.out = n), 
                     Presence = c(rep(0,30), rep(0:1,20), rep(1,30)))
)
mdata <- lapply(data, function(x) list(N = nrow(x),
                                       Presence = x$Presence,
                                       Environment = x$Environment))
load("../distribution_save/SingleModel.Rdata")
preds <- lapply(names(fits), function(model){
  lapply(as.list(names(data)), function(type)
    cbind(type = type, data[[type]],
          mu = apply(as.matrix(fits[[model]][[type]], pars = "theta"), 2, mean),
          t(apply(as.matrix(fits[[model]][[type]], pars = "theta"), 2, 
                  quantile, probs = c(0.05, 0.95))))) %>% 
    bind_rows() %>% 
    mutate(model = model)
}) %>% bind_rows()
preds %>% 
  mutate(model = recode_factor(model, B0 = "B[0]", Balpha = "B[alpha]",
                               Balpha2 = "B[alpha][.][alpha][2]",
                               Bbeta = "B[alpha][.][beta]",
                               Bbeta2 = "B[alpha][.][beta][2]",
                               Bbeta3 = "B[alpha][.][beta][3]")) %>% 
  ggplot(aes(x = Environment)) +
  geom_point(aes(y = Presence, col = as.factor(Presence))) +
  geom_point(aes(y = mu)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'red', alpha = 0.2) +
  geom_line(aes(y = `5%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "red", alpha = 1, size = 0.5, linetype = "dashed") +
  facet_grid(model ~ type, scales = "free", labeller = label_parsed) +
  ylab("Habitat suitabtility") + scale_color_discrete("Presence")
```

```{r ComplexPar}
fits <- list()
for(complex in c("Parvifolia", "Licania", "Iryanthera", "Talisia", "Symphonia")){
  load(paste0("../distribution_save/complexes", complex, ".Rdata"))
  fits[[complex]] <- fit
}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = c("alpha", "beta", "gamma"), 
                  droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "complex") %>% 
  mutate(descriptor = gsub("([[:alpha:]])", "", term)) %>% 
  dplyr::select(complex, term, descriptor, estimate) %>% 
  mutate(descriptor = gsub("([[:punct:]])", "", descriptor)) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "_{TWI}$", `2` = "_{NCI}$")) %>%
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(term = recode_factor(term, alpha = "$\\alpha$", 
                              beta = "$\\beta", 
                              gamma = "$\\gamma")) %>% 
  mutate(term = paste0(term, descriptor)) %>% 
  reshape2::dcast(complex ~ term, value.var = "estimate") %>% 
  kable(caption = "Model parameters for each complex with scaled descriptors.",
        escape = F, digits = 3, format = "pandoc") %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1)  
```

```{r Speciespar}
Complexes <- bind_rows(
  data.frame(Complex = "Parvifolia", Genus = "Eschweilera",
             Species = c("pedicellata", "coriacea", "decolorans", "sagotiana",
                         "wachenheimii", "grandiflora_form2")),
  data.frame(Complex = "Licania", Genus = "Licania",
             Species = c("alba", "membranacea", "canescens", "micrantha",
                         "ovalifolia", "sprucei", "densiflora",
                         "laxiflora", "parvifructa")),
  data.frame(Complex = "Iryanthera", Genus = "Iryanthera",
             Species = c("hostmannii", "sagotiana")),
  data.frame(Complex = "Talisia", Genus = "Talisia",
             Species = c("hexaphylla", "praealta", "simaboides")),
  data.frame(Complex = "Symphonia", Genus = "Symphonia",
             Species = c("globulifera", "sp.1")))
complexes <- unique(Complexes$Complex)
d <- 20
trees <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(Plot %in% c(1, 6, 11, 13:16)) %>% # control and biodiv only
  filter(CensusYear == 2015) %>%
  filter(Xfield > d, Xfield < 250-d, Yfield > d, Yfield < 250-d) %>% 
  filter(Species != "Indet.") %>% 
  mutate(DBH = CircCorr/pi) %>% 
  collect()
data <- trees %>% 
  left_join(env) %>% 
  left_join(Complexes) %>% 
  left_join(Competition) %>% 
  filter(!is.na(Plot)) %>% 
  filter(!is.na(Complex)) %>% 
  group_by(Complex, Species) %>% 
  filter(n() > 10) %>%
  ungroup()
mdata <- lapply(complexes, function(complex) {
  data<- filter(data, Complex == complex)
  list(N = nrow(data),
       S = length(unique(data$Species)),
       K = 2,
       Y = sapply(levels(as.factor(data$Species)),
                  function(sp) as.numeric(data$Species == sp)),
       X = dplyr::select(data, TWI) %>%
         mutate_all(scale, center = F) %>%
         as.matrix())
})
names(mdata) <- complexes
species <- lapply(mdata, function(data) 
  data.frame(species = 1:length(colnames(data$Y)),
             Species = colnames(data$Y))
  ) %>% bind_rows(.id = "complex") %>% 
  mutate(SpeciesLong = paste(substr(complex, 1, 1), Species))
fits <- list()
for(complex in c("Parvifolia", "Licania", "Iryanthera", "Talisia", "Symphonia")){
  load(paste0("../distribution_save/species", complex, ".Rdata"))
  fits[[complex]] <- fit
}
lapply(fits, function(fit)
  broom::tidyMCMC(fit, pars = c("alpha", "beta", "gamma"), 
                  droppars = NULL, rhat = T)) %>%
  bind_rows(.id = "complex") %>% 
  mutate(index = gsub("([[:alpha:]])", "", term)) %>% 
  mutate(index = gsub("([[:punct:]])", "", index)) %>% 
  mutate(term = gsub("([[:digit:]])", "", term)) %>% 
  mutate(term = gsub("([[:punct:]])", "", term)) %>% 
  mutate(species = as.numeric(substr(index, 1, 1))) %>% 
  left_join(species) %>% 
  mutate(SpeciesLong = gsub("P ", "E ", SpeciesLong)) %>% 
  mutate(descriptor = substr(index, 2, 2)) %>% 
  dplyr::select(complex, term, SpeciesLong, descriptor, estimate) %>% 
  mutate(descriptor = recode_factor(descriptor, `1` = "{TWI}$", `2` = "{NCI}$")) %>% 
  mutate(term = recode_factor(term, alpha = "$\\alpha$", 
                              beta = "$\\beta_", 
                              gamma = "$\\gamma_")) %>% 
  mutate(term = paste0(term, descriptor)) %>% 
  reshape2::dcast(SpeciesLong ~ term, value.var = "estimate") %>%
  kable(caption = "Model parameters for each complex with scaled descriptors.",
        escape = F, digits = 3, format = "pandoc") %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:3)
```

*Supplementary Material S6: Stan code for the inference of species complex presence*

```{stan singlemodel, eval=F, echo=T, cache=F, output.var="m1"}
functions {
  real scale(vector x) {
    return sqrt(sum(x .* x)/(num_elements(x) - 1)) ;
  }
}
data {
  int<lower=1> N ;                // # of individuals
  int<lower=1> K ;                // # of environmental descriptors
  int<lower=0, upper=1> Y[N] ;    // individuals presence or absence (0-1)
  matrix[N,K] X ;                 // environmental descriptors
  int<lower=1> N_pred ;
  matrix[N_pred,K] X_pred ;            
}
transformed data {
  matrix[N,K] X_std ;
  matrix[N_pred,K] X_pred_std ;
  for(k in 1:K) {
    X_std[,k] = X[,k] / scale(X[,k]) ;  
    X_pred_std[,k] = X_pred[,k] / scale(X[,k]) ;  
  }
}
parameters {
  real alpha ;       // intercept
  vector[K] beta ;   // sigmoidal slope
  vector[K] gamma ;  // quadratic form
}
model {
  alpha ~ normal(0, 1) ;
  beta ~ normal(0, 1) ;
  gamma ~ normal(0, 1) ;
  Y ~ bernoulli_logit(alpha + X_std * beta + (X_std .* X_std) * gamma) ;
}
generated quantities {
  vector[N_pred] Y_pred ;
  Y_pred = inv_logit(alpha + X_pred_std * beta + (X_pred_std .* X_pred_std) * gamma) ;
}
```

*Supplementary Material S7: Stan code for the inference of species joint-presence*

```{stan jointmodel, eval=F, echo=T, cache=F, output.var="m2"}
functions {
  real dirichlet_multinomial_lpmf(int[] y, vector alpha) {
    real alpha_plus = sum(alpha);
    return lgamma(alpha_plus) + sum(lgamma(alpha + to_vector(y)))
                - lgamma(alpha_plus+sum(y)) - sum(lgamma(alpha));
  }
  real scale(vector x) {
    return sqrt(sum(x .* x)/(num_elements(x) - 1)) ;
  }
}
data {
  int<lower = 1> N  ;       // # of individuals
  int<lower = 1> S  ;       // # of species
  int<lower = 1> K  ;       // # of environmental descriptors
  int<lower = 0, upper=1> Y[N, S]  ; // individuals presence or absence for each species
  matrix[N,K] X ;           // environmental descriptors
  int<lower=0> N_pred ;
  matrix[N_pred,K] X_pred ;
}
transformed data {
  matrix[N,K] X_std ;
  matrix[N_pred,K] X_pred_std ;
  for(k in 1:K) {
    X_std[,k] = X[,k] / scale(X[,k]) ;  
    X_pred_std[,k] = X_pred[,k] / scale(X[,k]) ;  
  }
}
parameters {
  vector[S] alpha ; // intercept
  matrix[S,K] beta ; // sigmoidal slope
  matrix[S,K] gamma ; // quadratic form
}
model {
  alpha ~ normal(0,1) ;  
  for(k in 1:K){
    beta[,k] ~ normal(0,1) ;
    gamma[,k] ~ normal(0,1) ;
  }
  for (n in 1:N)
    Y[n] ~ dirichlet_multinomial(softmax(alpha + beta*to_vector(X_std[n,]) + gamma*to_vector(X_std[n,] .* X_std[n,])))  ; // likelihood
}
generated quantities {
  matrix[N_pred,S] Y_pred ;
  for(n in 1:N_pred)
    Y_pred[n,] = to_row_vector(softmax(alpha + beta*to_vector(X_pred_std[n,]) + gamma*to_vector(X_pred_std[n,] .* X_pred_std[n,]))) ;
}
```

```{r thesis, eval=F, echo=T}
save(list = ls(), file = "../../thesis/rdata/Ch1SupMat.Rdata")
```
