---
title: "Competition model Plot 1"
author: Sylvain Schmitt, Bruno Hérault
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup_BA, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(raster)
library(tidyverse)
library(rstan)
library(bayesplot)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
theme_set(bayesplot::theme_default())
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
```

```{r computeBA}
computeBA <- function(trees, # all trees from the plot
                      idTreeVal, # from the tree for which BA is computed
                      CensusYearVal,
                      XfieldVal, # from the tree for which BA is computed
                      YfieldVal, # from the tree for which BA is computed
                      d # maximum distance to be taken into account
){
  trees %>% 
    mutate(Surface = (CircCorr)^2/(4*pi)) %>% 
    filter(CensusYear == CensusYearVal) %>%
    filter(idTree != idTreeVal) %>%
    group_by(idTree) %>% 
    mutate(distance = sqrt((XfieldVal-Xfield)^2+(YfieldVal-Yfield)^2)) %>% 
    filter(distance < d) %>% 
    ungroup() %>% 
    select(Surface) %>% 
    summarise(BA = sum(Surface)) %>% 
    unlist()
}
```

```{r data}
trees <- src_sqlite(file.path(path, "trees/Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(Plot == 1) %>% 
  filter(CodeAlive == 1) %>% 
  collect()
# symphonia <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv')) %>%
#   filter(CodeAlive == 1) %>%
#   filter(Plot == 1) %>%
#   filter(Xfield > 20, Xfield < max(Xfield)-20,
#          Yfield > 20, Yfield < max(Yfield)-20)
# symphonia <- symphonia %>%
#   group_by(CensusYear, idTree) %>%
#   mutate(BA10 = computeBA(trees, idTree, CensusYear, Xfield, Yfield, d = 10)) %>%
#   mutate(BA20 = computeBA(trees, idTree, CensusYear, Xfield, Yfield, d = 20))
# save(symphonia, file = "./Competition_save/symphoniaBA_P1.Rdata")
load("./Competition_save/symphoniaBA_P1.Rdata")
symphonia <- symphonia %>%
  mutate(dbh = CircCorr/pi) %>% 
  arrange(CensusYear) %>%
  group_by(idTree) %>%
  mutate(dt = CensusYear - lag(CensusYear)) %>% 
  mutate(ddbh = dbh - lag(dbh)) %>% 
  mutate(dt = ifelse(dt <= 0, NA, dt)) %>% 
  mutate(ddbh = ifelse(ddbh < 0, NA, ddbh)) %>% 
  mutate(agr = ddbh/dt) %>% 
  ungroup() %>% 
  mutate(BA10 = BA10*10^-4, BA20 = BA20*10^-4) %>% # cm2 to m2
  mutate(BA10 = BA10/(pi*10^2/10^4), BA20 = BA20/(pi*20^2/10^4)) # divided by surface in ha
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
symphoniaXY <- symphonia
coordinates(symphoniaXY) <- ~Xutm + Yutm
proj4string(symphoniaXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
symphoniaXY <- spTransform(symphoniaXY, CRSobj = crs)
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
wtd <- spTransform(wtd, CRSobj = crs)
symphonia$wtd <- raster::extract(wtd, symphoniaXY)$TypeWatEN

wetness <- raster(path, "/topography/env/wetness.tiff")
dem <- raster(file.path(path, "topography", "MNT_ParacouAvril2009_5m.tif")) # for CRS
projection(wetness) <- projection(dem)
wetness <- projectRaster(wetness, crs = crs)
sympho$wetness <- raster::extract(wetness, symphoXY)
rm(crs, symphoniaXY, wtd)
```

# Modelling distribution

## Data

```{r ModelData, fig.cap="Evolution of local basal area (BA) through census years."}
symphonia %>% 
  ggplot(aes(wtd, BA20, col = Species, alpha = dbh, size = dbh)) +
  geom_point(aes(size = dbh)) + 
  geom_line(aes(group = idTree)) +
  xlab("Census year") +
  ylab("Local basal area (BA in m2/ha)") +
  scale_size(range = c(0.2, 3))
```

## Model 

Local basal area is not dependant on the tree individual and not on the species, we will simply assume each patch to follow the same evolution trend linearly with a null intercept year $year0_i$ corresponding to a major treefall for the gap and we will assume the trend coefficient to be either equal or not for each patch $i$ thus ignoring intermediate treefall events. **Idea to check on the full dataset, BA decrease after reaching a maximum.**

$$BA20_{patch} \sim \mathcal{N}(year0_{patch}+\beta_{patch}*year,\sigma) $$
$$\beta_{patch} \sim \mathcal{N}^P(\beta,\sigma_{patch})~or~\beta_{patch}=\beta$$

```{stan M1, output.var="M1", echo=T, eval=F}
  data {
    int N ; // # obs
    real BA20[N] ;
    real year[N] ;
    int P ; // # of patch
    int<lower=1,upper=P> patch[N] ;
  }
  parameters {
    real year0[P] ;
    real beta_patch[P] ;
    real beta ;
    real<lower=0> sigma ;
    real<lower=0> sigma_patch ;
  }
  model {
    beta_patch ~ normal(beta, sigma_patch) ;
    for(n in 1:N)
      BA20[n] ~ normal(year0[patch[n]] + beta_patch[patch[n]]*year[n], sigma) ;
  }
  generated quantities {
    real BA20pred[N] ;
    for(n in 1:N)
      BA20pred[n] = year0[patch[n]] + beta_patch[patch[n]]*year[n] ;
  }
```

```{r fitM1}
# fit <- sampling(M1,
#                 data = list(N = nrow(symphonia),
#                             BA20 = symphonia$BA20,
#                             year = symphonia$CensusYear,
#                             P = length(unique(symphonia$idTree)),
#                             patch = as.numeric(as.factor(symphonia$idTree))))
# save(fit, file = "./Competition_save/M1_P1.Rdata")
load("./Competition_save/M1_P1.Rdata")
pars <- c("beta", "sigma_patch", "sigma")
broom::tidyMCMC(fit, pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>% 
  kable(caption = "Summary table of the model")
```

```{r betap, fig.cap="Coefficient for each patch."}
mcmc_intervals(as.array(fit), regex_pars = "beta_p")
```

```{r year0, fig.cap="Intercept for each patch."}
mcmc_intervals(as.array(fit), regex_pars = "year0")
```

```{r T1, fig.cap="Markov chains trace plot after warmup for model Mnull"}
mcmc_trace(as.array(fit), pars = c(pars, "lp__"),
           facet_args = list(labeller = label_parsed))
```

```{r 21, fig.cap="Markov chains pairs plot after warmup for model Mnull"}
mcmc_pairs(as.array(fit), pars = c(pars, "lp__"))
```

```{r Y1, fig.cap="Predictions for model M0"}
cbind(
  symphonia,
  pred = apply(as.matrix(fit, pars = "BA20pred"), 2, mean),
  t(apply(as.matrix(fit, pars = "BA20pred"), 2, quantile, probs = c(0.05, 0.95)))
) %>% 
  filter(idTree %in% sample(unique(symphonia$idTree), 9)) %>% 
  ggplot(aes(CensusYear, group = idTree, col = Species, 
             alpha = dbh, size = dbh)) +
  geom_point(aes(y = BA20)) +
  geom_line(aes(y = BA20),
            alpha = 1, size = 0.2, linetype = "dotted") +
  geom_line(aes(y = pred)) +
  geom_line(aes(y = `5%`), col = "black", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "black", alpha = 1, size = 0.5, linetype = "dashed") +
  xlab("Census year") +
  ylab("Local basal area (BA in m2/ha)") +
  scale_size(range = c(0.2, 3)) +
  facet_wrap(~idTree, labeller = "label_both", scales = "free")
```
