---
title: "Competition Plot 1"
author: Sylvain Schmitt, Bruno Hérault
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup_BA, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(raster)
library(tidyverse)
library(rstan)
library(bayesplot)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
theme_set(bayesplot::theme_default())
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
```

# Introduction

> Ce que je veux dire c’est que pour quantifier l’évolution l’environnement lumineux et son évolution, tu n’as pas besoin de paramétrer un modèle de croissance. Tu peux simplement calculer année après année l’évolution de la surface terrière sur un cercle de 10 ou 20m autour de chaque arbre cible. La valeur de base et la pente de cette évolution te permet de voir le niveau de compétition initial et si l’arbre a subit une augmentation ou une diminution de compétition. Et du coup de voir s’il y a des génotypes/phénotypes particuliers associés à ces différents environnements lumineux. *Bruno Hérault, 22/08/2018*

We assume biotic environment through neighbors interaction to play an important role on tree establishment and growth and more especially with putative differences between genotypes, phenotypes and closely related species due to sympatric fine-scale local adaptation.

In order to test that we want to study light and competition environment evolution through time for individual tree in relation to their morphotype / genotype.

Finally, we wish to use evidenced relation to build back light and competition environment through past time.

# Material and methods

## Study site

We will use Paracou censuses from 1984 to 2017. For computation purposes we will use in a first time an undisturbed control plot, the plot P1.

## Light and competition environment

Light and competition environment will be first approached with local basal area through time. We will test for local basal area with neighboring trees at a distance of $d=10m$ and $d=20m$, respectively $BA_{10}$ and $BA_{20}$ in $m^2.ha^-1$. Local basal area is computed as follow:

$${BA_d}_i = \frac{\sum_{j ~\forall~ d_{i,j}<d} SBH_j}{\pi*d^2}$$

where local basal area $BA$ for individual $i$ with neighbors at distance $d$ equal to the sum of neighbors $j$ surface at breast height $SBH_j$ for individuals at distance $d_{i,j}$ inferior to $d$ divided by area surface. We computed local basal area for all individuals from the species complex *Symphonia globulifera* with morphotype *S. globulifera* and *S. sp1* growing in plot 1. To avoid edge effects we excluded *Symphonia globulifera* individuals less than 20 meters ahead from a plot border.

```{r computeBA}
computeBA <- function(trees, # all trees from the plot
                      idTreeVal, # from the tree for which BA is computed
                      CensusYearVal,
                      XfieldVal, # from the tree for which BA is computed
                      YfieldVal, # from the tree for which BA is computed
                      d # maximum distance to be taken into account
){
  trees %>% 
    mutate(Surface = (CircCorr)^2/(4*pi)) %>% 
    filter(CensusYear == CensusYearVal) %>%
    filter(idTree != idTreeVal) %>%
    group_by(idTree) %>% 
    mutate(distance = sqrt((XfieldVal-Xfield)^2+(YfieldVal-Yfield)^2)) %>% 
    filter(distance < d) %>% 
    ungroup() %>% 
    select(Surface) %>% 
    summarise(BA = sum(Surface)) %>% 
    unlist()
}
```

```{r data}
trees <- src_sqlite(file.path(path, "trees/Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(Plot == 1) %>% 
  filter(CodeAlive == 1) %>% 
  collect()
# symphonia <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv')) %>%
#   filter(CodeAlive == 1) %>%
#   filter(Plot == 1) %>%
#   filter(Xfield > 20, Xfield < max(Xfield)-20,
#          Yfield > 20, Yfield < max(Yfield)-20)
# symphonia <- symphonia %>%
#   group_by(CensusYear, idTree) %>%
#   mutate(BA10 = computeBA(trees, idTree, CensusYear, Xfield, Yfield, d = 10)) %>%
#   mutate(BA20 = computeBA(trees, idTree, CensusYear, Xfield, Yfield, d = 20))
# save(symphonia, file = "./Competition_save/symphoniaBA_P1.Rdata")
load("./Competition_save/symphoniaBA_P1.Rdata")
symphonia <- symphonia %>%
  mutate(dbh = CircCorr/pi) %>% 
  arrange(CensusYear) %>%
  group_by(idTree) %>%
  mutate(dt = CensusYear - lag(CensusYear)) %>% 
  mutate(ddbh = dbh - lag(dbh)) %>% 
  mutate(dt = ifelse(dt <= 0, NA, dt)) %>% 
  mutate(ddbh = ifelse(ddbh < 0, NA, ddbh)) %>% 
  mutate(agr = ddbh/dt) %>% 
  ungroup() %>% 
  mutate(BA10 = BA10*10^-4, BA20 = BA20*10^-4) %>% # cm2 to m2
  mutate(BA10 = BA10/(pi*10^2/10^4), BA20 = BA20/(pi*20^2/10^4)) # divided by surface in ha
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
symphoniaXY <- symphonia
coordinates(symphoniaXY) <- ~Xutm + Yutm
proj4string(symphoniaXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
symphoniaXY <- spTransform(symphoniaXY, CRSobj = crs)
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
wtd <- spTransform(wtd, CRSobj = crs)
symphonia$wtd <- raster::extract(wtd, symphoniaXY)$TypeWatEN
rm(crs, symphoniaXY, wtd)
```

## Light and competition effect analysis

In a first approach, we looked at evolution of local basal area through time and to the evolution of diameter at breast height $dbh$ in $cm$ and annual growth rate $AGR$ in $cm.year^-1$ through growing local basal area. Annual growth rate as been computed as follow:

$$AGR_i = \frac{\Delta dbh_i}{\Delta t_i}$$
where annual growth rate $AGR$ from individual $i$ equal the difference of diameter between two census years divided by the time between the two years.

# Results

## General statistics

We studied `r length(unique(symphonia$idTree))` *Symphonia globulifera* individuals in the plot 1 over `r max(symphonia$CensusYear) - min(symphonia$CensusYear)` years. Local basal area ranged from 10 to 100 $m^2.ha^-1$ (figure \@ref(fig:GeneralStatistics)). *Symphonia sp1* morphotype grow in significantly higher local basal area than *Symphonia globulifera* morphotype.

```{r GeneralStatistics, fig.cap="Local basal area (BA) distribution by morphotype for BA computed in a radius of 10m (BA10, right subgraph) and in a radius of 20m (BA20, left subgraph)."}
symphonia %>% 
  dplyr::select(BA10, BA20, Species) %>% 
  reshape2::melt(id.var = "Species") %>% 
  ggplot(aes(Species, value, fill = Species)) +
  geom_boxplot() +
  facet_wrap(~variable, scales = "free") +
  ylab("Local basal area (BA in m2/ha)") +
  scale_fill_discrete("Symphonia") +
  ggpubr::stat_compare_means(method = "anova")
```

## Habitat

**BA is not linked to the habitat (important for interpretations) !**

```{r Habitat, fig.cap="Evolution of local basal area (BA) through census years. Figure presented for BA computed in a radius of 10m (BA10, right subgraph) and in a radius of 20m (BA20, left subgraph)."}
symphonia %>% 
  dplyr::select(Species, idTree, BA10, BA20, CensusYear, wtd, dbh) %>% 
  reshape2::melt(id.vars = c("Species", "idTree", "CensusYear", "wtd", "dbh"),
                 variable.name = "BAdist", value.name = "BA") %>% 
  ggplot(aes(CensusYear, BA, col = Species, alpha = dbh, size = dbh)) +
  geom_point(aes(size = dbh)) + 
  geom_line(aes(group = idTree)) +
  facet_grid(BAdist ~ wtd, scale = "free") +
  xlab("Census year") +
  ylab("Local basal area (BA in m2/ha)") +
  scale_size(range = c(0.2, 3))
```

## Census year

Most of individuals showed increasing local basal area through time (figure \@ref(fig:CensusYear)), but few individual showed basal area decrease at certain years indicating neighboring tree death. Increase coefficient of local basal area through time is similar for most individuals besides some individuals show higher rates of increase after a basal area decrease. Logically at broader scale of 20m radius, neighboring *Symphonia globulifera* individuals showed more similar patterns.

```{r CensusYear, fig.cap="Evolution of local basal area (BA) through census years. Figure presented for BA computed in a radius of 10m (BA10, right subgraph) and in a radius of 20m (BA20, left subgraph)."}
symphonia %>% 
  select(Species, idTree, BA10, BA20, CensusYear) %>% 
  reshape2::melt(id.vars = c("Species", "idTree", "CensusYear"),
                 variable.name = "BAdist", value.name = "BA") %>% 
  ggplot(aes(CensusYear, BA, col = Species)) +
  geom_point() + 
  geom_line(aes(group = idTree)) +
  facet_wrap(~ BAdist, scale = "free") +
  scale_color_discrete("Symphonia") +
  xlab("Census year") +
  ylab("Local basal area (BA in m2/ha)")
```

## Diameter at breast height

Diameter at breast height (dbh) variation seems to be dependent on the local basal area values, more especially higher dbh values seems allowed only for intermediate local basal area values especially for *S. globulifera* morphotype (figure \@ref(fig:dbh)). In addition *S. globulifera* morphotype seems to grow in a smaller range of local basal area which decrease with increasing dbh value.

```{r dbh, fig.cap="Observed diameter at breast height (dbh) depending on local basal area (BA) through census years. Figure presented for BA computed in a radius of 10m (BA10, right subgraph) and in a radius of 20m (BA20, left subgraph)."}
symphonia %>% 
  select(Species, idTree, BA10, BA20, dbh) %>% 
  reshape2::melt(id.vars = c("Species", "idTree", "dbh"),
                 variable.name = "BAdist", value.name = "BA") %>% 
  ggplot(aes(dbh, BA, col = Species, alpha = dbh, size = dbh)) +
  geom_point(aes(size = dbh)) + 
  geom_line(aes(group = idTree)) +
  xlab("Dimater at breast height (dbh in cm)") +
  ylab("Local basal area (BA in m2/ha)") +
  scale_size(range = c(0.2, 3)) +
  facet_wrap(~ BAdist, scale = "free", ncol = 2)
```

## Annual growth rate

We found similar results of observed annual growth rate evolution along increasing local basal area to the evolution of diameter at breast height (figure \@ref(fig:agr)). The pattern is only less visible.

```{r agr, fig.cap="Observed annual growth rate (AGR) depending on local basal area (BA) through census years. Figure presented for BA computed in a radius of 10m (BA10, right subgraph) and in a radius of 20m (BA20, left subgraph)."}
symphonia %>% 
  select(Species, idTree, BA10, BA20, agr, dbh) %>% 
  reshape2::melt(id.vars = c("Species", "idTree", "agr", "dbh"),
                 variable.name = "BAdist", value.name = "BA") %>% 
  ggplot(aes(BA, agr, col = Species, alpha = dbh, size = dbh)) +
  geom_point(aes(size = dbh)) +
  ylab("Annual growth rate (AGR in cm/year)") +
  xlab("Local basal area (BA in m2/ha)") +
  scale_size(range = c(0.2, 3)) +
  facet_wrap(~ BAdist, scale = "free", nrow = 2)
```

# Discussion

We found that generally *S. sp1* morphotype was growing in higher local basal area than *S. globulifera*. More especially, looking at diameter at breast height and annual growth rate values, we found that generally bigger individuals were found in a smaller range of intermediate local basal area values than small individuals. This phenomenon was increased for *S. globulifera* morphotype which seems to grow and survive only in local environment with 20 to 30 $m^2.ha^{-1}$, besides some individuals are recruited to environment reaching only 10 $m^2.ha^{-1}$. On the opposite, *S. sp1* wasn't recruited below 20 $m^2.ha^{-1}$ but could grow in denser habitat over 40 $m^2.ha^{-1}$. And *S. sp1* kept a wide range of habitat density when growing in diameter. Finally, *S. globulifera* reached bigger diameters than *S. sp1* (more than 60 $cm$ against only 50 $cm$, respectively).

It seems that *S. sp1* morphotype is more shade tolerant than *S. globulifera*, when growing. But regarding recruitment *S. globulifera* seems to tolerate lighter environment than *S. sp1*. And besides they grow around the same light and competition optimal, it seems that *S. sp1* tolerate a wider range of condition to grow but with lower maximum diameter (trade-off ?).

Regarding our objective to model past light and competition environment, those patterns seems important but not strong enough to model back past biotic environment. Effectively, if we look at evolution of local basal area through time, we surely have a global trend of constant increase. But sudden decrease in local basal area seems unpredictable besides they might have a huge role in morphotype ecology. Consequently it would be useless to model the general trend without accounting those brutal changes.
