# Ontgeny only

```{r setup_No, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(rstan)
library(bayesplot)
library(tidyverse)
mpath <- "./growth_models"
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
rstan_options(auto_write = T)
options(mc.cores = parallel::detectCores())
load("./growth_save/model_data.Rdata")
```

## Introduction

First model takes only into account ontogeny. This is simply the model proposed by @Herault2010:

\begin{equation}
  log(AGR_i+1) \sim \mathcal{N}(AGR_{max}*e^{-\frac{1}{2}*(log(\frac{dbh_i}{d_{opt}})/k_s)^2},\sigma)
  (AGR_{max}, d_{opt}, k_s,\sigma) \sim \mathcal{\Gamma}^4(10^{-2},10^{-2})
  (\#eq:No)
\end{equation}

## Model 1

```{r No}
# fit <- stan(file = file.path(mpath, 'No.stan'), data = data, chains = 2)
# save(fit, file = file.path(mpath, 'No.Rdata'))
load(file.path(mpath, 'No.Rdata'))
pars <- c("AGRmax", "Dopt", "Ks", "sigma")
```

### Diagnostic

Diagnostic seems okay. Summary table \@ref(tab:So) shows a correct $\hat{R}$ indicating good convergence as well as trace plots \@ref(fig:To). But if we look closely to the pair plot \@ref(fig:2o) we might have an apparent relation between parameters $AGR_{max}$ and $K_s$.

```{r So}
broom::tidyMCMC(fit, c(pars,"lp__"), droppars = NULL, rhat = T) %>% 
  kable(caption = "Summary table of the model \"Ontogeny only\"")
```

```{r To, fig.cap="Markov chains trace plot after warmup for model \"Ontogeny only\""}
mcmc_trace(as.array(fit), pars = c(pars, "lp__"),
           facet_args = list(labeller = label_parsed))
```

```{r 2o, fig.cap="Markov chains pairs plot after warmup for model \"Ontogeny only\""}
mcmc_pairs(as.array(fit), pars = pars)
```

### Predicitons & conclusion

Predictions seems okay in figure \@ref(fig:Yo), but we will still try to deccorelate parameters $AGR_{max}$ and $K_s$ to not have further issues in more complex models.

```{r Yo, fig.cap="Predictions for model \"Ontogeny only\""}
cbind(
  data$trees[c("dbh_c", "agr", "espece")],
  mean = apply(as.matrix(fit, pars = "AGRpred"), 2, mean),
  t(apply(as.matrix(fit, pars = "AGRpred"), 2, quantile, probs = c(0.05, 0.95)))
) %>% 
  ggplot(aes(dbh_c, agr)) +
  geom_point(aes(col = espece)) +
  geom_line(aes(y = mean)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'grey', alpha = 0.2)
```

We investigated different relations between $AGR_{max}$ and $K_s$ in figure \@ref(fig:Do), and kept the cube relation with $AGR_{max} = AGR_{max}'*{K_s}^3$.

```{r Do, fig.cap="Test of different relations between correlated parameters."}
decorrelate <- function(fit, par1, par2){
  par1 <- enquo(par1)
  par2 <- enquo(par2)
  as.data.frame(fit) %>% 
    select(!!par1, !!par2) %>% 
    mutate(identity = (!!par1)/(!!par2)) %>%
    mutate(pow1.5 = (!!par1)/(!!par2)^(3/2)) %>%
    mutate(square = (!!par1)/(!!par2)^2) %>%
    mutate(cube = (!!par1)/(!!par2)^3) %>%
    dplyr::select(-!!par2) %>% 
    reshape2::melt(id.var = as.character(par1)[2]) %>% 
    ggplot(aes_string("value", as.character(par1)[2])) +
    geom_point() +
    facet_wrap(~ variable)
}
decorrelate(fit, AGRmax, Ks)
```

## Model 2

```{r No2}
# fit <- stan(file = file.path(mpath, 'No2.stan'), data = data, chains = 2)
# save(fit, file = file.path(mpath, 'No2.Rdata'))
load(file.path(mpath, 'No2.Rdata'))
pars <- c("AGRmax", "Dopt", "Ks", "sigma")
```

### Diagnostic

Diagnostic seems okay. Summary table \@ref(tab:So2) shows a correct $\hat{R}$ a bit higher than previous model indicating good convergence as well as trace plots \@ref(fig:To2). And the pair plot \@ref(fig:2o2) seems okay, besides a relation could still exist between $AGR_{max}$ and $K_s$.

```{r So2}
broom::tidyMCMC(fit, c(pars,"lp__"), droppars = NULL, rhat = T) %>% 
  kable(caption = "Summary table of the model \"Ontogeny only\"")
```

```{r To2, fig.cap="Markov chains trace plot after warmup for model \"Ontogeny only\""}
mcmc_trace(as.array(fit), pars = c(pars, "lp__"),
           facet_args = list(labeller = label_parsed))
```

```{r 2o2, fig.cap="Markov chains pairs plot after warmup for model \"Ontogeny only\""}
mcmc_pairs(as.array(fit), pars = pars)
```

### Predicitons & conclusion

Predictions seems okay in figure \@ref(fig:Yo2). We will thus keep this model for ontogeny only.

```{r Yo2, fig.cap="Predictions for model \"Ontogeny only\""}
cbind(
  data$trees[c("dbh_c", "agr", "espece")],
  mean = apply(as.matrix(fit, pars = "AGRpred"), 2, mean),
  t(apply(as.matrix(fit, pars = "AGRpred"), 2, quantile, probs = c(0.05, 0.95)))
) %>% 
  ggplot(aes(dbh_c, agr)) +
  geom_point(aes(col = espece)) +
  geom_line(aes(y = mean)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), color = 'grey', alpha = 0.2)
```

## Conclusion

We have a stable basis of simple ontgeny only growth model, with decorrelated parameters, to be complexified with other models, including random effects and additional reductors.
