# Species and ontogeny

```{r setup_Nso, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(rstan)
library(bayesplot)
library(tidyverse)
mpath <- "./growth_models"
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
rstan_options(auto_write = T)
options(mc.cores = parallel::detectCores())
load("./growth_save/model_data.Rdata")
```

## Introduction

First model takes only into account ontogeny. This is simply the model proposed by @Herault2010:

\begin{equation}
  log(AGR_i+1) \sim \mathcal{N}(AGR_{max_{sp}}*e^{-\frac{1}{2}*(log(\frac{dbh_i}{d_{opt_{sp}}})/k_{s_{sp}})^2},\sigma_{sp})
  (AGR_{max_{sp}}, d_{opt_{sp}}, k_{s_{sp}}, \sigma_{sp}) \sim \mathcal{\Gamma}^{4*N_{sp}}(10^{-2},10^{-2})
  (\#eq:No)
\end{equation}

## Data 1

```{r Nso}
# fit <- stan(file = file.path(mpath, 'Nso.stan'), data = data, chains = 2)
# save(fit, file = file.path(mpath, 'Nso.Rdata'))
load(file.path(mpath, 'Nso.Rdata'))
pars <- c("AGRmax_sp", "Dopt_sp", "Ks_sp", "sigma_sp")
```

### Diagnostic

Diagnostic seems okay. Summary table \@ref(tab:Sso) shows a correct $\hat{R}$ indicating good convergence as well as trace plots \@ref(fig:Tso). But if we look to the pair plot \@ref(fig:2so) we have a lot of relations for *S. globulifera* morphotype whereas non with *S. indet* !

```{r Sso}
broom::tidyMCMC(fit, c(pars,"lp__"), droppars = NULL, rhat = T) %>% 
  kable(caption = "Summary table of the model \"Species and ontogeny\"")
```

```{r Tso, fig.cap="Markov chains trace plot after warmup for model \"Species and ontogeny\""}
mcmc_trace(as.array(fit), regex_pars = c(pars, "lp__"),
           facet_args = list(labeller = label_parsed))
```

```{r 2so, fig.cap="Markov chains pairs plot after warmup for model \"Species and ontogeny\""}
mcmc_pairs(as.array(fit), pars = paste0(pars, "[1]"))
mcmc_pairs(as.array(fit), pars = paste0(pars, "[2]"))
```

### Predicitons & conclusion

Predictions seems okay in figure \@ref(fig:Yso), but we need to fix correlations issue for morphotype *S. globulifera*. Maybe we have this issue due to a lack of data, so we will try with the second dataset.

```{r Yso, fig.cap="Predictions for model \"Species and ontogeny\""}
data.frame(
  mean_sp = apply(as.matrix(fit, pars = "AGRpred_sp"), 2, mean),
  t(apply(as.matrix(fit, pars = "AGRpred_sp"), 2, quantile, probs = c(0.05, 0.95))),
  data$trees
) %>% 
  ggplot(aes(dbh_c, agr, col = espece, fill = espece, group = espece)) +
  geom_point(alpha = 0.4) +
  geom_ribbon(aes(ymin = X5., ymax = X95.), alpha = 0.5)
```

## Data 2

```{r Nso2}
# fit <- stan(file = file.path(mpath, 'Nso.stan'), data = data2, chains = 2)
# save(fit, file = file.path(mpath, 'Nso2.Rdata'))
load(file.path(mpath, 'Nso2.Rdata'))
pars <- c("AGRmax_sp", "Dopt_sp", "Ks_sp", "sigma_sp")
```

### Diagnostic

Diagnostic seems okay. Summary table \@ref(tab:Sso2) shows a correct $\hat{R}$ indicating good convergence as well as trace plots \@ref(fig:Tso2). But if we look to the pair plot \@ref(fig:2so2) we have a lot of relations for *S. globulifera* morphotype whereas non with *S. indet* !

```{r Sso2}
broom::tidyMCMC(fit, c(pars,"lp__"), droppars = NULL, rhat = T) %>% 
  kable(caption = "Summary table of the model \"Species and ontogeny\"")
```

```{r Tso2, fig.cap="Markov chains trace plot after warmup for model \"Species and ontogeny\""}
mcmc_trace(as.array(fit), regex_pars = c(pars, "lp__"),
           facet_args = list(labeller = label_parsed))
```

```{r 2so2, fig.cap="Markov chains pairs plot after warmup for model \"Species and ontogeny\""}
mcmc_pairs(as.array(fit), pars = paste0(pars, "[1]"))
mcmc_pairs(as.array(fit), pars = paste0(pars, "[2]"))
mcmc_pairs(as.array(fit), pars = paste0(pars, "[3]"))
```

### Predicitons & conclusion

Predictions seems okay in figure \@ref(fig:Yso2), but we need to fix correlations issue for morphotype *S. globulifera*. Maybe we have this issue due to a lack of data, so we will try with the second dataset.

```{r Yso2, fig.cap="Predictions for model \"Species and ontogeny\""}
data.frame(
  mean_sp = apply(as.matrix(fit, pars = "AGRpred_sp"), 2, mean),
  t(apply(as.matrix(fit, pars = "AGRpred_sp"), 2, quantile, probs = c(0.05, 0.95))),
  data2$trees
) %>% 
  ggplot(aes(dbh_c, agr, col = espece, fill = espece, group = espece)) +
  geom_point(alpha = 0.4) +
  geom_ribbon(aes(ymin = X5., ymax = X95.), alpha = 0.5)
```

## Conclusion
