---
title: "Eschweilera data cleaning"
date: '`r Sys.Date()`'
author: Anne Baranger & Sylvain Schmitt
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(tidyverse)
library(stringdist)
library(googlesheets)
library(knitr)
library(kableExtra)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
path <- "../../../data/Eschweilera_Paracou/Measures/"
path1 <- "../../../data/Eschweilera_Paracou/Scans/"
```

# Introduction

Those documents aims to correct all *Eschweilera* raw data in order to prepare functional analysis. For each table, we need to join files and check IDs correspondance and data values. Those operations concern following tables: (i) Individuals, (ii) LeafFresh, (iii) LeafDry and (iv) LeafArea. Finally we will need to prepare the two final tables from all the previous ones:

* __AllMeasures__ with all corrected measures
* __AllTraits__ All corrected and computed functional traits

# Individuals

## Data formatting

The idea is to transform FTH data to the general formatting, i.e. :

* Date (YYYY/MM/DD)
* Time (HHHMM)
* Plot
* SubPlot
* FieldTreeNum
* Dawkins
* Position
* TrunkForm
* BarkColor
* BarkRuggedness
* BarkAspect
* Lenticels
* BladeForm
* LeafConsistency
* LeafSurface
* LeafBase
* LeafMargin
* HyperAcuminated
* BentPrincpalN


```{r ind}
indFTH <- read_delim(file.path(path, "data_FTH.csv"), delim = ",") %>% 
  rename(TrunkForm = `Trunk form`, BarkColor = `Bark color`, BarkRuggedness = `Bark ruggedness`, 
         BarkAspect = `Bark aspect`, BladeForm = `Blade form`, 
         LeafConsistency = `Leaf consistency`, LeafSurface = `Leaf surface`, 
         LeafBase = `Leaf base`, LeafMargin = `Leaf margin`,
         HyperAcuminated = `Longue pointe`, BentPrincipalN = `Pli sur nervure principale`) %>% 
  separate(Date, c("Month", "Day", "Year"), "/") %>% # Date
  mutate(Date = paste0(Year, "/0", Month, "/", Day)) %>% 
  select(-Year, -Month, -Day) %>% 
  mutate(Time = gsub(":", "H", substr(as.character(Time), 1, 5))) %>% # Time
  mutate(Dawkins = toupper(Dawkins)) %>% # Dawkins
  mutate(Position = recode_factor(Position, bas = "LOW", haut = "TOP", milieu = "MID",
                                  `milieu lumière` = "MID/LIGHT", `milieu/haut` = "MID",
                                  `bas lumière` = "LOW/LIGHT", `haut lumière` = "TOPLIGHT",
                                  `bas/lumière` = "LOW/LIGHT", `milieu/bas` = "MID")) %>% # Position
  mutate(TrunkForm = recode_factor(TrunkForm, `B--` = "B-", `B (petits)` = "B-", `B+` = "B",
                                  BS = "B/S", tordu = "N", C = "B")) %>% # TrunkForm
  mutate(BarkColor = recode_factor(BarkColor, `B/BR` = "B/RB", BG = "B/G", D = "DB",
                                   `DB+` = "DB", `DB/B` = "B/DB", GB = "B/G")) %>% # BarkColor
  mutate(BarkAspect = recode_factor(BarkAspect, `?` = "N", `F+` = "F", `F/D` = "D/F",
                                    `F/D-` = "D-/F", `F-HM` = "F-/HM", `F/HM+` = "F/HM",
                                    `F+/HM` = "F/HM", `F/S/HM` = "F/HM/S", 
                                    `HM/D (?)` = "D/HM", `HM/F` = "F/HM",
                                    `S+/D` = "D/S+", `S/F/HM`	= "F/HM/S", `S-/HM`	= "HM/S-",
                                    `S/HM`	= "HM/S", `R/HM` = "F/HM")) %>% # BarkAspect
  mutate(Lenticels = recode_factor(Lenticels, N = "A", `P+` = "P", `P-` = "P")) %>% # Lenticels
  mutate(BladeForm = recode_factor(BladeForm, `E+` = "E", `Oblan/E` = "E/Oblan")) %>% # BladeForm
  mutate(LeafConsistency = recode_factor(LeafConsistency, `Char -` = "Char-", 
                                         `Char+` = "Char")) %>% # LeafConsistency
  mutate(LeafBase = recode_factor(LeafBase, `O/A` = "A/O", `O-R` = "O/R", `R/A` = "A/R",
                                  RO = "O/R", `R/O` = "O/R")) %>% # LeafBase
  mutate(LeafMargin = recode_factor(LeafMargin, `E/C-` = "C-/E")) %>% # LeafMargin
  select(Date, Time, Plot, SubPlot, FieldTreeNum, Dawkins, Position,
         TrunkForm, BarkColor, BarkRuggedness, BarkAspect, Lenticels,
         BladeForm, LeafConsistency, LeafSurface, LeafBase, LeafMargin,
         HyperAcuminated, BentPrincipalN, Comment) %>% 
  unique()
# read_delim(file.path(path, "Indviduals.csv"), delim = ",", col_types = cols(Date = col_character())) %>%
#   bind_rows(indFTH, .) %>%
#   write_delim(file.path(path, "Indviduals_all.csv"), delim = ",")
```

## IDs check

```{r IndData}
individuals <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("Individuals") %>% 
  dplyr::select(-X1) %>% 
  mutate(CodeParacou = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum)))
trees <- src_sqlite(file.path("../../../data/Paracou/", 
                                 "trees/Paracou.sqlite")) %>% 
              tbl("Paracou") %>% 
              mutate(CodeParacou = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum))) %>% 
              filter(CodeParacou %in% individuals$CodeParacou) %>% 
              filter(Plot %in% c(1, 6, 11, 13:15) & CensusYear == 2017 |
                       Plot == 16 & CensusYear == 2015) %>% 
              mutate(DBH = CircCorr*pi) %>% 
              mutate(SpeciesLong = paste(Genus, Species)) %>% 
              collect()
```

Wrong IDs have been corrected.

```{r IndParacou}
ref <- setdiff(trees$CodeParacou, individuals$CodeParacou)
cor <- setdiff(individuals$CodeParacou, trees$CodeParacou)
data.frame(FreshID = cor,
           Suggestion = ref[stringdist::amatch(cor, ref, maxDist = Inf)]) %>% 
  group_by(Suggestion) %>% 
  mutate(TimeSuggested = n()) %>% 
  mutate(TimeSuggested = cell_spec(TimeSuggested, bold = ifelse(TimeSuggested > 1, T, F))) %>% 
  kable(format = "html", escape = F,
        caption = "Individual to be corrected in Individuals table.") %>%
  kable_styling("striped", full_width = F)
```

# Leaf Fresh

## Join files

Fresh measurements have three different format of files :

1. the FTH file (*data_FTH.csv*) mixing individual data and fresh measurement which might be differently encoded
1. Tom file (*sylvainESCHWEILERA.csv*) mixing individual data and fresh measurement with some errors:
1. All others files (*2018...ods*), except the november ones which are still FTH files, in the correct format which might integrate typo issues

Consequently for the ODS files we will only use those collected in October with pattern 2018.10.*.ods .

```{bash, eval=F}
cd ~/Documents/BIOGECO/PhD/data/Eschweilera_Paracou/Measures
mkdir csv
for file in ./2018.10.*.ods
do
  echo $file
  filename=$(basename "$file")
  filename="${filename%.*}"
  ssconvert -S --export-type=Gnumeric_stf:stf_csv "${filename}".ods ./csv/"${filename}".csv
done
```


```{r FreshMeasurements, eval=F}
FTH <- read_delim(file.path(path, "data_FTH.csv"), delim = ",") %>% 
  rename(Leaf = Numero_feuille, FreshWeight = Fresh_Weight, SPAD1 = `SPAD 1`) %>% 
  select(IdGenetic, Plot, SubPlot, FieldTreeNum, Leaf, 
         SPAD1, SPAD2, SPAD3, FreshWeight, LT1, LT2, LT3) %>% 
  mutate_at(vars(starts_with("SPAD")),as.numeric) %>% 
  mutate_at(vars(starts_with("LT")),as.numeric) %>% 
  mutate(Leaf = as.numeric(Leaf))
Tom <- read_delim(file.path(path, "sylvainESCHWEILERA.csv"), delim = ",") %>% 
  rename(IdGenetic = IdGentic) %>% 
  fill(IdGenetic, Plot, SubPlot, FieldTreeNum) %>% 
  mutate(Plot = as.numeric(gsub('([[:alpha:]])', "", Plot)))
All <- sapply(list.files(file.path(path, "csv"), full.names = T), function(file)
  read.csv(file, dec = ","), simplify = F) %>% 
  bind_rows() %>% 
  select(IdGenetic, Plot, SubPlot, FieldTreeNum, Leaf, 
         SPAD1, SPAD2, SPAD3, FreshWeight, LT1, LT2, LT3) %>% 
  mutate(IdGenetic = ifelse(IdGenetic == "", NA, IdGenetic)) %>% 
  filter(!is.na(Leaf)) %>% 
  fill(IdGenetic, Plot, SubPlot, FieldTreeNum)
bind_rows(All, Tom, FTH) %>% 
  write_delim(file.path(path, "LeafFresh_all.csv"), delim = ",")
```

## IDs check

Now we need to check and correct individual codes. 

```{r FreshIndData}
individuals <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("Individuals") %>% 
  dplyr::select(-X1) %>% 
  mutate(CodeParacou = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum)))
fresh <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("LeafFresh") %>% 
  dplyr::select(-X1) %>% 
  mutate(CodeParacou = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum)))
```

The number of individuals now correctly match between the two tables.

```{r FreshIndCount}
data.frame(Table = c("Individual", "LeafFresh"),
           N = c(length(unique(individuals$CodeParacou)), 
                 length(unique(fresh$CodeParacou)))) %>% 
  kable(caption = "Individual count in Individual and LeafFresh tables.")
```

Wrong IDs have been corrected.

```{r FreshInd}
ref <- setdiff(individuals$CodeParacou, fresh$CodeParacou)
cor <- setdiff(fresh$CodeParacou, individuals$CodeParacou)
data.frame(FreshID = cor,
           Suggestion = ref[stringdist::amatch(cor, ref, maxDist = Inf)]) %>% 
  group_by(Suggestion) %>% 
  mutate(TimeSuggested = n()) %>% 
  mutate(TimeSuggested = cell_spec(TimeSuggested, bold = ifelse(TimeSuggested > 1, T, F))) %>% 
  kable(format = "html", escape = F,
        caption = "Individual to be corrected in LeafFresh table.") %>%
  kable_styling("striped", full_width = F)
```

## Values check

```{r FreshValues}
fresh %>% 
  filter_at(vars(starts_with("SPAD")), any_vars(. > 100)) %>% 
  select(IdGenetic, CodeParacou, Leaf, SPAD1, SPAD2, SPAD3) %>% 
  kable(format = "html", escape = F,
        caption = "SPAD to be corrected in LeafFresh table.") %>%
  kable_styling("striped", full_width = F)
fresh %>% 
  filter_at(vars(starts_with("LT")), any_vars(. > 500)) %>% 
  select(IdGenetic, CodeParacou, Leaf, LT1, LT2, LT3) %>% 
  kable(format = "html", escape = F,
        caption = "LT to be corrected in LeafFresh table.") %>%
  kable_styling("striped", full_width = F)
fresh %>% 
  filter(FreshWeight > 13) %>% 
  select(IdGenetic, CodeParacou, Leaf, FreshWeight) %>% 
  kable(format = "html", escape = F,
        caption = "FreshWeight to be corrected in LeafFresh table.") %>%
  kable_styling("striped", full_width = F)
```

# Leaf Dry

## Join files

Dry measurements have three different format of files :

1. the FTH file (*data_FTH.csv*) mixing individual data and fresh measurement which might be differently encoded
1. Anne file (*LeafDry_Anne.csv*) in the correct format which might integrate typo issues

```{r DryMeasurements, eval=F}
FTH <- read_delim(file.path(path, "data_FTH.csv"), delim = ",") %>% 
  mutate(TreeFieldNum = FieldTreeNum, Leaf = Numero_feuille,
         DryWeight = Dry_Weight) %>% 
  select(IdGenetic, Plot, SubPlot, TreeFieldNum, Leaf, DryWeight) %>% 
  mutate_at(vars(Leaf, DryWeight), as.numeric)
Anne <- read.csv(file.path(path, "LeafDry_Anne.csv"), 
         dec = ",", stringsAsFactors = F) %>% 
  select(-X) %>% 
  mutate(IdGenetic = ifelse(IdGenetic == "", NA, IdGenetic)) %>% 
  fill(IdGenetic, Plot, SubPlot, TreeFieldNum)
bind_rows(Anne, FTH) %>% 
  write_delim(file.path(path, "LeafDry_all.csv"), delim = ",")
```

## IDs check

Now we need to check and correct individual codes. 

```{r DryIndData}
individuals <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("Individuals") %>% 
  dplyr::select(-X1) %>% 
  mutate(CodeParacou = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum)))
dry <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("LeafDry") %>% 
  dplyr::select(-X1) %>% 
  mutate(CodeParacou = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum)))
```

The number of individuals now correctly match between the two tables.

```{r DryIndCount}
data.frame(Table = c("Individual", "LeafDry"),
           N = c(length(unique(individuals$CodeParacou)), 
                 length(unique(dry$CodeParacou)))) %>% 
  kable(caption = "Individual count in Individual and LeafDry tables.")
```

Wrong IDs have been corrected.

```{r DryInd}
ref <- setdiff(individuals$CodeParacou, dry$CodeParacou)
cor <- setdiff(dry$CodeParacou, individuals$CodeParacou)
data.frame(DryID = cor,
           Suggestion = ref[stringdist::amatch(cor, ref, maxDist = Inf)]) %>% 
  group_by(Suggestion) %>% 
  mutate(TimeSuggested = n()) %>% 
  mutate(TimeSuggested = cell_spec(TimeSuggested, bold = ifelse(TimeSuggested > 1, T, F))) %>% 
  kable(format = "html", escape = F,
        caption = "Individual to be corrected in LeafDry table.") %>%
  kable_styling("striped", full_width = F)
```

Three individuals are lacking dry measurements and are currently in the oven for 48H in waiting for new measurements.

```{r DryMissing}
individuals %>% 
  anti_join(dry, by = "CodeParacou") %>% 
  select(CodeParacou, Date, Time) %>% 
  kable(format = "html", escape = F,
        caption = "Missing individuals in LeafDry table.") %>%
  kable_styling("striped", full_width = F)
```

## Values check

```{r DryValues}
dry %>% 
  filter(DryWeight > 5) %>% 
  select(IdGenetic, CodeParacou, DryWeight) %>% 
  kable(format = "html", escape = F,
        caption = "DryWeight to be corrected in LeafFresh table.") %>%
  kable_styling("striped", full_width = F)
```

# Leaf Area

## Join files

Leaf Areas have two different types of files :

1. the Holes included files (*Leafarea_holesincluded.csv*) where each leaves' area were measured including holes of the limb
1. the holes excluded files (*Leafarea_holesexcluded.csv*) where no corrections were made to the area measurement

Thus, for each days of sampling (19 days), those two types of files where created. The LeafArea table needs to join data of areas for each leaves, and to bind all sampling values.

```{r LeafAreaData}
pathsHI<-list.files(path = path1, pattern="holesincluded.csv",full.names=TRUE,recursive = TRUE)
LAHI <-read_delim(pathsHI[1],delim=",")
for(path0 in pathsHI[2:length(pathsHI)]){
  LAHI <- bind_rows(LAHI,read_delim(path0,delim=","))}
pathsHE <- list.files(path=path1, pattern="holesexcluded.csv",full.names=TRUE,recursive = TRUE)
LAHE <-read_delim(pathsHE[1],delim=",")
for(path0 in pathsHE[2:length(pathsHE)]){
  LAHE <- bind_rows(LAHE,read_delim(path0,delim=","))}
LA<-full_join(LAHI,LAHE,by=c(" ","Label")) %>% 
  rename(LeafAreaHI = `Area.x`, LeafAreaHE = `Area.y`, StdDevHI = `StdDev.x`, StdDevHE = `StdDev.y`) #%>% 
  #write_delim(file.path("./","Indviduals.csv"), delim = ",")
```

Now we need to organize data and to give each leaf a number:

```{r DataOrganization}
LA <- LA %>%
  separate(Label,c("Date","Tree"),":") %>% 
  select(-Date) %>% 
  separate(Tree, c("Plot", "SubPlot", "TreeFieldNum"), "-", convert = T) %>% 
  mutate(Plot=str_to_upper(Plot)) %>% 
  mutate(Plot = as.numeric(gsub("P", "", Plot)))  %>% 
  mutate(FieldTreeNum0 = TreeFieldNum) %>% 
  mutate(TreeFieldNum = as.numeric(gsub('([[:alpha:]])', "", FieldTreeNum0))) %>% 
  mutate(Scan = gsub('[0-9]+', "", FieldTreeNum0)) %>% 
  select(-FieldTreeNum0) %>% 
  group_by(Plot, SubPlot, TreeFieldNum) %>% 
  arrange(Plot, SubPlot,TreeFieldNum,Scan) %>% 
  mutate(Leaf_order = (1:6)[1:n()]) #%>% 
  #write_delim(file.path("./","Indviduals_all.csv"), delim = ",")
```

## IDs check

Now we need to check and correct individual codes. 

```{r labelsvalues, warning=TRUE}
## Plot
if(!all(LA$Plot %in% 1:16))
  warning("Lines ", paste(which(!(LA$Plot %in% 1:16)), collapse = ', '),
          " have a wrong Plot number.\n")
## SubPlot
if(!all(LA$SubPlot %in% 1:25)){
  warning("Lines ", paste(which(!(LA$SubPlot %in% 1:25)), collapse = ', '),
          " have a wrong SubPlot number.\n")}
## TreeFieldNum
if(!all(LA$TreeFieldNum %in% 1:4000))
  warning("Lines ", paste(which(!(LA$TreeFieldNum %in% 1:4000)), collapse = ', '),
          " have a wrong TreeFieldNum number.\n")
```

```{r DataIndCheck, warning=TRUE}
individuals <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("Individuals") %>% 
  select(Date,Time,Plot,SubPlot,TreeFieldNum)
(Lacks <- anti_join(individuals,LA,by=c("Plot","SubPlot","TreeFieldNum")) %>% 
  select(Plot,SubPlot,TreeFieldNum) %>% 
  distinct() %>% 
  unite(Lacks,Plot,SubPlot,TreeFieldNum,sep="-") %>% 
  kable())
(Lacks2 <- anti_join(LA,individuals,by=c("Plot","SubPlot","TreeFieldNum")) %>%
  select(Plot,SubPlot,TreeFieldNum) %>% 
  distinct() %>%
  unite(Lacks,Plot,SubPlot,TreeFieldNum,sep="-") %>% 
  kable())
#for (x in Lacks[,1]){
#  print(Lacks2[amatch(x,as.character(Lacks2[,1]),maxDist=Inf),])}
#  warning(Lacks2[amatch(as.character(x),Lacks2,maxDist=Inf)],"is not a misspel of", x, "?\n")
```

Individuals were corrected and matched with individuals table.

```{r TableCorrected}
LAcor <- read_delim(file.path(path1,"IndividualsUnsortedLeaves.csv"),delim=",") %>% 
  select(-Leaf_order) %>% 
  group_by(Plot, SubPlot, TreeFieldNum) %>% 
  arrange(Plot, SubPlot,TreeFieldNum,Scan) %>% 
  mutate(Leaf_order = (1:6)[1:n()]) #%>% 
  #write_delim(file.path("./","Individuals_all.csv"), delim = ",")
```

3 types of files:
1. Individuals_all : after joining and binding all data
1. IndividualsUnsortedLeaves: after correcting column names and individuals ID
1. LeafArea: after correcting leaves order

## Values check

Here we checked wether leaves measures were missing or not.

```{r LeafNumber, warning=TRUE}
if(dim(distinct(select(LAcor,Plot,SubPlot,TreeFieldNum)))[1]!=dim(individuals)[1])
   warning(paste(dim(distinct(select(LAcor,Plot,SubPlot,TreeFieldNum)))[1],collapse=",")," individuals in LAcor wheareas ",paste(dim(individuals)[1],collapse=",")," in Individuals")
```

After ordering leaves, we still can observe some individuals where leaves number isn't equal to 5:
```{r LeavesNumberPerInd,warning=TRUE}
read_delim(file.path(path1,"Individuals_all.csv"),delim=",") %>% 
  group_by(Plot,SubPlot,TreeFieldNum) %>% 
  count() %>% 
  filter(n!=5) %>% 
  kable(format = "html", escape = F,
        caption = "Individuals with less then 5 leaves measured") %>%
  kable_styling("striped", full_width = F)
```


## Leaf attribution

```{r Aattribution, eval=F}
LA <- LA %>% 
  separate(Label, c("Plot", "SubPlot", "FieldTreeNum"), 
           "-", convert = T) %>% 
  mutate(Plot = as.numeric(gsub("P", "", Plot))) %>% 
  mutate(FieldTreeNum0 = FieldTreeNum) %>% 
  mutate(FieldTreeNum = as.numeric(gsub('([[:alpha:]])', "", FieldTreeNum0))) %>% 
  mutate(Scan = gsub('[0-9]+', "", FieldTreeNum0)) %>% 
  select(-FieldTreeNum0) %>% 
  arrange(Plot, SubPlot, FieldTreeNum, Area_exclude) %>% 
  group_by(Plot, SubPlot, FieldTreeNum) %>% 
  mutate(Leaf_order = (1:6)[1:n()])
data <- data %>% 
  arrange(Plot, SubPlot, FieldTreeNum, Fresh_Weight) %>% 
  group_by(Plot, SubPlot, FieldTreeNum) %>% 
  mutate(Leaf_order = (1:6)[1:n()])
data <- data %>% 
  left_join(LA)
# data %>% 
#   write_csv("./data.csv")
```

# All measures

AllMeasures table contains the follwoing fields with table from origin in bracket:

* idTree _(Paracou)_
* IdGenetic _(LeafFresh & LeafDry)_
* Plot _(All)_
* SubPlot _(All)_
* TreeFieldNum _(All)_
* Leaf _(LeafFresh, LeafDry and LeafArea)_
* Genus _(Paracou)_
* Species _(Paracou)_
* DBH _(Paracou)_
* Dawkins _(Individuals)_
* LAexclude _(LeafArea)_
* LAinclude _(LeafArea)_
* FreshWeight _(LeafFresh)_
* DryWeight _(LeafDry)_
* SPADmean _(LeafFresh, mean(SPAD1, SPAD2, SPAD3))_
* LTmean _(LeafFresh, mean(LT1, LT2, LT3))_

```{r AllMeasuresData, eval=F}
individuals <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("Individuals") %>% 
  dplyr::select(-X1) %>% 
  mutate(CodeParacou = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum)))
paracou <- src_sqlite(file.path("../../../data/Paracou/", 
                                "trees/Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  mutate(CodeParacou = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum))) %>% 
  filter(CodeParacou %in% individuals$CodeParacou) %>% 
  select(-CodeParacou) %>% 
  filter(Plot %in% c(1, 6, 11, 13:15) & CensusYear == 2017 |
           Plot == 16 & CensusYear == 2015) %>% 
  mutate(DBH = CircCorr*pi) %>% 
  collect()
individuals <- select(individuals, -CodeParacou)
fresh <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("LeafFresh") %>% 
  dplyr::select(-X1)
dry <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("LeafDry") %>% 
  dplyr::select(-X1)
area <- googlesheets::gs_title("Measures_Eschweilera") %>%
  googlesheets::gs_read("LeafArea") %>%
  dplyr::select(-X1)
```

```{r AllMeasuresJoin, eval=F}
paracou %>% 
  left_join(individuals, by = c("Plot", "SubPlot", "TreeFieldNum"), 
            suffix = c("", ".individuals")) %>% 
  left_join(fresh, by = c("Plot", "SubPlot", "TreeFieldNum"),
            suffix = c("", ".fresh")) %>% 
  group_by(idTree, IdGenetic, Plot, SubPlot, TreeFieldNum, Leaf) %>% 
  mutate(SPADmean = mean(c(SPAD1, SPAD2, SPAD3), na.rm = T)) %>% 
  mutate(LTmean = mean(c(LT1, LT2, LT3), na.rm = T)) %>% 
  ungroup() %>% 
  left_join(dry, by = c("Plot", "SubPlot", "TreeFieldNum", "Leaf"),
  suffix = c("", ".dry")) %>%
  left_join(area, by = c("Plot", "SubPlot", "TreeFieldNum", "Leaf"),
            suffix = c("", ".area")) %>%
  select(idTree, IdGenetic, Plot, SubPlot, TreeFieldNum, Leaf, Genus, Species, DBH,
         Dawkins, LAexclude, LAinclude, FreshWeight, DryWeight, SPADmean, LTmean) %>% 
  write_delim(file.path(path, "AllMeasures.csv"), delim = ",")
```

# All traits

AllTraits have fields similar to AllMeasures (idTree, IdGenetic, Plot, SubPlot, TreeFieldNum, Leaf, Genus, Species, DBH, Dawkins) and following final functional traits:

* $SLA = \frac{DryWeight}{LAexclude}$
* $LDMC = \frac{DryWeight}{FreshWeight}$
* $LT = LTmean$
* $LA = LAexclude$
* $CC = \frac{117.1*SPAD}{148.84*SPAD}$ Homographic model for all species from @Coste2010

```{r AllTraits, eval=F}
googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("AllMeasures") %>% 
  mutate(SLA = LAexclude/DryWeight) %>% 
  mutate(LDMC = DryWeight/FreshWeight) %>% 
  mutate(LT = LTmean) %>% 
  mutate(LT = ifelse(is.nan(LT), NA, LT)) %>% 
  mutate(LA = LAexclude) %>% 
  mutate(CC = (117.1*SPADmean)/(148.84-SPADmean)) %>% 
  mutate(CC = ifelse(is.nan(CC), NA, CC)) %>% 
  select(IdTree, IdGenetic, Plot, SubPlot, TreeFieldNum, Leaf, Genus, Species, DBH,
         Dawkins, SLA, LDMC, LT, LA, CC) %>% 
  write_delim(file.path(path, "AllTraits.csv"), delim = ",")
```

# Conclusion

Data are ready, so enjoy analysis now ;)

# References


