---
title: "Eschweilera outliers"
date: '`r Sys.Date()`'
author: Sylvain Schmitt & Anne Baranger
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
linestretch: 1.5
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(ggfortify)
library(plotly)
library(kableExtra)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
path <- "../../../data/Paracou/"
```

```{r data}
data <- googlesheets::gs_title("Measures_Eschweilera") %>% 
  googlesheets::gs_read("AllTraits") %>% 
  filter(idTree!=141909) #outliers removed
# paracou <- src_sqlite(file.path(path, "trees", "Paracou.sqlite")) %>% 
#   tbl("Paracou") %>% 
#   filter(idTree %in% data$idTree) %>% 
#   filter(Plot %in% c(1, 6, 11, 13:15) & CensusYear == 2017 |
#            Plot == 16 & CensusYear == 2015) %>% 
#   collect()
# data <- left_join(data, paracou) ; rm(paracou)
```

# Introduction

Subsequent analysis aimed to detect leaf measurements outliers within functional traits (either individually with boxplots or together with principal component analysis). Then, we checked if outliers were natural or measurement errors. If measurment errors were detected they have been further cleaned either by removing them, making the good link or reproducing the measure.

# Outliers

## Identified

Thanks to subsequent boxplot (fig \@ref(fig:outliersBoxplot) and principal component analysis (PCA, fig \@ref(fig:outliersPCA), we detected following outliers:

* __P14-1-760__ all leaves, is a natural outlier identified in the PCA due to an anormally high SLA due to the fact that the sample comse from a tree resprout grown on a beheaded tree. Was removed from the datatable for further outliers detection and analysis.
* __P13-2-881__ seemed to be offset on the PCA, but after checking value all traits were in a normal range except for the CC. Leaves sampled seemed to be young ones and wich could explain low SPAD values. On the other hand, the indivdual was a *E. pedicellata* with different values from the rest of the genus, spanning fully the range of CC. Consequently, we kept the observation.
* __P16-3-260__ seemed to be offset on the PCA, what could be due to abnormal values of LDMC and SLA. We discovered spurious value of dry weight that were repeated and corrected.
* __P16-7-185-5__ have a particularly high SLA regarding other leaves of the same individual. LA values were badly attributed regarding leaves scan. Subsequently, LA data table has been corrected and updated.

## Boxplot

Boxplots (fig \@ref(fig:outliersBoxplot) allow to identify outliers individuall on each functional traits. Table \@ref(tab:outliersTable) present outliers outside of the 5-95% interval (the value can be changed in easily in the Rmarkdown).

```{r outliersBoxplot, fig.height=6, fig.width=12, fig.cap="Functional Traits Outliers Detection through Boxplots with Plotly."}
data %>% 
  mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>% 
    mutate(Code = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum),
                              "-", as.character(Leaf))) %>% 
  select(Code, SpeciesLong, SLA, LDMC, LT, LA, CC) %>% 
  reshape2::melt(id.vars = c("Code", "SpeciesLong"),
                 variable.name = "Trait") %>% 
  group_by(Trait) %>% 
  mutate(value = scale(value)) %>% 
  plot_ly(x = ~Trait, y = ~value, color = ~SpeciesLong,
          text = ~Code, type = "box") %>%
  layout(boxmode = "group")
```

```{r outliersTable}
o <- 0.05
data %>% 
  mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>% 
    mutate(Code = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum),
                              "-", as.character(Leaf))) %>% 
  select(Code, SpeciesLong, SLA, LDMC, LT, LA, CC) %>% 
  reshape2::melt(id.vars = c("Code", "SpeciesLong"),
                 variable.name = "Trait") %>% 
  group_by(Trait, SpeciesLong) %>% 
  filter(!is.na(value)) %>% 
  filter(value < quantile(value, o) - 1.5*IQR(value) | value > quantile(value, (1-0)) + 1.5*IQR(value)) %>% 
  select(Trait, SpeciesLong, Code, value) %>% 
  arrange(Trait, desc(value)) %>% 
  kable(caption = "Outliers outside of the 5%-95% confidenc interval.",
        format = "html", escape = F) %>% 
  kable_styling("striped", full_width = F) %>% 
  collapse_rows(1)
```

## Principal Component Analysis

Principal Component Analysis (PCA, fig \@ref(fig:outliersPCA) allow to identify outliers with spurious traits relations. Table below allow to search for outliers in the whole dataset (search engine set to P14-1-760 for instance).

```{r outliersPCA, fig.cap="Functional Traits Outliers Detection through PCA with Plotly."}
data %>% 
  mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>% 
    mutate(Code = paste0("P", as.character(Plot), 
                              "-", as.character(SubPlot),
                              "-", as.character(TreeFieldNum),
                              "-", as.character(Leaf))) %>% 
  select(Code, SpeciesLong, DBH, SLA, LDMC, LT, LA, CC) %>% 
  na.omit() %>% 
  fortify(princomp(~ SLA + LDMC + LT + LA + CC, data = ., cor = T), .) %>% 
  plot_ly(x = ~Comp.1 , y = ~Comp.2, text = ~Code,
          type = "scatter", mode = "markers", color = ~Code,
          opacity = 0.5, maxdisplayed = 100) %>% 
  layout(showlegend = FALSE) %>% 
  layout(xaxis = list(title = "PC 1"), yaxis = list(title = "PC 2"))
```

```{r DataTable}
data %>% 
  mutate(SpeciesLong = paste(substr(Genus, 1, 1), Species)) %>% 
  mutate(Code = paste0("P", as.character(Plot), 
                       "-", as.character(SubPlot),
                       "-", as.character(TreeFieldNum),
                       "-", as.character(Leaf))) %>% 
  select(Code, SpeciesLong, SLA, LDMC, LT, LA, CC) %>% 
  DT::datatable(options = list(
    searchHighlight = TRUE, search = list(search = 'P13-2-881'),
    initComplete = htmlwidgets::JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}")))
```

# Conclusion

Finally, we need to remove sampled species from outgroups that does not belong to *Eschweilera clade Parvifolia* and the dataset will be fully ready for functional analysis in addition to the *Symphonia globulifera* dataset (traits covariation by principal component analysis and trait-environment relation by bayesian inferences).
