---
title: "Sampling"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(flexdashboard)
library(rgdal)
library(raster)
library(leaflet)
library(starmie)
library(rstan)
library(bayesplot)
library(abind)
library(ggrepel)
library(tidyverse)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
```

# General

##

### To Do

1. Material command:
    + $\sim 10$ Sac de lancer Libre 250g
    + $2-4$ Fil de lancer shaoLINE Teufelberger 60m
    + $\sim 2$ Élastique BIG SHOT
    + $\sim 6$ Gants fins ?
    + $\sim 4$ casques ?
1. Sampling design
    + Plots: Control & Biodiversity ?
    + Individuals: 400-500 ?
    + Traits
        * Leaf (SLA, LDMC, LT, LA, SPAD) 
        * Branch (BrWD, BrBD, BrBT)
        * Trunk (WD, BT) ?
        * Anatomy ?
        * Physiology ?
1. FTH module design
    + Pre-sampling
    + Plots
    + Question / Hypothesis / Analysis

### Docs

* [Field protocol](https://docs.google.com/document/d/1yjYXe50mYG95mUjw61SujkdDBxzIDkbvcJbXg8w6fPo/edit?usp=sharing)
* [Fresh measurements protocol](https://docs.google.com/document/d/1S-0O5ndADJhtXlrtz6qgRfYZtfS9f1qOgvFlaEF6P8M/edit?usp=sharing)
* [Symphonia individuals](https://docs.google.com/spreadsheets/d/13BKXG1ETpTvQwcpYcGoTK9eTU002PSWa4Cjyox-OanA/edit?usp=sharing)
* [Symphonia leaves](https://docs.google.com/spreadsheets/d/1QycHaAmsbBGd76Ma088RXxvFrWT5c_9Md3Eqqob286Q/edit?usp=sharing)
* [Dawkins index](https://drive.google.com/file/d/0B-lAe8V-ri2SUVZrRGJpQUhUVkU/view?usp=sharing)

##

### Planning

1. 31.07-03.08 Material command
1. 21.09-14.09 pre-FTH field (2-3 days)
1. 17.09-21.09 FTH field work (4-5 days)
1. 24.09-28.09 FTH analysis and defense
1. 03.08-05.08 post-CEBA field work w Myriam (2-3 days)
1. +2-4 weeks of field sampling (depending on individuals objective)

### Maps

* [Parcelle 1](./sampling_save/maps/Parcelle1.png)
* [Parcelle 5](./sampling_save/maps/Parcelle5.png)
* [Parcelle 9](./sampling_save/maps/Parcelle9.png)
* [Parcelle 13](./sampling_save/maps/Parcelle13.png)
* [Parcelle 14](./sampling_save/maps/Parcelle14.png)
* [Parcelle 15](./sampling_save/maps/Parcelle15.png)
* [Parcelle 16](./sampling_save/maps/Parcelle16.png)


# Samples

```{r Par_trees}
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
trees <- read.csv(file.path(path, 'trees/Eschweilera_Paracou.csv'),
                  dec=".", stringsAsFactors = F)
trees <- trees %>% 
  filter(CensusYear == 2015) %>% 
  filter(CodeAlive == 1) %>% 
  filter(CodeMeas == 0) %>% 
  filter(Plot != 18) %>% 
  mutate(Treatment = c(rep(c('C', 'T1', 'T2', 'T3'), 3), rep('B', 4))[Plot]) %>% 
  mutate(Treatment_simp = ifelse(Treatment == "C", "Control", ifelse(Treatment == "B", "Biodiv", "Treatment"))) %>% 
  mutate(Clade = ifelse(Species %in% c("congestiflora", "simiorum"),
                        "Other", "Parvifolia")) %>% 
  mutate(SpeciesType = ifelse(Clade == "Other", "outgroup", NA)) %>%
  mutate(SpeciesType = ifelse(Species == "sagotiana", "sagotiana", SpeciesType)) %>%
  mutate(SpeciesType = ifelse(Species == "coriacea", "coriacea", SpeciesType)) %>%
  mutate(SpeciesType = ifelse(Species == "decolorans", "decolorans", SpeciesType)) %>%
  mutate(SpeciesType = ifelse(Species %in% c("pedicellata", "Indet.", "wachenheimii",
                                             "grandiflora_form2", "collina", 
                                             "grandiflora"), "other", SpeciesType)) %>%
  mutate(SpeciesType = ifelse(Species %in% c("micrantha", "parviflora", "squamata"), 
                              "rare", SpeciesType)) %>% 
  filter(Treatment %in% c("B", "C"))
```

```{r filters}
# Total of 4.5 plates = 4.5*96 = 432
sagotiana <- trees %>%
  filter(Species == "sagotiana") %>% 
  sample_n(144) # 1/3 = 144
coriacea <- trees %>%
  filter(Species == "coriacea") %>% 
  sample_n(144) # 1/3 = 144
decolorans <- trees %>%
  filter(Species == "decolorans") # 1/6 = 72 but 66
rares <- trees %>%
  filter(Species %in% c("pedicellata", "wachenheimii")) %>%
  group_by(Species) %>%
  sample_n(22) %>%   # 1/6 = 72 - others (2*3) - Lecythis (3) - rare(3) - grandiflora (2) - collina (5) = 60
  ungroup() %>%
  bind_rows(trees %>%
              filter(Species %in% c("collina", "grandiflora_form2", "grandiflora"))) %>%
    bind_rows(trees %>%
              filter(SpeciesType == "rare"))
outgroup <- trees %>% 
  filter(Clade == "Other") %>% 
  group_by(Species) %>% 
  sample_n(3) %>% 
  ungroup()
lecythis <- read.csv(file.path(path, 'trees/Lecithis_Paracou.csv'),
                  dec=".", stringsAsFactors = F) %>% 
    filter(CensusYear == 2015) %>% 
  filter(CodeAlive == 1) %>% 
  filter(CodeMeas == 0) %>% 
  filter(Plot != 18) %>% 
  mutate(Treatment = c(rep(c('C', 'T1', 'T2', 'T3'), 3), rep('B', 4))[Plot]) %>% 
  filter(Treatment %in% c("B", "C")) %>% 
  sample_n(3)
selection <- bind_rows(sagotiana, coriacea, decolorans, rares, outgroup, lecythis) %>% 
  mutate(selection = 1)
trees <- trees %>% 
  full_join(selection)
```

##

> Original headcounts

### Species headcount

```{r species}
trees %>% 
  filter(Species != "Indet.") %>% 
  reshape2::dcast(Clade + SpeciesType + Genus + Species ~ 1) %>% 
  arrange(desc(Clade), desc(`1`)) %>% 
  kable()
```

### Species and treatment headcount

```{r treatment}
trees %>% 
  filter(Species != "Indet.") %>% 
  reshape2::dcast(Clade + SpeciesType + Genus + Species ~ Treatment) %>% 
  arrange(desc(Clade), desc(B)) %>% 
  kable()
```

###  Species and plot headcount

```{r plot}
trees %>% 
  filter(Species != "Indet.") %>% 
  reshape2::dcast(Clade + SpeciesType + Genus + Species ~ Plot) %>% 
  arrange(desc(Clade), desc(`16`)) %>% 
  kable()
```

###  Species and diameters

```{r diam}
trees %>% 
  filter(Species != "Indet.") %>% 
  ggplot(aes(CircCorr, fill = Species)) + 
  geom_histogram(position = "dodge")
```

##

> Number of trees selected: `r sum(trees$selection, na.rm = T)`

### Selected species headcount

```{r speciessel}
trees %>% 
  filter(selection == 1) %>% 
  reshape2::dcast(Clade + SpeciesType + Genus + Species ~ 1) %>% 
  arrange(desc(Clade), desc(`1`)) %>% 
  kable()
```

### Species and treatment headcount

```{r treatmentsel}
trees %>% 
  filter(selection == 1) %>% 
  reshape2::dcast(Clade + SpeciesType + Genus + Species ~ Treatment) %>% 
  arrange(desc(Clade), desc(B)) %>% 
  kable()
```

###  Species and plot headcount

```{r plotsel}
trees %>% 
  filter(selection == 1) %>% 
  reshape2::dcast(Clade + SpeciesType + Genus + Species ~ Plot) %>% 
  arrange(desc(Clade), desc(`16`)) %>% 
  kable()
```

###  Species and diameters

```{r diamsel}
trees %>% 
  filter(selection == 1) %>% 
  ggplot(aes(CircCorr, fill = Species)) + 
  geom_histogram(position = "dodge")
```

# Maps

### Dynamic map of selected individuals

```{r selected}
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
limits <- readOGR(file.path(path, 'limits'), "OverallPlots", verbose = F)
limits <- spTransform(limits, CRSobj = crs)
sel_pal <- colorBin(c('transparent', 'black'), trees$selection)
treesXY <- trees
coordinates(treesXY) <- ~Xutm + Yutm
proj4string(treesXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
treesXY <- spTransform(treesXY, CRSobj = crs)
leaflet() %>%
  addPolygons(data = limits, fillOpacity = 0,
              opacity = 0.5, col = 'black') %>% 
  addCircles(data = treesXY, col = 'black', fillOpacity = 1,
             fillCol = sel_pal(trees$selection), radius= ~CircCorr/25,
             label = ~as.character(TreeFieldNum)) %>% 
  addLegend(colors = c('transparent', 'black'), 
            labels = c('not selected', 'selected'),
            title = 'Sampling')
```

```{r maps functions, eval=F}
rotateProj = function(spobj, angle) {
    boxpts = SpatialPoints(t(bbox(spobj)), proj4string = CRS(proj4string(spobj)))
    boxLL = bbox(spTransform(boxpts, CRS("+init=epsg:4326")))
    llc = apply(boxLL, 1, mean)
    prj = paste0("+proj=omerc +lat_0=", llc[2], " +lonc=", llc[1], " +alpha=", 
        angle, " +gamma=0.0 +k=1.000000 +x_0=0.000 +y_0=0.000 +ellps=WGS84 +units=m ")
    CRS(prj)
}
find_angle <- function(spobj){
  spobj <- subset(spobj, Subplot == unique(spobj$Subplot)[1])
  spobj <- fortify(spobj) %>% 
    filter(long == min(long) | lat == min(lat)) %>% 
    select(long, lat) %>% 
    distinct()
  row.names(spobj) <- c('A', 'B')
  spobj['C',] <- apply(spobj, 2, min)
  Pab <- sqrt(sum(apply(spobj[c('A', 'B'),], 2, diff)^2))
  Pac <- sqrt(sum(apply(spobj[c('A', 'C'),], 2, diff)^2))
  Pbc <- sqrt(sum(apply(spobj[c('B', 'C'),], 2, diff)^2))
  angle <- acos((Pab^2 + Pac^2 - Pbc^2) / (2 * Pab * Pac))
  angle <- angle*180/pi
  angles <- c(angle, angle-90)
  angle <- angles[which(sapply(c(angle, angle-90), abs) == min(sapply(c(angle, angle-90), abs)))]
  return(angle)
}
make_map <- function(limits, treesXY, contour, p){
  limits_plot <-  subset(limits, Plot == p)
  if(p != 16)
    limits_plot <-  subset(limits_plot, TypeSub == "Subplots")
  else
    limits_plot <-  subset(limits_plot, TypeSub == "SubplotsP16")
  newProj <- rotateProj(limits_plot, find_angle(limits_plot))
  limits_plot <- spTransform(limits_plot, newProj)
  treesXY_plot <- subset(treesXY, Plot == p)
  treesXY_plot <- spTransform(treesXY_plot, newProj)
  contour_plot <- spTransform(contour, newProj)
  contour_plot <- crop(contour_plot, extent(limits_plot))
  g <- ggplot() +
    geom_polygon(data = fortify(limits_plot),
                 aes(long, lat, group=group),
                 fill = "transparent", colour = "black", alpha = 0.2) +
    geom_polygon(data = fortify(contour_plot),
                 aes(long, lat, group=group),
                 colour = "grey", alpha = 0) +
    geom_point(data = as.data.frame(treesXY_plot),
               aes(x = Xutm, y = Yutm, size=CircCorr/pi*100,
                   col = selection)) +
    geom_text_repel(data = as.data.frame(treesXY_plot),
                    aes(Xutm, Yutm, 
                        label = TreeFieldNum)) +
    ggtitle(paste('Parcelle', p)) +
    coord_equal() +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    scale_color_continuous(palette = colorBin( c('lightgrey', 'black'), domain = c(0,1))) +
    ggsn::north(fortify(limits_plot), symbol = 12, 
                scale = 0.05, anchor = c(x= 125, y = 140))
  return(g)
}
```

```{r maps, eval=F}
contour <- readOGR(file.path(path, 'topography'), "ContourLinePlots", verbose = F)
contour <- spTransform(contour, CRSobj = crs)
treesXY$selection[is.na(treesXY$selection)] <- 0
maps <- sapply(sort(unique(trees$Plot)), 
               function(x) make_map(limits, treesXY, contour, x),
               simplify = F)
names(maps) <- paste0('P', sort(unique(trees$Plot)))
lapply(maps, function(x)
  ggsave(x, file = paste0(gsub(' ', '', x$labels$title), '.png'),
         path = './sampling_save/maps',
         width = 297, height = 420, unit = 'mm',
         dpi = 300))
```

```{r table, eval=F}
trees %>%
  filter(selection == 1) %>%
  arrange(Plot, SubPlot, TreeFieldNum) %>% 
  write.csv("./sampling_save/trees_sel.csv",
            row.names = F)
```

