---
output:
  pdf_document: default
  html_document: default
---
```{r setup_func_sel, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(parallel)
library(Biostrings)
library(tidyverse)
library(kableExtra)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/"
```

# Fuctional region selection


We used open reading frames (ORF) to target genes within scaffolds. ORFs have been detected with `transdecoder` on assembled transcripts. First, we filtered ORFs including a start codon(figure \@ref(fig:orfFilterGraph)). Then, we aligned ORFs on pre-selected and merged genomic scaffolds with `blat`. We obtained 7 744 aligned scaffolds (table \@ref(tab:blatCovORF) and figure \@ref(fig:blatFigORF)). Thanks to the alignment, we removed overlapping genes (figure \@ref(fig:orfDist)) and obtained 4 076 pre-selected genes with a total length of 757 hbp (figure \@ref(fig:genesPlot)). Finally we used transcript differential expression to select all genes differentially expressed between *Symphonia globulifera* and *Symphonia sp1* (figure \@ref(fig:diffExprFig)). We selected 1150 sequences of 500 to 1-kbp representing 1 063 Mbp (table \@ref(tab:functionalSelectionTab)). To validate our final target set, we aligned with `bwa` raw reads from one library from Scotti *et al. (in prep)*. ...


## Open Reading Frames (ORFs) filtering

173 828 ORFs including a start codon (Methyonin, M) were detected (over 231 883, 75%, see figure \@ref(fig:orfFilterGraph).

```{r orfFilterGraph, fig.cap="Open Reading Frames left and right peptides."}
src_sqlite(file.path("~/Documents/BIOGECO/PhD/data/Symphonia_Niklas/Trinotate/",
                           "symphonia.trinity500.trinotate.sqlite")) %>% 
  tbl("ORF") %>% 
  dplyr::rename(orf = orf_id, trsc = transcript_id, width = length) %>% 
  collect() %>% 
  mutate(left = substr(peptide, 1, 1)) %>% 
  mutate(right = substr(peptide, nchar(peptide), nchar(peptide))) %>% 
  select(orf, strand, left, right) %>% 
  reshape2::melt(id.vars = c("orf", "strand"),
                 variable.name = "side",
                 value.name = "peptide") %>% 
  ggplot(aes(peptide, fill = (peptide %in% c("M", "*")))) +
  geom_bar() +
  facet_grid(strand~side, labeller = "label_both") +
  scale_fill_discrete(guide = F)
```

```{r orfFilterBed, echo=T}
orf <- src_sqlite(file.path("~/Documents/BIOGECO/PhD/data/Symphonia_Niklas/Trinotate/",
                           "symphonia.trinity500.trinotate.sqlite")) %>% 
  tbl("ORF") %>% 
  dplyr::rename(orf = orf_id, trsc = transcript_id,
                orfSize = length) %>% 
  filter(substr(peptide, 1, 1) == "M") %>% 
  select(-peptide, -strand) %>%
  collect() %>% 
  rowwise() %>% 
  mutate(orfStart = min(as.numeric(lend), as.numeric(rend)),
         orfEnd = max(as.numeric(lend), as.numeric(rend))) %>% 
  select(-lend, -rend) %>% 
  select(trsc, orfStart, orfEnd, orf) %>% 
  mutate_if(is.numeric, as.character) %>% 
  write_tsv(path = file.path(path, "functional_selection2", "orf.all.bed"),
            col_names = F)
```

## ORF alignment on genomics scaffolds

7 744 scaffolds matched with ORFs (10.5% for 15.4 Mbp, see table \@ref(tab:blatCovORF) and figure \@ref(fig:blatFigORF)).

```{bash orfAlign, eval=F, echo=T}
cd ~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/functional_selection2
cat orf.all.bed | sort -k 1,1 -k2,2n >  orf.all.sorted.bed
trsc=/home/sylvain/Documents/BIOGECO/PhD/data/Symphonia_Niklas/filtered_transcripts.fasta
bedtools getfasta -name -fi $trsc -bed orf.all.sorted.bed -fo orf.fasta
scf=~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/neutral_selection/merged.fasta
orf=./orf.fasta
blat $scf $orf alignment.psl
```

```{r blatResORF}
alignment <- read_tsv(file.path(path, "functional_selection2", "alignment.psl"), skip = 4, 
         col_names = c("matches", "misMatches", "repMatches", "nCount", 
                       "qNumInsert", "qBaseInsert",
                       "tNumInsert", "tBaseInsert", "strand", 
                       "qName", "qSize", "qStart", "qEnd", 
                       "tName", "tSize", "tStart", "tEnd", 
                       "blockCount", "blockSizes", "qStarts", "tStarts")) %>% # psl3 generics
  dplyr::rename(orf = qName, orfSize = qSize, orfStart = qStart, orfEnd = qEnd, 
         scf = tName, scfSize = tSize, scfStart = tStart, scfEnd = tEnd) %>% 
  mutate(alnSize = matches + misMatches + repMatches) %>% 
  select(alnSize,
         orf, orfSize, orfStart, orfEnd,
         scf, scfSize, scfStart, scfEnd)
```

```{r  blatCovORF}
total_length <- sum(width(readDNAStringSet(file.path(path, "neutral_selection", "merged.fasta"))))
n_total <- length(readDNAStringSet(file.path(path, "neutral_selection", "merged.fasta")))
data.frame(name = c("aligned sequence", "selected scaffold", "total"),
           n = c(nrow(alignment), length(unique(alignment$scf)), n_total),
           width = c(sum(alignment$alnSize), 
                     sum(unique(select(alignment, scf, scfSize))$scfSize), 
                     total_length)) %>% 
  mutate(coverage = width/total_length*100) %>% 
  mutate(width = width / 10^6) %>% 
  kable(caption = "Alignment coverage of  Tysklind et al. (in prep) ORFs over genomic scaffolds with blat.",
        col.names = c("", "N", "Width (Mbp)", "Coverage (%)"),
        format.args = list(big.mark = " "))
```

```{r blatFigORF, fig.cap="Alignment result of Tysklind et al. (in prep) ORFs over genomic scaffolds with blat. Left graph represents the overlap distribution. Right graph represent the selected and deduplicated scaffolds distribution."}
g1 <- ggplot(alignment, aes(alnSize)) +
  geom_histogram() +
  scale_x_log10() +
  xlab("Overlap length of selected transcript\n(from Tysklind et al. (in prep) over genomic scaffolds") +
  coord_flip()
scf <- readDNAStringSet(file.path(path, "Ivan_2018", "transcript_alignment", "unique_selected_scaffolds.fa"))
g2 <- alignment %>% 
  select(scf, scfSize) %>% 
  unique() %>% 
  ggplot(aes(scfSize)) +
  geom_histogram() +
  scale_x_log10() +
  xlab("Selected scaffolds width (bp)") +
  coord_flip()
cowplot::plot_grid(g1, g2)
```

## Overlaping genes filtering

995 genes were overlapping and filtered out (figure \@ref(fig:orfDist)).

```{r genesFilter, echo=T}
alignment %>% 
  separate(orf, into = c("gene", "isoform", "geneNumber", "orfNumber"),
           sep = "::", remove = F) %>% 
  select(gene, orf, orfStart, orfEnd, scf, scfStart, scfEnd) %>% 
  unique() %>% 
  select(scf, scfStart, scfEnd, orf, gene) %>% 
  arrange(scf, scfStart, scfEnd) %>% 
  mutate_if(is.numeric, as.character) %>% 
  write_tsv(path = file.path(path, "functional_selection2", "genes.all.bed"),
            col_names = F)
```

```{bash orfBedMerged, eval=F, echo=T}
cd ~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/functional_selection2
cat genes.all.bed | sort -k 1,1 -k2,2n >  genes.all.sorted.bed
bedtools merge -i genes.all.sorted.bed -c 5 -o collapse > genes.merged.bed
```

```{r orfDist, fig.cap="Genes overlap", fig.height=4}
read_tsv(file.path(path, "functional_selection2", "genes.merged.bed"), 
                col_names = c("scf",  "start",  "end",  "gene")) %>% 
  separate_rows(gene, sep = ",") %>% 
  unique() %>% 
  group_by(scf, start, end) %>% 
  summarise(genes = n())  %>% 
  ggplot(aes(genes, fill = (genes == 1))) + 
  geom_histogram(binwidth = 1) +
  xlab("Number of overlapping genes\n(1 = no overlap)") +
  xlim(0, 12) +
  scale_fill_discrete(guide = F)
```

## Pre-selected genes

We obtained 4 076 genes pre-selected for a total length of 757 kbp (figure \@ref(fig:genesPlot)).

```{r genesFilter2}
overlapping_genes <- read_tsv(file.path(path, "functional_selection2", "genes.merged.bed"), 
                col_names = c("scf",  "start",  "end",  "gene")) %>% 
  separate_rows(gene, sep = ",") %>% 
  unique() %>% 
  group_by(scf, start, end) %>% 
  filter(n() > 1) %>% 
  ungroup() %>% 
  select(gene) %>% unique() %>% unlist()
genes <- alignment %>% 
  separate(orf, into = c("gene", "isoform", "geneNumber", "orfNumber"),
           sep = "::", remove = F) %>% 
  select(gene, orf, orfStart, orfEnd, scf, scfSize, scfStart, scfEnd) %>% 
  unique() %>% 
  filter(!(gene %in% overlapping_genes)) %>% 
  group_by(scf, scfSize, gene) %>% 
  summarise(geneStart = min(scfStart), geneEnd = max(scfEnd)) %>% 
  mutate(geneSize = geneEnd - geneStart) %>% 
  ungroup()
```


```{r genesPlot, fig.cap="Available genes for target sequences design."}
genes %>% 
  ggplot(aes(scfSize/1000, geneSize)) + 
  geom_hline(yintercept = 1000, col = "red", lwd = 1) +
  geom_point(alpha = 0.5) +
  xlab("Scaffold width (kbp)") +
  ylab("Gene width (kbp)") +
  scale_y_log10() +
  ggtitle(paste(nrow(genes), "genes"), 
          paste(format(sum(genes$geneSize)/1000, big.mark = " "), "kbp"))
```

## Differential Expression (DE) of genes

Figure \@ref(fig:diffExprFig) shows genes differential expression. First circle represent genes with isoforms not enriched whereas second and third circle represents respectivelly genes with isoforms *S. sp1* and *S. globulifera* enriched. Relatively few genes contained enriched isoforms, and most of them were *S. globulifera* enriched.

```{r diffExprFig, fig.cap="Genes differential expression."}
de <- read_delim(file.path(path, "..", "Symphonia_Niklas", "edgeR",
                     "juv_sympho_43k_retained_genes_DecisionTest.txt"), 
           delim = " ") %>% 
  select(-X) %>% 
  dplyr::rename(trsc = genes, deg = X.0.5.glo_HT..0.5.glo_SF.0.5.sp1_HT.0.5.sp1_SF) %>%
  mutate(deg = ifelse(deg == 1, "Eglo", deg)) %>% 
  mutate(deg = ifelse(deg == -1, "Esp", deg)) %>% 
  mutate(deg = ifelse(deg == 0, "NE", deg)) %>% 
  mutate(gene = gsub("_i[0-9]", "", trsc)) %>% 
  filter(gene %in% genes$gene) %>% 
  select(gene, deg) %>% 
  unique()
de %>% 
  count(gene, deg) %>% 
  tidyr::spread(deg, n, fill=0) %>% 
  mutate(NE, NE = recode(NE, "1"="Not enriched")) %>% 
  mutate(Eglo, Eglo = recode(Eglo, "1"="Globulifera enriched")) %>% 
  mutate(Esp, Esp = recode(Esp, "1"="Sp1 enriched")) %>% 
  mutate(deg = paste(NE, Esp, Eglo, sep = "-")) %>% 
  count(deg) %>% 
  select(deg, n) %>% 
  sunburstR::sunburst(count = T)
```

## Final subset of selected functional scaffolds

We finally selected 1150 sequences of 500 to 1-kbp with 100 bp before geneStart and a maximum of 900 bp after, resulting in 1 063, 544 kbp of targets. All differentially expressed genes between morphotypes were selected (159). And the rest of sequences were selected among non differentially expressed genes randomly (1001).

```{r finalSubset, echo=T}
targets <- genes %>% 
  left_join(de) %>% 
  filter(deg %in% c("Eglo", "Esp")) %>% 
  rbind(genes %>% 
          left_join(de) %>% 
          filter(deg %in% c("DE", NA)) %>% 
          sample_n(1309)) %>% 
  group_by(scf, gene) %>% 
  filter(n() < 2) %>% 
  mutate(targetStart = max(0, geneStart - 100)) %>% 
  mutate(targetEnd = min(targetStart + 1000, scfSize)) %>% 
  mutate(targetSize = targetEnd - targetStart) %>% 
  filter(targetSize > 500) %>% 
  mutate(target = paste0(gene, "_on_", scf)) %>% 
  ungroup()
targets %>% 
  select(scf, targetStart, targetEnd, target) %>% 
  arrange(scf, targetStart, targetEnd, target) %>% 
  mutate_if(is.numeric, as.character) %>% 
  write_tsv(path = file.path(path, "functional_selection2", "targets.all.bed"),
            col_names = F)
```

```{bash targetFasta, eval=F, echo=T}
cd ~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/functional_selection2
cat targets.all.bed | sort -k 1,1 -k2,2n >  targets.all.sorted.bed
scf=~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/neutral_selection/merged.fasta
bedtools getfasta -name -fi $scf -bed targets.all.sorted.bed -fo targets.fasta
```

```{r functionalSelectionTab}
scf <- readDNAStringSet(file.path(path, "functional_selection2", "targets.fasta"))
data.frame(n = length(scf),
           width = sum(width(scf))/10^6,
           Nperc = sum(letterFrequency(scf, "N"))/sum(width(scf))) %>% 
    kable(caption = "Selected functional targets",
        col.names = c("N", "Width (Mbp)", "Mask (%N)"),
        format.args = list(big.mark = " "))
```

## Repetitive regions final check

Last but not least, we do not want to include repetitive regions in our targets for baits design. We consequently aligned raw reads from one library from Scotti *et al. (in prep)* on our targets with `bwa`. 

```{bash funcBWAcheck, eval=F, echo=T}
cd ~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/functional_selection2
reference=targets.fasta
query=~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/Ivan_2018/raw_reads/globu1_symphonia_globulifera_CTTGTA_L001_R1_001.fastq.gz
bwa index $reference
bwa mem -M $reference $query > raw_read_alignment.sam
picard=~/Tools/picard/picard.jar
java -Xmx4g -jar $picard SortSam I=raw_read_alignment.sam O=raw_read_alignment.bam SORT_ORDER=coordinate
bedtools bamtobed -i raw_read_alignment.bam > raw_read_alignment.bed
cat raw_read_alignment.bed | sort -k 1,1 -k2,2n >  raw_read_alignment.sorted.bed
bedtools merge -i raw_read_alignment.sorted.bed -c 1 -o count > raw_read_alignment.merged.bed
```

```{r funcRepeatFilter1}
bed <- read_tsv(file.path(path, "functional_selection2", "raw_read_alignment.merged.bed"), 
                col_names = c("target", "start", "end", "coverage"))
```

We obtained a continuous decreasing distribution of read coverage across our scaffolds regions (figure \@ref(fig:funcRepeatFilter2)). We fitted a $\Gamma$ distribution with positive parameters for scaffolds regions with a coverage under 5 000 (non continuous distribution with optimization issues). We obtained a distribution with a mean of 309 reads per region and a $99^{th}$ quantile of 2 606. We decided to mask regions with a coverage over the $99^{th}$ quantile and remove scaffolds with a mask superior to 75% of its total length (figure \@ref(fig:funcRepeatFilter3)).

```{r funcRepeatFilter2, fig.cap="Read coverage distribution."}
data <- bed %>%
  filter(coverage < 5000)
fit.params <- MASS::fitdistr(data$coverage, 
                             "gamma", lower = c(0, 0))
alpha = fit.params$estimate["shape"]
beta = fit.params$estimate["rate"]
ggplot(data) +
  geom_histogram(aes(coverage, ..density..), binwidth = 100) +
  geom_line(aes(coverage, y=dgamma(data$coverage, alpha, beta)),
            color="red", size = 1) +
  geom_vline(xintercept = alpha/beta, lwd = 1.4) +
  geom_vline(xintercept = qgamma(0.99, alpha, beta), 
             lwd = 1.4, linetype = 2) +
  ylim(NA, 0.005) + xlim(1,5000) +
  ggtitle(paste("Y ~ Gamma(", round(alpha,3), ",", round(beta,3), ")"),
          paste("E(Y) = ", round(alpha/(beta)), 
                ", q99(Y) = ", round(qgamma(0.99, alpha, beta))))
```

```{r funcRepeatFilter3, echo=T}
repetitive_target <- bed %>% 
  filter(coverage > qgamma(0.99, alpha, beta)) %>% 
  mutate(size = end - start) 
repetitive_target %>% 
  select(target, start, end) %>% 
  arrange(target, start, end) %>% 
  mutate_if(is.numeric, as.character) %>% 
  write_tsv(path = file.path(path, "functional_selection2", "repetitive_targets.bed"),
            col_names = F)
```

```{r funcRepeatFilter4, fig.cap="target regions with a coverage over the 99th quantile of the fitted Gamma distribution (2606)."}
ggplot(repetitive_target, aes(size)) + 
  geom_histogram() +
  ggtitle(paste(nrow(repetitive_target), "targets regions to remove (",
                sum(repetitive_target$size)/10^3, "kbp)"),
          paste(length(unique(repetitive_target$target)), "targets over 1165"))
```

```{bash funcRepeatFilter5, eval=F, echo=T}
cd ~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/functional_selection2
cat repetitive_targets.bed | sort -k 1,1 -k2,2n >  repetitive_targets.sorted.bed
bedtools maskfasta -fi targets.fasta -bed repetitive_targets.sorted.bed -fo targets.masked.fasta
```

```{r funcRepeatFilter6, echo = T}
targets <- readDNAStringSet(file.path(path, "functional_selection2", "targets.masked.fasta"))
writeXStringSet(targets[which(letterFrequency(targets, "N")/width(targets) < 0.65)],
                file.path(path, "functional_selection2", "targets.filtered.masked.fasta"))
```

```{r funcRepeatFilter7}
targets <- readDNAStringSet(file.path(path, "functional_selection2", 
                                      "targets.filtered.masked.fasta"))
data.frame(n = length(targets),
           width = sum(width(targets))/10^6,
           Nperc = sum(letterFrequency(targets, "N"))/sum(width(targets))) %>% 
    kable(caption = "Selected, masked and filtered funcional targets.",
        col.names = c("N", "Width (Mbp)", "Mask (%N)"),
        format.args = list(big.mark = " "))
```

```{bash funcRepeatFilter8}
cd ~/Documents/BIOGECO/PhD/data/Symphonia_Genomes/functional_selection2
cp targets.filtered.masked.fasta functional.targets.20180717.fasta
```

