```{r setup_plates, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(parallel)
library(tidyverse)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
path <- "~/Documents/BIOGECO/PhD/data/Symphonia_Paracou/"
```

# Plates

## Organization

```{r plates, fig.cap="Extraction plates organization"}
plates <- read_csv(file.path(path, "Symphonia_sequence_capture - Plates.csv"))
bd <- read_csv(file.path(path, "Base_Données_Symphonia.csv"))

plates %>% 
  mutate(row = substr(Position, 1, 1)) %>% 
  mutate(col = substr(Position, 2, 3)) %>% 
  group_by(Plate) %>% 
  select(Plate, Position, row, col, ID_genetic) %>% 
  arrange(Plate, row, col) %>% 
  ggplot(aes(y = factor(row, levels = rev(LETTERS[1:8])),
             x = factor(col, levels = 1:12))) + 
  # geom_point(shape=21, size =10)  +
  geom_text(aes(label = ID_genetic), size = 2.5) +
  facet_wrap(~ Plate, ncol = 2) +
  theme_bw() +
  labs(x=NULL, y = NULL)
```

## Quantifiaction 

### NanoDrop

```{r nanodrop, fig.cap="Extraction plate NanoDrop concentration"}
nanodrop <- do.call("rbind", sapply(list.files(file.path(path, "NanoDrop"), full.names = T),
                                  read.delim2, simplify = F)) %>% 
  mutate(row = substr(Well, 1, 1)) %>% 
  mutate(col = substr(Well, 2, 3)) %>% 
  group_by(`Plate.ID`) %>% 
  arrange( Well, -Time) %>% 
  filter(!duplicated(Well)) %>% 
  mutate(Conc. = ifelse(Conc. < 0, NA, Conc.)) %>% 
  mutate(Conc. = ifelse(Conc. > 1000, NA, Conc.)) %>% 
  dplyr::rename(plate = `Plate.ID`, well = Well,
                concentration = Conc., r260.280 = X260.280, 
                r260.230 = X260.230)
nanodrop %>% 
  select(plate, well, row, col, concentration) %>% 
  arrange(plate, row, col) %>% 
  ggplot(aes(y = factor(row,levels = rev(LETTERS[1:8])),
             x = factor(col,levels = 1:12))) + 
  geom_text(aes(label = concentration), size = 2.5) +
  facet_wrap(~ plate, nrow = 3) +
  theme_bw() +
  labs(x=NULL, y = NULL)
```

### Electrophoresis

### QBit

```{r qbitTable}
qbit <- read_csv(file.path(path, "Symphonia_sequence_capture - QBit.csv"))
qbit %>%
  select(Plate, Position, Quality_Electrophoresis, Concentration_Final) %>% 
  kable(col.names = c("Plate", "Position", "Electrophoresis", "Concentraion (ng/microL)"),
        fig.cap = "Qbit concentrations.")
```

```{stan qbitModel, output.var="qbitModel", echo=T, eval=F}
  data {
    int N ;
    real Concentration_Qbit[N] ;
    real Concentration_NanoDrop[N] ;
  }
  parameters {
    real beta ;
    real sigma ;
  }
  model {
    for(n in 1:N)
      Concentration_Qbit[n] ~ normal(beta*Concentration_NanoDrop[n], sigma) ;
  }
```

```{r fit}
data <- filter(qbit, Concentration_NanoDrop < 250)
# fit <- sampling(qbitModel,
#                 data = list(N = nrow(data),
#                             Concentration_Qbit = data$Concentration_Final,
#                             Concentration_NanoDrop = data$Concentration_NanoDrop))
# save(fit, file = "./probes_save/MQbit.Rdata")
load("./probes_save/MQbit.Rdata")
broom::tidyMCMC(fit, droppars = NULL, rhat = T) %>%
  kable(caption = "Summary table of the model")
```

```{r pred}
beta <- broom::tidyMCMC(fit, droppars = NULL, rhat = T) %>% 
  filter(term == "beta") %>% select(estimate) %>% unlist()
qbit %>% 
  ggplot(aes(Concentration_NanoDrop, Concentration_Final)) +
  geom_point(aes(color = Quality_Electrophoresis)) +
  geom_abline(intercept = 0, slope = beta) +
  xlab("Concentration NanoDrop (ng/microL)") +
  ylab("Concentration Qbit (ng/microL)") +
  scale_color_discrete("Electrophoresis") +
  ggtitle(paste("Y =", round(beta,2), "* X"))
```

## Library plates design

```{r data}
data <- plates %>% 
  filter(!is.na(ID_genetic)) %>% 
  left_join(nanodrop %>% 
              select(plate, well, concentration) %>% 
              dplyr::rename(Plate = plate, Position = well, nanodrop = concentration)) %>% 
  mutate(concentration_est = beta*nanodrop) %>%
  mutate(ID_paracou = paste0("P", Plot, "_", Subplot, "_", Number)) %>% 
  mutate(ID = ifelse(!is.na(Plot), ID_paracou, ID_genetic))
```

### Pool

We will pool all individuals  with 2 extraction and with a nanodrop concentration inferior to 25 $ng.\mu L^{-1}$. Those wiht only one extraction will be concentrated.

```{r pool}
data %>%
  group_by(ID) %>% 
  filter(nanodrop < 25 | is.na(nanodrop)) %>% 
  filter(n() > 1) %>% 
  arrange(ID) %>% 
  select(ID, Plate, Position, nanodrop) %>% 
  kable(caption = "Samples to be pooled. From plates 1, 2 and 3 to plate 5, 6 and 7") %>% 
  kable_styling(full_width = F) %>% 
  kableExtra::collapse_rows(1)
```


### New nanodrop

Pooled individuals and new individuals from Itubera Brazil (n = 3), La Selva Costa Rica (n = 2) and Baro Colorado Island Panama (n = 2) will be quantified with the nanodrop again.

```{r nanodrop2, fig.cap="Extraction plate NanoDrop concentration"}
america <- read.delim2(file.path(path, "NanoDrop2", "sylvain_symphonia_america.txt")) %>% 
  dplyr::select(Sample.ID, Conc.) %>% 
  dplyr::rename(ID = Sample.ID, nanodrop = Conc.) %>% 
  mutate(Plate = "America", Position = NA)
itubera <- read.delim2(file.path(path, "NanoDrop2", "sylvain_symphonia_Itubera.txt")) %>% 
  dplyr::select(Plate.ID, Well, Conc.) %>% 
  dplyr::rename(Plate = Plate.ID, Position = Well, nanodrop = Conc.) %>% 
  mutate(Plate = "Itubera") %>% 
  mutate(ID = paste0("IT_", Position))
pull <- read.delim2(file.path(path, "NanoDrop2", "sylvain_symphonia_pull.txt")) %>% 
  dplyr::select(Sample.ID, Conc.) %>% 
  dplyr::rename(ID = Sample.ID, nanodrop = Conc.) %>% 
  mutate(Plate = "Pull", Position = NA)
rbind(america, itubera, pull) %>% 
  kable(caption = "New nanodrops.")
```

### Concentration plates

```{r newData}
pull <- pull %>% 
  dplyr::rename(nanodrop2 = nanodrop) %>% 
  select(ID, nanodrop2) %>% 
  mutate(ID = gsub("-", "_", ID))
america <- america %>%
  mutate(Origin = substr(ID, 1,2)) %>% 
  mutate(Country = ifelse(Origin == "LS", "Costa Rica", "Panama")) %>% 
  mutate(Origin = ifelse(Origin == "LS", "La Selva", "BCI")) %>% 
  mutate(Species = "S. globulifera",
          ID_genetic = ID, ID_paracou = NA, 
         Plot = NA, Subplot = NA, Number = NA) %>% 
  arrange(ID, Plate, Position, nanodrop,
          Species, Origin, Country,
          ID_genetic, ID_paracou, Plot, Subplot, Number) %>% 
  group_by(Origin) %>% 
  arrange(desc(nanodrop)) %>% 
  slice(1:2)
itubera <- itubera %>% 
  mutate(Species = "S. globulifera", Origin = "Itubera", Country = "Brazil",
          ID_genetic = ID, ID_paracou = NA, 
         Plot = NA, Subplot = NA, Number = NA) %>% 
  arrange(ID, Plate, Position, nanodrop,
          Species, Origin, Country,
          ID_genetic, ID_paracou, Plot, Subplot, Number) %>% 
  mutate_if(is.factor, as.character) %>% 
  arrange(desc(nanodrop)) %>% 
  slice(1:3)
data <- data %>% 
  filter(!(Plate %in% 1:2)) %>% 
  group_by(ID) %>% 
  arrange(ID, Plate) %>% 
  filter(row_number() == n()) %>% 
  left_join(pull, by = "ID") %>% 
  mutate(nanodrop = ifelse(is.na(nanodrop2), nanodrop, nanodrop2)) %>% 
  select(-nanodrop2, -concentration_est, -Comment) %>% 
  arrange(ID, Plate, Position, nanodrop,
          Species, Origin, Country,
          ID_genetic, ID_paracou, Plot, Subplot, Number) %>% 
  mutate(Plate = as.character(Plate)) %>%
  bind_rows(america) %>%
  bind_rows(itubera) %>% 
  mutate(concentration = beta*nanodrop)
```

```{r concPlate}
data <- data %>% 
  select(ID, Plate, Position, concentration) %>% 
  arrange(concentration)
data$new_Plate <- rep(1:5, each = 8*12)[1:432]
data$new_Position <- rep(unlist(sapply(1:12, function(x) paste0(LETTERS[1:8], x))), 5)[1:432]
# data %>% 
#   kable(caption = "New plates design by concentration") %>% 
#   kable_styling(full_width = F)
```

```{r concPlateFig}
data %>% 
  mutate(row = substr(new_Position, 1, 1)) %>% 
  mutate(col = substr(new_Position, 2, 3)) %>% 
  group_by(new_Plate) %>% 
  mutate(past_Position = ifelse(!is.na(Position) & Plate != "Itubera", 
                                paste0("P", Plate, ".", Position), ID)) %>% 
  select(new_Plate, row, col, past_Position, Plate) %>% 
  arrange(new_Plate, row, col) %>% 
  ggplot(aes(y = factor(row,levels = rev(LETTERS[1:8])),
             x = factor(col,levels = 1:12),
             col = Plate)) + 
  geom_text(aes(label = past_Position), size = 2.5) +
  facet_wrap(~ new_Plate, nrow = 3) +
  theme_bw() +
  labs(x=NULL, y = NULL)
```

### Concentration

All individuals with an estimated concetration inferior to 20 $ng.\mu L^{-1}$ will be dried in the speed vacuum centrifuge. And corresponding volume of miliQ water will be added to reach a concentration of 20 $ng.\mu L^{-1}$.


### Dilution

All individuals with an estimated concetration superior to 20 $ng.\mu L^{-1}$ will be be diluted adding corresponding volume of miliQ water will be to reach a concentration of 20 $ng.\mu L^{-1}$.

