---
title: "Data cleaning & exploration"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(tidyverse)
library(stringdist)
library(googlesheets)
library(knitr)
library(kableExtra)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
```

# Symphonia

```{r SymphoLoad}
ind <- gs_title("Measures_Symphonia") %>% 
  gs_read("Individuals", range = cell_cols(2:13))
SPAD <- gs_title("Measures_Symphonia") %>% 
  gs_read("LeafSPAD", range = cell_cols(2:9))
fresh <- gs_title("Measures_Symphonia") %>% 
  gs_read("LeafFresh", range = cell_cols(2:10))
# dry <- gs_title("Measures") %>% 
#   gs_read("Dry", range = cell_cols(2:7))
# leaves <- SPAD %>% 
#   left_join(fresh, by = c('Plot', 'SubPlot', 'Numéro', 'Feuille')) %>% 
#   left_join(dry, by = c('Plot', 'SubPlot', 'Numéro', 'Feuille'))
# rm(SPAD, fresh, dry)
```

## Data check

```{r SymphoIndividuals check, warning=TRUE}
IDs <- list(ind = with(ind, paste0('P', Plot, '-', SubPlot, '-', TreeFieldNum)),
            SPAD =  unique(with(SPAD, paste0('P', Plot, '-', SubPlot, '-', TreeFieldNum))),
            fresh = unique(with(fresh, paste0('P', Plot, '-', SubPlot, '-', TreeFieldNum))))

# Number of individuals
if(!all(unlist(lapply(IDs, length))[-1] == unlist(lapply(IDs, length))[1])){
  warning("Tables do not have the same number of individuals.\n")
  invalids <- names(which(unlist(lapply(IDs, length))[-1] != unlist(lapply(IDs, length))[1]))
  warning(invalids, " table(s) is(are) invalid(s).\n")
  for(invalid in invalids)
    diff <- length(IDs[[1]]) - length(IDs[[invalid]])
    if(diff > 0)
      warning(invalid, " table is missing ", abs(diff)," individual(s).\n")
    else
      warning(invalid, " table have ", abs(diff)," extra individual(s).\n")
}
  
# Individual names 
## SPAD table
diff <- max(length(sort(with(IDs, setdiff(ind,SPAD)))),
length(sort(with(IDs, setdiff(SPAD,ind)))))
if(diff > 0)
  warning("There are ", diff, " non matching individuals between individual and SPAD tables.\n")
ind_IDs_diff <- sort(with(IDs, setdiff(ind,SPAD)))
SPAD_IDs_diff <- sort(with(IDs, setdiff(SPAD,ind)))
for(id in ind_IDs_diff)
  warning(SPAD_IDs_diff[amatch(id, SPAD_IDs_diff, maxDist = Inf)],
          " is not a misspel of ", id, " ?\n")
## fresh table
diff <- max(length(sort(with(IDs, setdiff(ind,fresh)))),
length(sort(with(IDs, setdiff(fresh,ind)))))
if(diff > 0)
  warning("There are ", diff, " non matching individuals between individual and fresh tables.\n")
ind_IDs_diff <- sort(with(IDs, setdiff(ind,fresh)))
fresh_IDs_diff <- sort(with(IDs, setdiff(fresh,ind)))
for(id in ind_IDs_diff)
  warning(fresh_IDs_diff[amatch(id, fresh_IDs_diff, maxDist = Inf)],
          " is not a misspel of ", id, " ?\n")
```

```{r SymphoIndValuesCheck, warning=TRUE}
## Date
Date <- ind %>% 
  dplyr::select(Date) %>% 
  separate(Date, c("Day", "Month", "Year"))
if(all(Date$Year > 2017))
  warning("There is a wrong year in one date.\n")
if(!all(Date$Month %in% 10:12))
  warning("There is a wrong month in one date.\n")
if(!all(as.numeric(Date$Day) %in% 1:31))
  warning("There is a wrong month in one date.\n")
## Time
Time <- ind %>% 
  dplyr::select(Time) %>% 
  separate(Time, c('Hour', 'Minute'), 'H')
if(!all(Time$Hour %in% 1:24))
  warning("Lines ", paste(which(!(Time$Hour %in% 1:24)), collapse = ', '),
          " have a wrong hour format.\n")
if(!all(as.numeric(Time$Minute) %in% 0:59))
  warning("Lines ", paste(which(!(as.numeric(Time$Minute) %in% 0:59)), collapse = ', '),
          " have a wrong minutes format.\n")
## Plot
if(!all(ind$Plot %in% 1:16))
  warning("Lines ", paste(which(!(ind$Plot %in% 1:16)), collapse = ', '),
          " have a wrong Plot number.\n")
## SubPlot
if(!all(ind$SubPlot %in% 1:25))
  warning("Lines ", paste(which(!(ind$SubPlot %in% 1:25)), collapse = ', '),
          " have a wrong SubPlot number.\n")
## TreeFieldNum
if(!all(ind$TreeFieldNum %in% 1:4000))
  warning("Lines ", paste(which(!(ind$TreeFieldNum %in% 1:4000)), collapse = ', '),
          " have a wrong TreeFieldNum number.\n")
## Dawkins
Dawkins <- c(NA, '1A', '1B', '2A', '2B', '2C', '3A',
             '3B', '3C', '4A', '4B', '4C', '5A', '5B')
if(!all(ind$Dawkins %in% Dawkins))
  warning("Lines ", paste(which(!(ind$Dawkins %in% Dawkins)), collapse = ', '),
          " have a wrong Dawkins index.\n")
## Position
Positions <- c(NA, 'TOP', 'MID-TOP', 'MID', 'MID-LOW', 'LOW',
             'CANOPY', 'REJET')
if(!all(ind$Position %in% Positions))
  warning("Lines ", paste(which(!(ind$Position %in% Positions)), collapse = ', '),
          " have a wrong Position code.\n")
## Pneumatophores
if(!all(ind$Pneumatophores %in% 0:1))
  warning("Lines ", paste(which(!(ind$Pneumatophores %in% 0:1)), collapse = ', '),
          " have a wrong Pneumatophores code.\n")
## Buttress
if(!all(ind$Buttress %in% 0:300))
  warning("Lines ", paste(which(!(ind$Buttress %in% 0:300)), collapse = ', '),
          " have a wrong Buttress value.\n")
## Stilts
if(!all(ind$Stilts %in% 0:1000))
  warning("Lines ", paste(which(!(ind$Stilts %in% 0:300)), collapse = ', '),
          " have a wrong Stilts value.\n")
## Bark
Barks <- c('S', 'G', 'SG')
if(!all(ind$Bark %in% Barks))
  warning("Lines ", paste(which(!(ind$Bark %in% Barks)), collapse = ', '),
          " have a wrong Bark code.\n")
## Wood
if(!all(ind$Wood %in% c(NA, 0:1)))
  warning("Lines ", paste(which(!(ind$Wood %in% c(NA, 0:1))), collapse = ', '),
          " have a wrong Wood code.\n")
```

```{r SymphoSPAD values check, warning=TRUE}
## Plot
if(!all(SPAD$Plot %in% 1:16))
  warning("Lines ", paste(which(!(SPAD$Plot %in% 1:16)), collapse = ', '),
          " have a wrong Plot number.\n")
## SubPlot
if(!all(SPAD$SubPlot %in% 1:25))
  warning("Lines ", paste(which(!(SPAD$SubPlot %in% 1:25)), collapse = ', '),
          " have a wrong Carré number.\n")
## TreeFieldNum
if(!all(SPAD$TreeFieldNum %in% 1:4000))
  warning("Lines ", paste(which(!(SPAD$TreeFieldNum %in% 1:4000)), collapse = ', '),
          " have a wrong TreeFieldNum number.\n")
## Leaf
if(!all(SPAD$Leaf %in% 1:4000))
  warning("Lines ", paste(which(!(SPAD$Leaf %in% 1:4000)), collapse = ', '),
          " have a wrong Leaf number.\n")
## SPAD
test <- function(x) ((x > 20 & x < 100) | is.na(x))
if(!all(apply(SPAD[5:7], 1, function(x) all(sapply(x, test)))))
  warning("Lines ", paste(which(!apply(SPAD[5:7], 1, function(x) all(sapply(x, test)))), 
                          collapse = ', '), " have a wrong SPAD value.\n")
```

```{r SymphoFreshValuesCheck, warning=TRUE}
## Plot
if(!all(fresh$Plot %in% 1:16))
  warning("Lines ", paste(which(!(fresh$Plot %in% 1:16)), collapse = ', '),
          " have a wrong Plot number.\n")
## SubPlot
if(!all(fresh$SubPlot %in% 1:25))
  warning("Lines ", paste(which(!(fresh$SubPlot %in% 1:25)), collapse = ', '),
          " have a wrong SubPlot number.\n")
## TreeFieldNum
if(!all(fresh$TreeFieldNum %in% 1:4000))
  warning("Lines ", paste(which(!(fresh$TreeFieldNum %in% 1:4000)), collapse = ', '),
          " have a wrong TreeFieldNum number.\n")
## Leaf
if(!all(fresh$Leaf %in% 1:4000))
  warning("Lines ", paste(which(!(fresh$Leaf %in% 1:4000)), collapse = ', '),
          " have a wrong Leaf number.\n")
## LT
test <- function(x) ((x > 50 & x < 600) | is.na(x))
if(!all(apply(fresh[5:7], 1, function(x) all(sapply(x, test)))))
  warning("Lines ", paste(which(!apply(fresh[5:7], 1, function(x) all(sapply(x, test)))), 
                          collapse = ', '), " have a wrong LT value.\n")
## FreshWeight
test <- function(x) ((x > 0.1 & x < 5) | is.na(x))
if(!all(sapply(fresh$FreshWeight, test)))
  warning("Lines ", paste(which(!(sapply(fresh$FreshWeight, test))), 
                          collapse = ', '), " have a wrong Poids frais value.\n")
```

## Exploration

```{r SymphoExplore, fig.cap="Symphonia functional traits exploration."}
path <- "~/Documents/BIOGECO/PhD/data/Symphonia_Paracou/"
data <- read_csv2(file.path(path, "Full_traits_data.csv"))
data  %>% 
  dplyr::select(morphotype, Dry_Weight, Area_exclude, Wood_infra_density) %>% 
  mutate(SLA = Area_exclude / Dry_Weight) %>% 
  mutate(WD = Wood_infra_density) %>% 
  na.omit() %>% 
  ggplot(aes(WD, SLA, col = morphotype)) +
  geom_point() +
  stat_ellipse()
```

# Eschweilera

# Data cleaning

## Individuals

The idea is to transform FTH data to the general formatting, i.e. :

* Date (YYYY/MM/DD)
* Time (HHHMM)
* Plot
* SubPlot
* FieldTreeNum
* Dawkins
* Position
* TrunkForm
* BarkColor
* BarkRuggedness
* BarkAspect
* Lenticels
* BladeForm
* LeafConsistency
* LeafSurface
* LeafBase
* LeafMargin
* HyperAcuminated
* BentPrincpalN


```{r MahoInd}
indFTH <- read_delim(file.path("~/Documents/BIOGECO/PhD/data/Eschweilera_Paracou/Measures/",
                     "data_FTH.csv"), delim = ",") %>% 
  rename(TrunkForm = `Trunk form`, BarkColor = `Bark color`, BarkRuggedness = `Bark ruggedness`, 
         BarkAspect = `Bark aspect`, BladeForm = `Blade form`, 
         LeafConsistency = `Leaf consistency`, LeafSurface = `Leaf surface`, 
         LeafBase = `Leaf base`, LeafMargin = `Leaf margin`,
         HyperAcuminated = `Longue pointe`, BentPrincipalN = `Pli sur nervure principale`) %>% 
  separate(Date, c("Month", "Day", "Year"), "/") %>% # Date
  mutate(Date = paste0(Year, "/0", Month, "/", Day)) %>% 
  select(-Year, -Month, -Day) %>% 
  mutate(Time = gsub(":", "H", substr(as.character(Time), 1, 5))) %>% # Time
  mutate(Dawkins = toupper(Dawkins)) %>% # Dawkins
  mutate(Position = recode_factor(Position, bas = "LOW", haut = "TOP", milieu = "MID",
                                  `milieu lumière` = "MID/LIGHT", `milieu/haut` = "MID",
                                  `bas lumière` = "LOW/LIGHT", `haut lumière` = "TOPLIGHT",
                                  `bas/lumière` = "LOW/LIGHT", `milieu/bas` = "MID")) %>% # Position
  mutate(TrunkForm = recode_factor(TrunkForm, `B--` = "B-", `B (petits)` = "B-", `B+` = "B",
                                  BS = "B/S", tordu = "N", C = "B")) %>% # TrunkForm
  mutate(BarkColor = recode_factor(BarkColor, `B/BR` = "B/RB", BG = "B/G", D = "DB",
                                   `DB+` = "DB", `DB/B` = "B/DB", GB = "B/G")) %>% # BarkColor
  mutate(BarkAspect = recode_factor(BarkAspect, `?` = "N", `F+` = "F", `F/D` = "D/F",
                                    `F/D-` = "D-/F", `F-HM` = "F-/HM", `F/HM+` = "F/HM",
                                    `F+/HM` = "F/HM", `F/S/HM` = "F/HM/S", 
                                    `HM/D (?)` = "D/HM", `HM/F` = "F/HM",
                                    `S+/D` = "D/S+", `S/F/HM`	= "F/HM/S", `S-/HM`	= "HM/S-",
                                    `S/HM`	= "HM/S", `R/HM` = "F/HM")) %>% # BarkAspect
  mutate(Lenticels = recode_factor(Lenticels, N = "A", `P+` = "P", `P-` = "P")) %>% # Lenticels
  mutate(BladeForm = recode_factor(BladeForm, `E+` = "E", `Oblan/E` = "E/Oblan")) %>% # BladeForm
  mutate(LeafConsistency = recode_factor(LeafConsistency, `Char -` = "Char-", 
                                         `Char+` = "Char")) %>% # LeafConsistency
  mutate(LeafBase = recode_factor(LeafBase, `O/A` = "A/O", `O-R` = "O/R", `R/A` = "A/R",
                                  RO = "O/R", `R/O` = "O/R")) %>% # LeafBase
  mutate(LeafMargin = recode_factor(LeafMargin, `E/C-` = "C-/E")) %>% # LeafMargin
  select(Date, Time, Plot, SubPlot, FieldTreeNum, Dawkins, Position,
         TrunkForm, BarkColor, BarkRuggedness, BarkAspect, Lenticels,
         BladeForm, LeafConsistency, LeafSurface, LeafBase, LeafMargin,
         HyperAcuminated, BentPrincipalN, Comment) %>% 
  unique()
# read_delim(file.path("~/Documents/BIOGECO/PhD/data/Eschweilera_Paracou/Measures/",
#                      "Indviduals.csv"), delim = ",", col_types = cols(Date = col_character())) %>% 
#   bind_rows(indFTH) %>% 
#   write_delim(file.path("~/Documents/BIOGECO/PhD/data/Eschweilera_Paracou/Measures/",
#                      "Indviduals_all.csv"), delim = ",")
```

## Leaf attribution in LA

```{r MahoLAsummary, eval=F}
LA %>% 
  select(commentaire) %>% 
  group_by(commentaire) %>% 
  summarise(n = n()) %>% 
  kable()
```

```{r MahoLA, eval=F}
LA <- LA %>% 
  separate(Label, c("Plot", "SubPlot", "FieldTreeNum"), 
           "-", convert = T) %>% 
  mutate(Plot = as.numeric(gsub("P", "", Plot))) %>% 
  mutate(FieldTreeNum0 = FieldTreeNum) %>% 
  mutate(FieldTreeNum = as.numeric(gsub('([[:alpha:]])', "", FieldTreeNum0))) %>% 
  mutate(Scan = gsub('[0-9]+', "", FieldTreeNum0)) %>% 
  select(-FieldTreeNum0) %>% 
  arrange(Plot, SubPlot, FieldTreeNum, Area_exclude) %>% 
  group_by(Plot, SubPlot, FieldTreeNum) %>% 
  mutate(Leaf_order = (1:6)[1:n()])
data <- data %>% 
  arrange(Plot, SubPlot, FieldTreeNum, Fresh_Weight) %>% 
  group_by(Plot, SubPlot, FieldTreeNum) %>% 
  mutate(Leaf_order = (1:6)[1:n()])
data <- data %>% 
  left_join(LA)
# data %>% 
#   write_csv("./data.csv")
```
