---
title: Symphonia preliminary anlaysis
author: Sylvain SCHMITT
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r set, message=FALSE, warning=FALSE, echo=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
path <- "~/Documents/BIOGECO/PhD/data/growth_Paracou/"
mpath <- "~/Documents/BIOGECO/PhD/documents/preliminary analysis/preliminary_models/"
library(knitr)
library(readr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(rstan)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
```

# Introduction

View [PhD coneptual document online](https://docs.google.com/document/d/1ws7OHgMXHwdHyw5Se_nb_xQk1eDelew8lF4viN7wCpQ/edit?usp=sharing).

# Model M0

```{r opening}
# sp <- read_delim(file.path(path, 'Paracou_species.csv'), ',') %>%
#   filter(Genre == "Symphonia") %>%
#   rename(idSp = idTaxon, Family = Famille, Genus = Genre, species = espece) %>%
#   dplyr::select(-Family)
# data <- rbind(read_delim(file.path(path, 'R_PreExploit_1984-85-86.csv'), ';'),
#            read_delim(file.path(path, 'R_PostExploit_1987-88.csv'), ';'),
#            read_delim(file.path(path, 'R_Paracou_1988-2016.csv'), ';')) %>%
#   filter(code_vivant == 'VRAI') %>%
#   filter(idTaxon %in% sp$idSp) %>%
#   dplyr::select(n_parcelle, n_carre, n_arbre, idArbre, campagne, Xutm, Yutm, circonf, idTaxon) %>%
#   mutate(circonf = circonf*10 / pi) %>%
#   rename(plot = n_parcelle, square = n_carre, id = n_arbre, idPR = idArbre,
#          census = campagne, X = Xutm, Y = Yutm, dbh = circonf, idSp = idTaxon) %>% 
#   filter(!is.na(dbh))
# rm(sp)
# save(data, file = './preliminary_save/data.Rdata')
load('./preliminary_save/data.Rdata')
```

We gathered data from Paracou for *Symphonia* genus resulting in `r length(unique(data$idPR))` individuals. We computed the tree annual growth ($AGR$, in $mm.year^{-1}$) as the difference of diameter at breast height ($dbh$, in $mm$) divided by the time step (from 1 to 2 year depending on the censuses). We related $AGR$ to diameter at breast height with the Canham model following @herault_functional_2011 :
$$log(AGR_i) = max * exp(-\frac{1}{2}*log(\frac{dbh_i}{\frac{dopt}{ks}}))$$
Where $max$ represents the maximum value of $log(AGR_i +1)$, $dopt$ the optimal $dbh$ when the maximum is reached and $ks$ the kurtosis of the hump shaped distribution.

```{r M0 data}
data <- data %>% 
  select(idPR, census, dbh, idSp) %>% 
  group_by(idPR) %>% 
  arrange(census) %>% 
  mutate(AGR = (dbh - lag(dbh, default=first(dbh)))/(census - lag(census, default=first(census)))) %>% 
  filter(!is.na(AGR)) %>% 
  filter(AGR > 0) %>% 
  filter(AGR < 200)
data %>%
  ggplot(aes(dbh, log(AGR))) +
  geom_point(alpha = 0.3) +
  xlab('diameter at breast height (dbh in mm)') +
  ylab(expression('logarithm of annual growth rate (AGR in'~mm.year^{-1}~")"))
```

```{r}
fit <- stan(file = file.path(mpath, 'M0.stan'),
            data = list(N = dim(data)[1],
                        dbh = data$dbh,
                        AGR = data$AGR,
            chains = 1))
save(fit, file = file.path(mpath, 'M0.Rdata'))
load(file.path(mpath, 'M0.Rdata'))
g <- fitgraph(fit)
```


# References
