---
title: "Competition model Plot 1"
author: Sylvain Schmitt, Bruno Hérault
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup_BA, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(raster)
library(tidyverse)
library(rstan)
library(bayesplot)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
theme_set(bayesplot::theme_default())
path <- "~/Documents/BIOGECO/PhD/data/Paracou/"
```

```{r computeBA}
computeBA <- function(trees, # all trees from the plot
                      idTreeVal, # from the tree for which BA is computed
                      CensusYearVal,
                      XfieldVal, # from the tree for which BA is computed
                      YfieldVal, # from the tree for which BA is computed
                      d # maximum distance to be taken into account
){
  trees %>% 
    mutate(Surface = (CircCorr)^2/(4*pi)) %>% 
    filter(CensusYear == CensusYearVal) %>%
    filter(idTree != idTreeVal) %>%
    group_by(idTree) %>% 
    mutate(distance = sqrt((XfieldVal-Xfield)^2+(YfieldVal-Yfield)^2)) %>% 
    filter(distance < d) %>% 
    ungroup() %>% 
    select(Surface) %>% 
    summarise(BA = sum(Surface)) %>% 
    unlist()
}
```

```{r data}
trees <- src_sqlite(file.path(path, "trees/Paracou.sqlite")) %>% 
  tbl("Paracou") %>% 
  filter(Plot == 1) %>% 
  filter(CodeAlive == 1) %>% 
  collect()
# symphonia <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv')) %>%
#   filter(CodeAlive == 1) %>%
#   filter(Plot == 1) %>%
#   filter(Xfield > 20, Xfield < max(Xfield)-20,
#          Yfield > 20, Yfield < max(Yfield)-20)
# symphonia <- symphonia %>%
#   group_by(CensusYear, idTree) %>%
#   mutate(BA10 = computeBA(trees, idTree, CensusYear, Xfield, Yfield, d = 10)) %>%
#   mutate(BA20 = computeBA(trees, idTree, CensusYear, Xfield, Yfield, d = 20))
# save(symphonia, file = "./Competition_save/symphoniaBA_P1.Rdata")
load("./Competition_save/symphoniaBA_P1.Rdata")
symphonia <- symphonia %>%
  mutate(dbh = CircCorr/pi) %>% 
  arrange(CensusYear) %>%
  group_by(idTree) %>%
  mutate(dt = CensusYear - lag(CensusYear)) %>% 
  mutate(ddbh = dbh - lag(dbh)) %>% 
  mutate(dt = ifelse(dt <= 0, NA, dt)) %>% 
  mutate(ddbh = ifelse(ddbh < 0, NA, ddbh)) %>% 
  mutate(agr = ddbh/dt) %>% 
  ungroup() %>% 
  mutate(BA10 = BA10*10^-4, BA20 = BA20*10^-4) %>% # cm2 to m2
  mutate(BA10 = BA10/(pi*10^2/10^4), BA20 = BA20/(pi*20^2/10^4)) # divided by surface in ha
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0' # global crs definition
symphoniaXY <- symphonia
coordinates(symphoniaXY) <- ~Xutm + Yutm
proj4string(symphoniaXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
symphoniaXY <- spTransform(symphoniaXY, CRSobj = crs)
wtd <- shapefile(file.path(path, "hydrology", "WaterTable.shp"))
wtd <- spTransform(wtd, CRSobj = crs)
symphonia$wtd <- raster::extract(wtd, symphoniaXY)$TypeWatEN
rm(crs, symphoniaXY, wtd)
```

# Introduction

> Ce que je veux dire c’est que pour quantifier l’évolution l’environnement lumineux et son évolution, tu n’as pas besoin de paramétrer un modèle de croissance. Tu peux simplement calculer année après année l’évolution de la surface terrière sur un cercle de 10 ou 20m autour de chaque arbre cible. La valeur de base et la pente de cette évolution te permet de voir le niveau de compétition initial et si l’arbre a subit une augmentation ou une diminution de compétition. Et du coup de voir s’il y a des génotypes/phénotypes particuliers associés à ces différents environnements lumineux. *Bruno Hérault, 22/08/2018*

We assume biotic environment through neighbors interaction to play an important role on tree establishment and growth and more especially with putative differences between genotypes, phenotypes and closely related species due to sympatric fine-scale local adaptation.

In order to test that we want to study light and competition environment evolution through time for individual tree in relation to their morphotype / genotype.

Finally, we wish to use evidenced relation to build back light and competition environment through past time.

# Modelling local basal area

## Data

I tried to directly model basal area, besides previous results suggested that death stochasticity was too unpredictable to do it. I tried to look at local basal area evolution through time to estimate in which light environment each tree has been recruited. I started fromm the graph below assuming a linear evolution for each individual.

```{r ModelData, fig.cap="Evolution of local basal area (BA) through census years."}
symphonia %>% 
  ggplot(aes(CensusYear, BA20, col = Species, alpha = dbh, size = dbh)) +
  geom_point(aes(size = dbh)) + 
  geom_line(aes(group = idTree)) +
  xlab("Census year") +
  ylab("Local basal area (BA in m2/ha)") +
  scale_size(range = c(0.2, 3))
```

## Model 

Local basal area is not dependant on the tree individual and not on the species, we will simply assume each patch to follow the same evolution trend linearly with a null intercept year $year0_i$ corresponding to a major treefall for the gap and we will assume the trend coefficient to be either equal or not for each patch $i$ thus ignoring intermediate treefall events.

$$BA20_{patch} \sim \mathcal{N}(year0_{patch}+\beta_{patch}*year,\sigma) $$
$$\beta_{patch} \sim \mathcal{N}^P(\beta,\sigma_{patch})$$

```{stan M1, output.var="M1", echo=T, eval=F}
  data {
    int N ; // # obs
    real BA20[N] ;
    real year[N] ;
    int P ; // # of patch
    int<lower=1,upper=P> patch[N] ;
  }
  parameters {
    real year0[P] ;
    real beta_patch[P] ;
    real beta ;
    real<lower=0> sigma ;
    real<lower=0> sigma_patch ;
  }
  model {
    beta_patch ~ normal(beta, sigma_patch) ;
    for(n in 1:N)
      BA20[n] ~ normal(year0[patch[n]] + beta_patch[patch[n]]*year[n], sigma) ;
  }
  generated quantities {
    real BA20pred[N] ;
    for(n in 1:N)
      BA20pred[n] = year0[patch[n]] + beta_patch[patch[n]]*year[n] ;
  }
```

Models seems to have correctly converged. Like expected $\beta$ coefficients are really weak with a variance $\sigma_{patch}$ pretty high. But looking closer into intercepts ($year0$) and individual coefficients ($\beta$), we noticed a wide diversity of case with some coefficients non different from 0 or negatives.

```{r fitM1}
# fit <- sampling(M1,
#                 data = list(N = nrow(symphonia),
#                             BA20 = symphonia$BA20,
#                             year = symphonia$CensusYear,
#                             P = length(unique(symphonia$idTree)),
#                             patch = as.numeric(as.factor(symphonia$idTree))))
# save(fit, file = "./Competition_save/M1_P1.Rdata")
load("./Competition_save/M1_P1.Rdata")
pars <- c("beta", "sigma_patch", "sigma")
broom::tidyMCMC(fit, pars = c(pars, "lp__"), droppars = NULL, rhat = T) %>% 
  kable(caption = "Summary table of the model")
```

```{r betap, fig.cap="Coefficient for each patch."}
mcmc_intervals(as.array(fit), regex_pars = "beta_p")
```

```{r year0, fig.cap="Intercept for each patch."}
mcmc_intervals(as.array(fit), regex_pars = "year0")
```

```{r T1, fig.cap="Markov chains trace plot after warmup for model Mnull"}
mcmc_trace(as.array(fit), pars = c(pars, "lp__"),
           facet_args = list(labeller = label_parsed))
```

```{r 21, fig.cap="Markov chains pairs plot after warmup for model Mnull"}
mcmc_pairs(as.array(fit), pars = c(pars, "lp__"))
```

Finally, looking at individual predictions for few patches / individuals we get a better idea of what happens. Individual models are working properly as soon as there is no local basal area loss due to mortality events. But when mortality events occured and basal area decrease with disturbances, the model does not work anymore. Worst, there are a diversity of case of basal area loss, which seems to important to be predicted !

```{r Y1, fig.cap="Predictions for model M0"}
cbind(
  symphonia,
  pred = apply(as.matrix(fit, pars = "BA20pred"), 2, mean),
  t(apply(as.matrix(fit, pars = "BA20pred"), 2, quantile, probs = c(0.05, 0.95)))
) %>% 
  filter(idTree %in% sample(unique(symphonia$idTree), 9)) %>% 
  ggplot(aes(CensusYear, group = idTree, col = Species, 
             alpha = dbh, size = dbh)) +
  geom_point(aes(y = BA20)) +
  geom_line(aes(y = BA20),
            alpha = 1, size = 0.2, linetype = "dotted") +
  geom_line(aes(y = pred)) +
  geom_line(aes(y = `5%`), col = "black", alpha = 1, size = 0.5, linetype = "dashed") +
  geom_line(aes(y = `95%`), col = "black", alpha = 1, size = 0.5, linetype = "dashed") +
  xlab("Census year") +
  ylab("Local basal area (BA in m2/ha)") +
  scale_size(range = c(0.2, 3)) +
  facet_wrap(~idTree, labeller = "label_both", scales = "free")
```
