# Introduction {-}

```{r setup_Intro, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
load("./growth_save/model_data.Rdata")
```

## Models {-}

I wanted to test for an eventual effect of logging through light and disturbance on Symphonia individuals growth. I combined the disturbance indice of model proposed by @Herault2010 with the potential reductor model proposed by @herault_functional_2011:

\begin{equation}
  log(AGR_i+1) \sim \mathcal{N}(\mu*\sum_j^J(e^{-\alpha*d_{i,j}}*S_j^\beta);\sigma)~|~i\in[1:I]~,~j\in[1:J]
  log(AGR_i+1) \sim \mathcal{N}(AGR_{max}*e^{-\frac{1}{2}*log(\frac{dbh_i}{d_{opt}/k_s})};\sigma)~|~i\in[1:I]
  log(AGR_i+1) \sim \mathcal{N}(AGR_{max}*e^{-\frac{1}{2}*log(\frac{dbh_i}{d_{opt}/k_s})}*e^{-\sum_j^Je^{-\alpha*d_{i,j}}*S_j^\beta};\sigma)~|~i\in[1:I]~,~j\in[1:J]
  (\#eq:intro)
\end{equation}

I declined the model in different versions adding indiviuals and species random effects:

```{r, echo=F}
data.frame(
  Name = c('Ontogeny only', 'Individual and ontogeny', 'Species and ontogeny', 'Gaps only', 'Gaps and ontogeny'),
  Model = c(
    '$log(AGR_i+1) \\sim \\mathcal{N}(AGR_{max}*e^{-\\frac{1}{2}*(log(\\frac{dbh_i}{d_{opt}})/k_s)^2};\\sigma)$',
    '$log(AGR_i+1) \\sim \\mathcal{N}({AGR_{max}}_{ind}*e^{-\\frac{1}{2}*(log(\\frac{dbh_i}{d_{opt}})/k_s)^2};\\sigma)$',
    '$log(AGR_i+1) \\sim \\mathcal{N}({AGR_{max}}_{sp}*e^{-\\frac{1}{2}*(log(\\frac{dbh_i}{{d_{opt}}_{sp}})/{k_s}_{sp})^2};\\sigma)$',
    '$log(AGR_i+1) \\sim \\mathcal{N}(AGR_{max}*e^{-\\sum_j^Je^{-\\alpha*d_{i,j}}*S_j^\\beta};\\sigma)$',
    '$log(AGR_i+1) \\sim \\mathcal{N}(AGR_{max}*e^{-\\frac{1}{2}*(log(\\frac{dbh_i}{d_{opt}})/k_s)^2}*e^{-\\sum_j^Je^{-\\alpha*d_{i,j}}*S_j^\\beta};\\sigma)$'
  )
) %>% 
knitr::kable(caption = "Summary of growth models.")
```

## Data  {-}

Data encompass all *Symphonia* annual growth rate over time in plots 1 to 16 in Paracou. Annual growth rate for an individual $AGR_i$ in $cm.yr^-1$ is computed a the variation of diameter at breast height $\Delta dbh_i$ between two dates divided by the time variation between the two dates $\Delta t_i$ for this individual: $AGR_i = \frac{\Delta dbh_i}{\Delta t_i}$. Data include also diameter at breast height $dbh$ in $cm$ over time, individuals id and individuals morphotype (*S. globulifera*, *S. sp1*, *S. indet*). We those data we built two datasets:

* **Data 1** includes only the plot 2 (for quicker computations)
* **Data 2** includes only the individual having both $dbh$ inferior to 20 cm and superior to 30 cm (to have a longer growth trajectory)

```{r datasets}
rbind(
data$trees %>% 
  mutate(dataset = "data1"),
data2$trees %>% 
  mutate(dataset = "data2")
) %>% 
  ggplot(aes(dbh_c, log(agr+1))) + 
  geom_point(aes(shape = espece)) +
  geom_line(aes(group = idArbre, col = as.factor(idArbre)), alpha = 0.2) +
  facet_wrap(~dataset, nrow = 2) +
  scale_color_discrete(guide = "none") +
  xlab("diameter at breast height (dbh)") +
  ylab("log of annual growth rate (log(AGR+1))") +
  scale_shape_discrete("morphotype", labels = c("S. globulifera", "S. indet", "S. sp1"))
```


