---
title: "Like mother, like daughter: the fine spatio-temporal dynamics of forest gaps drive a genotypic trade-off between growth and competition for the tropical tree species complex *Symphonia globulifera*"
date: '`r Sys.Date()`'
output: 
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::word_document2:
    reference_docx: ./template/template.docx
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

> Proverb. like mother, like daughter. (idiomatic) A daughter will have traits similar to her mother upon reaching adulthood.

# Take-home message

Tree species decrease their risk of a stochastic local extinction because individual genotypes display differential succession strategies in response to fine spatio-temporal dynamics of forest gaps. Consequently, forest gaps dynamics effect on genetic adaptations among tropical tree species is conserved within species for the species complex *Symphonia globulifera*.

# Outlines

The heterogeneity of resource distribution in space and time defines fine-scale habitat structure where species and individuals can grow and coexist [@lortie_rethinking_2004-1; @weiher_assembly_1995]. 
For instance, topography spatially drives water and nutrient distribution in tropical forests [@ferry2010higher].
Topography drives pervasive habitat preference differentiation among species [@Allie2015; Schmitt et al., in prep] and functional responses among and within species [@Schmitt2020].
Nevertheless, in natural tropical forests, above-ground competition for light due to forest gap dynamics [@Hubbell1999; @Molino2001] is often greater than the effect of below-ground competition for water and nutrients, even in early successional stages [@VanBreugel2012].
Forest gap dynamics is characterised by successional niches with a wide variety of demographic response depending on species and individual functional strategy [@Herault2010; @Ruger2009].
After a treefall, pioneer species with fast growth, thanks to productive leaf tissue and light wood, rapidly colonize gaps with high light level and decreased competition from adult trees, before late successional species with slow growth and more conservative tissue establish in shaded environment with increased competition [@Craven2015].

But “nothing in biology makes sense except in the light of evolution” [@Dobzhansky1973]. 
Topography and abiotic habitats have been shown to drive adaptation among species, with adaptive radiations [@Paun2016; @Pillon2014], and within species, with microgeographic adaptations [@Brousseau2015; @Team2013].
But despite the fact that sympatric species grow differently depending on access to light [@Yamasaki2013], and despite the strong impact of forest gap dynamics on tree ecology [@VanBreugel2012], to our knowledge, only few studies have explored the possible role of forest gaps dynamics on adaptation within and among species.

Eco-evolutionary processes have been documented for a while [@Tutt1896] and can play an important role in the dynamics of species [@Bailey2009]. 
However, the assumption that evolution and ecology play at very different time and spatial scales might have been an impediment to understand the role of forest gaps dynamics in species eco-evolutionary dynamics [@Pelletier2009]. 
Spatially, evidence exists for eco-evolutionary processes at microgeographic scale, e.g. within the dispersal neighborhood of the organism [@Richardson2014].
For instance, microgeographic divergent selection is driven by fine-scale habitat structure of the Neotropical tree *Eperua falcata* [@Brousseau2015; @Team2013].
Temporally, overlapping generations can maintain genetic variation to answer strong selection of a fluctuating environment on a specific stade [@Ellner1994].
The storage [@Chesson1983] of genotypes in long-lived stages, on which the fluctuating selection does not act, maintain genetic diversity.
Nevertheless, genetic variation is maintained only if bet hedging mixed-strategy is prevented by physiological or adaptation constraints.
For instance, overlapping generations with alternance of short high-recruitment periods and long population-decline periods explain the spatial genetic structure of the Neotropical tree *Jacaranda copaia* [@Jones2006].

The dynamics of forest gaps strongly determine the survival and recruitment of young tropical trees with competition and light access; whereas once in the canopy, adult trees are less vulnerable to successional niches [ref].
Adult trees can therefore act as a reservoir of genotypes against fine spatio-temporal dynamics of forest gaps. 
Generations of tropical trees broadly overlap with a maturity starting around 20cm of diameter at breast height [@Hardy2006] for individuals growing above 1m of diameter at breast height. 
And the existence of mixed-strategy genotypes to answer the spectrum of successional niches encountered along forest gaps dynamics seems unrealistic.
Consequently, we hypothesized forest gap dynamics to be a major driver of the structuring of genotypic diversity within tropical tree species.

Since the dynamics of forest gaps oppose rapid growth in a bright environment to slow growth in a shaded environment, we expected that individual genotypes would be adapted to a compromise between growth and competition.
Following our hypothesis, we expected that the genotypic trade-off would be spatially structured in the area where a tree falls.
Finally, we expected the genotypic trade-off to be structured in time.
Because of recent selection, genotypic adaptation to competition and growth should be stronger in the intermediate stages, compared to the younger stages under selection and the older stages less vulnerable to forest gap dynamics.

In the present study, we assessed genotypic diversity within closely-related sympatric tree species belonging to the widespread tropical tree species complex *Symphonia globulifera*. 
We addressed the fine-scale spatial and temporal adaptation of individual growth along forest gap dynamics.
Combining tree inventories and single nucleotide polymorphisms (SNPs), we used population genomics, environmental association analyses, genome wide association studies and growth modelling to address the following questions:

* Are individual genotypes adapted to the neighbourhood crowding gradient resulting from forest gap dynamics?
* Is the growth of individuals determined by genotypes?
* Can we identify a trade-off between genotypic adaptations to crowding and growth?
* How are genotypic adaptations to crowding and growth structured in space and time? 

We sampled leaf from 402 individuals previously morphologically identified as *Symphonia globulifera* and *Symphonia sp. 1* in the Paracou field station, where individuals have been censused for more than 30 years. 
We designed and conducted a targeted sequence capture of genomic regions for each individual.
We called, filtered and annotated single nucleotide polymorphisms (SNPs) of sequenced libraries against the references used to build the targeted sequences. 
We analyzed population genetic structure, genetic diversity, and demographic history. 
We then measured species and genotypes association to mean neighbor crowding and individual growth potential over the last 30 years with Bayesian inferences of Animal and growth models.
Finally, we investigated the relation between genotypic associations to neighbor crowding and individual growth and their structure in space and time with respectively spatial autocorrelogram and decomposed relations over diameter classes.

*Symphonia globulifera* was locally composed of three sympatric species corresponding to three morphotypes differentiated along topography [Schmitt et al. in prep]. 
Individual genotypes explained a significant variation in individual neighbor crowding (NCI, $R_m^2−R_c^2=0.18$) and individual growth potential (Gmax, $R_m^2−R_c^2=0.43$, Fig. \@ref(fig:VarPart)).
Spatially, genotypic adaptation to neighbor crowding ($gNCI$) and individual growth potential ($gGmax$) were structured up to 47m ($Moran’s~I, p−value<0.01$, Fig. \@ref(fig:SpCor)).
Moreover, a negative significant correlation (Pearson's $R=−0.3$, $p=4.5.10^{−9}$) was observed between genotypic adaptation to neighbourhood crowding ($gNCI$) and individual growth potential ($gGmax$) across diameter classes (Fig. \@ref(fig:gCor)), with a stronger signal for intermediate diameter classes than small or large individuals.

Following our hypothesis, we found that individual tree genotypes are differentially adapted to regenerate and thrive in response to the fine spatio-temporal dynamics of forest gaps. 
A genetic trade-off exists between intra-specific adaptive strategies for competition vs. for growth. 
The genetic trade-off is spatially structured in the area where a tree falls.
The genetic trade-off is also temporally structured with a stronger effect in the intermediate stages, than young stages under selection or immune older stages.
Therefore, species can grow in a diversity of successional niches, thanks to these genotypic adaptations to forest gap dynamics, decreasing overall the risk of a stochastic local extinction.
Consequently, forest gaps dynamics effect on genetic adaptations among tropical tree speciesis [@Herault2010; @Ruger2009] is conserved within species for the species complex *Symphonia globulifera*.
These results question the influence of forest gaps dynamics on genotypes in species complexes and more generally widespread species of tropical forest trees.

# Figures

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(tidyverse)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 4, fig.width = 6,
               cache = T, cache.lazy = F)
load(file = "../symcapture_save/outlines.Rdata")
```

```{r VarPart, fig.cap="Variance partitioning for neighbor crowding index (NCI) and individual maximum growth potential (Gmax). Variation of each variable has been partitioned into between-species (red), between-genotype (green), and residual (blue) variation."}
lapply(list(NCI = fitNCI, Gmax = fitGmaxGeno), 
       mcmc_intervals_data, regex_pars = c("Vp", "Vg", "Vr")) %>% 
  bind_rows(.id = "variable") %>% 
  mutate(parameter = recode(parameter, "Vp" = "Species", 
                            "Vg" = "Genotype", "Vr" = "Residual")) %>%
  mutate(variable = factor(variable, levels = c("NCI", "Gmax"))) %>% 
  group_by(variable) %>% 
  mutate(pct = paste0(round(m / sum(m) * 100), "%")) %>% 
  ggplot(aes(x = variable, fill = parameter)) +
  geom_col(aes(y = m)) +
  geom_text(aes(y = m, label = pct), col = "white", position = position_stack(vjust = .5)) +
  facet_wrap(~ variable, scales = "free") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank(),
        legend.position = "bottom") +
  scale_fill_discrete(expression(sigma^2)) 
```

```{r SpCor, fig.cap="Spatial autocorrelogram for raw variable and genotypic values of neighbor crowding index (NCI) and individual maximum growth potential (Gmax). The spatial autoccorelogram shows Moran's I value for different distance classes with significant values represented by a filled points against transparent non-significant values. Lines and points colors represent neighbor crowding index (blue) and individual maximum growth potential (red). Linetype represents raw variable (dashed) and genotypic values (continuous)."}
cor %>% 
  filter(dist.class < 10^3) %>% 
  filter(variable %in% c("Gmax", "NCI", "gGmax", "gNCI")) %>% 
  mutate(type = ifelse(substr(variable, 1, 1) == "g", "genotype", "variable")) %>% 
  mutate(variable = gsub("^g", "", variable)) %>% 
  ggplot(aes(x = dist.class, y = coef, col = variable, linetype = type)) + 
  geom_hline(yintercept = 0) +
  geom_point(aes(alpha = p.value < 0.01)) + 
  geom_line() +
  scale_alpha_manual(guide = "none", values = c(0.3, 1)) + 
  scale_linetype("") +
  scale_color_discrete("") +
  scale_x_log10() +
  geom_vline(xintercept = c(27, 53), col = "lightgrey", linetype = "dashed") +
  geom_vline(xintercept = 150, col = "lightgrey", linetype = "dotted") +
  xlab("Distance (m)") + ylab("Moran\'s I") +
  theme(legend.position = "bottom")
```

```{r gCor, fig.cap="Relation between geontypic values for individual growth potential (Gmax) and neighbor wroding index (NCI) for different class of diameters. Regression lines represent a linear model of form y ~ x. Annotations gives for each diameter class the Pearson's R correlation coefficient and the associated p-value.", fig.width=8}
n <- 6
t %>% 
  arrange(DBHtoday) %>% 
  mutate(DBHclass = rep(1:n, each = ceiling(nrow(.)/n))[1:nrow(.)]) %>% 
  group_by(DBHclass) %>% 
  mutate(DBHclassCm = paste0(round(min(DBHtoday)), " - ", round(max(DBHtoday)), " cm")) %>% 
  ggplot(aes(aGmax, aNCI)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ DBHclassCm, scales = "free") +
  geom_smooth(method = "lm", col = "black") +
  scale_size_continuous(guide = "none") +
  ggpubr::stat_cor(method = "pearson", label.x = 0.5, label.y = 0.85, ) +
  xlab("Genotypic value for individual growth potential (Gmax)") +
  ylab("Genotypic value for neighbor crowding index (NCI)")
```

# References

