---
title: "**Like mother, like daughter: forest gap dynamics drives divergent adaptive growth strategies within tropical tree species**"
author: "*Bruno Hérault, Myriam Heuertz, Sylvain Schmitt, Niklas Tysklind*"
date: '`r Sys.Date()`'
output: 
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
  bookdown::word_document2:
    reference_docx: ./template/template.docx
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

**In tropical forests, natural gap dynamics shape ecosystem functioning through a wide variety of demographic responses of trees, each with their individual ecological strategies. Demographic responses to forest gap dynamics have long been studied between species, neglecting genetic variability within species. However, overlapping generations within species provide a mechanism to maintain the genetic variation necessary for temporally-variable but strong selection to act on specific life stages. Tropical trees are long-lived and produce many seedlings, whose growth and survival is determined by the dynamics of forest gaps. Here we provide evidence that mature-forest tree species can grow in a diversity of successional niches thanks to genotypic adaptive growth strategies to fine spatio-temporal dynamics of forest gaps. These diverse adaptive strategies allow to broaden species’ niches which decrease the overall risk of a stochastic local extinction.**

The heterogeneity of resource distribution in space and time defines fine-scale habitat structure where species and individuals can grow and coexist [@Lortie2004a; @weiher_assembly_1995].
For instance, topography drives the spatial distribution of water and nutrient availability in tropical forests [@Ferry2010]. 
Therefore, topography drives pervasive differentiation in habitat preference among species [@Engelbrecht2007]
and in functional responses among and within species [@Schmitt2020]. 
In natural tropical forests, competition for light due to forest gap dynamics [@Hubbell1999; @Molino2001]
is often greater than the effect of competition for water and nutrients, even in early successional stages [@VanBreugel2012].
Forest gap dynamics is characterised by successional niches that trigger a wide variety of demographic responses depending on the ecological  strategies of individuals and species [@Herault2010; @Ruger2009].
After a treefall, pioneer species with fast growth, thanks to productive leaf tissue and light wood, rapidly colonize gaps with high light availability and decreased competition from adult trees; then, late successional species with slow growth and more conservative tissue progressively establish in a more shaded environment [@Craven2015].

But “nothing in biology makes sense except in the light of evolution” [@Dobzhansky1973].
Topography and abiotic habitats have been shown to drive differential adaptation of species, 
in the case of adaptive radiations [@Paun2016; @Pillon2014],
and within species, with microgeographic adaptations within the dispersal neighbourhood of the organism [@Brousseau2015; @Richardson2014; @Team2013].
However, despite the strong impact of forest gap dynamics on tree ecology [@VanBreugel2012],
to our knowledge, no studies have explored the possible role of forest gap dynamics on adaptation within species.

The assumption that evolution and ecology play at very different times and spatial scales might have been an impediment to understanding the role of forest gap dynamics in the eco-evolutionary dynamics of species [@Pelletier2009].
An increased number of studies in the last couple of decades have shown that evolutionary processes can be fast enough to impact ecological processes on a contemporary time scale [@Hairston2005; @Rudman2017].
Theoretical work demonstrated that overlapping generations can maintain the genetic variation necessary for strong temporally variable selection to act on a specific life stage [@Ellner1994]
while long-lived stages in which selection is relaxed can act as a reservoir to maintain genetic diversity [@Chesson1983]. 
For instance, overlapping generations with oscillation between short high-recruitment periods and long population-decline periods explain the spatial genetic structure of the Neotropical tree *Jacaranda copaia* [@Jones2006].

In tropical forests, the dynamics of forest gaps strongly determines the survival and recruitment of young trees through competition for access to light;
whereas once in the canopy, adult trees with full access to light are free from the light-driven selection of forest gap dynamics. 
Adult trees can therefore act as a reservoir of genotypes against fine spatio-temporal dynamics of forest gaps.
Generations of tropical trees broadly overlap, 
for instance mature-forest tree species with a maturity starting around 20cm grow up to 1m of diameter at breast height [@Hardy2006].   
Tropical trees thus disperse many seedlings over time into a mosaic of light conditions that fluctuates spatially and temporally due to the dynamics of forest gaps. 
Individual offspring encounter mainly shady and sometimes bright environments in which they must grow to survive. 
Since forest gap dynamics opposes rapid growth in a bright environment to slow growth in a shaded environment, 
we expected that individual genotypes within species typical of mature forests would be adapted to divergent growth strategies between light and shaded environments. 
Adaptation to slow growth in shade would allow individuals to persist with conservative tissues; 
whereas adaptation to fast growth in a bright environment would allow individuals to outgrow other seedlings and maintain access to light through an acquisition strategy with cheaper tissues but with less defence against herbivores and pathogens. 
Consequently, we hypothesized that forest gap dynamics is a major driver of the spatial and temporal structuring of genotypic diversity within tropical tree species.

Following our hypothesis, we expected that the divergent adaptive growth strategies would be spatially structured in the area where a tree falls. 
We also expected the divergent adaptive growth strategies to be structured in time. 
Because of recent selection, genotypic adaptations to growth in different light conditions are expected to be more pronounced in the intermediate stage. Younger stages may be mostly affected by  very stochastic survival. 
Conversely, older stages have been recruited from diverse past successional niches and have reached the canopy, 
thus they are now free from the light-driven selection of forest gap dynamics.

In the present study, we assessed genotypic diversity within closely-related sympatric tree species belonging to the widespread tropical tree species complex *Symphonia globulifera*.
We addressed the fine-scale spatial and temporal genetic adaptations of individuals through differential growth strategies in response to forest gap dynamics.
Combining tree diameter censuses, indirect measures of light environment of the near past and present, and single nucleotide polymorphisms (SNPs), 
we used population genomics, environmental association analyses, genome wide association studies and growth modelling to address the following questions:

* Are individual genotypes structured by the light access gradient resulting from forest gap dynamics?
* Is the growth of individuals determined by genotypes?
* Can we identify a trade-off between genotypic adaptations to light access and growth?
* How are genotypic adaptations to light access and growth structured in space and across life stages?

*Quick methods*

We sampled leaf tissue from 402 individuals previously morphologically identified as *Symphonia globulifera* and Symphonia *sp. 1* in the Paracou field station,
where individuals have been censused for more than 30 years. 
We designed and conducted a targeted sequence capture experiment of genomic regions obtaining sequence information for each individual. 
We called, filtered and annotated single nucleotide polymorphisms (SNPs) of sequenced libraries against the references used to build the experiment. 
We analyzed population genetic structure, genetic diversity, and demographic history. 
We then measured association of species and genotypes to mean neighbourhood crowding, an indirect measurement of access to light,
and individual growth potential over the last 30 years with Bayesian inferences of animal and growth models. 
Finally, we investigated the relationships between genotypic associations to neighbourhood crowding and individual growth and their structure in space and time with respectively spatial autocorrelogram and decomposed relations over diameter classes.

*Main results*

*Symphonia globulifera* was locally composed of three sympatric species corresponding to three morphotypes with differentiated topographic niches (Schmitt et al., in prep). 
Individual genotypes explained a significant variation in individual neighbourhood crowding (NCI, $R_m^2-R_c^2=0.18$)
and individual growth potential (Gmax, $R_m^2-R_c^2=0.43$, Fig. \@ref(fig:VarPart)),
indicating an adaptive genetic response of *Symphonia* genotypes to access to light and an adaptive component to growth potential. 
Genotypic adaptations to neighbourhood crowding (NCI) and individual growth potential (Gmax) 
were spatially structured up to 47m ($Moran’s~I, p-value<0.01$, Fig. \@ref(fig:SpCor)).
Moreover, as predicted by our hypothesis, 
intermediate life stages showed a stronger signature of adaptive response to light access through their growth strategy: 
a significant negative correlation (Pearson's $R=-0.3$, $p=4.5.10^{-9}$)
was observed between genotypic values  for neighbourhood crowding (NCI) and those for individual growth potential (Gmax) across diameter classes (Fig. \@ref(fig:gCor)),
with a stronger and a more significant signal for individuals in intermediate diameter classes than for individuals with small or large diameters. 
As a result, individual genotypes within *Symphonia* species are adapted to a trade-off between rapid growth potential and a preference for high light 
access, as opposed to slow growth potential and low light tolerance, structured in space and in diameter classes.

*Quick discussion*

In agreement with our hypothesis, we found that individual tree genotypes are differentially adapted to regenerate and thrive in response to the fine spatio-temporal dynamics of forest gaps.
A genetic trade-off exists between divergent adaptive growth strategies under different light conditions.
The genetic trade-off is spatially structured in the area where a tree falls. 
The genetic trade-off is also temporally structured with a stronger effect in the intermediate stages,
than in young stages under diverse selection pressures or in older stages recruited in diverse past environments and free from the light-driven selection of forest gap dynamics.
Therefore, a species uses differentiated growth strategies that allow its individuals to grow in a diversity of successional niches,
thanks to these divergent adaptive growth strategies to forest gap dynamics, decreasing overall the risk of a stochastic local extinction. 
Consequently, the effect of forest gap dynamics on genetic adaptations among tropical tree species [@Herault2010; @Ruger2019] is conserved within species in the species complex *Symphonia globulifera*.
Our results observed in a complex containing three species question the influence of forest gap dynamics on genotypes in species complexes and more generally tree species of tropical mature forests.

# References

<div id="refs"></div>

\newpage 

<!-- # Figures -->

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(tidyverse)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 4, fig.width = 6,
               cache = T, cache.lazy = F)
load(file = "../symcapture/symcapture_save/outlines.Rdata")
```

```{r VarPart, fig.cap="Variance partitioning for neighbourhood crowding index (NCI), an indirect measurement of access to light, and individual maximum growth potential (Gmax). Variation of each variable has been partitioned into between-species (red), between-genotype (green), and residual (blue) variation."}
lapply(list(NCI = fitNCI, Gmax = fitGmaxGeno), 
       mcmc_intervals_data, regex_pars = c("Vp", "Vg", "Vr")) %>% 
  bind_rows(.id = "variable") %>% 
  mutate(parameter = recode(parameter, "Vp" = "Species", 
                            "Vg" = "Genotype", "Vr" = "Residual")) %>%
  mutate(variable = factor(variable, levels = c("NCI", "Gmax"))) %>% 
  group_by(variable) %>% 
  mutate(pct = paste0(round(m / sum(m) * 100), "%")) %>% 
  ggplot(aes(x = variable, fill = parameter)) +
  geom_col(aes(y = m)) +
  geom_text(aes(y = m, label = pct), col = "white", position = position_stack(vjust = .5)) +
  facet_wrap(~ variable, scales = "free") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank(),
        legend.position = "bottom") +
  scale_fill_discrete(expression(sigma^2)) 
```

\newpage 

```{r SpCor, fig.cap="Spatial autocorrelogram for raw variable and genotypic values of neighbourhood crowding index (NCI), an indirect measurement of access to light, and individual maximum growth potential (Gmax). The spatial autocorrelogram shows Moran’s I value for different distance classes with significant values represented by filled points against transparent non-significant values. Colours of lines and points represent neighbourhood crowding index (blue) and individual maximum growth potential (red). Linetype represents raw variable (dashed) and associated genotypic values (continuous)."}
cor %>% 
  filter(dist.class < 10^3) %>% 
  filter(variable %in% c("Gmax", "NCI", "gGmax", "gNCI")) %>% 
  mutate(type = ifelse(substr(variable, 1, 1) == "g", "genotype", "variable")) %>% 
  mutate(variable = gsub("^g", "", variable)) %>% 
  ggplot(aes(x = dist.class, y = coef, col = variable, linetype = type)) + 
  geom_hline(yintercept = 0) +
  geom_point(aes(alpha = p.value < 0.01)) + 
  geom_line() +
  scale_alpha_manual(guide = "none", values = c(0.3, 1)) + 
  scale_linetype("") +
  scale_color_discrete("") +
  scale_x_log10() +
  geom_vline(xintercept = c(27, 53), col = "lightgrey", linetype = "dashed") +
  geom_vline(xintercept = 150, col = "lightgrey", linetype = "dotted") +
  xlab("Distance (m)") + ylab("Moran\'s I") +
  theme(legend.position = "bottom")
```

\newpage 

```{r gCor, fig.cap="Relation between genotypic values for individual growth potential (Gmax) and neighbourhood crowding index (NCI), an indirect measurement of access to light, for different classes of diameters. Regression lines represent a linear model of form y ~ x. Annotations give for each diameter class the Pearson’s R correlation coefficient and the associated p-value.", fig.width=8}
n <- 6
t %>% 
  arrange(DBHtoday) %>% 
  mutate(DBHclass = rep(1:n, each = ceiling(nrow(.)/n))[1:nrow(.)]) %>% 
  group_by(DBHclass) %>% 
  mutate(DBHclassCm = paste0(round(min(DBHtoday)), " - ", round(max(DBHtoday)), " cm")) %>% 
  ggplot(aes(aGmax, aNCI)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ DBHclassCm, scales = "free") +
  geom_smooth(method = "lm", col = "black") +
  scale_size_continuous(guide = "none") +
  ggpubr::stat_cor(method = "pearson", label.x = 0.5, label.y = 0.85, ) +
  xlab("Genotypic value for individual growth potential (Gmax)") +
  ylab("Genotypic value for neighbor crowding index (NCI)")
```
