---
title: "Symphonia growth"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(flexdashboard)
library(rgdal)
library(raster)
library(leaflet)
library(starmie)
library(rstan)
library(bayesplot)
library(abind)
library(tidyverse)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
```

# Intro

```{r data}
source("~/Documents/BIOGECO/PhD/scripts/dbh_correction.R")
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
trees <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv'),
                  dec=",", stringsAsFactors = F)
trees <- trees %>% # dbh correction
  filter(n_parcelle %in% 1:16) %>% 
  mutate(treatment = c(rep(c('C', 'T1', 'T2', 'T3'), 3), rep('B', 4))[n_parcelle]) %>% 
  mutate(dbh = circonf/pi) %>% 
  group_by(idArbre) %>% 
  mutate(dbh_c = correction(dbh, campagne, code_vivant, code_mesure)) %>% 
  ungroup()
trees <- trees %>% # tree selection
  filter(treatment != 'B') %>% 
  arrange(campagne) %>% 
  group_by(idArbre) %>% 
  filter(any(campagne == 1988)) %>% 
  filter(any(campagne == 1992)) %>% 
  filter(campagne %in% c(1988, 1992)) %>% 
  mutate(growth = diff(dbh_c)) %>% 
  filter(growth > 0) %>% 
  filter(campagne == 1988) %>% 
  select(idArbre, n_parcelle, n_carre, n_arbre, Xutm, Yutm, dbh_c, growth)
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
# global crs definition
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
# gaps
gaps <- shapefile("/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/logging/Gaps.shp")
gaps <- spTransform(gaps, CRSobj = crs)
gaps$area <- area(gaps)
# trees
treesXY <- trees
coordinates(treesXY) <- ~Xutm + Yutm
proj4string(treesXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
treesXY <- spTransform(treesXY, CRSobj = crs)
# gap distance
d <- raster(extent(treesXY), resolution = 10^-4, crs = crs)
d <- setValues(d, 0)
d <-  mask(d, gaps)
d <- distance(d)
trees$dgaps <- raster::extract(d, treesXY)
D <- rgeos::gDistance(spTransform(gaps, "+proj=utm +zone=22 +units=m"),
                      spTransform(treesXY, "+proj=utm +zone=22 +units=m"),
                      byid = T)
mdata <- list(I = dim(trees)[1], # Nb of trees
              growth = trees$growth, # growth vector
              dbh = trees$dbh_c/max(trees$dbh_c), # dbh in 1988 vector
              J = dim(gaps)[1], # Nb of gaps
              S = gaps$area/max(gaps$area), # gaps surface vector
              D = D/max(D)) #tree-gaps distance matrix
mpath <- "/home/sylvain/Documents/BIOGECO/PhD/documents/Symphonia/symphonia_models/"
rm(crs, path, correction, detect)
```

## Models

> I wanted to test for an eventual effect of logging through light and disturbance on Symphonia individuals growth. I used control and treatment plots between 1988 and 1992. I kept only trees already present in 1988 and still alive in 1992, and calculate their growth during this time. I then tried to look at the effect of both distance to the closest logging gaps ($d_{gaps}$ in $m$), original dbh of inidividuals in 1998 ($dbh_{1998}$ in $cm$), and their interaction on growth ($growth$ in $cm$). I used a bayesian model following a log normal law for growth :
$$M_{log}:~~growth \sim log \mathcal{N}(\alpha*log(d_{gaps}+1) + \beta*dbh_{1998} + \gamma*log(d_{gaps}+1)*dbh_{1998}, \sigma) $$

> Then I used the model proposed by @Herault2010:
$$growth_i \sim \mathcal{N}(\mu*\sum_j^J(e^{-\alpha*d_{i,j}}*S_j^\beta);\sigma)~|~i\in[1:I]~,~j\in[1:J]$$
where the growth between 1988 and 1992 ($growth$ in $cm$) of an individual $i$ depend on its distance to logging gaps $j$ ($d_{i,j}$ in $m$) and the logging gap surface ($S_j$ in $m$). $\mu$, $\alpha$, and $\beta$ represents are the disturbance parameters. A nul $\alpha$ or $\beta$ indicates a nul effect of logging gaps distance or surface. That i declined in different versions :

### Table 1: Models summary.

```{r}
mtab <- data.frame(
  M = c('$M_{\\mu, \\alpha, \\beta}$', 
        '$M_{\\alpha, \\beta}$', 
        '$M_{\\mu, \\alpha, \\beta, \\omega}$'),
  Model = c(
    '$growth_i \\sim \\mathcal{N}(\\mu*\\sum_j^J(e^{-\\alpha*d_{i,j}}*S_j^\\beta);\\sigma)$',
    '$growth_i \\sim \\mathcal{N}(\\sum_j^J(e^{-\\alpha*d_{i,j}}*S_j^\\beta);\\sigma)$',
    '$growth_i \\sim \\mathcal{N}(\\mu*\\sum_j^J(e^{-\\alpha*d_{i,j}}*S_j^\\beta)+\\omega;\\sigma)$'
  )
)
mtab$M <- as.character(mtab$M)
mtab$Model <- as.character(mtab$Model)
row.names(mtab) <- c('Mmab', 'Mab', 'Mmabo')
kable(mtab, row.names = F)
```

## Map

### Map of Symphonia growth after logging (between 1988 and 1992) and distance to logging gaps. *dots size and label represents the growth in cm between 1988 and 1992*

```{r map}
leaflet() %>%
  addRasterImage(log(d+1),
                 opacity = 0.5) %>% 
  addCircles(data = treesXY, radius = ~growth, 
             label = paste("+", round(treesXY$growth,1), "cm")) %>% 
  addLegend(pal = colorNumeric("Spectral", values(log(d+1))), 
            values = values(log(d+1)), position = "bottomleft",
    title = "log of distance\nto logging gaps")
rm(d, treesXY)
```


# $M_{log}$

## Column 1 {data-width=200}

> The model seems to have correctly converged besides log likelyhood seems to be constrained to a maximal value (see markov chain plot), and parameters seems not much correlated besides a small link between $\beta$ and $\gamma$ (see parameters pairs plot). So if we consider the model as valid, parameter posterior distribution seems to indicate a **strong effect for the logarithm distance to the gap** ($\alpha$ distribution don't overlap 0 with a mean at $\alpha_m = 0.18$). The growth of Symphonia individuals being increased close to gaps by almost 14% ($e^{\alpha_m}=1.14$).

```{r M0}
mdatalog <- trees %>% filter(!is.na(dgaps))
# fit <- stan(file = file.path(mpath, 'M1.stan'),
#             data = list(N = dim(mdata)[1],
#                         dbh = mdatalog$dbh_c,
#                         dgaps = mdatalog$dgaps,
#                         growth = mdatalog$growth),
#             chains = 1)
# save(fit, file = file.path(mpath, 'M1.Rdata'))
load(file.path(mpath, 'M1.Rdata'))
pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('alpha', 'beta', 'gamma', 'sigma')]
```

### Model Predictions.

```{r M0 pred}
P <- function(dgaps, dbh, pars)
  rlnorm(1000, 
         meanlog = (pars[1]*log(dgaps+1) + pars[2]*dbh + pars[3]*log(dgaps+1)*dbh),
         sdlog = pars[4])
# pred<- apply(mdatalog[c('dgaps', 'dbh_c')], 1, function(o)
#   apply(as.matrix(fit)[,c('alpha', 'beta', 'gamma', 'sigma')], 1, function(p)
#     P(o[1], o[2], p)))
# pred <- data.frame(t(apply(pred, 2, function(x) quantile(x, probs = c(0.05, 0.95)))))
# pred$mean <- exp(pars[1]*log(mdatalog$dgaps+1) + pars[2]*mdatalog$dbh_c + pars[3]*log(mdatalog$dgaps+1)*mdatalog$dbh_c)
# pred$gaps <- mdatalog$dgaps
# pred$dbh <- mdatalog$dbh_c
# names(pred)[1:2] <- c('5%', '95%')
# save(pred, file = file.path(mpath, 'M1p.Rdata'))
load(file.path(mpath, 'M1p.Rdata'))
ggplot(mdatalog, aes(dgaps, growth)) +
  geom_ribbon(aes(x = pred$gaps, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$gaps, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$gaps, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$gaps, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  xlab('distance to the closest logging gap (m)') +
  ylab('growth between 1988 and 1992 (cm)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
```

### Parameters markov chains.

```{r C0}
mcmc_trace(as.matrix(fit), facet_args = list(labeller = label_parsed))
```

## Column 2 {data-width=200}

> $$growth = e^{`r round(pars['alpha'],2)`*log(d_{gaps}+1) + `r round(pars['beta'],2)`*dbh_{1988} + `r round(pars['gamma'],2)`*log(d_{gaps}+1)*dbh_{1988}} $$

### Parameters posterior ditribution. *Light blue area represents 80% confidence interval, and vertical blue line the mean.*

```{r R0}
mcmc_areas(as.matrix(fit)[,1:3],  prob = 0.8)
```

### Parameters pairs plot. *Parameters density distribution, pairs plot and Pearson's coefficient of correlation.*

```{r P0}
GGally::ggpairs(data.frame(as.matrix(fit)[,1:3]))
```

# `r mtab['Mmab',1]`

```{r M2}
# fit <- stan(file = file.path(mpath, 'M2.stan'), data = mdata, chains = 1)
# save(fit, file = file.path(mpath, 'M2.Rdata'))
load(file.path(mpath, 'M2.Rdata')) #
pars <- c("mu", "alpha", "beta", "sigma")
pars_val <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),pars]
```

## Column 1 {data-width=200}

> 

### Model Predictions.

```{r M2 pred}
P <- function(d, s, dmax, smax, pars_val)
  rnorm(1000, 
         mean = pars_val[1] + sum(exp(-pars_val[2]*(d/dmax))*(s/smax)^pars_val[3]),
         sd = pars_val[4])
# pred<- sapply(seq_len(mdata$I), function(i)
#   apply(as.matrix(fit)[,pars], 1, function(p)
#     P(D[i,], gaps$area, max(D), max(gaps$area), p)))
# pred <- data.frame(t(apply(pred, 2, function(x) quantile(x, probs = c(0.05, 0.95)))))
# pred$mean <- sapply(seq_len(mdata$I), function(i)
#                     pars_val[1] + sum(exp(-pars_val[2]*(D[i,]/max(D)))*(gaps$area/max(gaps$area))^pars_val[3]))
# pred$gaps <- apply(D, 1, min)
# pred$dbh <- mdata$dbh
# names(pred)[1:2] <- c('5%', '95%')
# save(pred, file = file.path(mpath, 'M2p.Rdata'))
load(file.path(mpath, 'M2p.Rdata'))
ggplot(data.frame(dgaps = apply(D, 1, min), 
                  growth = trees$growth),
       aes(dgaps, growth)) +
  geom_ribbon(aes(x = pred$gaps, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$gaps, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$gaps, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$gaps, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  xlab('distance to the closest logging gap (m)') +
  ylab('growth between 1988 and 1992 (cm)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
```

### Parameters markov chains.

```{r C2}
mcmc_trace(as.matrix(fit)[,c(pars, "lp__")], facet_args = list(labeller = label_parsed))
```

## Column 2 {data-width=200}

> `r paste(mtab['Mmab',], collapse = ': ')`

### Parameters posterior ditribution. *Light blue area represents 80% confidence interval, and vertical blue line the mean.*

```{r R2}
mcmc_areas(as.matrix(fit)[,pars],  prob = 0.8)
```

### Parameters pairs plot. *Parameters density distribution, pairs plot and Pearson's coefficient of correlation.*

```{r P2}
GGally::ggpairs(data.frame(as.matrix(fit)[,pars]))
```


# `r mtab['Mab',1]`

```{r M3}
# fit <- stan(file = file.path(mpath, 'M3.stan'), data = mdata, chains = 1)
# save(fit, file = file.path(mpath, 'M3.Rdata'))
load(file.path(mpath, 'M3.Rdata'))
pars <- c("alpha", "beta", "sigma")
# pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('alpha', 'beta', 'gamma', 'sigma')]
```

## Column 1 {data-width=200}

### Parameters markov chains.

```{r C3}
mcmc_trace(as.matrix(fit)[,c(pars, "lp__")], facet_args = list(labeller = label_parsed))
```

## Column 2 {data-width=200}

> `r paste(mtab['Mab',], collapse = ': ')`

### Parameters posterior ditribution. *Light blue area represents 80% confidence interval, and vertical blue line the mean.*

```{r R3}
mcmc_areas(as.matrix(fit)[,pars],  prob = 0.8)
```

### Parameters pairs plot. *Parameters density distribution, pairs plot and Pearson's coefficient of correlation.*

```{r P3}
GGally::ggpairs(data.frame(as.matrix(fit)[,pars]))
```

# `r mtab['Mmabo',1]`

```{r M4}
# fit <- stan(file = file.path(mpath, 'M4.stan'), data = mdata, chains = 1)
# save(fit, file = file.path(mpath, 'M4.Rdata'))
load(file.path(mpath, 'M4.Rdata'))
pars <- c("mu", "alpha", "beta", "omega", "sigma")
# pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('alpha', 'beta', 'gamma', 'sigma')]
```

## Column 1 {data-width=200}

> 

### Parameters markov chains.

```{r C4}
mcmc_trace(as.matrix(fit)[,c(pars, "lp__")], facet_args = list(labeller = label_parsed))
```

## Column 2 {data-width=200}

> `r paste(mtab['Mmabo',], collapse = ': ')`

### Parameters posterior ditribution. *Light blue area represents 80% confidence interval, and vertical blue line the mean.*

```{r R4}
mcmc_areas(as.matrix(fit)[,pars],  prob = 0.8)
```

### Parameters pairs plot. *Parameters density distribution, pairs plot and Pearson's coefficient of correlation.*

```{r P4}
GGally::ggpairs(data.frame(as.matrix(fit)[,pars]))
```

# References
