---
title: "Symphonia"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(flexdashboard)
library(rgdal)
library(raster)
library(leaflet)
library(starmie)
library(rstan)
library(bayesplot)
library(abind)
library(tidyverse)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
```

Paracou
==================

Headcount {data-width=200}
-------------------------------------

> If we are only using Paracou Biodiversity and Control plots we lack of *Symphonia globulifera* morphotype to have a balanced stratified sampling ($\frac{99}{400}$ instead of $\frac{99}{400}$, see Table 1.1), and we only have 6 *S. indet* which means we can't hope they are in fact *S. globulifera*. Including treatment plots only allow us to reach 168 *S. globulifera* morphotype but with a lot of *S. indet* in which we might find other *S. globulifera* and thus reached a balanced sampling. We observed a strong relation to be tested between water table depth and *S. globulifera* presence (see Figure 1.1). So we may estimate with the layer the number of *S. globulifera* alive in Paracou. But including treatment plots necessitate to control for the light variable due to anthropic gaps of the logging. We thus need to test the effect of the treatments and gaps on individuals growth, mortality, and recruitment to evaluate the effect of the treatment on *Symphonia* Paracou population. So in parrallel we are looking at other Guyafor plots looking for *Symphonia globulifera* with both diameters inventories for at least three censuses and environmental data. 

### Table 1.1 Alive Symphonia headcounts in Paracou by species and treatment. B stand for Biodiviersity plots, C for control plots, and T1, T2, and T3 for treatment plots.

```{r Par_trees}
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
trees <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv'),
                  dec=",", stringsAsFactors = F)
trees <- trees %>% 
  filter(campagne == 2015) %>% 
  filter(code_vivant == 1) %>% 
  filter(code_mesure == 0) %>% 
  filter(n_parcelle != 18)
trees %>% 
  mutate(treatment = c(rep(c('C', 'T1', 'T2', 'T3'), 3), rep('B', 4))[n_parcelle]) %>% 
  dplyr::select(espece, treatment) %>% 
  reshape2::dcast(espece ~ treatment) %>% 
  mutate(Total = B + C + T1 + T2 + T3) %>% 
  kable()
```

Map {data-width=200}
-------------------------------------

### Figure 1.1 Symphonia distribution in Paracou.

```{r Par_map}
# global crs definition
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
# limits
limits <- readOGR(file.path(path, 'limits'), "OverallPlots", verbose = F)
limits <- spTransform(limits, CRSobj = crs)
# topography
topo <- readOGR(file.path(path, 'topography'), "ContourLinePlots", verbose = F)
topo <- spTransform(topo, CRSobj = crs)
# hydrology
hydro <- readOGR(file.path(path, 'hydrology'), "WaterTable", verbose = F)
hydro <- spTransform(hydro, CRSobj = crs)
hydro_pal <- colorFactor('Blues', as.factor(hydro$TypeWatEN), reverse = T)
# Trees to plot
treesXY <- trees
coordinates(treesXY) <- ~Xutm + Yutm
proj4string(treesXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
treesXY <- spTransform(treesXY, CRSobj = crs)
tree_pal <- colorFactor(c('darkgreen', 'black', 'firebrick'), as.factor(treesXY$espece))
# Plot
leaflet() %>%
  addPolylines(data = limits, group = 'Limits', color = 'black') %>%
  addPolylines(data = topo, group = 'Contour',
               color = 'grey') %>%
  addPolygons(data = hydro, group = 'WaterTable',
              opacity = 0, fillOpacity = 0.7, fill = T, 
              fillColor = ~hydro_pal(TypeWatEN), label = ~TypeWatEN) %>%
  addLayersControl(overlayGroups = c('Contour', 'Limits', 
                                     'WaterTable', 'Symphonia'),
                   options = layersControlOptions(collapsed = F)) %>% 
  addLegend(pal = hydro_pal, values = levels(as.factor(hydro$TypeWatEN)), 
            title = 'Water table', position = "bottomright") %>% 
  addCircles(data = treesXY, radius = ~circonf/pi/10, color = ~tree_pal(espece),
             label = paste("Symphonia", treesXY$espece, treesXY$n_arbre),
             group = 'Symphonia') %>% 
  addLegend(pal = tree_pal, values = levels(as.factor(treesXY$espece)), 
            title = 'Symphonia') 
rm(list = ls())
```

Guyafor
==================

Text {data-width=200}
-------------------------------------

> Table 2.1 shows few individuals in other Guyafor plots. It is due to the fact that most of Guyafor network plots include few bottom lands (Pascal Petronelli, personnal communication). Consequently it seems if we want both good quality diameter inventories and environment data we will have to sample all individuals in Paracou. The question stay to sample or not in treatment plots ? And more generally the question, is about to know if we will have sufficient environmental data (gaps map, lidar) to control anthropic and natural gaps effects ?

Headcount {data-width=200}
-------------------------------------

### Table 2.1 Symphonia headcounts among Guyafor network by species. Including all individuals, meaning dead ones too.

```{r Guy_trees}
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/trees/"
trees <- read.csv(file.path(path, 'Symphonia_Guyafor.csv'),
                  dec=",", stringsAsFactors = F)
trees %>% 
  dplyr::select(NomForet, espece, n_parcelle, n_carre, n_arbre) %>% 
  distinct() %>% 
  reshape2::dcast(NomForet ~ espece) %>% 
  arrange(desc(globulifera)) %>% 
  filter(globulifera > 0) %>% 
  kable()
rm(list = ls())
```

Genotypes
==================

Text {data-width=200}
-------------------------------------

> We gathered genetic material (ddRADseq) of french Guiana from Torroba-Balmori unpublished data (Paracou and Regina). We cleaned `fastq` files after a check with `fastQCheck` allowing us to correct two sequences by removing theim for individuals $PR_{49}$ and $RG_1$. We used `ipyrad ` for the interactive assembly of ddRADseq data sets on genotoul cluster (with denovo assembly, AATT and AT restriction overhang, 85% clustering threshold and a minimum of 48 sample per locus).

> We used `vcfR` to load SNPs data into R, and we transform it in genligh object for `adegenet`. We related indivdual IDs to their population and coordinates with *links* table. We coded population in 4 subset for *Symphonia globulifera* and *sp1* in both Paracou and Régina ($PR_{gl}$, $PR_{sp}$, $RG_{gl}$, $RG_{sp}$). Population definition was used to transform vcf file to structure file with PGDspider for further genetic structure analysis with STRUCTURE software. We corrected and transformed in UTM coordinates to compute kinship distance matrix with SPAGEDI.

```{r Gen_data}
# path <- "~/Documents/BIOGECO/PhD/data/GbS_Symphonia/assembly/symphoGbS2_outfiles/"
# data <- read.vcfR(file.path(path, 'symphoGbS2.vcf'), verbose = F)
# data <- vcfR2genlight(data)
# ploidy(data) <- 2
# save(data, file = './preliminary_save/assembly.Rdata')
load('./preliminary_save/assembly.Rdata')
ID0 <- data@ind.names
data@ind.names[which(data@ind.names %in% c('PR_49c', 'RG_1c'))] <- c('PR_49', 'RG_1')
path <- "~/Documents/BIOGECO/PhD/data/links/"
pop <- data.frame(ID0 = ID0, pop_num = data$ind.names) %>% 
  left_join(read_csv(file.path(path, 'Symphonia_ind_names.csv'))) %>% 
  left_join(read_csv(file.path(path, 'Symphonia_ind_coords.csv'),
                     col_types = cols(Latitude = col_character(), 
        Longitude = col_character())),
            by = c("ID" = "Ind_ID")) %>% 
  dplyr::select(ID0, pop_num, Pop, Species, Latitude, Longitude) %>%
  separate(Species, c("Genus", "species"), sep = " ") %>%
  mutate(species = substr(species, 0,2)) %>%
  mutate(Pop = paste0(Pop, '_', species)) %>%
  dplyr::select(ID0,pop_num, Pop, Latitude, Longitude) %>%
  rename(ID = pop_num) %>% 
  mutate(Latitude = as.numeric(gsub(',', '.', Latitude)),
         Longitude = as.numeric(gsub(',', '.', Longitude)))
rm(ID0)
pop(data) <- pop$Pop
```

### Figure 3.1 Symphonia population structure in Paracou.

```{r structure}
path <- '~/Documents/BIOGECO/PhD/data/GbS_Symphonia/structure/results_f/'
data <- list.files(path, recursive = T, pattern = "_run1_f")
data <- c(data[-1], data[1])
data <- lapply(data, function(d) loadStructure(file.path(path, d)))
data <- do.call('structList', data)
plotMultiK(data, pop[c('ID', 'Pop')]) +
  scale_fill_brewer(palette = "Paired")
```

Map {data-width=200}
-------------------------------------

```{r GBS}
path <- '~/Documents/BIOGECO/PhD/data/GbS_Symphonia/structure/results_f/k2/'
data <- list.files(path)
data <- sample(data, 1)
data <- starmie::loadStructure(file.path(path, data))
path <- "~/Documents/BIOGECO/PhD/data/links/"
ref <- read_csv(file.path(path, 'Symphonia_ind_names_complete-1.csv')) %>% 
  filter(Pop == 'PR') %>%
  dplyr::select(ID, pop_num, X16)
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
trees <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv'),
                  dec=",", stringsAsFactors = F)
trees <- trees %>% 
  filter(campagne == 2015) %>% 
  filter(code_vivant == 1) %>% 
  filter(code_mesure == 0) %>% 
  filter(n_parcelle != 18)
data <- data$ancest_df %>% 
  filter(Pop %in% 1:2) %>% 
  dplyr::select(Label, `Cluster 1`) %>% 
  rename(Globu = `Cluster 1`) %>% 
  mutate(Label = gsub('c', '', Label)) %>% 
  left_join(ref, by = c('Label' = 'pop_num')) %>% 
  tidyr::separate(X16, c('Pop', 'n_parcelle', 'n_carre', 'n_arbre'), '-', convert = T) %>% 
  dplyr::select(-Pop) %>% 
  left_join(trees)
```

### Figure 3.2 Symphonia genotype distribution in Paracou.

```{r GbS_map}
# global crs definition
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
# limits
limits <- readOGR(file.path(path, 'limits'), "OverallPlots", verbose = F)
limits <- spTransform(limits, CRSobj = crs)
# topography
topo <- readOGR(file.path(path, 'topography'), "ContourLinePlots", verbose = F)
topo <- spTransform(topo, CRSobj = crs)
# hydrology
hydro <- readOGR(file.path(path, 'hydrology'), "WaterTable", verbose = F)
hydro <- spTransform(hydro, CRSobj = crs)
hydro_pal <- colorFactor('Blues', as.factor(hydro$TypeWatEN), reverse = T)
# Trees to plot
treesXY <- data[-which(is.na(data$X)),]
coordinates(treesXY) <- ~Xutm + Yutm
proj4string(treesXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
treesXY <- spTransform(treesXY, CRSobj = crs)
tree_pal <- colorNumeric("RdYlBu", domain = treesXY$Globu)
# Plot
leaflet() %>%
  addPolylines(data = limits, group = 'Limits', color = 'black') %>%
  addPolylines(data = topo, group = 'Contour',
               color = 'grey') %>%
  addPolygons(data = hydro, group = 'WaterTable',
              opacity = 0, fillOpacity = 0.7, fill = T, 
              fillColor = ~hydro_pal(TypeWatEN), label = ~TypeWatEN) %>%
  addLayersControl(overlayGroups = c('Contour', 'Limits', 
                                     'WaterTable', 'Symphonia'),
                   options = layersControlOptions(collapsed = F)) %>% 
  addLegend(pal = hydro_pal, values = levels(as.factor(hydro$TypeWatEN)), 
            title = 'Water table', position = "bottomright") %>% 
  addCircles(data = treesXY, radius = ~circonf/pi/10, color = ~tree_pal(Globu),
             label = paste("Symphonia", treesXY$espece, treesXY$n_arbre),
             group = 'Symphonia') %>% 
  addLegend(pal = tree_pal, values = treesXY$Globu, title = "Globulifera\nGene\nPool")
rm(list = ls())
```

Environment
==================

> For the moment habitat association are tested only with Water Table Depth but should further be tested for different environmental variable. We can see that *S. globulifera* is significantly positively associated to water table depth under 60 cm on most of cases (11 plots over 15 for morphotype and 2 plots for genotypes, knowing that the 4 lacking plots for morphotype are certainly due to too coarse rasterization of the habitat shapefile). And *S. globulifera* is also significantly negatively associated to water table depth over 100 cm (12 plots over 15 for morphotypes and 2 plots for genotypes). Last but not least *S. sp.1* does not show significant association for genotypes and only 3 for morphotype (2 positive association to water table depth over 100 cm and one negative association to water table depth under 60 cm). Finally, absence of a lot of significant association for genotypes might not be due to an absence of association but more to a lack of power. Indeed only individuals associated more than 90% to a genotype in STRUCTURE analysis were kept resulting in relatively few individuals.

Morphotype association {data-width=200}
-------------------------------------

```{r opening}
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
# global crs definition
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
# limits
limits <- readOGR(file.path(path, 'limits'), "OverallPlots", verbose = F)
limits <- spTransform(limits, CRSobj = crs)
limits <- subset(limits, TypeSub %in% c('Subplots', 'SubplotsP16'))
# hydrology
hydro <- readOGR(file.path(path, 'hydrology'), "WaterTable", verbose = F)
hydro <- spTransform(hydro, CRSobj = crs)
# trees
trees <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv'),
                  dec=",", stringsAsFactors = F) %>% 
  filter(campagne == 2015) %>% 
  filter(code_vivant == 1) %>% 
  filter(code_mesure == 0) %>% 
  filter(n_parcelle != 18)
treesXY <- trees
coordinates(treesXY) <- ~Xutm + Yutm
proj4string(treesXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
treesXY <- spTransform(treesXY, CRSobj = crs)
```

```{r functions}
# grid
make.grid <- function(limits, hydro, treesXY, plot){
  limits <- subset(limits, Plot == plot)
  sppt <- spsample(limits, 400, "regular")
  grd <- points2grid(sppt, tolerance=0.76, round=1)
  grd <- SpatialGrid(grd, proj4string = crs(limits))
  grd_id <- as.numeric((grd %over% limits)$Project)
  grd_id[!is.na(grd_id)] <- 1:length(grd_id[!is.na(grd_id)])
  grd$id <- grd_id
  grd$hydro <- (grd %over% hydro)$TypeWater
  treesXY$quadrat <- treesXY %over% grd %>% 
    select(id) %>% 
    unlist()
  treecount <- treesXY@data %>% 
    reshape2::dcast(quadrat ~ espece) %>% 
    filter(!is.na(quadrat))
  if(!('Indet.' %in% names(treecount)))
    treecount$Indet. <- 0
  if(!('sp.1' %in% names(treecount)))
    treecount$sp.1 <- 0
  names(treecount)[1] <- 'id'
  grd@data <- grd@data %>% 
    left_join(treecount)
  # plot(raster(grd, 'sp.1'))
  # plot(stack(grd))  
  return(grd)
}

# new habitat
newhab <- function(grd, x, y, res = 20){
  data <- grd@data %>% 
    filter(!is.na(id)) %>% 
    mutate_at(c('globulifera', 'Indet.', 'sp.1'), funs(ifelse(is.na(.), 0, .)))
  habmat <- matrix(as.numeric(data$hydro), nrow = res)
  newhab <- matrix(0,res,res)
  x_new <- c(1:res);y_new <- c(1:res)
  if(y==0 & x==0){x_new <- x_new; y_new <- y_new}
  if(y==0 & x>0) x_new <- c((res-x+1):res,1:(res-x))
  if(x==0 & y>0) y_new <- c((res-y+1):res,1:(res-y))
  if(x >0 & y>0) {
    x_new <- c((res-x+1):res,1:(res-x))
    y_new <- c((res-y+1):res,1:(res-y))
  }
  newhab <- habmat[y_new,x_new]
  newhab <- c(newhab)
  newhab <- levels(grd$hydro)[newhab]
  newhab <- data.frame(id = data$id[1:length(newhab)],
                       hydro = newhab)
  newgrd <- grd
  newgrd@data <- grd@data %>% 
    left_join(newhab, by = 'id') %>% 
    select(-hydro.x) %>% 
    rename(hydro = hydro.y)
  return(newgrd)
}

# count
count <- function(grd){
  res <- grd@data %>% 
    filter(!is.na(id)) %>% 
    filter(!is.na(hydro)) %>% 
    mutate_at(c('globulifera', 'Indet.', 'sp.1'), funs(ifelse(is.na(.), 0, .))) %>% 
    select(-id) %>% 
    group_by(hydro) %>% 
    summarise_all(sum)
  res <- data.frame(res)
  row.names(res) <- res$hydro
  return(as.matrix(res[-1]))
}

# torus
torus <- function(grd, res){
  grd_null <- do.call('abind',
                  c(sapply(c(1:(res-1)), function(y)
                 do.call('abind', 
                  c(sapply(c(1:(res-1)), function(x) count(newhab(grd, x, 0)), simplify = F),
                    along = 3)),
                 simplify = F),
                 along = 3))
  apply(abind(count(grd), grd_null, along = 3), c(1,2), 
        function(x) (rank(x)/length(x))[1])
}

# stars
stars.two.tail <- function(X){
  sapply(X, function(x){
     s <- 'n.s.'
  if(x < 0.025)
    s <- paste('-', '*')
  if(x < 0.005)
    s <- paste('-', '**')
  if(x < 0.0005)
    s <- paste('-', '***')
  if(x > 0.975)
    s <- paste('+', '*')
  if(x > 0.995)
    s <- paste('+', '**')
  if(x > 0.9995)
    s <- paste('+', '***')
  return(s) 
  })
}
```

### Torus test of morphotypes association to water table depth. *two.tail.test column summarizes test significativity*

```{r association test}
# grd <- lapply(1:15, function(p) make.grid(limits, hydro, treesXY, p))
# grd_res <- suppressWarnings(lapply(grd, torus, res = 20))
# save(grd_res, file = './preliminary_save/torus_morpho.Rdata')
load('./preliminary_save/torus_morpho.Rdata')
do.call('abind', c(grd_res, along = 3)) %>% 
  reshape2::melt(varnames = c('WTD', 'species', 'plot'),
                 value.name = 'pvalue') %>% 
  mutate(two.tail.test = stars.two.tail(pvalue)) %>% 
  filter(two.tail.test != 'n.s.') %>% 
  arrange(species, WTD, plot) %>% 
  kable()
```

Genotype association {data-width=200}
-------------------------------------

```{r opening 2}
path <- '~/Documents/BIOGECO/PhD/data/GbS_Symphonia/structure/results_f/k2/'
data <- list.files(path)
data <- sample(data, 1)
data <- starmie::loadStructure(file.path(path, data))
path <- "~/Documents/BIOGECO/PhD/data/links/"
ref <- read_csv(file.path(path, 'Symphonia_ind_names_complete-1.csv')) %>% 
  filter(Pop == 'PR') %>%
  dplyr::select(ID, pop_num, X16)
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
trees <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv'),
                  dec=",", stringsAsFactors = F)
trees <- trees %>% 
  filter(campagne == 2015) %>% 
  filter(code_vivant == 1) %>% 
  filter(code_mesure == 0) %>% 
  filter(n_parcelle != 18)
data <- data$ancest_df %>% 
  filter(Pop %in% 1:2) %>% 
  dplyr::select(Label, `Cluster 1`) %>% 
  rename(Globu = `Cluster 1`) %>% 
  mutate(Label = gsub('c', '', Label)) %>% 
  left_join(ref, by = c('Label' = 'pop_num')) %>% 
  tidyr::separate(X16, c('Pop', 'n_parcelle', 'n_carre', 'n_arbre'), '-', convert = T) %>% 
  dplyr::select(-Pop) %>% 
  left_join(trees)
# species morphotype to genotype
data <- data %>%
  filter(Globu > 0.9, Globu > 0.1) %>% 
  mutate(espece = ifelse(Globu > 0.9, 'globulifera', 'sp.1'))
treesXY <- data
coordinates(treesXY) <- ~Xutm + Yutm
proj4string(treesXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
treesXY <- spTransform(treesXY, CRSobj = crs)
```

### Torus test of genotype association to water table depth. *two.tail.test column summarizes test significativity*

```{r association test 2}
# grd <- lapply(unique(data$n_parcelle), function(p) make.grid(limits, hydro, treesXY, p))
# grd_res <- suppressWarnings(lapply(grd, torus, res = 20))
# save(grd_res, file = './preliminary_save/torus_morpho2.Rdata')
load('./preliminary_save/torus_morpho2.Rdata')
do.call('abind', c(grd_res, along = 3)) %>% 
  reshape2::melt(varnames = c('WTD', 'species', 'plot'),
                 value.name = 'pvalue') %>% 
  mutate(two.tail.test = stars.two.tail(pvalue)) %>% 
  filter(two.tail.test != 'n.s.') %>% 
  arrange(species, WTD) %>% 
  kable()
```

Functional
==================

> We still do not have BRIDGE raw data too compare intraindividual versus interindividual functional trait variations.

### Figure 5.1 Symphonia functional triat variation in BRIDGE data ($n=23$).

```{r BRIDGE}
path <- "~/Documents/BIOGECO/PhD/data/BRIDGE/"
data <- read.csv("~/Documents/BIOGECO/PhD/data/BRIDGE/bridge.csv",
         dec=",", stringsAsFactors = F) %>%
  filter(Genus == "Symphonia")
data %>%
  select_if(is.numeric) %>%
  dplyr::select(-X, -Y, -n_lvs, -surete, -taxon_code, -ID_arbre, -ID_ram, -Ht_sample) %>%
  reshape2::melt() %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() + 
  facet_wrap(~variable, scales = "free")
rm(list = ls())
```

Growth
==================

Column 1 {data-width=200}
-------------------------------------

```{r dbh correction}
source("~/Documents/BIOGECO/PhD/scripts/dbh_correction.R")
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
trees <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv'),
                  dec=",", stringsAsFactors = F)
trees <- trees %>% 
  filter(n_parcelle %in% 1:16) %>% 
  mutate(treatment = c(rep(c('C', 'T1', 'T2', 'T3'), 3), rep('B', 4))[n_parcelle]) %>% 
  mutate(dbh = circonf/pi) %>% 
  group_by(idArbre) %>% 
  mutate(dbh_c = correction(dbh, campagne, code_vivant, code_mesure)) %>% 
  ungroup()
```

```{r AGR}
trees <- trees %>% 
  group_by(idArbre) %>% 
  mutate(ddbh = (dbh_c - lag(dbh_c, default = first(dbh_c)))) %>% 
  mutate(dt = (campagne - lag(campagne, default = first(campagne)))) %>% 
  mutate(agr = ddbh/dt) %>% 
  filter(agr > 0) %>% 
  ungroup()
```

> I wanted to test for an eventual effect of logging through light and disturbance on Symphonia individuals growth. I used control and treatment plots between 1988 and 1992. I kept only trees already present in 1988 and still alive in 1992, and calculate their growth during this time. I then tried to look at the effect of both distance to the closest logging gaps ($d_{gaps}$ in $m$), original dbh of inidividuals in 1998 ($dbh_{1998}$ in $cm$), and their interaction on growth ($growth$ in $cm$). I used a bayesian model following a log normal law for growth :
$$growth \sim log \mathcal{N}(\alpha*log(d_{gaps}+1) + \beta*dbh_{1998} + \gamma*log(d_{gaps}+1)*dbh_{1998}, \sigma) $$
The model seems to have correctly converged besides log likelyhood seems to be constrained to a maximal value (see markov chain plot), and parameters seems not much correlated besides a small link between $\beta$ and $\gamma$ (see parameters pairs plot). So if we consider the model as valid, parameter posterior distribution seems to indicate a **strong effect for the logarithm distance to the gap** ($\alpha$ distribution don't overlap 0 with a mean at $\alpha_m = 0.18$). The growth of Symphonia individuals being increased close to gaps by almost 14% ($e^{\alpha_m}=1.14$).

### Map of Symphonia growth after logging (between 1988 and 1992) and distance to logging gaps. *dots size and label represents the growth in cm between 1988 and 1992*

```{r gap}
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
# global crs definition
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
# gaps
gaps <- shapefile("/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/logging/Gaps.shp")
gaps <- spTransform(gaps, CRSobj = crs)
gaps$area <- area(gaps)
# trees
treesXY <- trees %>% 
  filter(treatment != 'B') %>% 
  arrange(campagne) %>% 
  group_by(idArbre) %>% 
  filter(any(campagne == 1988)) %>% 
  filter(any(campagne == 1992)) %>% 
  filter(campagne %in% c(1988, 1992)) %>% 
  mutate(growth = diff(dbh_c)) %>% 
  filter(campagne == 1988) %>% 
  select(idArbre, n_parcelle, n_carre, n_arbre, Xutm, Yutm, dbh_c, growth)
coordinates(treesXY) <- ~Xutm + Yutm
proj4string(treesXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
treesXY <- spTransform(treesXY, CRSobj = crs)
# gap distance
d <- raster(extent(treesXY), resolution = 10^-4, crs = crs)
d <- setValues(d, 0)
d <-  mask(d, gaps)
d <- distance(d)
treesXY$dgaps <- raster::extract(d, treesXY)
leaflet() %>%
  addRasterImage(log(d+1),
                 opacity = 0.5) %>% 
  addCircles(data = treesXY, radius = ~growth, 
             label = paste("+", round(treesXY$growth,1), "cm")) %>% 
  addLegend(pal = colorNumeric("Spectral", values(log(d+1))), 
            values = values(log(d+1)), position = "bottomleft",
    title = "log of distance\nto logging gaps")
```

```{r M0}
mpath <- "/home/sylvain//Documents/BIOGECO/PhD/documents/Symphonia/preliminary_models/"
mdata <- treesXY@data %>%
  filter(!is.na(dgaps)) %>%
  filter(growth > 0) # Should be corrected before
# fit <- stan(file = file.path(mpath, 'M1.stan'),
#             data = list(N = dim(mdata)[1],
#                         dbh = mdata$dbh_c,
#                         dgaps = mdata$dgaps,
#                         growth = mdata$growth),
#             chains = 1)
# save(fit, file = file.path(mpath, 'M1.Rdata'))
load(file.path(mpath, 'M1.Rdata'))
pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('alpha', 'beta', 'gamma', 'sigma')]
```

### Model Predictions.
```{r M0 pred}
P <- function(dgaps, dbh, pars) exp(pars[1]*log(dgaps+1) + pars[2]*dbh + pars[3]*log(dgaps+1)*dbh)
pred <- apply(as.matrix(fit), 1, function(x)
  sapply(seq_len(dim(mdata)[1]), function(i)
    P(mdata$dgaps[i], mdata$dbh_c[i], c(x['alpha'], x['beta'], x['gamma']))
  ))
pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.05, 0.95)))))
names(pred) <- c('5%', '95%')
pred$mean <- exp(pars[1]*log(mdata$dgaps+1) + pars[2]*mdata$dbh_c + pars[3]*log(mdata$dgaps+1)*mdata$dbh_c)
pred$gaps <- mdata$dgaps
pred$dbh <- mdata$dbh_c
ggplot(mdata, aes(dgaps, growth)) +
  geom_ribbon(aes(x = pred$gaps, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$gaps, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$gaps, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$gaps, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  xlab('distance to the closest logging gap (m)') +
  ylab('growth between 1988 and 1992 (cm)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
```

Column 2 {data-width=200}
-------------------------------------

> $$growth = e^{`r round(pars['alpha'],2)`*log(d_{gaps}+1) + `r round(pars['beta'],2)`*dbh_{1988} + `r round(pars['gamma'],2)`*log(d_{gaps}+1)*dbh_{1988}} $$

### Parameters posterior ditribution. *Light blue area represents 80% confidence interval, and vertical blue line the mean.*

```{r R0}
mcmc_areas(as.matrix(fit)[,1:3],  prob = 0.8)
```

### Parameters markov chains.

```{r C0}
mcmc_trace(as.matrix(fit), facet_args = list(labeller = label_parsed))
```

### Parameters pairs plot. *Parameters density distribution, pairs plot and Pearson's coefficient of correlation.*

```{r P0}
GGally::ggpairs(data.frame(as.matrix(fit)[,1:3]))
rm(list = ls()) ; gc()
```

Growth 2
==================

Column 1 {data-width=200}
-------------------------------------

> To test the effect of gaps on individual tree growth inside *Symphonia* species complex I used the model proposed by @Herault2010:
$$growth_i \sim \mathcal{N}(\mu*\sum_j^J(e^{-\alpha*d_{i,j}}*S_j^\beta);\sigma)~|~i\in[1:I]~,~j\in[1:J]$$
where the growth between 1988 and 1992 ($growth$ in $cm$) of an individual $i$ depend on its distance to logging gaps $j$ ($d_{i,j}$ in $m$) and the logging gap surface ($S_j$ in $m$). $\mu$, $\alpha$, and $\beta$ represents are the disturbance parameters. A nul $\alpha$ or $\beta$ indicates a nul effect of logging gaps distance or surface.

```{r M2}
# fit <- stan(file = file.path(mpath, 'M2.stan'), data = mdata, chains = 1)
# save(fit, file = file.path(mpath, 'M2.Rdata'))
load(file.path(mpath, 'M2.Rdata'))
pars <- c("mu", "alpha", "beta", "sigma") #
# pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('alpha', 'beta', 'gamma', 'sigma')]
```

Column 2 {data-width=200}
-------------------------------------

>

### Parameters posterior ditribution. *Light blue area represents 80% confidence interval, and vertical blue line the mean.*

```{r R2}
mcmc_areas(as.matrix(fit)[,pars],  prob = 0.8)
```

### Parameters markov chains.

```{r C2}
mcmc_trace(as.matrix(fit)[,c(pars, "lp__")], facet_args = list(labeller = label_parsed))
```

### Parameters pairs plot. *Parameters density distribution, pairs plot and Pearson's coefficient of correlation.*

```{r P2}
GGally::ggpairs(data.frame(as.matrix(fit)[,pars]))
```


Growth 3
==================

Column 1 {data-width=200}
-------------------------------------
> To test the effect of gaps on individual tree growth inside *Symphonia* species complex I used the model proposed by @Herault2010:
$$growth_i \sim \mathcal{N}(\sum_j^J(e^{-\alpha*d_{i,j}}*S_j^\beta);\sigma)~|~i\in[1:I]~,~j\in[1:J]$$
where the growth between 1988 and 1992 ($growth$ in $cm$) of an individual $i$ depend on its distance to logging gaps $j$ ($d_{i,j}$ in $m$) and the logging gap surface ($S_j$ in $m$). $\alpha$, and $\beta$ represents are the disturbance parameters. A nul $\alpha$ or $\beta$ indicates a nul effect of logging gaps distance or surface.

```{r M3}
# fit <- stan(file = file.path(mpath, 'M3.stan'), data = mdata, chains = 1)
# save(fit, file = file.path(mpath, 'M3.Rdata'))
load(file.path(mpath, 'M3.Rdata'))
pars <- c("alpha", "beta", "sigma")
# pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('alpha', 'beta', 'gamma', 'sigma')]
```

Column 2 {data-width=200}
-------------------------------------

>

### Parameters posterior ditribution. *Light blue area represents 80% confidence interval, and vertical blue line the mean.*

```{r R3}
mcmc_areas(as.matrix(fit)[,pars],  prob = 0.8)
```

### Parameters markov chains.

```{r C3}
mcmc_trace(as.matrix(fit)[,c(pars, "lp__")], facet_args = list(labeller = label_parsed))
```

### Parameters pairs plot. *Parameters density distribution, pairs plot and Pearson's coefficient of correlation.*

```{r P3}
GGally::ggpairs(data.frame(as.matrix(fit)[,pars]))
```

Growth 4
==================

Column 1 {data-width=200}
-------------------------------------
> To test the effect of gaps on individual tree growth inside *Symphonia* species complex I used the model proposed by @Herault2010:
$$growth_i \sim \mathcal{N}(\mu*\sum_j^J(e^{-\alpha*d_{i,j}}*S_j^\beta) + \omega;\sigma)~|~i\in[1:I]~,~j\in[1:J]$$
where the growth between 1988 and 1992 ($growth$ in $cm$) of an individual $i$ depend on its distance to logging gaps $j$ ($d_{i,j}$ in $m$) and the logging gap surface ($S_j$ in $m$). $\mu$, $\alpha$, and $\beta$ represents are the disturbance parameters. A nul $\alpha$ or $\beta$ indicates a nul effect of logging gaps distance or surface.

```{r M4}
# fit <- stan(file = file.path(mpath, 'M4.stan'), data = mdata, chains = 1)
# save(fit, file = file.path(mpath, 'M4.Rdata'))
load(file.path(mpath, 'M4.Rdata'))
pars <- c("mu", "alpha", "beta", "omega", "sigma")
# pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('alpha', 'beta', 'gamma', 'sigma')]
```

Column 2 {data-width=200}
-------------------------------------

>

### Parameters posterior ditribution. *Light blue area represents 80% confidence interval, and vertical blue line the mean.*

```{r R4}
mcmc_areas(as.matrix(fit)[,pars],  prob = 0.8)
```

### Parameters markov chains.

```{r C4}
mcmc_trace(as.matrix(fit)[,c(pars, "lp__")], facet_args = list(labeller = label_parsed))
```

### Parameters pairs plot. *Parameters density distribution, pairs plot and Pearson's coefficient of correlation.*

```{r P4}
GGally::ggpairs(data.frame(as.matrix(fit)[,pars]))
```

References
==================

