---
title: "Sampling"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(flexdashboard)
library(rgdal)
library(raster)
library(leaflet)
library(starmie)
library(rstan)
library(bayesplot)
library(abind)
library(ggrepel)
library(tidyverse)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
```

# General

##

### To Do

* Order 50 mL centrifuge tubes (to replace those from the lab used for wood sample)
* Order Eppendorf PCR tube 0.5 mL  (to replace those from the lab used for latex samples)
* Book the car for following weeks next week (mail Carole and GHP)
* Buy/Find journal, touc, porte vue, enveloppes, plastic boxes

### Docs

* [Field protocol](https://docs.google.com/document/d/1yjYXe50mYG95mUjw61SujkdDBxzIDkbvcJbXg8w6fPo/edit?usp=sharing)
* [Fresh measurements protocol](https://docs.google.com/document/d/1S-0O5ndADJhtXlrtz6qgRfYZtfS9f1qOgvFlaEF6P8M/edit?usp=sharing)
* [Symphonia individuals](https://docs.google.com/spreadsheets/d/13BKXG1ETpTvQwcpYcGoTK9eTU002PSWa4Cjyox-OanA/edit?usp=sharing)
* [Symphonia leaves](https://docs.google.com/spreadsheets/d/1QycHaAmsbBGd76Ma088RXxvFrWT5c_9Md3Eqqob286Q/edit?usp=sharing)
* [Dawkins index](https://drive.google.com/file/d/0B-lAe8V-ri2SUVZrRGJpQUhUVkU/view?usp=sharing)

##

### Planning

* Starting from $30^{th}$ of October
* During 25 to 30 days
* End between $4^{th}$ of December and $11^{th}$ of December
* Sampling from 7 AM until 3 PM
* Fresh measurement and herbarium preparation to finsih the day

### Maps

* [Parcelle 1](./symphonia_save/maps/Parcelle1.png)
* [Parcelle 2](./symphonia_save/maps/Parcelle2.png)
* [Parcelle 3](./symphonia_save/maps/Parcelle3.png)
* [Parcelle 4](./symphonia_save/maps/Parcelle4.png)
* [Parcelle 5](./symphonia_save/maps/Parcelle5.png)
* [Parcelle 6](./symphonia_save/maps/Parcelle6.png)
* [Parcelle 7](./symphonia_save/maps/Parcelle7.png)
* [Parcelle 8](./symphonia_save/maps/Parcelle8.png)
* [Parcelle 9](./symphonia_save/maps/Parcelle9.png)
* [Parcelle 10](./symphonia_save/maps/Parcelle10.png)
* [Parcelle 11](./symphonia_save/maps/Parcelle11.png)
* [Parcelle 12](./symphonia_save/maps/Parcelle12.png)
* [Parcelle 13](./symphonia_save/maps/Parcelle13.png)
* [Parcelle 14](./symphonia_save/maps/Parcelle14.png)
* [Parcelle 15](./symphonia_save/maps/Parcelle15.png)
* [Parcelle 16](./symphonia_save/maps/Parcelle16.png)

# Samples

```{r trees}
path <- "/home/sylvain//Documents/BIOGECO/PhD/data/Paracou/"
trees <- read.csv(file.path(path, 'trees/Symphonia_Paracou.csv'),
                  dec=",", stringsAsFactors = F)
trees <- trees %>% 
  filter(campagne == 2015) %>% 
  filter(code_vivant == 1) %>% 
  filter(code_mesure == 0) %>% 
  filter(n_parcelle != 18) %>% 
  mutate(treatment = c(c('C', 'T1', 'T2', 'T3', 
                         'T2', 'C', 'T1', 'T3',
                         'T1', 'T2', 'C', 'T3'), 
                       rep('B', 4))[n_parcelle])
```

```{r env}
# global crs definition
crs <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
# limits
limits <- readOGR(file.path(path, 'limits'), "OverallPlots", verbose = F)
limits <- spTransform(limits, CRSobj = crs)
# plot(limits)
# contour
contour <- readOGR(file.path(path, 'topography'), "ContourLinePlots", verbose = F)
contour <- spTransform(contour, CRSobj = crs)
# plot(contour)
# water table
WT <- readOGR(file.path(path, 'hydrology'), "WaterTable", verbose = F)
WT <- spTransform(WT, CRSobj = crs)
WT_pal <- colorFactor('Blues', as.factor(WT$TypeWatEN), reverse = T)
# plot(WT, col = WT_pal(WT$TypeWatEN))
# topo
topo <- readOGR(file.path(path, 'hydrology'), "TopographicLevels", verbose = F)
topo <- spTransform(topo, CRSobj = crs)
topo_pal <- colorFactor('Set3', as.factor(topo$TypeTopoEN), reverse = T)
# plot(topo, col = topo_pal(topo$TypeTopoEN))
# gaps
gaps <- readOGR(file.path(path, 'logging'), "Gaps", verbose = F)
gaps <- spTransform(gaps, CRSobj = crs)
# plot(gaps, col = 'green')
# Trees to plot
treesXY <- trees
coordinates(treesXY) <- ~Xutm + Yutm
proj4string(treesXY) <- '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
treesXY <- spTransform(treesXY, CRSobj = crs)
tree_pal <- colorFactor(c('darkgreen', 'black', 'firebrick'), as.factor(treesXY$espece))
# plot(treesXY, col = tree_pal(treesXY$espece), pch = 16)
```

```{r interpolation}
trees <- trees %>% 
  mutate(WT = as.character((treesXY %over% WT)$TypeWatEN)) %>% 
  mutate(WT = ifelse(is.na(WT), 'unknwon', WT)) %>% 
  mutate(topo = as.character((treesXY %over% topo)$TypeTopoEN)) %>% 
  mutate(topo = ifelse(is.na(topo), 'unknwon', topo)) %>% 
  mutate(gap = as.numeric(!(is.na((treesXY %over% gaps)$Plot))))
```

```{r previous sampling}
path <- "/home/sylvain/Documents/BIOGECO/PhD/data/links/"
SymphoGrowth <- read.csv(file.path(path, 'Symphonia_SymphoGrowth.csv'),
                  dec=",", stringsAsFactors = F) %>% 
  mutate(individual = gsub('P', '', individual)) %>% 
  separate(individual, c("n_parcelle", "n_carre", "n_arbre" ), '-', convert = T) %>% 
  distinct() %>% 
  mutate(SymphoGrowth = 1)
GbS <- read.csv(file.path(path, 'Symphonia_Match_btw_Gbs_otherDatasets.csv'),
                  dec=",", stringsAsFactors = F) %>% 
  select(Code_Par) %>% 
  mutate(Code_Par = gsub('P', '', Code_Par)) %>% 
  separate(Code_Par, c("n_parcelle", "n_carre", "n_arbre" ), '-', convert = T) %>% 
  distinct() %>% 
  mutate(GbS = 1)
CROPS <- read.csv(file.path(path, 'Symphonia_CROPS.csv'),
                  dec=",", stringsAsFactors = F) %>% 
  mutate(individual = gsub('P', '', individual)) %>% 
  separate(individual, c("n_parcelle", "n_carre", "n_arbre" ), '-', convert = T) %>% 
  distinct() %>% 
  mutate(CROPS = 1)
DROUGHT <- read.csv(file.path(path, 'Symphonia_DROUGHT.csv'),
                  dec=",", stringsAsFactors = F) %>% 
  select(Plot, Carre, Arbre) %>% 
  rename(n_parcelle = Plot, n_carre = Carre, n_arbre = Arbre) %>% 
  distinct() %>% 
  mutate(DROUGHT = 1)
trees <- trees %>% 
  left_join(GbS) %>% 
  mutate(GbS = ifelse(is.na(GbS), 0, GbS)) %>% 
  left_join(SymphoGrowth) %>% 
  mutate(SymphoGrowth = ifelse(is.na(SymphoGrowth), 0, SymphoGrowth)) %>% 
  left_join(CROPS) %>% 
  mutate(CROPS = ifelse(is.na(CROPS), 0, CROPS)) %>% 
  left_join(DROUGHT) %>% 
  mutate(DROUGHT = ifelse(is.na(DROUGHT), 0, DROUGHT))
rm(SymphoGrowth, GbS, CROPS, DROUGHT)
```

##

> Original headcounts

### Species headcount

```{r species}
trees %>% 
  reshape2::dcast(espece ~ 1) %>%
  kable()
```

### Species and gaps headcount

```{r gap}
trees %>% 
  reshape2::dcast(gap ~ espece) %>%
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

### Species and water table depth headcount

```{r WT}
trees %>% 
  reshape2::dcast(WT ~ espece) %>% 
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

### Species and topography headcount

```{r topo}
trees %>% 
  reshape2::dcast(topo ~ espece) %>% 
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

### Species and treatment headcount

```{r treatment}
trees %>% 
  reshape2::dcast(treatment ~ espece) %>% 
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

###  Species and plot headcount

```{r plot}
trees %>% 
  reshape2::dcast(n_parcelle ~ espece) %>% 
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

##

```{r selection}
trees <- trees %>% 
  mutate(sel = 0) %>% 
  mutate(sel = ifelse(SymphoGrowth == 1, 1, sel)) %>% 
  mutate(sel = ifelse(GbS == 1, 1, sel)) %>% 
  mutate(sel = ifelse(CROPS == 1, 1, sel)) %>% 
  mutate(sel = ifelse(DROUGHT == 1, 1, sel)) %>% 
  mutate(sel = ifelse(espece == "globulifera", 1, sel)) %>% 
  mutate(sel = ifelse(gap == 1, 1, sel)) %>% 
  mutate(sel = ifelse(WT == "unknwon", 1, sel)) %>% 
  mutate(sel = ifelse(WT == "0 : accompanying water-table of the stream", 1, sel)) %>% 
  mutate(sel = ifelse(WT == "Watertable between 0 and 60 cm", 1, sel)) %>% 
  mutate(sel = ifelse(WT == "Watertable between 60 and 100 cm", 1, sel)) %>% 
  mutate(sel = ifelse(topo == "unknwon", 1, sel))
P <- row.names(trees[trees$topo == 'Plateau' & trees$sel != 1,])  
S <- row.names(trees[trees$topo == 'Slope' & trees$sel != 1,])
nb <- (400-sum(trees$sel))/2
P <- sample(P, floor(nb))
S <- sample(S, ceiling(nb))
trees[P,]$sel <- 1
trees[S,]$sel <- 1
treesXY@data$sel <- trees$sel
rm(nb, P, S)
```

> Number of trees selected: `r sum(trees$sel)`

### Selected species headcount

```{r speciessel}
trees %>% 
  filter(sel == 1) %>% 
  reshape2::dcast(espece ~ 1) %>%
  kable()
```

### Selected species and gaps headcount

```{r gapsel}
trees %>% 
  filter(sel == 1) %>% 
  reshape2::dcast(gap ~ espece) %>%
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

### Selected species and water table depth headcount

```{r WTsel}
trees %>% 
  filter(sel == 1) %>% 
  reshape2::dcast(WT ~ espece) %>% 
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

### Selected species and topography headcount

```{r toposel}
trees %>% 
  filter(sel == 1) %>% 
  reshape2::dcast(topo ~ espece) %>% 
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

### Selected species and treatment headcount

```{r treatmentsel}
trees %>% 
  filter(sel == 1) %>% 
  reshape2::dcast(treatment ~ espece) %>% 
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

###  Selected species and plot headcount

```{r plotsel}
trees %>% 
  filter(sel == 1) %>% 
  reshape2::dcast(n_parcelle ~ espece) %>% 
  mutate(Total = globulifera + Indet. + sp.1) %>% 
  kable()
```

# Maps

### Dynamic map of selected individuals

```{r selected}
sel_pal <-colorBin(c('transparent', 'black'), trees$sel)
leaflet() %>%
  addPolygons(data = limits, fillOpacity = 0,
              opacity = 0.5, col = 'black') %>% 
  addCircles(data = treesXY, col = 'black', fillOpacity = 1,
             fillCol = sel_pal(trees$sel), radius= ~circonf/25,
             label = ~as.character(n_arbre)) %>% 
  addLegend(colors = c('transparent', 'black'), 
            labels = c('not selected', 'selected'),
            title = 'Sampling')
```

```{r maps functions}
rotateProj = function(spobj, angle) {
    boxpts = SpatialPoints(t(bbox(spobj)), proj4string = CRS(proj4string(spobj)))
    boxLL = bbox(spTransform(boxpts, CRS("+init=epsg:4326")))
    llc = apply(boxLL, 1, mean)
    prj = paste0("+proj=omerc +lat_0=", llc[2], " +lonc=", llc[1], " +alpha=", 
        angle, " +gamma=0.0 +k=1.000000 +x_0=0.000 +y_0=0.000 +ellps=WGS84 +units=m ")
    CRS(prj)
}
find_angle <- function(spobj){
  spobj <- subset(spobj, Subplot == unique(spobj$Subplot)[1])
  spobj <- fortify(spobj) %>% 
    filter(long == min(long) | lat == min(lat)) %>% 
    select(long, lat) %>% 
    distinct()
  row.names(spobj) <- c('A', 'B')
  spobj['C',] <- apply(spobj, 2, min)
  Pab <- sqrt(sum(apply(spobj[c('A', 'B'),], 2, diff)^2))
  Pac <- sqrt(sum(apply(spobj[c('A', 'C'),], 2, diff)^2))
  Pbc <- sqrt(sum(apply(spobj[c('B', 'C'),], 2, diff)^2))
  angle <- acos((Pab^2 + Pac^2 - Pbc^2) / (2 * Pab * Pac))
  angle <- angle*180/pi
  angles <- c(angle, angle-90)
  angle <- angles[which(sapply(c(angle, angle-90), abs) == min(sapply(c(angle, angle-90), abs)))]
  return(angle)
}
make_map <- function(limits, treesXY, contour, p){
  limits_plot <-  subset(limits, Plot == p)
  if(p != 16)
    limits_plot <-  subset(limits_plot, TypeSub == "Subplots")
  else
    limits_plot <-  subset(limits_plot, TypeSub == "SubplotsP16")
  newProj <- rotateProj(limits_plot, find_angle(limits_plot))
  limits_plot <- spTransform(limits_plot, newProj)
  treesXY_plot <- subset(treesXY, n_parcelle == p)
  treesXY_plot <- spTransform(treesXY_plot, newProj)
  contour_plot <- spTransform(contour, newProj)
  contour_plot <- crop(contour_plot, extent(limits_plot))
  g <- ggplot() +
    geom_polygon(data = fortify(limits_plot),
                 aes(long, lat, group=group),
                 fill = "transparent", colour = "black", alpha = 0.2) +
    geom_polygon(data = fortify(contour_plot),
                 aes(long, lat, group=group),
                 colour = "grey", alpha = 0) +
    geom_point(data = as.data.frame(treesXY_plot),
               aes(x = Xutm, y = Yutm, size=circonf/pi*100,
                   col = sel)) +
    geom_text_repel(data = as.data.frame(treesXY_plot),
                    aes(Xutm, Yutm, 
                        label = n_arbre)) +
    ggtitle(paste('Parcelle', p)) +
    coord_equal() +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    scale_color_continuous(palette = colorBin( c('lightgrey', 'black'), domain = c(0,1))) +
    ggsn::north(fortify(limits_plot), symbol = 12, 
                scale = 0.05, anchor = c(x= 125, y = 140))
  return(g)
}
```

```{r maps}
maps <- sapply(sort(unique(trees$n_parcelle)), 
               function(x) make_map(limits, treesXY, contour, x),
               simplify = F)
names(maps) <- paste0('P', sort(unique(trees$n_parcelle)))
# lapply(maps, function(x)
#   ggsave(x, file = paste0(gsub(' ', '', x$labels$title), '.png'),
#          path = './symphonia_save/maps/',
#          width = 297, height = 420, unit = 'mm',
#          dpi = 300))
```
