---
title: The extractable water drives a leaf thinning of tropical trees during the rainy season
date: 'Journal? `r Sys.Date()` '
output:
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::word_document2:
    reference_docx: ./template/template.docx
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, fig.height = 6, fig.width = 8, cache = F, cache.lazy = F)
```

Authors (alphabetic):
Émilie Ducouret$^6$, 
Géraldine Derroire$^6$, 
Bruno Hérault$^{3,4,5}$,
Myriam Heuertz$^1$,
Sylvain Schmitt$^1$, 
Niklas Tysklind$^2$, 
...

*$^1$ Univ. Bordeaux, INRAE, BIOGECO, 69 route d’Arcachon, 33610 Cestas France;* 
*$^2$ INRAE, UMR EcoFoG (Agroparistech, CNRS, Cirad, Université des Antilles, Université de la Guyane), Campus Agronomique, 97310 Kourou, French Guiana;*
*$^3$ CIRAD, UPR Forêts et Sociétés, Yamoussoukro, Côte d'Ivoire;*
*$^4$ Forêts et Sociétés, Univ Montpellier, CIRAD, Montpellier, France;*
*$^5$ Institut National Polytechnique Félix Houphouët-Boigny, INP-HB, Yamoussoukro, Côte d'Ivoire;*
*$^6$ Cirad, UMR EcoFoG (Agroparistech, CNRS, INRAE, Université des Antilles, Université de la Guyane), Campus Agronomique, 97310 Kourou, French Guiana;*

* **Corresponding author**: Sylvain Schmitt, +33 6 49 19 32 63, sylvain.m.schmitt@gmail.com

\newpage 

# Abstract

**To be written**

## Keywords

Keyword 1; ...

# Introduction 

* Functional traits
* Leaf variation
* Weather effects
* Variation in leaf thickness + weather data + diameter data + bayesian modelling
* How does weather influence leaf thickness in tropical trees ?
* Hypothesis

# Material and Methods

## Study site

**Paracou.**

## Plant material

Four hundreds and one individuals of the *Symphonia* genus were sampled between October and December 2017.
Sampling overlapped dry and rainy seasons.
The genus *Symphonia* includes the well-recognised species *Symphonia globulifera* (Clusiaceae)
that include two locally recognized morphotypes,
*S. globulifera sensu stricto* (*S. globulifera* hereafter) and *Symphonia sp.1*.
The two morphotypes occurs in sympatry but in differentiated habitats, 
with *S. globulifera* preferentially growing in valley bottoms 
and *S. sp1* preferentially exploiting a variety of drier habitats [Schmitt et al. in prep; @Allie2015]. 
Individuals tres with diameter at breast height larger than 10 cm (DBH > 10 cm) were randomly selected across all plots,
spanning the natural distribution of tree diameters and topographic habitats.
For each tree five mature and healthy leaves were sampled at the top of the crown 
using a slingshot sampling device (BIG SHOT SHERILLtree), 
and kept in humidified ziplock bags with CO2-enriched air in darkness until measurement within the next six hours following a standard protocol [@Perez-Harguindeguy2013]. 
Access to light for each sampled tree was assessed using the Dawkins index [@Dawkins1958].

## Leaf thickness measurement

We measured leaf thickness (LT, $\mu m$) on all fresh leaves, without rehydratation, with a micrometer of precision 1 $\mu m$.
The average thickness of the leaf was calculated on three points on both sides of the blade, while avoiding the midrib.
The petiole was removed for all measurements.

## Leaf thickness descriptors

The realtive extractable water [REW, @Wagner2011] was selected among several interdependent abiotic descriptors of weather (Fig. S1).
**Further describe REW.**

The diameter at breast height (DBH, cm) was chosen to control for tree size [@OBrien1995; @Zhang2004]. 
DBH values of sampled individuals were retrieved from the 2017 inventory of the Paracou permanent plots.
We log-transformed DBH to have a linear effect of DBH on LT, 
because leaf thickness have been shown to follow a Michaelis-Menten trajectory with tree size [@Schmitt2020].

## Analyses

We used a linear mixed models with REW, the logarithm of DBH and species fixed effects on the intercept and the slope of REW combined with individuals random effect to explain leaf thickness among several tested models (Mat. Sup. S2).

Leaf thickness $LT_{l,i,s}$ for leaf $l$ belonging individual $i\in[1;I]$ in species $s\in[1;S]$ 
was inferred with a lognormal distribution with the following formula:

$$LT_{l,i,s} \sim \mathcal{logN}(log[\alpha_s+ \gamma_i+\beta_{REW}.REW_i+\beta_{DBH}.logDBH_i],\sigma)$$
$$with~\gamma_i \sim \mathcal{N}(0,\sigma_I)$$

where $\alpha_s$ is the mean leaf thickness of species $s$,
$\gamma_i$ is the individual random effect centered on 0 of variance $\sigma_I$,
and $\beta_{REW}$ and $\beta_{DBH}$ are the slope of REW and logDBH effects.
REW and the logartihm of DBH were normalised to ease model inference and enable comparison of covariates.
A Bayesian method was used to infer parameters using stan language [@Carpenter2017, stan code available in Model S1] and rstan package [@Team2018] in the R environment [@RCoreTeam2020].

# Results

DBH and REW showed both a significant but opposed effect on leaf thickness (Tab. \@ref(tab:pars) and Fig. \@ref(fig:preds)).
REW had a negative effect on leaf thickness (Fig. \@ref(fig:preds)) 
for both *S. globulifera* ($-45 \mu m$ of leaf thickness for one standard unit of REW, Tab. \@ref(tab:pars)) 
and *S. sp1* ($-41 \mu m$ of leaf thickness for one standard unit of REW, Tab. \@ref(tab:pars) and ).
DBH had a positive effect on leaf thickness ($31 \mu m$ of leaf thickness for one standard unit of DBH, Tab. \@ref(tab:pars)) .
In other words, increased extractible water drives leaf thinning and increased tree size drives leaves thckening.

```{r pars}
load("../thickness_save/fits.Rdata")
broom.mixed::tidyMCMC(fits$LogRxSDSI, 
                pars =  c("alpha", "beta_rew", "beta_dbh", "sigmaI", "sigma"),
                droppars = NULL, ess = F, rhat = F, estimate.method = "median") %>% 
  dplyr::select(term, estimate, std.error) %>% 
  mutate(term = recode(term,
                       "alpha[1]" = "$\\alpha_{S.~globulifera}$",
                       "alpha[2]" = "$\\alpha_{S.~sp 1}$",
                       "beta_rew[1]" = "$\\beta_{REW,S.~globulifera}$",
                       "beta_rew[2]" = "$\\beta_{REW,S.~sp 1}$",
                       "beta_dbh" = "$\\beta_{DBH}$",
                       "sigmaI" = "$\\sigma_I$",
                       "sigma" = "$\\sigma$")) %>% 
  kable(caption = "Parameters estimate.",
        col.names = c("Parameter", "Estimate", "Standard error")) 
```

```{r preds, message=FALSE, fig.cap="Model predictions."}
data <- read_tsv("../thickness_save/data.tsv") %>% 
  mutate(Species = ifelse(Species == "Indet.",
                          c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G",
                                                                      "sp.1" =  "S")], Species)) %>% 
  filter(!is.na(Species)) %>% 
  mutate(SpeciesNum = as.numeric(as.factor(Species))) %>% 
  mutate(IndNum = as.numeric(as.factor(idTree))) %>%
  mutate(Species = recode(Species,
                        "globulifera"= "S. globulifera",
                        "sp.1" = "S. sp1"))
# load("../thickness_save/fits.Rdata")
load("../thickness_save/fitLT.Rdata")
as.data.frame(fitLTLogRxSDSI, pars = "predicted") %>% 
  reshape2::melt(NULL) %>% 
  group_by(variable) %>% 
  summarise(LTm = median(value), LTq5 = quantile(value, 0.05), LTq95 = quantile(value, 0.95)) %>% 
  separate(variable, c("X1", "LeafNum", "SpeciesNum"), convert = T) %>% 
  dplyr::select(-X1) %>% 
  left_join(mutate(data, LeafNum = 1:n()) %>% 
              dplyr::select(LeafNum, REW)) %>% 
  left_join(dplyr::select(data, SpeciesNum, Species) %>% unique()) %>% 
  unique() %>% 
  ggplot(aes(x = REW, col = Species)) +
  geom_point(aes(y = LT), data = data, alpha = 0.3) +
  geom_ribbon(aes(ymin = LTq5, ymax = LTq95, group = Species), fill = "lightgrey", col = NA, alpha = 0.5) +
  geom_line(aes(y = LTm)) +
  scale_color_manual("", values = c("#0000EE", "#CD2626")) +
  theme(legend.position = "bottom", legend.text = element_text(face = "italic")) +
  xlab("Relative extractable water (REW)") +
  ylab("Leaf thickness (LT, µm)") +
  geom_text(y = 200, label = expression(R^2==0.45), col = "black", data = data.frame(REW = 0.5))
```

# Discussion

* The extractable water drives a leaf thinning of tropical trees during the rainy season.
* Why ?
* Increased thickness with closed stomata: "When the stomata close at low air humidity the water content of the apricot leaves increases. The stomata open at high air humidity in spite of a decrease in leaf water content." [@Schulze1972] ?

# Acknowledgements

We thank the University of Bordeaux for a PhD grant to Sylvain Schmitt.
We are grateful to Pascal Petronelli and the CIRAD inventory team for their work on tree inventories and botanical identification.
Special thanks go to Saint-Omer Cazal, Anne Baranger, Josselin Cazal, Ilke Gelaldi, Fabien Lehuede, Adeline Adam, Agathe Benfredj Zaleski, Numa Faucherre, and David Zipper for their assistance during sampling in Paracou station.
This study was partially funded by an Investissement d’Avenir grant of the ANR: CEBA (ANR-10-LABEX-0025).


# Authors’ contributions

**To be written**

# References

<div id="refs"></div>
