---
title: Supporting information
output:
  bookdown::word_document2: 
      reference_docx: ./template/template.docx
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, fig.height = 6, fig.width = 8, cache = F, cache.lazy = F)
```

Article title: **The extractable water drives a leaf thinning of tropical trees during the rainy season**

Authors (alphabetic): Émilie Ducouret, Géraldine Derroire, Bruno Hérault, Myriam Heuertz, Niklas Tysklind, Sylvain Schmitt, ...


The following Supporting Information is available for this article:

**Model S1.** Stan code for the linear mixed model.

**Fig. S\@ref(fig:cor).** Correlations among leaf thickness descriptors. 

**Fig. S\@ref(fig:lp).** Log-likelhood of teted models.

\newpage

# Model S1: Stan code for the linear mixed model.

```{stan, echo=T, eval=F, output.var=""}
data {
  int<lower=1> L ;
  int<lower=1> S ;
  int<lower=1> I ;
  vector[L] LT ;
  vector[L] REW ;
  vector[L] logDBH ;
  int<lower=1, upper=S> species[L] ;
  int<lower=1, upper=I> individual[L] ;
}
parameters {
  vector<lower=0>[S] alpha ;
  vector[I] gamma ;
  vector[S] beta_rew ;
  real beta_dbh ;
  real<lower=0> sigma ;
  real<lower=0> sigmaI ;
}
model {
  LT ~ lognormal(log(alpha[species] + gamma[individual] + beta_rew[species] .* REW + beta_dbh * logDBH), sigma) ;
  gamma ~ normal(0, sigmaI) ;
}
```

\newpage

```{r cor, fig.cap="Correlations among leaf thickness descriptors. Pearson's r below the diagonal and corresponding color above the diagonal."}
read_tsv("../thickness_save/data.tsv") %>% 
  dplyr::select(LT, WMA, DBH, T55, T30, Hr55, Hr30, Patm, windaspect, windspeed, 
                rain, rainday, rainfivedays, vpd55, PET, rainfall, REW) %>% 
  na.omit() %>%
  cor() %>% 
  corrplot::corrplot.mixed(tl.pos = "lt")
```

\newpage

```{r lp, fig.cap="Log-likelhood of teted models."}
load("../thickness_save/fits.Rdata")
lapply(fits, mcmc_intervals_data, pars = "lp__") %>% 
  bind_rows(.id = "model") %>% 
  filter(!(model %in% c("RxSDSI", "LogRxSDSI"))) %>% 
  mutate(model = recode(model,
                        "R"= "REW",
                        "RS" = "REW + Species",
                        "RxS" = "REW.Species",
                        "RxSS" = "REW.Species + Species",
                        "D"= "DBH",
                        "DS" = "DBH + Species",
                        "DxS" = "DBH.Species",
                        "DxSS" = "DBH.Species + Species",
                        "RxSDS" = "REW.Species + DBH + Species"
  )) %>% 
  mutate(low = ifelse(m < 0, "low", "up")) %>% 
  ggplot(aes(x = model, 
             xend = model,
             color = model, 
             fill = model)) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() +
  datatools::scale_x_reordered() +
  xaxis_title(F) +
  ylab("Log-likelihood") +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")
```
